---
title: "{{< var course.short >}}: Tips and Tricks"
execute:
  freeze: true # Done with AY-2025-FALL offering
---

```{r}
#| echo: false
#| purl: true
#| warning: false
library(yaml)
variables <- read_yaml("https://raw.githubusercontent.com/michaelweylandt/STA9750/refs/heads/main/_variables.yml")
course_repo  <- variables$course$repo
course_short <- variables$course$short
```

## Miscellaneous `RStudio` and `Git` Advice

1. <a name="visualeditor"></a>
   When you are writing `quarto` documents in `RStudio`, 
   **do not use** the "Visual" editor mode. It will seem
   tempting at first, but it will lead to difficult
   and hard to debug errors down the road. You should
   disable it in your `RStudio` settings (if possible)
   and include `editor: source` in the header of `quarto`
   documents to keep it from opening automatically. 
2. <a name="gitmisc"></a>
   `git` supports a huge variety of programming workflow
   and was originally designed to allow large teams of
   programmers to work together. This flexibility leads
   to a few "sharp edges", however. 
   
   You will have the best experience with `git` in this
   course if: 
   
   1) <a name="commitoften"></a>
      You `commit` early and often. `git` commits are
      essentially free (40 bytes - less than a text!) and
      give you the ability to roll back changes if you get
      into a bad situation. 
      
   2) <a name="singlecomputer"></a>
      You do all your work on a single computer. Keeping
      files and `git` synchronized across multiple machines
      is possible, but trickier. If, for whatever reason,
      you need to use multiple computers, reach out to
      course staff in office hours for advice.
      
      In particular, even though it may be tempting to edit files
      directly through the GitHub web interface, resist! You will
      get into situations where your local copy and the copy on
      GitHub are out of sync and you will have issues pushing
      your work onto GitHub. 
      
      (It is possible to fix this, but you'll have to learn a lot
      more about how `git` works, far beyond what is expected of
      you in this course.)
      
   3) <a name="pushoften"></a>
      You `git push` at the end of every work session. This
      is the stage when things are most likely to go wrong,
      so you want to be `push`ing early and often, giving
      you the best chance to spot and fix an issues early. 
      
   4) <a name="filesystem"></a>
      You work in a "regular" file system. Synchronization
      and back up tools like DropBox or OneDrive perform 
      magic behind the scenes that doesn't always play well
      with `git`. You don't need to use these tools with
      `git` since `git` (used properly and in conjunction
      with `GitHub`) will give you automatic history and 
      backups. 
      
   5) <a name="antivirus"></a>
      You avoid anti-virus issues. If you are on a Windows
      machine, tell your anti-virus software to ignore the
      directory containing your work for this course. Both
      `git` and `quarto` create and delete lots of little
      files as they execute; certain anti-virus tools
      interpret this as evidence of a cyber attack and
      attempt to lock the file system, leading to a variety
      of subtle and hard to explain errors. 
      
   6) <a name="gitadd"></a>
      You are careful about what files you `add` in `git`.
      You should generally only include your `qmd` files and
      the rendered outputs in the `docs` folder. The 
      `build_site` script included in [Mini-Project
      #00](./miniprojects/mini00.html) is helpful, but not
      bullet-proof, here. 

3. <a name="filenames"></a>
   Use the file names and directory structures specified in
   the course instructions. I use a variety of automated
   processing tools to coordinate this course and if you do
   not follow instructions closely, your assignments will
   not be processed. I will typically manually handle your
   submissions, but with a penalty applied to your grade.
   It is better for all parties if you avoid this. 
   
   The helper scripts below can be used to confirm whether
   your submissions are properly formatted. 
4. <a name="nospaces"></a>Avoid using file or directory names with spaces in 
   them. Some software, particularly on Windows computers, does not handle
   spaces in directory or file names properly. 

## Helper Scripts

The following helper scripts can be used to help
automate several students tasks in this course. To use
any of them, run

> `source("https://michael-weylandt.com/STA9750/load_helpers.R")`

and then run the name of the relevant script. 

### Mini-Projects

#### Start a New Mini-Project {#mp_start}

This function will create a new file for a mini-project and pre-populate it
with a basic skeleton and some licensing information. Note that this function
should only be used for the graded Mini-Projects (MP#01 to MP#04) and not for
MP#00. 

```{r}
#| label: mp_start
mp_start <- function(N, github_id){
    if(!require("yaml")) install.packages("yaml"); library(yaml)
    if(!require("gh")) install.packages("gh"); library(gh)
    if(!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
    if(!require("rvest")) install.packages("rvest"); library(rvest)
    if(!require("glue")) install.packages("glue"); library(glue)
    if(!require("httr2")) install.packages("httr2"); library(httr2)
    
    if(missing(N)){
        N <- menu(title="Which Mini-Project would you like to submit on GitHub?", 
                  choices=c(1, 2, 3, 4))
    }
    
    `%not.in%` <- Negate(`%in%`)
    
    if(N %not.in% c(1, 2, 3, 4)){
        stop("Mini-Project Number not Recognized")
    }
    
    
    fname <- glue("mp0{N}.qmd")
    
    
    if(file.exists(fname)){
        stop(paste0("A ", sQuote(fname), " file already exists in this directory."))
    }
    
    
    mp_url <- glue("https://michael-weylandt.com/STA9750/miniprojects/mini0{N}.html")
    
    if(missing(github_id)){
        github_id <- readline("What is your GitHub ID? ")
    }
    
    variables <- read_yaml("https://raw.githubusercontent.com/michaelweylandt/STA9750/refs/heads/main/_variables.yml")
    course_repo  <- variables$course$repo
    course_short <- variables$course$short
    
    if(!endsWith(getwd(), course_repo)){
        stop(paste("This command should be run in your", course_repo, "directory only."))
    }
    
    mp_skeleton_url <- "https://michael-weylandt.com/STA9750/mp_template.qmd"
    mp_instructions_url <- glue("https://raw.githubusercontent.com/michaelweylandt/STA9750/refs/heads/main/miniprojects/mini0{N}.qmd")
    
    mp_instructions <- readLines(mp_instructions_url)
    
    dash_lines <- which(str_equal(mp_instructions, "---"))
    
    mp_header <- mp_instructions[(dash_lines[1] + 1):(dash_lines[2]-1)]
    
    mp_variables <- yaml.load(mp_header)
    
    mp_skeleton <- mp_skeleton_url |> 
        readLines() |> 
        str_replace_all(fixed("{GITHUB_ID}"), github_id) |>
        str_replace_all(fixed("{N}"), N) |>
        str_replace_all(fixed("{MP_TITLE}"), mp_variables$mp_title) |>
        str_replace_all(fixed("{MP_SKILLS}"), mp_variables$mp_skills |> str_split(", ")  |> pluck(1) |> paste(collapse = "/")) |>
        str_replace_all(fixed("{MP_DATA}"), mp_variables$mp_application |> str_split(", ")  |> pluck(1) |> paste(collapse = "/"))
    
    writeLines(mp_skeleton, fname)
    
    if("tools:rstudio" %in% search()){
        cat("Opening miniproject file")
        invisible(rstudioapi::documentOpen(fname))
    } else {
        cat("You can now open", sQuote(fname), "and begin the mini-project.")
    }
}
```

#### Push Mini-Project to GitHub

`TODO`

#### Check Mini-Project Prior to Submission {#mp_submission_ready}

When you are about ready to submit a mini-project, the following script can be
used to make sure all relevant files have been uploaded. In particular, it 
verifies that the source `mp0N.qmd` and the rendered `docs/mp0N.html` files
are on GitHub, then it also checks for the `css`, `js` and image files, alerting
if these cannot be loaded. Note that this check is only 'first-order' (so a `css`
imported by another `css` file will not be verified) but this is enough to catch
most problems.

```{r}
#| label: mp_submission_ready
mp_submission_ready <- function(N, github_id){
    if(!require("yaml")) install.packages("yaml"); library(yaml)
    if(!require("gh")) install.packages("gh"); library(gh)
    if(!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
    if(!require("rvest")) install.packages("rvest"); library(rvest)
    if(!require("glue")) install.packages("glue"); library(glue)
    if(!require("httr2")) install.packages("httr2"); library(httr2)  
    
    if(missing(N)){
        N <- menu(title="Which Mini-Project would you like to check was properly submitted on GitHub?", 
                  choices=c(0, 1, 2, 3, 4))
    }
    
    mp_url <- glue("https://michael-weylandt.com/STA9750/miniprojects/mini0{N}.html")
    
    mp_text <- read_html(mp_url) |> html_element("#submission-text") |> html_text()
    
    if(missing(github_id)){
        github_id <- readline("What is your GitHub ID? ")
    }
    
    variables <- read_yaml("https://raw.githubusercontent.com/michaelweylandt/STA9750/refs/heads/main/_variables.yml")
    course_repo  <- variables$course$repo
    course_short <- variables$course$short
BASE_URL <- glue("https://{github_id}.github.io/{course_repo}/")

page_resp <- request(BASE_URL) |>
    req_url_path_append(glue("mp0{N}.html")) |>
    req_error(is_error = \(r) FALSE) |> 
    req_perform()

if(resp_is_error(page_resp)){
    stop(glue("I was unable to access your submission at {page_resp$url}. Please make sure that your content has been successfully uploaded to GitHub."))
}

cat(glue("I was able to access your submission at {page_resp$url}, now checking JS and CSS files."), "\n")

SUPPORT_FILES <- page_resp |> resp_body_html() |> html_elements("link") |> html_attr("href")

SUPPORT_STATUS <- map_lgl(SUPPORT_FILES, function(s){
    supp_resp_error <- request(BASE_URL) |> 
        req_url_path_append(s) |>
        req_error(is_error = \(r) FALSE) |>
        req_perform() |>
        resp_is_error()
    
    if(supp_resp_error){
        cat(glue("- I could not find the file {s} that your site attempts to load.\n  Please make sure docs/{s} has been uploaded via Git to your GitHub.\n"))
        cat("\n")
    }
    !supp_resp_error
}) 

if(!all(SUPPORT_STATUS)){
    stop("Necessary support files could not be verified. Please visually inspect all elements of your site before submission.")
    browseURL(page_resp$url)
}

    IMAGE_FILES <- page_resp |> resp_body_html() |> html_elements("img") |> html_attr("src")
    
    IMAGE_STATUS <- map_lgl(IMAGE_FILES, function(s){
        img_resp_error <- request(BASE_URL) |> 
            req_url_path_append(s) |>
            req_error(is_error = \(r) FALSE) |>
            req_perform() |>
            resp_is_error()
        
        if(img_resp_error){
            cat(glue("- I could not find the image file {s} that your site attempts to load.\n  Please make sure docs/{s} has been uploaded via Git to your GitHub.\n"))
            cat("\n")
        }
        !img_resp_error
    }) 
    
    if(!all(IMAGE_STATUS)){
        stop("Necessary image files could not be verified. Please visually inspect all elements of your site before submission.")
        browseURL(page_resp$url)
    }
    
    if(N == 0){
        raw_url <- glue("https://raw.githubusercontent.com/{github_id}/{course_repo}/refs/heads/main/index.qmd")
    } else {
        raw_url <- glue("https://raw.githubusercontent.com/{github_id}/{course_repo}/refs/heads/main/mp0{N}.qmd")
    }
    
    resp_raw <- request(raw_url) |> 
        req_error(is_error = \(r) FALSE) |>
        req_perform()
    
    if(resp_is_error(resp_raw)){
        ## Try again with 'master' instead of 'main' branch
        raw_url_master <- str_replace(raw_url, "main", "master")
        resp_raw <- request(raw_url_master) |> 
            req_error(is_error = \(r) FALSE) |>
            req_perform()
    }
    
    if(resp_is_error(resp_raw)){
        cat("I cannot find the source qmd document at", raw_url, "or", 
            raw_url_master, ".\n",
            "Please confirm it was correctly submitted and try again.\n",
            "This document is needed for automated code quality checks.\n")
        
        stop("Miniproject not ready for submission.")
    }
    
    cat("Congratulations! Your mini-project appears to be ready for submission!\n")
    cat("Visually verify it one more time before submission.\n")
    cat("Then run the following command to submit on GitHub.\n")
    cat("\n")
    cat(glue('source("https://michael-weylandt.com/STA9750/load_helpers.R"); mp_submission_create(N={N}, github_id="{github_id}")'), "\n")
    cat("Then upload a PDF of your submission to Brightspace.\n")
    invisible(TRUE)
}
```

#### Submit Mini-Project via GitHub Issue {#mp_submission_create}

The following functon will create the GitHub issue necessary to submit a 
mini-project: 

```{r}
#| label: mp_submission_create
mp_submission_create <- function(N, github_id){
    if(!require("yaml")) install.packages("yaml"); library(yaml)
    if(!require("gh")) install.packages("gh"); library(gh)
    if(!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
    if(!require("rvest")) install.packages("rvest"); library(rvest)
    if(!require("glue")) install.packages("glue"); library(glue)
    if(!require("httr2")) install.packages("httr2"); library(httr2)
    
    if(missing(N)){
        N <- menu(title="Which Mini-Project would you like to submit on GitHub?", 
                  choices=c(0, 1, 2, 3, 4))
    }
    
    mp_url <- glue("https://michael-weylandt.com/STA9750/miniprojects/mini0{N}.html")
    
    mp_text <- read_html(mp_url) |> html_element("#submission-text") |> html_text()
    
    if(missing(github_id)){
        github_id <- readline("What is your GitHub ID? ")
    }
    
    variables <- read_yaml("https://raw.githubusercontent.com/michaelweylandt/STA9750/refs/heads/main/_variables.yml")
    course_repo  <- variables$course$repo
    course_short <- variables$course$short
    
    title <- glue("{course_short} {github_id} MiniProject #0{N}")
    
    body <- mp_text |> str_replace("<GITHUB_ID>", github_id)
    
    r <- request("https://github.com/") |>
        req_url_path_append("michaelweylandt") |>
        req_url_path_append(course_repo) |>
        req_url_path_append("issues/new") |>
        req_url_query(title=title, 
                      body=body) 
    
    browseURL(r$url)
}
```

#### Verify Mini-Project Submission {#mp-submission-verify}

The following function will confirm that the submission issue is properly
formatted. 

```{r}
#| label: mp_submission_verify
mp_submission_verify <- function(N, github_id){
    if(!require("yaml")) install.packages("yaml"); library(yaml)
    if(!require("gh")) install.packages("gh"); library(gh)
    if(!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
    if(!require("rvest")) install.packages("rvest"); library(rvest)
    if(!require("glue")) install.packages("glue"); library(glue)
    if(!require("httr2")) install.packages("httr2"); library(httr2)  
    
    if(missing(N)){
        N <- menu(title="Which Mini-Project would you like to check was properly submitted on GitHub?", 
                  choices=c(0, 1, 2, 3, 4))
    }
    
    mp_url <- glue("https://michael-weylandt.com/STA9750/miniprojects/mini0{N}.html")
    
    mp_text <- read_html(mp_url) |> html_element("#submission-text") |> html_text()
    
    if(missing(github_id)){
        github_id <- readline("What is your GitHub ID? ")
    }
    
    variables <- read_yaml("https://raw.githubusercontent.com/michaelweylandt/STA9750/refs/heads/main/_variables.yml")
    course_repo  <- variables$course$repo
    course_short <- variables$course$short
    
    title <- glue("{course_short} {github_id} MiniProject #0{N}")
    
    body <- mp_text |> str_replace("<GITHUB_ID>", github_id) |> str_squish()
    
    page <- 1
    
    issues <- list()
    
    while((length(issues) %% 100) == 0){
        new_issues <- gh("/repos/michaelweylandt/{repo}/issues?state=all&per_page=100&page={page}", 
                         repo=course_repo, 
                         page=page)
        
        if(length(new_issues) == 0){
            break
        }
        
        issues <- c(issues, new_issues)
        page <- page + 1
    }
    
    issue_names <- vapply(issues, function(x) str_squish(x$title), "")
    
    name_match <- which(issue_names == title)
    case_insensitive_match <- which(tolower(issue_names) == tolower(title))
    
    if((length(name_match) == 0) & (length(case_insensitive_match) == 0)){
        cat("I could not find a unique issue with the title:\n", 
            "    ", sQuote(title),"\n",
            "The issues I found had the following titles:\n",
            paste(c("", issue_names), collapse="\n - "), "\n",
            "If something on that list looks correct, please check",
            "capitalization and punctuation.\n")
        stop("MINIPROJECT NOT SUBMITTED CORRECTLY.")
    } else if((length(name_match) == 0) & (length(case_insensitive_match) >= 0)) {
        cat("I have found a partial match (differing by case only).\n",
            "- Searching for: ", title, , "\n", 
            "- Found: ", issue_names[case_insensitive_match], ".\n",
            "Attempting to proceed with partial match issue")
        name_match <- case_insensitive_match
    } else if(length(name_match) > 1){
        cat("I found multiple issues with the title:\n", 
            "    ", sQuote(title),"\n",
            "Please change the names of issues so that there is only", 
            "issue with the desired name. (Note that closing an issue is",
            "not sufficient.)\n")
        
        stop("MINIPROJECT NOT SUBMITTED CORRECTLY.")
    }
    
    issue <- issues[[name_match]]
    
    issue_num <- issue$number
    issue_url <- issue$html_url
    issue_body <- issue$body |> str_squish()
    
    cat("Identified Issue #", issue_num, "at", issue_url, "as possible candidate.\n")
    
    if(issue$state != "open"){
        cat("Issue does not appear to be in 'open' status. Please",
            "confirm the issue is open and try again.\n")
        stop("MINIPROJECT NOT SUBMITTED CORRECTLY.")
    }
    
    if(!str_equal(issue_body, body, ignore_case = TRUE)){
        cat("Issue does not appear to exactly match the correct body text.",
            "This is not necessarily an issue since the expected link is present",
            "but you may want to confirm all relevant",
            "information is included.\n")
    }
    
    # This isn't quite general enough, but I think it covers everything
    # we'll see in this course. 
    URL_REGEX <- "http[A-Za-z0-9:/\\-\\.]*"
    
    urls <- str_match_all(paste(issue_body, "\n Thanks!\n", sep=""), 
                          URL_REGEX)[[1]] |> unique()
    expected_url <- str_extract(body, URL_REGEX)
    
    if(length(urls) > 1){
        cat("The following URLs were found in the issue text:",
            paste(c("", urls), collapse="\n - "), 
            "Only one URL was expected; please remove any extras.\n")
        stop("MINIPROJECT NOT SUBMITTED CORRECTLY.")
    }
    
    submitted_url <- urls[1]
    
    if(!str_equal(submitted_url, expected_url, ignore_case = TRUE)){
        cat("The submitted URL does not match the Mini-Project instructions.\n", 
            "Expected:\n - ", expected_url, "\n", 
            "Submitted:\n - ", submitted_url, "\n", 
            "Please correct any differences and try again.\n")
        stop("MINIPROJECT NOT SUBMITTED CORRECTLY.")
    }
    
    resp <- request(submitted_url) |> 
        req_error(is_error = \(r) FALSE) |>
        req_perform()
    
    if(resp_is_error(resp)){
        cat("Something appears to be incorrect at the URL: ", 
            submitted_url, 
            "\n Please confirm that it is working as expected and try again.")
        browseURL(submitted_url)
        stop("MINIPROJECT NOT SUBMITTED SUCCESSFULLY.")
    }
    
    if(N == 0){
        raw_url <- glue("https://raw.githubusercontent.com/{github_id}/{course_repo}/refs/heads/main/index.qmd")
    } else {
        raw_url <- glue("https://raw.githubusercontent.com/{github_id}/{course_repo}/refs/heads/main/mp0{N}.qmd")
    }
    
    resp_raw <- request(raw_url) |> 
        req_error(is_error = \(r) FALSE) |>
        req_perform()
    
    if(resp_is_error(resp_raw)){
        ## Try again with 'master' instead of 'main' branch
        raw_url_master <- str_replace(raw_url, "main", "master")
        resp_raw <- request(raw_url_master) |> 
            req_error(is_error = \(r) FALSE) |>
            req_perform()
    }
    
    if(resp_is_error(resp_raw)){
        cat("I cannot find the source qmd document at", raw_url, "or", 
            raw_url_master, ".\n",
            "Please confirm it was correctly submitted and try again.\n",
            "This document is needed for automated code quality checks.\n")
        
        stop("MINIPROJECT NOT SUBMITTED SUCCESSFULLY.")
    }
    
    cat("Congratulations! Your mini-project appears to have been submitted correctly!\n")
    invisible(TRUE)
}
```

### Peer Feedback

#### Find Peer Feedback Assigned to Me {#mp_feedback_locate}

```{r}
#| label: mp_feedback_locate
mp_feedback_locate <- function(N, github_id){
    if(!require("yaml")) install.packages("yaml"); library(yaml)
    if(!require("gh")) install.packages("gh"); library(gh)
    if(!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
    if(!require("rvest")) install.packages("rvest"); library(rvest)
    if(!require("glue")) install.packages("glue"); library(glue)
    if(!require("httr2")) install.packages("httr2"); library(httr2)
    
    if(missing(N)){
        N <- menu(title="Which Mini-Project's Peer Feedback cycle is it currently?", 
                  choices=c(0, 1, 2, 3, 4))
    }
    
    if(missing(github_id)){
        github_id <- readline("What is your GitHub ID? ")
    }
    
    page <- 1
    issues <- list()
    
    variables <- read_yaml("https://raw.githubusercontent.com/michaelweylandt/STA9750/refs/heads/main/_variables.yml")
    course_repo  <- variables$course$repo
    course_short <- variables$course$short
    
    while((length(issues) %% 100) == 0){
        new_issues <- gh("/repos/michaelweylandt/{repo}/issues?state=all&per_page=100&page={page}", 
                         repo=course_repo, 
                         page=page)
        
        if(length(new_issues) == 0){
            break
        }
        
        issues <- c(issues, new_issues)
        page <- page + 1
    }
    
    pf_urls <- issues |> 
        keep(~ str_detect(.x$title, glue("#0{N}"))) |>
        keep(~ !str_equal(.x$user$login, github_id, ignore_case=TRUE)) |>
        map(~data.frame(html=.x$html_url, comments=.x$comments_url, body=.x$body)) |>
        keep(~any(str_detect(map_chr(gh(.x$comments), "body"), github_id) | 
                  str_detect(.x$body, github_id)
                  )) |>
        map("html") |>
        list_c()
    
    cat(glue("I have found several MP#0{N} issues that may be assigned to you.\n"),
        "Please review the following:\n")
    for(pf in pf_urls){
        cat(" - ", pf, "\n")
    }
    
    to_browser <- menu(title="Would you like me to open these in your web browser?", 
                       choices=c("Yes", "No"))
    
    if(to_browser == 1){
        for(pf in pf_urls){
            browseURL(pf)
        }
    }
    invisible(TRUE)
}
```


#### Create Peer Feedback Comment {#mp_feedback_submit}

```{r}
#| label: mp_feedback_submit
mp_feedback_submit <- function(N, peer_id){
    if(!require("yaml")) install.packages("yaml"); library(yaml)
    if(!require("gh")) install.packages("gh"); library(gh)
    if(!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
    if(!require("rvest")) install.packages("rvest"); library(rvest)
    if(!require("glue")) install.packages("glue"); library(glue)
    if(!require("clipr")) install.packages("clipr"); library(clipr)
    if(!require("httr2")) install.packages("httr2"); library(httr2)
    
    if(missing(N)){
        N <- menu(title="Which Mini-Project's Peer Feedback would you like to check was properly submitted on GitHub?", 
                  choices=c(0, 1, 2, 3, 4))
    }
    
    if(missing(peer_id)){
        peer_id <- readline("What is your Peer's GitHub ID? ")
    }
    
    variables <- read_yaml("https://raw.githubusercontent.com/michaelweylandt/STA9750/refs/heads/main/_variables.yml")
    course_repo  <- variables$course$repo
    course_short <- variables$course$short
    
    title <- glue("{course_short} {peer_id} MiniProject #0{N}")
    
    page <- 1
    
    issues <- list()
    
    while((length(issues) %% 100) == 0){
        new_issues <- gh("/repos/michaelweylandt/{repo}/issues?state=all&per_page=100&page={page}", 
                         repo=course_repo, 
                         page=page)
        
        if(length(new_issues) == 0){
            break
        }
        
        issues <- c(issues, new_issues)
        page <- page + 1
    }
    
    issue_names <- vapply(issues, function(x) str_squish(x$title), "")
    
    name_match <- which(issue_names == title)
    
    if(length(name_match) != 1){
        cat("I could not find a unique issue with the title:\n", 
            "    ", sQuote(title),"\n",
            "I'm afraid I can't verify whether the peer feedback was submitted properly\n", 
            "but the issue likely lies with the submittor, not with your feedback.")
        stop("PEER FEEDBACK NOT VERIFIED.")
    } 
    
    issue <- issues[[name_match]]
    
    issue_url <- issue$html_url
    
    template_url <- "https://michael-weylandt.com/STA9750/miniprojects.html"
    
    template_text <- read_html(template_url) |> 
        html_element("#peer-feedback-template") |> 
        html_text() |>
        str_trim() 
    
    template_categories <- c(
        "Written Communication", 
        "Project Skeleton", 
        "Formatting & Display", 
        "Code Quality", 
        "Data Preparation", 
        "Extra Credit"
    )
    
    overall <- readline("Overall Comments: (Hit enter if no comments) ")
    
    template_text <- template_text |> str_replace("OPTIONAL TEXT", overall)
    
    for(category in template_categories){
        has_score <- FALSE
        
        while(!has_score){
            score <- readline(glue("On a scale of 0 to 10, how many points does {peer_id} deserve for {category}? "))
            
            score <- as.integer(score)
            
            if(!is.na(score) & (score >= 0) & (score <= 10)){
                has_score <- TRUE
            }
        }
        
        has_score <- FALSE
        
        text <- readline(glue("Why did you give {peer_id} {score} points for {category}? "))
        
        template_text <- template_text |>
            str_replace("NN", as.character(score)) |>
            str_replace("TEXT TEXT TEXT", as.character(text))
    }
    
    write_clip(template_text)
    
    readline(paste0(
        "After you hit Enter / Return, you will be taken to a GitHub page.\n",
        "Paste from the clipboard into the comment box at the bottom to\n",
        "pre-populate your peer feedback. Then hit 'Comment' to finish your\n",
        "submission. "))
    
    browseURL(issue_url)
    
    cat(glue("You can now use `mp_feedback_verify({N}, peer_id={peer_id})`\n",
             "to confirm your submission was properly formatted.\n"))
}
```


#### Verify Peer Feedback Properly Formatted {#mp_feedback_verify}

```{r}
#| label: mp_feedback_verify
mp_feedback_verify <- function(N, github_id, peer_id){
    if(!require("yaml")) install.packages("yaml"); library(yaml)
    if(!require("gh")) install.packages("gh"); library(gh)
    if(!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
    if(!require("rvest")) install.packages("rvest"); library(rvest)
    if(!require("glue")) install.packages("glue"); library(glue)
    
    if(missing(N)){
        N <- menu(title="Which Mini-Project's Peer Feedback would you like to check was properly submitted on GitHub?", 
                  choices=c(0, 1, 2, 3, 4))
    }
    
    if(missing(github_id)){
        github_id <- readline("What is your GitHub ID? ")
    }
    
    if(missing(peer_id)){
        peer_id <- readline("What is your Peer's GitHub ID? ")
    }
    
    variables <- read_yaml("https://raw.githubusercontent.com/michaelweylandt/STA9750/refs/heads/main/_variables.yml")
    course_repo  <- variables$course$repo
    course_short <- variables$course$short
    
    template_url <- "https://michael-weylandt.com/STA9750/miniprojects.html"
    
    template_text <- read_html(template_url) |> 
        html_element("#peer-feedback-template") |> 
        html_text() |>
        str_trim() |>
        str_replace_all("OPTIONAL TEXT", "(.*)") |>
        str_replace_all("NN", "(.*)") |>
        str_replace_all("TEXT TEXT TEXT", "(.*)")
    
    title <- glue("{course_short} {peer_id} MiniProject #0{N}")
    
    page <- 1
    
    issues <- list()
    
    while((length(issues) %% 100) == 0){
        new_issues <- gh("/repos/michaelweylandt/{repo}/issues?state=all&per_page=100&page={page}", 
                         repo=course_repo, 
                         page=page)
        
        if(length(new_issues) == 0){
            break
        }
        
        issues <- c(issues, new_issues)
        page <- page + 1
    }
    
    issue_names <- vapply(issues, function(x) str_squish(x$title), "")
    
    name_match <- which(issue_names == title)
    
    if(length(name_match) != 1){
        cat("I could not find a unique issue with the title:\n", 
            "    ", sQuote(title),"\n",
            "I'm afraid I can't verify whether the peer feedback was submitted properly\n", 
            "but the issue likely lies with the submittor, not with your feedback.")
        stop("PEER FEEDBACK NOT VERIFIED.")
    } 
    
    issue <- issues[[name_match]]
    
    issue_url <- issue$html_url
    issue_comment_url <- issue$comments_url
    
    comments <- gh(issue_comment_url)
    
    commenters <- vapply(comments, function(x) x$user$login, "")
    
    comment_num <- which(commenters == github_id)
    
    if(length(comment_num) != 1){
        cat("I cannot identify your comment on the issue at", 
            sQuote(issue_url), ". Please verify that this is the correct link.\n")
        
        if(length(comment_num) == 0){
            cat("You do not appear to have commented on this issue.")
        } else {
            cat("You have left multiple comments on this issue.",
                "Please consolidate your feedback")
        }
        stop("PEER FEEDBACK NOT VERIFIED.")
    }
    
    comment_body <- comments[[comment_num]]$body
    
    cat("I have found your comment, with the following text:\n-----\n")
    
    cat(comment_body)
    
    cat("\n-----\n")
    
    comment_body  <- comment_body |> str_squish()
    template_text <- template_text |> str_squish()
    
    matches <- str_match_all(comment_body, template_text)[[1]][-1]
    
    if(anyNA(matches)){
        cat("I couldn't match the template to your comment. Please modify and try again.")
        stop("PEER FEEDBACK NOT VERIFIED.")
    }
    
    cat(glue("Congratulations! Your peer feedback on {peer_id}'s MP #0{N} appears properly formatted.\nThank you for your contributions to {course_short}!\n"))
    
    invisible(TRUE)
    
}
```

### Count Words {#count_words}

The following script can be used to count words on a
Quarto-generated web page. 

```{r}
#| label: count_words
count_words <- function(url){
    library(rvest)
    library(stringi)
    
    # Note that this includes code inside an inline block, but omits
    # full-sized code blocks
    read_html(url) |>
        html_elements("main p") |>
        html_text() |>
        stri_count_words() |>
        sum()
}
```

### Lint Submission

The following will automatically run the `lintr` package on a
submitted `qmd` document. 

```{r}
#| label: lint_submission
lint_submission <- function(N, peer_id){
    if(!require("yaml")) install.packages("yaml"); library(yaml)
    if(!require("gh")) install.packages("gh"); library(gh)
    if(!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
    if(!require("rvest")) install.packages("rvest"); library(rvest)
    if(!require("glue")) install.packages("glue"); library(glue)
    if(!require("lintr")) install.packages("lintr"); library(lintr)
    if(!require("httr2")) install.packages("httr2"); library(httr2)
    
    if(missing(N)){
        N <- menu(title="Which Mini-Project submission would you like to lint?", 
                  choices=c(1, 2, 3, 4))
    }
    
    if(missing(peer_id)){
        peer_id <- readline("What is your Peer's GitHub ID? ")
    }
    
    variables <- read_yaml("https://raw.githubusercontent.com/michaelweylandt/STA9750/refs/heads/main/_variables.yml")
    course_repo  <- variables$course$repo
    course_short <- variables$course$short
    
    raw_url <- glue("https://raw.githubusercontent.com/{peer_id}/{course_repo}/refs/heads/main/mp0{N}.qmd")
    
    cat("Attempting to access qmd source at", raw_url, ".\n")
    
    resp <- request(raw_url) |> 
        req_error(is_error = \(r) FALSE) |>
        req_perform()
    
    if(resp_is_error(resp)){
        cat("I could not access the raw qmd document. Attempting to open in browser...\n")
        browseURL(resp)
        return(FALSE)
    } else {
        cat("I was able to access the raw qmd document. Beginning to lint.\n")
        
        tf <- tempfile(pattern=glue("mp0{N}_{peer_id}_lint_document.qmd"))
        
        writeLines(resp_body_string(resp), tf)
        
        lintr::lint(tf)
    }
}
```

## Advice for Learning to Program

Learning a programming language is in many ways like learning a new natural
(human) language. We can look at best practices for learning spoken languages
and use them to inform how you should best approach learning `R` in this course: 

i) **Immersion Learning**: Immersion programs-intensive environments where 
   students have to live their entire life in the language they are learning-are
   the best way to learn a language. Similarly, there really is no more effective
   way to learn to program than forcing yourself to use the tools of this course
   to _do something_. Small random isolated exercises will not build speed or
   fluency, just like speaking a single sentence is a poor substitute for actual
   conversation. In this course, I attempt to replicate this 'immersive' 
   experience through realisitic [mini-projects](miniprojects.qmd). You can get
   more experience by just committing to use `R` for every day tasks outside 
   this class: say, *e.g.*, you need to rename a bunch of files on your computer.
   Rather than doing this manually, force yourself to use `R` to do the same task.
   It may take a bit longer the first few times, but you will quickly develop
   additional programming fluency that will serve you well. 
   
ii) **Don't Memorize the Dictionary**. New programmers often try to sit down and
    learn _every function_ and _every detail_ before attempting a task. This is 
    impractical - there are tens of thousands of functions in just the small set
    of `R` we cover in this course - and honestly rather dull. It is far more
    efficient to take on a task and look up functions as you need them. (You can
    always Google "How do I XYZ in R?") This strategy will naturally focus your
    attention on the most useful functions - since you will look them up many
    more times - and let you learn the additional bells-and-whistles of each 
    function as you need it. In computer science education, this is often called
    *JIT* (Just-in-Time) learning. 
    
    You wouldn't sit down to learn French by reading the dictionary front-to-back
    and you shouldn't try to do it for `R` either.
    
iii) **Accept that You Will Feel Foolish**. You are learning a new language, one
     that comes with its own grammar (syntax), nouns and pronouns (variables), 
     verbs (functions), and idioms. You won't get it all right the first time-no
     matter how hard you work, developing fluency simply takes time. Thankfully
     for you, you have a 'conversation partner' who is willing to spend as much
     time with you as you want twenty-four hours a day, seven days a week. And
     unlike a human conversation partner, `R` won't tell anyone if you 
     say anything embarrassing. 
     
iv) **Native Speakers Can be a Bit Judgmental** `R` is picky about getting the 
     grammar right. This can be frustrating - why can't it just know what you 
     want! - but it is ultimately to your benefit as it gives you an opportunity
     for immediate feedback and improvement. R's error message are 
     systematic, but are not always perfectly clear. (Writing clear error messages
     is actually quite difficult, as they only occur when something has already
     unexpectedly gone wrong.) As you start seeing the same error messages over
     and over, you will learn to make sense of them, but whenever you get a new
     one, simply Google it. You are almost certainly not the first person to have
     made any mistake. And never be embarrassed to ask the instructor or other
     course staff for help making sense of an error: you would never judge someone
     for using the English past progressive tense incorrectly after only 10 weeks
     of speaking English - give yourself the same charity. 
     
v) **You Won't Think in the Foreign Language at First**. When facing an analytics
   question, don't immediately think "how can I code this in `R`?" Take a moment,
   figure out the steps you need to execute, and then translate your problem 
   step by step from English to `R`. You will develop fluency quickly, but always
   make sure you know *what you want to do* and *the steps you need to take* 
   before you jump straight into writing code. 
   
   If you take a moment to clarify for yourself - in English or any other 
   language - what you want the computer to do first, then the only 
   "`R`-thinking" you have to do is translating your steps into `R`. This is 
   often much easier than trying to write in `R` on a "blank page." Compare the
   difficulty of translating a sentence you already have written from one language 
   to another word-by-word to the much harder task of writing an essay from scratch.

### Computers are Very Fast and Very Dumb

Your computer is a powerful tool, but it is fundamentally a very simple one. 
It executes millions or billions of simple steps every second, but it cannot 
read your mind. You have to be very precise in giving commands, as small 
omissions or errors will immediately lead to errors and other problems. As you
fix these, you will often feel a bit silly for having made those mistakes
in the first place. Don't be embarrassed - working with a computer is like giving
instructions to an enthusiastic toddler: if something goes wrong, it's almost
always because you assumed the toddler has more background knowledge and common
sense than is actually there. 

There are really only two fundamental programming problems you will come across 
in this course:

- You don't know what you want the computer to do. 
- You know what you want the computer to do, but you told it incorrectly. 

Technologically, we are at a point in time where errors of that second category
are becoming less common, but they are certainly still very present and you'll
encounter them repeatedly in this course. That said, whenever you're facing a 
problem, try to categorize it as one of these two classes (or maybe a bit of 
each) and address them appropriately. The best way to deal with a "What should 
I do?" question is not the same as the best way to deal with a "How do I do it?"
question. Being able to distinguish which type of question you are answering
is the first step to being able to answer it effectively. 


```{r}
#| eval: false
#| include: false
# Run this to update helper scripts
knitr::purl("tips.qmd",
            documentation=0L,
            output="docs/load_helpers.R")
```

