---
title: "{{< var course.short >}} Week 7 Pre Assignment: Lots of Plots"
---

```{r}
#| message: false
#| echo: false
#| warning: false
library(tidyverse)
NUMBER <- 7
due_date_str <- read_csv("key_dates.csv") |> 
    filter(`Course Element` == "Pre-Assignments", 
           `Item Number` == NUMBER, 
           str_detect(Details, "Due")) |> 
    mutate(dt = parse_date_time(paste(Date, Time), "%Y-%m-%d %H:%M%p")) |>
    pull(dt) |>
    format("%Y-%m-%d (%A) at %I:%M%P")
```

**Due Date**: `{r} due_date_str`

**Submission**: [CUNY Brightspace](https://brightspace.cuny.edu/)

This week we begin to explore _statistical_ visualizations. Visualizations
play several interrelated roles in statistical practice: we use visualization
to explore new data sets, to see how well models fit to data, and to communicate
the results of analyses to new audiences. Compared with the 'point summary' tools
we have discussed to date, visualizations are far more flexible and more
powerful: we can extract novel insights from data visualizations, but we can
also deceive ourselves and others.


As you review this document, also watch how I iterate and refine each figure
until I have something I'm finally happy with. This is quite typical of how
working data scientists produce plots: you rarely know exactly what you want,
particularly before you begin to explore your data. You should adapt a similar
pattern of "take a sad plot and make it better" as you create plots for your
mini-projects and, ultimately, your final project.

## Grammar of Graphics

Computer graphics is a field of essentially infinite possibility - anything
you dream can be represented in the digital domain. At one limit, we have
the option to work on a "per pixel" basis, telling the computer exactly what
to draw and where. Because this is clearly overwhelmingly monotonous, very
little work is actually done at such fine detail and _abstraction layers_ are
provided to automate the low-level detail work. We will use an abstraction
model known as the _Grammar of Graphics_. 

Designed by Leland Wilkinson and popularized by Hadley Wickham, the _Grammar
of Graphics_ poses a set of rules for visualizing (tidy) data. It draws its
name from the concept of linguistic grammar - the rules that dictate how
basic elements (nouns, verbs, and adjectives) may and may not be combined into
clear and meaningful sentences The Grammar of Graphics provides specifications
for combining different plot elements (legends, data, axes, *etc.*) into clear
and meaningful statistical graphics. Taking the metaphor too far, we might
say that the Grammar of Graphics is the Strunk and White to the Ska that are
"Infographics." 

The grammar of graphics has several interconnected components, which are
combined to form a meaningful graphic. We assume that we have a "tidy" data
set we want to visualize; recall that, by "tidy", we mean that our data is

- Organized in a rectangular array
- Homogenously-typed within a column
- One observation per row
- One value per cell

Given this form of tidy data, the grammar of graphics provides us the 
"parts of speach" necessary to convert tidy data to a visual representation.
We'll only cover the basic components here, leaving more advanced tools for
class session: 

- The _aesthetics_ are mappings between columns of the data and aspects
  of the visualization. For instance, "put the grade column on the $y$ axis"
  or "use color to represent the course ID". 
- _Scales_ convert data values to the aesthetics: scales may be quite trivial,
  *e.g.* placing continuous values on the $x$ axis in proper order, or more
  advanced, *e.g.*, binning values and converting them to a sequence of 
  perceptually-ordered colors. 
)
- _Geometric elements_ or `geom`s specify how the data are represented on
  the page through the scales. `geom`s include basic representations, like
  points for a scatter plot or lines for a trend plot, as well as more
  complex objects like boundaries on a map.
- _Guides_ provide interpretational assistance to the viewer. Most guides
  take the form of legends. 
  
That's all a bit abstract, so let's put it into practice. For now, you shouldn't
worry so much about what each of these really mean; it's just useful to have
a rough sense of what "knob" you want to turn to modify plots. 

## Getting started with `ggplot2`

The leading implementation of the grammar of graphics is the `ggplot2` package
in `R` (`gg` = Grammar of Graphics). It comes to us from Hadley Wickham and
the `tidyverse` team, who also developed the `dplyr` and `tidyr` tools we have
been using for the past several weeks. 

Let's begin by using `ggplot2` to explore our `penguins` data: 

```{r}
library(ggplot2)
library(palmerpenguins)
ggplot(penguins)
```

Somewhat underwhelming... 

`ggplot2`, like many of the tools in this course, do exactly what we ask, and 
no more. Because we have not specified any of the Grammar of Graphics elements,
we only get a blank canvas. Let's now begin by adding an _aesthetic_ to map
some of the elements of our data to aspects of our plot. 

Specifically, suppose we want to see how flipper length correlates with
body mass. Let's make a scatter plot with flipper length on the $x$-axis
and body mass on the $y$-axis.[^1]

```{r}
ggplot(penguins, aes(x=flipper_length_mm, y=body_mass_g))
```

Ok - this is perhaps a bit better. We can see that the columns
`flipper_lenth_mm` and `body_mass_g` have been placed on the $x$- and $y$-axes
as we wanted, but we still don't see anything. 

We need a `geom` to actually put "ink to paper". The simplest `geom` is 
a point, useful for making scatter plots. 

```{r}
ggplot(penguins, 
       aes(x=flipper_length_mm, y=body_mass_g)) + 
    geom_point()
```

Pretty nifty! Before we go forward, note that `ggplot2` _adds_ elements to
create ever more complex plots. This is different than `dplyr` where we 
"piped" data from one step to the next, refining it along the way. 

How can we improve the plot above? Before anything else, let's clean up
the $x$ and $y$ axis labels. While the default behavior of showing variable
names is helpful for exploratory data analysis, we never want to let variable
names "leak" in plots we intend to share with others. We should instead use
meaningful (and attractive) axis labels. 

```{r}
ggplot(penguins, 
       aes(x=flipper_length_mm, y=body_mass_g)) + 
    geom_point() + 
    xlab("Flipper Length (mm)") + 
    ylab("Body Mass (g)")
```

We can also take advantage of the `theme` mechanisms of `ggplot2` to change
the color of the "infrastructure" of our plot. The `theme` mechanism doesn't
change how the data itself is visualized, but it controls how things like
the background, font sizing, *etc* behave. I tend to prefer the black and
white theme over the grey default: 

```{r}
ggplot(penguins, 
       aes(x=flipper_length_mm, y=body_mass_g)) + 
    geom_point() + 
    xlab("Flipper Length (mm)") + 
    ylab("Body Mass (g)") + 
    theme_bw()
```

### Adding Color to Depict Species

Our data set three distinct species and the correlation between flipper
and body size may vary across species. Let's add some color to our plot:
since color maps a data element (`species`) to a graphical aspect (color)
we add it to our aesthetic mapping (`aes`). 

```{r}
ggplot(penguins, 
       aes(x=flipper_length_mm, 
           y=body_mass_g, 
           color=species)) + 
    geom_point() + 
    xlab("Flipper Length (mm)") + 
    ylab("Body Mass (g)") + 
    theme_bw()
```

We see here that a few things happened automatically for us: 

1) The color element was automatically propagated into `geom_point`. 
   By default, any "top-level" aesthetics are automatically applied
   to any `geom` that can handle them. 
2) A legend was created.
3) A color scale was chosen.

The [`geom_point`](https://ggplot2.tidyverse.org/reference/geom_point.html#aesthetics) help page tells us which aesthetics are required ($x, y$) and which
are optional for the point geom. We couldn't have gotten away without providing
$x, y$ coordinates, but until this point, we were just using the default (black)
color. 

To improve the look of the colors, we can choose a different color scale. 
I tend to like the colors of the [Color Brewer](https://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3) project,
though strictly speaking these are designed for use in maps, not scatter plots.
You can access these in `ggplot2` as follows: 

```{r}
ggplot(penguins, 
       aes(x=flipper_length_mm, 
           y=body_mass_g, 
           color=species)) + 
    geom_point() + 
    xlab("Flipper Length (mm)") + 
    ylab("Body Mass (g)") + 
    theme_bw() + 
    scale_color_brewer(type="qual", palette=2)
```

Here, I'm using a `type="qual"` (qualitative) palette because there is
no inherent ordering to the penguin types. I like the "bolder" colors of
the second palette in this set, but you can adjust the number to try different
schemes. 

Next, let's improve the look of the legend. As before, we see that it is
by default titled with the variable name (`species`). We can provide
a proper title instead: 

```{r}
ggplot(penguins, 
       aes(x=flipper_length_mm, 
           y=body_mass_g, 
           color=species)) + 
    geom_point() + 
    xlab("Flipper Length (mm)") + 
    ylab("Body Mass (g)") + 
    theme_bw() + 
    scale_color_brewer(type="qual", 
                       palette=2, 
                       name="Species")
```
I still don't like this legend on the side - it takes too much room, so let's
move it below the image instead. This involves changing a "non-data"
element of the plot, so we go through the theme machinery.
[`theme()`](https://ggplot2.tidyverse.org/reference/theme.html) allows an 
enormous number of possible changes, but here we want `legend.position`:

```{r}
ggplot(penguins, 
       aes(x=flipper_length_mm, 
           y=body_mass_g, 
           color=species)) + 
    geom_point() + 
    xlab("Flipper Length (mm)") + 
    ylab("Body Mass (g)") + 
    theme_bw() + 
    scale_color_brewer(type="qual", 
                       palette=2, 
                       name="Species") + 
    theme(legend.position="bottom")
```

Not so bad!

### Visualizing Statistical Relationships

Recall that our goal was to measure the correlation between Body Mass
and Flipper Length. We can visualize this correlation on the plot by adding
a regression line (recall that for univariate regression like this, the
slope of the regression line is $\hat{\beta} = \rho_{XY}\frac{\sigma_Y}{\sigma_X}$). 

This is a new geometric element, called a `smoother`. `ggplot2` allows
many possible smoothers, but let's use the `lm` (linear model) version, which
we specify by setting the `method` argument:

```{r}
ggplot(penguins, 
       aes(x=flipper_length_mm, 
           y=body_mass_g, 
           color=species)) + 
    geom_point() + 
    geom_smooth(method="lm") + 
    xlab("Flipper Length (mm)") + 
    ylab("Body Mass (g)") + 
    theme_bw() + 
    scale_color_brewer(type="qual", 
                       palette=2, 
                       name="Species") + 
    theme(legend.position="bottom")
```

By default, `ggplot2` is giving us confidence intervals around the linear trend.
These are sometimes useful, but perhaps a bit crowded for now, so let's turn
them off: 

```{r}
ggplot(penguins, 
       aes(x=flipper_length_mm, 
           y=body_mass_g, 
           color=species)) + 
    geom_point() + 
    geom_smooth(method="lm", 
                se=FALSE) + 
    xlab("Flipper Length (mm)") + 
    ylab("Body Mass (g)") + 
    theme_bw() + 
    scale_color_brewer(type="qual", 
                       palette=2, 
                       name="Species") + 
    theme(legend.position="bottom")
```

This is nicer, but maybe still a bit crowded. It would be nicer if we could
avoid the "overlaps" of the different species. Here, let's break out a 
"small multiples" plot: this is, in essence, a `group_by` for plotting. 

In `ggplot2` speak, this is called a faceted plot: 

```{r}
ggplot(penguins, 
       aes(x=flipper_length_mm, 
           y=body_mass_g, 
           color=species)) + 
    geom_point() + 
    geom_smooth(method="lm", 
                se=FALSE) + 
    xlab("Flipper Length (mm)") + 
    ylab("Body Mass (g)") + 
    theme_bw() + 
    scale_color_brewer(type="qual", 
                       palette=2, 
                       name="Species") + 
    theme(legend.position="bottom") + 
    facet_wrap(~species)
```

Not too shabby. But now it's a bit too hard to tell the lines from the points. 
Let's override the color used: 

```{r}
ggplot(penguins, 
       aes(x=flipper_length_mm, 
           y=body_mass_g, 
           color=species)) + 
    geom_point() + 
    geom_smooth(method="lm", 
                se=FALSE, 
                color="black") + 
    xlab("Flipper Length (mm)") + 
    ylab("Body Mass (g)") + 
    theme_bw() + 
    scale_color_brewer(type="qual", 
                       palette=2, 
                       name="Species") + 
    theme(legend.position="bottom") + 
    facet_wrap(~species)
```

These plots give us some idea of the correlation, but what if we want
actual numbers? We can't do this in "plain" `ggplot2`, but now we can take
advantage of the _enormous_ number of [`ggplot2` extension
packages](https://exts.ggplot2.tidyverse.org/). It turns out that the `ggpmisc`
package supports what we need, so let's download and install it: 

```{r}
#| warning: false
#| message: false
if(!require("ggpmisc")) install.packages("ggpmisc")
library(ggpmisc)
```

Now we have access to the various `geom_`, `scale_`, *etc* objects from
that package. We can now introduce a new category, `stat_`, that represents
statistical transformations or modeling. Generally, these are applied 
"automagically" for us, as in `geom_smooth`, but here we need to build
our regression models explicitly: 

```{r}
ggplot(penguins, 
       aes(x=flipper_length_mm, 
           y=body_mass_g, 
           color=species)) + 
    geom_point() + 
    stat_poly_line(se=FALSE, 
                   color="black") +
    stat_poly_eq() + 
    xlab("Flipper Length (mm)") + 
    ylab("Body Mass (g)") + 
    theme_bw() + 
    scale_color_brewer(type="qual", 
                       palette=2, 
                       name="Species") + 
    theme(legend.position="bottom") + 
    facet_wrap(~species)
```

See how, even though we're using functionality from outside `ggplot2`, the
structure of the "grammar" makes it easy for all these tools to work well
together.

### Final Polish

We are almost done, but every figure needs a bit of final polish. Firstly, 
we should add a title, using the `ggtitle` function. (`R` has a built-in
`title` function but that won't help us here)

```{r}
ggplot(penguins, 
       aes(x=flipper_length_mm, 
           y=body_mass_g, 
           color=species)) + 
    geom_point() + 
    stat_poly_line(se=FALSE, 
                   color="black") +
    stat_poly_eq() + 
    xlab("Flipper Length (mm)") + 
    ylab("Body Mass (g)") + 
    theme_bw() + 
    scale_color_brewer(type="qual", 
                       palette=2, 
                       name="Species") + 
    theme(legend.position="bottom") + 
    facet_wrap(~species) + 
    ggtitle("Correlation of Flipper Length and Body Mass across Penguin Species")
```

In good academic practice, we should always add a footnote citing the source
of our data. The [`palmerpenguins` site](https://allisonhorst.github.io/palmerpenguins/) has appropriate source
information: 

```{r}
ggplot(penguins, 
       aes(x=flipper_length_mm, 
           y=body_mass_g, 
           color=species)) + 
    geom_point() + 
    stat_poly_line(se=FALSE, 
                   color="black") +
    stat_poly_eq() + 
    xlab("Flipper Length (mm)") + 
    ylab("Body Mass (g)") + 
    theme_bw() + 
    scale_color_brewer(type="qual", 
                       palette=2, 
                       name="Species") + 
    theme(legend.position="bottom") + 
    facet_wrap(~species) + 
    ggtitle("Correlation of Flipper Length and Body Mass across Penguin Species") + 
    labs(caption="Data provided by Dr. K. Gorman and the Palmer Station, Antarctica LTER")
```

At this point, the bottom of our figure looks a bit crowded. To clear out
some space, let's remove the legend from the bottom, since it simply repeats
the facet labels: 

```{r}
ggplot(penguins, 
       aes(x=flipper_length_mm, 
           y=body_mass_g, 
           color=species)) + 
    geom_point() + 
    stat_poly_line(se=FALSE, 
                   color="black") +
    stat_poly_eq() + 
    xlab("Flipper Length (mm)") + 
    ylab("Body Mass (g)") + 
    theme_bw() + 
    scale_color_brewer(type="qual", 
                       palette=2, 
                       name="Species") + 
    theme(legend.position="bottom") + 
    facet_wrap(~species) + 
    ggtitle("Correlation of Flipper Length and Body Mass across Penguin Species") + 
    labs(caption="Data provided by Dr. K. Gorman and the Palmer Station, Antarctica LTER") + 
    guides(color="none")
```

Cleaner and no loss of information. In designing good _scientific_ graphics,
the concept of a "ink-to-information" ratio is useful: if you can remove
some ink without removing any (relevant) information, you should generally
do so. This makes it easier for the reader to identify the _important_
elements of the plot. 

To make our point even clearer, it is sometimes useful to add a short 
"summary" to a plot like this: 

```{r}
ggplot(penguins, 
       aes(x=flipper_length_mm, 
           y=body_mass_g, 
           color=species)) + 
    geom_point() + 
    stat_poly_line(se=FALSE, 
                   color="black") +
    stat_poly_eq() + 
    xlab("Flipper Length (mm)") + 
    ylab("Body Mass (g)") + 
    theme_bw() + 
    scale_color_brewer(type="qual", 
                       palette=2, 
                       name="Species") + 
    theme(legend.position="bottom") + 
    facet_wrap(~species) + 
    ggtitle("Correlation of Flipper Length and Body Mass across Penguin Species", 
            subtitle="Flipper Length and Body Mass are positively correlated across species\nGentoo penguins exhibit the strongest relationship at 70% correlation") + 
    labs(caption="Data provided by Dr. K. Gorman and the Palmer Station, Antarctica LTER")
```

This is not always a good idea - it requires "hard-coding" the insights
for your reader and takes up some space. In scientific writing, I generally
prefer to put this sort of summary in a figure caption, while I tend
to "say the point" verbally if the figure is destined for a presentation. 

Personally, I only use this sort of "here is the point" text if I expect
a figure to "escape beyond" my presentation and need it to stand fully on
its own. 

### Conclusions

`ggplot2` provides an exceptionally powerful and flexible set of tools for
creating statistical visualizations. We will explore it in more depth in 
class. For now, review the examples above and make sure you see how each
plot is created "piecewise" from its various components. 


To see more about what `ggplot2` can do, check out the [R Graphics
Gallery](https://r-graph-gallery.com/ggplot2-package.html). If you want to 
see the specifics of each `ggplot2` function, check out the [package reference
page](https://ggplot2.tidyverse.org/reference/). To go further with `ggplot2`,
you should also explore its [extension
gallery](https://exts.ggplot2.tidyverse.org/gallery/). 

[^1]: I once had a boss who refused to use the terms $x$ and $y$ axis unless
quantities called $x$ and $y$ were actually being plotted. Being Cantabrigian, 
he insisted I use the terms abscissa and ordinate. I will not inflict such
pedantry in this course. 
