---
title: "{{< var course.short >}} - Software Tools for Data Analysis"
filters:
  - conditional_sections.lua
---

```{r}
#| echo: false
#| message: false
source("render_key_dates.R")
```

Welcome to the course website for {{< var course.short >}} 
({{< var course.semester >}})!

{{< var course.short >}} is an _Introduction to R_ targeted at students in the 
[MS in Business Analytics](https://zicklin.baruch.cuny.edu/academic-programs/graduate/ms/masters-in-business-analytics/), 
[MS in Statistics](https://zicklin.baruch.cuny.edu/academic-programs/graduate/ms/statistics/), 
and [MS in Quantitative Methods](https://zicklin.baruch.cuny.edu/academic-programs/graduate/ms/quantitative-methods-modeling/) 
programs. 

This site hosts the [Course Syllabus](./syllabus.html), 
[Course Policies](./resources.html), and [Course Learning 
Objectives](./objectives.html). Copies of these documents may also be
found on your [CUNY Brightspace](https://brightspace.cuny.edu).

Required course [pre-assignments](./preassignments.qmd), [labs](./labs.qmd), and
[mini-projects](./miniprojects.qmd) can also be found on this site, as can
my lecture slides and information relevant to the [course project](./project.qmd).
Some helpful technical advice and tools can be found [here](./tips.qmd)

This offering of {{< var course.short >}} will be taught in a mixture of the
*flipped-classroom* and  *experiential-learning* formats. Roughly, this means
that most weeks, students will be asked to complete a small [_pre-assignment_](./preassignments.qmd) 
each week to introduce the core concept(s) covered in that week's lecture.
Each class period will be split between a brief lecture covering concepts in more
detail and an extended [_lab activity_](./labs.qmd) designed to build
familiarity and fluency with that week's subject matter. 

::: {multiple_sections="true"}

This semester, I am teaching both the *Tuesday* and *Thursday* sections of
{{< var course.short >}}. In order to provide maximal flexibility to students, 
students may attend either section as their schedule permits.[^double] All
assignments and deadlines are shared between both sections. See below for the 
schedule relevant to each section. This makes a few of the deadlines a bit
awkward[^deadline] (when one section has a holiday and the other doesn't) but the
advantages for student flexibility seem to be worth it.

[^deadline]: Note that all deadlines are set based on the appropriate due date
for the *later* section. If your section doesn't have a holiday, you essentially
get extra time to complete an assignment. While this may occasionally help in
a pinch, I **strongly** encourage you to start work as soon as you are able. It 
is often hard to estimate how long work for this course takes and waiting 
up to the deadline is asking for trouble. 

[^double]: Students who are struggling with the course material may even find
it useful to attend *both* sections on certain topics. Both sections will
use the same Zoom link.

:::

::: {multiple_sections="false"}

:::: {#calendar}

There are quite a few moving parts to this course, so this 
[key dates file](./key_dates.csv) or the list of upcoming course activities 
below may be useful:

```{r}
#| eval: false
#| echo: false
#| message: false
#| warning: false
library(yaml)
library(glue)
library(tidyverse)
if(fs::file_exists("_variables.yml")){
  variables <- read_yaml("_variables.yml")
} else {
  variables <- read_yaml("https://raw.githubusercontent.com/michaelweylandt/STA9750/refs/heads/main/_variables.yml")
}
course_repo  <- variables$course$repo
course_short <- variables$course$short

read_csv("key_dates.csv") |> 
    select(-`Significant Student Time`) |>
    filter_out(str_detect(Details, "Estimated Return")) |> 
    mutate(Time2 = Time |> str_replace("11:", "23:") |> str_replace("6:", "18:")) |>
    arrange(Date, Time2) |> 
    select(-Time2) |>
    DT::datatable(options=list(searching=FALSE,
                               paging=TRUE),
                  caption=glue("Upcoming Important Dates for {course_short}"), 
                  elementId="calendar_table",
                  rownames=FALSE)
```

A CSV file suitable for import into Google Calendar with all assignment 
deadlines can be found [here](./key_dates_students_gcal.csv).

Instructions for how to do so can be found
[here](https://support.google.com/calendar/answer/37118?hl=en&co=GENIE.Platform=Desktop).
If you use another calendar system that cannot accept CSV files in this format, 
let me know and I will try to provide additional file formats. 


<script>
// Update Calendar DataTables (created above in R) to show future events only
// 
// We set a one-time hook on each table (selected by ID) at its initial 'draw'
// event. This function checks the current date and compares it to the value
// in the fourth (date) column, deleting rows with past dates. After deletion,
// the table is re-drawn. 
//
// We use a one time hook (\$.one) as opposed to an everytime hook (\$.on) to 
// avoid an infinite loop. Since the table is not destroyed, just hidden, when
// moving from one tab to the other, this seems to be safe. 
$('#calendar_table').each(function(){
  $(this).one('draw.dt', function(){
    var filter_date = new Date();
    filter_date.setDate(filter_date.getDate() - 1);
    var filter_string = filter_date.toISOString().split('T')[0];
    $("table").DataTable().rows((idx, data) => data[3] < filter_string).remove().draw();
  });
});

</script>


::::

:::

::: {multiple_sections="true"}

:::: {.panel-tabset #calendar}

## Tuesday Section

There are quite a few moving parts to this course, so this [key dates file](./key_dates_tuesday.csv)
or the list of upcoming course activities below may be useful:

```{r}
#| echo: false
#| message: false
#| warning: false
library(yaml)
library(glue)
variables <- read_yaml("https://raw.githubusercontent.com/michaelweylandt/STA9750/refs/heads/main/_variables.yml")
course_repo  <- variables$course$repo
course_short <- variables$course$short

readr::read_csv("key_dates_tuesday.csv") |> 
    dplyr::arrange(Date) |> 
    filter_out(str_detect(Details, "Estimated Return")) |> 
    DT::datatable(options=list(searching=FALSE,
                               paging=TRUE),
                  caption=glue("Upcoming Important Dates for {course_short} - Tuesday Section"), 
                  elementId="tuesday_table",
                  rownames=FALSE)
```


A CSV file suitable for import into Google Calendar with all assignment 
deadlines can be found [here](./key_dates_students_gcal_tuesday.csv).

Instructions for how to do so can be found
[here](https://support.google.com/calendar/answer/37118?hl=en&co=GENIE.Platform=Desktop).
If you use another calendar system that cannot accept CSV files in this format, 
let me know and I will try to provide additional file formats. 

## Thursday Section

There are quite a few moving parts to this course, so this [key dates file](./key_dates_thursday.csv)
or the list of upcoming course activities below may be useful:

```{r}
#| echo: false
#| message: false
#| warning: false
library(yaml)
library(glue)
variables <- read_yaml("https://raw.githubusercontent.com/michaelweylandt/STA9750/refs/heads/main/_variables.yml")
course_repo  <- variables$course$repo
course_short <- variables$course$short

readr::read_csv("key_dates_thursday.csv") |> 
    dplyr::arrange(Date) |> 
    dplyr::filter_out(str_detect(Details, "Estimated Return")) |> 
    DT::datatable(options=list(searching=FALSE,
                               paging=TRUE),
                  caption=glue("Upcoming Important Dates for {course_short} - Thursday Section"), 
                  elementId="thursday_table",
                  rownames=FALSE)
```


A CSV file suitable for import into Google Calendar with all assignment 
deadlines can be found [here](./key_dates_students_gcal_thursday.csv).

Instructions for how to do so can be found
[here](https://support.google.com/calendar/answer/37118?hl=en&co=GENIE.Platform=Desktop).
If you use another calendar system that cannot accept CSV files in this format, 
let me know and I will try to provide additional file formats. 

::::

<script>
// Update Calendar DataTables (created above in R) to show future events only
// 
// We set a one-time hook on each table (selected by ID) at its initial 'draw'
// event. This function checks the current date and compares it to the value
// in the fourth (date) column, deleting rows with past dates. After deletion,
// the table is re-drawn. 
//
// We use a one time hook (\$.one) as opposed to an everytime hook (\$.on) to 
// avoid an infinite loop. Since the table is not destroyed, just hidden, when
// moving from one tab to the other, this seems to be safe. 
$('#tuesday_table,#thursday_table').each(function(){
  $(this).one( 'draw.dt', function(){
    var filter_date = new Date();
    filter_date.setDate(filter_date.getDate() - 1);
    var filter_string = filter_date.toISOString().split('T')[0];
    $("table").DataTable().rows((idx, data) => data[3] < filter_string).remove().draw();
  });
});

</script>

:::

Instructor: [Michael Weylandt](https://michael-weylandt.com)

