---
title: "{{< var course.short >}} - Course Syllabus"
filters: 
    - list-table
---

### Instructor

-   [Professor Michael Weylandt](https://michael-weylandt.com)
-   Department of Information Systems & Statistics
-   Zicklin School of Business
-   Baruch College, CUNY

### Course Meetings

#### Lecture / Lab Sessions

-   Virtual (Synchronous Online)
-   Thursdays 6:05pm-9:00pm
    -   Zoom link provided via Brightspace

#### Office Hours

-   In-Person
    -   [Baruch Main Campus (1 Bernard Baruch Way)](https://www.baruch.cuny.edu/campus-map-and-directions/)
    -   Newman Vertical Campus (NVC) 11-246
    -   Mondays 4:30-5:30pm
    -   Subject to periodic cancellation (communicated via Brightspace)
-   Virtual:
    -   Thursdays 4:30pm-5:30pm
    -   Zoom link provided via Brightspace

### Grading

-   24% [Weekly Pre-Assignments](./preassignments.html)
    -   3% Each (best eight of ten: lowest two dropped)
    -   Submission *via* Brightspace
-   36% [Mini Projects](./miniprojects.html)
    -   7% Each (four total)
    -   2% per Peer Feedback (Meta-Review)
-   40% [Course Project](./project.html)
    -   5% Proposal
    -   5% Mid-term Check-In
    -   10% Final Presentation
    -   7.5% Final Report
    -   12.5% Individual Evaluation

*Opportunities for Extra Credit:* Several opportunities for extra credit will be made available. These include:

1.  Participation in Course Discussion Board (Piazza)
2.  Correction of errors in published course materials (*via* Github pull request)
3.  Submission of homework from other STA courses completed using `R`. (Note that the assignment must not ordinarily require `R`: this is *extra* credit.)

Final course grades will be curved in accordance with relevant program, departmental, school, and college policies.[^1]

[^1]: Theoretically, this may result in scores equivalent to an `A` in an un-curved course receiving a lower grade in this course. In practice, the instructor will design course assessments to induce a range of scores and does not anticipate "down-curving" happening.

#### Regrading Policy

If you feel an assignment has been improperly graded, please contact the instructor by private message on the course discussion board within 48 hours of the graded assignment being returned. Note that the instructor will regrade the assignment *de novo*, so your grade may be adjusted upwards or downwards.

#### Late Work Policy

This course has *many* moving pieces and late work will not generally be accepted. Specifically, late submission slows down the peer review cycle used in thise course. All assignments can be submitted multiple times on Brightspace, so you are **strongly** encouraged to submit early and to submit often. Students are also encouraged to load the [course deadlines file](./key_dates_students_gcal.csv) into their personal calendar to better track key course dates.

Individual exceptions will be made i) with *prior written* approval of instructor; or ii) *ex post* with written note from the [Office of the Dean of Students](https://studentaffairs.baruch.cuny.edu/dean-of-students/).

### Tentative Course Schedule

```{r}
#| echo: false
#| message: false
library(tidyverse)
class_date <- function(n){
    read_csv("key_dates.csv") |>
        filter(`Course Element` == "Class Session") |>
        filter(`Item Number` == n) |>
        pull(Date) |>
        format("%Y-%m-%d")
}
```

```{r}
#| echo: false
#| message: false
library(tidyverse)
holiday_date <- function(n, details=FALSE){
    h <- read_csv("key_dates.csv") |>
            filter(`Course Element` == "Holiday") |>
            filter(`Item Number` == n)
    
    if(details){
        h |> pull(Details)
    } else {
        h |> pull(Date) |> format("%Y-%m-%d")
    }
}
```


```{r}
#| echo: false
#| message: false
library(tidyverse)
due_date_mini <- function(n, submission=TRUE, feedback=!submission){
    details <- if(feedback) "Mini-Project Peer Feedback Due" else "Mini-Project Due"
    
    read_csv("key_dates.csv") |>
        filter(`Course Element` == "Mini-Projects",
               `Details` == details) |>
        filter(`Item Number` == n) |>
        mutate(deadline = paste(format(Date, "%A %B %d, %Y"), "at", Time)) |>
        pull(deadline)
}
```






::: {.list-table width="1,1,1,1,1,1,1" aligns="c,c,l,l,l,l,l" header-rows=1 .hover}

   * - Week 
     - Lecture Date
     - Topics
     - Pre-Assignment
     - In-Class Lab
     - Mini-Projects
     - Additional Notes
     
   * - 1
     - `r class_date (01)`
     - Course Overview and Key Infrastructure:
       1. `R` and `RStudio`
       2. `git` and GitHub
     - None
     - [Lab #01](./labs/lab01.html)
     - [Mini-Project #00](./miniprojects/mini00.html) assigned. 
       Due `r due_date_mini(0)`
       
   * - 2
     - `r class_date (02)`
     - 1. Introduction to Reproducible Research:
          - Markdown and Quarto
          - GitHub Pages
       2. Getting Help
       3. Course Project Overview
     - [PA #02](./preassigns/pa02.html)
     - [Lab #02](./labs/lab02.html)

   * - 3
     - `r class_date (03)`
     - First Steps with `R`:
       1. Vectors
       2. Subsetting
       3. Control Flow
       4. Function Calls
       5. Data Frames
     - [PA #03](./preassigns/pa03.html)
     - [Lab #03](./labs/lab03.html)
     - [Mini-Project #01](./miniprojects/mini01.html) assigned. 
       Due `r due_date_mini(1)`
     - Peer Feedback on Mini Project #00 due 
       `r due_date_mini(0, feedback=TRUE)`

   * - 4
     - `r class_date (04)`
     - Tidy Data Manipulation I:
       1. Single Table Operations
       2. Factors
       3. Dates & Times
       
       Review of `git` fundamentals.
     - [PA #04](./preassigns/pa04.html)
     - [Lab #04](./labs/lab04.html)

   * - 5
     - `r class_date (05)`
     - Tidy Data Manipulation II:
       1. Multi-Table Operations (Joins)
       2. Pivots
       3. Working with Missing Data
       
       Debugging Techniques
     - [PA #05](./preassigns/pa03.html)
     - [Lab #05](./labs/lab03.html)
     - [Mini-Project #05](./miniprojects/mini05.html) assigned. 
       Due `r due_date_mini(2)`
     - Peer Feedback on Mini Project #01 due 
       `r due_date_mini(1, feedback=TRUE)`

   * - 
     - `r holiday_date(1)`
     <td colspan="5" style="text-align:center"> ***`r holiday_date(1, details=TRUE)`*** </td>

   * - 6
     - `r class_date (06)`
     <td colspan="5" style="text-align:center"> **In-Class Project Proposal Presentations** </td>

   * - 7
     - `r class_date (07)`
     - Plotting I:
       - Introduction to `ggplot2`
       - Customizing Formatting
       - Maps
     - [PA #07](./preassigns/pa07.html)
     - [Lab #07](./labs/lab07.html)

   * - 8
     - `r class_date (08)`
     - Tools for Interactive Data Analysis
       - `shiny`
       - `ggplot2` Extensions for Animated and Interactive Plots
     - [PA #08](./preassigns/pa03.html)
     - [Lab #08](./labs/lab03.html)
     - [Mini-Project #03](./miniprojects/mini03.html) assigned. 
       Due `r due_date_mini(3)`
     - Peer Feedback on Mini Project #02 due 
       `r due_date_mini(2, feedback=TRUE)`
     
   * - 9
     - `r class_date (09)`
     - Data Import
     - [PA #09](./preassigns/pa09.html)
     - [Lab #09](./labs/lab09.html)
     
   * - 10
     - `r class_date (10)`
     <td colspan="5" style="text-align:center"> **Mid-Semester Check-In Presentations** </td>

   * - 
     - `r holiday_date(2)`
     <td colspan="5" style="text-align:center"> ***`r holiday_date(2, details=TRUE)`*** </td>

   * - 11
     - `r class_date (11)`
     - Elements of Web Scraping
       - Introduction to HTTP and HTML
       - `httr2`
       - `rvest`
       - JSON
     - [PA #11](./preassigns/pa03.html)
     - [Lab #11](./labs/lab03.html)
     - [Mini-Project #04](./miniprojects/mini01.html) assigned. 
       Due `r due_date_mini(4)`
     - Peer Feedback on Mini Project #03 due 
       `r due_date_mini(3, feedback=TRUE)`

   * - 12
     - `r class_date (12)`
     - 
       - Strings & Regular Expressions
       - Classical Statistical Modeling in `R`
     - [PA #12](./preassigns/pa12.html)
     - [Lab #12](./labs/lab12.html)
    
   * - 13
     - `r class_date (13)`
     - Predictive Modeling in `R`
     - [PA #13](./preassigns/pa03.html)
     - [Lab #13](./labs/lab03.html)
     - 
     - Peer Feedback on Mini Project #04 due 
       `r due_date_mini(4, feedback=TRUE)`

   * - 14
     - `r class_date(14)`
     <td colspan="5" style="text-align:center"> **Final Presentations** </td>
::: 

**`All syllabus provisions subject to change with suitable advance notice.`**
