---
title: "{{< var course.short >}} - Course Syllabus"
filters: 
 - list-table
 - highlight-text
format:
    html: 
        page-layout: "full"
---

**`All syllabus and course schedule provisions subject to change with suitable advance notice.`**

### Instructor

-   [Professor Michael Weylandt](https://michael-weylandt.com)
-   Department of Information Systems & Statistics
-   Zicklin School of Business
-   Baruch College, CUNY

### Course Meetings

#### Lecture / Lab Sessions

-   Virtual (Synchronous Online)
-   Thursdays 6:05pm-9:00pm
    -   Zoom link provided via Brightspace

#### Office Hours

-   In-Person
    -   [Baruch Main Campus (1 Bernard Baruch Way)](https://www.baruch.cuny.edu/campus-map-and-directions/)
    -   Newman Vertical Campus (NVC) 11-246
    -   Wednesdays 5:00-5:45pm
-   Virtual:
    -   Thursdays 5:00pm-5:45pm
    -   Zoom link provided via Brightspace

### Grading

-   24% [Weekly Pre-Assignments](./preassignments.qmd)
    -   3% Each (best eight of ten: lowest two dropped)
    -   Submission *via* Brightspace
-   30% [Mini Projects #01-#04](./miniprojects.qmd) (Best 3 of 4)
    -   8% per submission
    -   2% per peer feedback (scored *via* meta-review)
-   46% [Course Project](./project.qmd)
    -   1% Team Contract
    -   5% Proposal
    -   5% Mid-term Check-In
    -   2.5% Midterm Peer Evaluation
    -   7.5% Final Presentation
    -   7.5% Group Summary Report
    -   12.5% Individual Technical Report
    -   5% Final Peer Evaluation

Points will be aggregated for a maximum class score of 1000, so *e.g.*,
each pre-assignment will have a max score of 30, while the final peer
evaluation will have a max score of 50. 

Final course grades will be curved in accordance with relevant program,
departmental, school, and college policies.[^1]

*Opportunities for Extra Credit:* Several opportunities for extra credit
will be made available. These include:

1. Participation in Course Discussion Board (Piazza)
2. Correction of errors in published course materials (*via* Github pull request)
3. Contributions to and enhancements of the course pre-assignments and in-class
   activities. These should be submitted as a Github pull request against
   [the Course Repository](https://github.com/michaelweylandt/STA9750). 
   
   Note that contributions to course materials don't have to be large to be
   valuable. You can simply clarify points that were not obvious to you, 
   add new auto-graded exercises (see existing materials for examples), create 
   new labs, or even add whole new topics. 
   
   You can take part in this extra credit even if you are very new to `R`: if
   anything, being a recent learner helps you approach this topic with new eyes
   and to identify un-clear or "dangerous" edges. 
   
   Obviously, you should try to make your contributions as accurate as 
   possible, but avoid letting yourself be paralyzed by concern: the 
   instructor will review any contributions before making anything "official."

<!-- 
4. Submission of homework from other Baruch {{< var course.semester >}} courses
   completed using `R`. (Note that the assignment must not ordinarily require
   `R`: this is *extra* credit.) 

   Submission of work from other courses must include: 

   1) the syllabus from the other course; 
   2) the original assignment text and other relevant materials; 
   3) the _graded_ returned submission. 
    
   Work submitted for extra credit must be submitted no later than the final
   day of class. Students may submit no more than 2 assignments for extra credit
   per semester, so think carefully about which assignments you wish to submit.

-->

Extra credit will be awarded at the instructor's sole discretion. 

[^1]: Theoretically, this may result in scores equivalent to an `A` in an
un-curved course receiving a lower grade in this course. In practice, the
instructor will design course assessments to induce a range of scores and
does not anticipate "down-curving" happening.

#### Regrading Policy

If you feel an assignment has been improperly graded, please contact the
instructor using the appropriate Brightspace quiz ("Regrade Request Form")
within 48 hours of the graded assignment being returned. Note that the 
instructor will regrade the assignment *de novo*, so your grade may be
adjusted upwards or downwards. Requests for re-grades after 48 hours will
not be honored outside of exceptional circumstances.

#### Late Work Policy

This course has *many* moving pieces and late work will not
be accepted. Specifically, late submission slows down the
peer review cycle used in this course. All assignments can
be submitted multiple times on Brightspace, so you are
**strongly** encouraged to submit early and to submit often.
Students are also encouraged to load the [course deadlines
file](./key_dates_students_gcal.csv) into their personal
calendar to better track key course dates.

Please note that work **must be submitted on Brightspace** unless 
*specifically* otherwise stated in the written course description or 
approved *in writing* by the instructor at least 48 hours before 
the deadline. Email submissions are not allowed.[^email]

[^email]: Though ubiquitous, email is a remarkably 'flaky' service, 
providing the sender no way to guarantee their message arrives
untampered and providing the recipient no way to guarantee the provenance 
of a message received. (This is not quite true: there are 
[tools for more secure email](https://gnupg.org) but they are somewhat 
more difficult to use and are not supported at CUNY.)
Brightspace is integrated with CUNY's Identity Verification Services and
allows students to guarantee correct submission. Note that Brightspace does
not, by default, send students an email confirming submission, but I believe
this is an option that can be enabled on the student's end. 

Technology issues are not an approved excuse for late submissions; in
particular, Brightspace issues are not an approved excuse for late submission
of the [Mini-Projects](./miniprojects.qmd). Brightspace support is provided by
[BCTC](https://bctc.baruch.cuny.edu/students/) and the [BCTC Help
Desk](https://bctc.baruch.cuny.edu/instructional-technology/services/service-now/). 
The Help Desk is not staffed 24 hours per day, so you are encouraged to submit
early and submit often to avoid issues. The course staff is able to provide 
additional support for personal computer issues, including `git` and `RStudio`
during regularly-scheduled Office Hours, but BCTC should be your first point
of contact.

Individual exceptions to the above policies will be made i) with *prior written*
approval of instructor *at least 48 hours* before the deadline; or ii) *ex post*
with written note from the [Office of the Dean of
Students](https://studentaffairs.baruch.cuny.edu/dean-of-students/). Please note
that, in particular: 

- The instructor is not able to provide *ex post* exceptions; these must be
  handled through [DoS's Notice of Absence
  procedures](https://studentaffairs.baruch.cuny.edu/dean-of-students/notice-of-absence/). 
- Verbal exceptions discussed in Office Hours or after class are not sufficient.
  You must follow-up on any discussions and get my written pre-approval by
  email or on the course discussion board. 

### Tentative Course Schedule

```{r}
#| echo: false
#| message: false
#| warning: false
library(tidyverse)
class_date <- function(n){
    read_csv(date_file) |>
        filter(`Course Element` == "Class Session") |>
        filter(`Item Number` == n) |>
        pull(Date) |>
        format("%Y-%m-%d")
}

class_topic <- function(n){
    read_csv(date_file) |>
        filter(`Course Element` == "Class Session") |>
        filter(`Item Number` == n) |>
        pull(Details)
}
```

```{r}
#| echo: false
#| message: false
library(tidyverse)
date_file <- "key_dates.csv"
holiday_date <- function(n, details=FALSE){
    h <- read_csv(date_file) |>
            filter(`Course Element` == "No Class") |>
            filter(`Item Number` == n)
    
    if(details){
        h |> pull(Details)
    } else {
        h |> pull(Date) |> format("%Y-%m-%d")
    }
}

due_date_mini <- function(n, submission=TRUE, feedback=!submission){
    details <- if(feedback) "Mini-Project Peer Feedback Due" else "Mini-Project Due"
    
    read_csv(date_file) |>
        filter(`Course Element` == "Mini-Projects",
               `Details` == details) |>
        filter(`Item Number` == n) |>
        mutate(deadline = paste(format(Date, "%A %B %d, %Y"), "at", Time)) |>
        pull(deadline)
}

syllabus_quiz <- function(){
    read_csv(date_file) |>
        filter(`Course Element` == "Syllabus Quiz") |>
        pull(Date) |> 
        format("%Y-%m-%d")
}

final_report_date <- function(){
    read_csv(date_file) |>
        filter(`Course Element` == "Course Project") |>
        pull(Date) |> 
        max() |>
        format("%Y-%m-%d")
}

final_grades_date <- function(){
    read_csv(date_file) |>
        filter(`Course Element` == "Grades Posted") |>
        pull(Date) |> 
        max() |>
        format("%Y-%m-%d")
}
```


::: {.list-table width="1,1,1,1,1,1,1,1" aligns="c,c,l,l,l,l,l,l" header-rows=1 .hover}

   * - Week 
     - Lecture Date
     - Topics
     - Pre-Assignment
     - In-Class Lab
     - Mini-Projects
     - Peer Feedback
     - Additional Notes
     
   * - 1
     - `r class_date (01)`
     - `r class_topic(01)`
       1. `R` and `RStudio`
       2. `git` and GitHub
     - None
     - [Lab #01](./labs/lab01.qmd)
     - [Mini-Project #00](./mini/mini00.qmd) assigned. 
       Due `r due_date_mini(0)`
       
   * - 2
     - `r class_date (02)`
     - **Introduction to Course Project**
     
       `r class_topic(02)`
     
       1. Introduction to Reproducible Research:
          - Markdown and Quarto
          - GitHub Pages
       2. Getting Help
       3. Interactive Use of `R`
          a. Use of the REPL
          b. Installing and Loading Packages
     - [PA #02](./pre/pa02.qmd) due at start of class (6:00pm)
     - [Lab #02](./labs/lab02.qmd)
     
   * - 
     - `r holiday_date(1)`
     <td colspan="4" style="text-align:center"> ***`r holiday_date(1, details=TRUE)`*** </td>
     - 
     - Syllabus Quiz / Verification of Enrollment due on `r syllabus_quiz()`

   * - 3
     - `r class_date (03)`
     - **Special Presentation: Library Data Science Resources**
     
     
        `r class_topic(03)`
     
       1. Vectors
       2. Subsetting
       3. Control Flow
       4. Function Calls
       5. Data Frames
       
     - [PA #03](./pre/pa03.qmd) due at start of class (6:00pm)
     - [Lab #03](./labs/lab03.qmd)
     - [Mini-Project #00](./mini/mini00.qmd) due `r due_date_mini(0)`.
     
       [Mini-Project #01](./mini/mini01.qmd) assigned; due `r due_date_mini(1)`
     - Peer Feedback on Mini Project #00 assigned; 
       due `r due_date_mini(0, feedback=TRUE)`
     - Project Team Rosters due at 11:59pm
     
       Professor will assign teams for 'unmatched'
       students at the end of class
       
   * - 4
     - `r class_date (04)`
     - <td colspan="4" style="text-align:center"> 
       **In-Class Project Proposal Presentations** <br> 
       *Optional Enrichment Topic: Additional R Basics* 
       </td>
     - Team Contract and Slides due at 6:00pm

   * - 5
     - `r class_date (05)`
     - `r class_topic(05)`
       1. Single Table Operations
       2. Factors
       3. Dates & Times
       
       Review of `git` fundamentals.
     - [PA #04](./pre/pa04.qmd) due at start of class (6:00pm)
     - [Lab #04](./labs/lab04.qmd)


   * - 6
     - `r class_date(06)`
     - `r class_topic(06)`
       1. Multi-Table Operations (Joins)
       2. Pivots
       3. Working with Missing Data
       
       Debugging Techniques
     - [PA #05](./pre/pa05.qmd) due at start of class (6:00pm)
     - [Lab #05](./labs/lab05.qmd)
     - [Mini-Project #01](./mini/mini01.qmd) due `r due_date_mini(1)`.
     
       [Mini-Project #02](./mini/mini02.qmd) assigned. 
       Due `r due_date_mini(2)`
     - Peer Feedback on Mini Project #01 assigned; due 
       `r due_date_mini(1, feedback=TRUE)`

   * - 7
     - `r class_date (07)`
     - `r class_topic(07)`
       - Introduction to `ggplot2`
       - Customizing Formatting
     - [PA #06](./pre/pa06.qmd) due at start of class (6:00pm)
     - [Lab #06](./labs/lab06.qmd)
     
   * - 8
     - `r class_date (08)`
     <td colspan="5" style="text-align:center"> 
     **Mid-Semester Check-In Presentations** <br> 
     *Optional Enrichment Topic: SQL* 
     </td>
     - Mid-Semester Peer Evaluation and Slides due at 6:00pm
     
   * - 
     - `r holiday_date(2)`
     <td colspan="3" style="text-align:center"> ***`r holiday_date(2, details=TRUE)`*** </td>
     - [Mini-Project #02](./mini/mini02.qmd) due `r due_date_mini(2)`.
     
       [Mini-Project #03](./mini/mini03.qmd) assigned. 
       Due `r due_date_mini(3)`
     - Peer Feedback on Mini Project #02 assigned; due 
       `r due_date_mini(2, feedback=TRUE)`

   * - 
     - `r holiday_date(3)`
     <td colspan="3" style="text-align:center"> ***`r holiday_date(3, details=TRUE)`*** </td>

   * - 9
     - `r class_date (09)`
     - `r class_topic(09)`
     - [PA #07](./pre/pa07.qmd) due at start of class (6:00pm)
     - [Lab #07](./labs/lab07.qmd)
     
     
   * - 10
     - `r class_date (10)` [**NB: Tuesday**]{color="#FF3333"}
     - `r class_topic(10)`
       - Maps
       - `shiny`
       - `ggplot2` Extensions for Animated and Interactive Plots
     - [PA #08](./pre/pa08.qmd) due at start of class (6:00pm)
     - [Lab #08](./labs/lab08.qmd)
     - 
     - 
   * - 11
     - `r class_date (11)`
     - `r class_topic(11)`
       - Introduction to HTTP and HTML
       - `httr2`
       - `rvest`
       - JSON
     - [PA #09](./pre/pa09.qmd) due at start of class (6:00pm)
     - [Lab #09](./labs/lab09.qmd)
     - [Mini-Project #04](./mini/mini04.qmd) assigned. 
       Due `r due_date_mini(4)`
     - Peer Feedback on Mini Project #03 assigned; due 
       `r due_date_mini(3, feedback=TRUE)`

   * - 12
     - `r class_date (12)`
     - `r class_topic(12)`
       - Strings & Regular Expressions
       - Classical Statistical Modeling in `R`
     - [PA #10](./pre/pa10.qmd) due at start of class (6:00pm)
     - [Lab #10](./labs/lab10.qmd)
    
   * - 13
     - `r class_date (13)`
     <td colspan="5" style="text-align:center"> 
       **Final Presentations** <br> 
       *Optional Enrichment Topic: Functional Programming in `R`*
     </td>
     - Slides due at 6:00pm
     
   * - 14
     - `r class_date(14)`
     - **Course Wrap-Up** </br>
     
       `r class_topic(14)`
     
     - [PA #11](./pre/pa11.qmd) due at start of class (6:00pm)
     - [Lab #11](./labs/lab11.qmd)
     - [Mini-Project #04](./mini/mini04.qmd) due `r due_date_mini(4)`.
     - Peer Feedback on Mini Project #04 assigned; due 
       `r due_date_mini(4, feedback=TRUE)`
     
   * - 
     - Finals Period
      <td colspan="6" style="text-align:center"> 
        <p>Group Report, Individual Report, Final Peer Evaluations 
           due `r final_report_date()` (Tentatively)</p> 
        </br> 
        <p>Final Grades posted to CUNYFirst by `r final_grades_date()`</p>
    </td>
      
::: 

### Syllabus Quiz

After the first day of class, a *syllabus quiz* will be available on
Brightspace. This quiz covers the key points of this syllabus and course
policies. You must pass this quiz with 100% to be allowed to remain in this
course. If you have not passed the syllabus quiz by the 
[Verification of Enrollment](https://enrollmentmanagement.baruch.cuny.edu/registrar/verification-of-enrollment/)
deadline, you may receive a `WN` grade and be removed from this course.

### Workload Expectations

The following _approximate_ breakdown of expected course workload is
intended to help you properly prepare for and schedule the out-of-class
work associated with this course. Note that, persuant to relevant
[Federal](https://www.ecfr.gov/current/title-34/part-600#p-600.2(Credit%20hour))
and [State](https://govt.westlaw.com/nycrr/Document/Ieca5c8abc22111dd97adcd755bda2840)
regulations, a 3-credit course taken over a 15 week semester should
require approximately 6 hours of out-of-class work from an _average_ 
student, or 90 hours total over the course of the semester.[^avg][^gc] This
time should be roughly allocated as: 

- Review of Lecture Materials (5 hours = 0.5 hours $\times$ 10 lectures)
- Weekly Pre-Assignments (10 hours = 1 hour $\times$ 10 Pre-Assignments)
- Mini-Projects: (45 hours total)
  - Mini-Project #00 (3 hours):
    - Initial Submission (2 hours)
    - Peer Feedback (1 hour)
  - Mini-Projects #01-#04 (42 hours total, assuming skipping one 3 MP)
    - Initial Submission (39 hours = 13 hours $\times$ 3 MPs)
    - Peer Feedback (3 hours = 1 hour $\times$ 3 MPs)
- Course Project (30 hours)
  - Team Formation & Proposal Presentation (2 hours)
  - Check-In Presentation & Peer Evaluation (2 hours)
  - Final Presentation (3 hours)
  - Group Report & Peer Evaluation (3 hours)
  - Individual Report (20 hours)

Note that, for the course project, the individual and group reports will
require work throughout the semester, not simply at the end, though the majority
of the effort is likely required in the latter half of the semester. As such,
the first two Mini-Projects are designed to take more effort than the final two.

[^avg]: For this course, an _average_ student is a student who enters the
course with: 

    i) Basic computer literacy, including use of the file system, plain text
       files and editors, *etc.*; 
    ii)  A small amount of programming experience, not necessarily in `R`; and
    iii) Fluency with statistics and data analysis at the level of (at least) 
         STA 9708, ideally STA 9700; 

    <br>and is earning a **B**-range grade. If you have less background or are 
    aiming for a higher grade, you should expect to commit proportionally
    more time to this course. 
    
    If you lack the prerequisite background listed above or
    simply wish to review it before the semester begins in
    earnest, please reach out to the instructor and I will be
    more than happy to provide supplementary readings. 

[^gc]: The CUNY Graduate Center has a useful [summary of these
expectations](https://gc.catalog.cuny.edu/general-information/credithourpolicy).
Baruch courses follow the same standards. See also [CUNY Central
Policy](https://www.cuny.edu/wp-content/uploads/sites/4/page-assets/about/administration/offices/registrar/resources/State-and-Federal-Credit-Hour-Definitions-08022019-2-1.pdf).

**`All syllabus and course schedule provisions subject to change with suitable advance notice.`**
