---
title: "{{< var course.short >}}: Tips and Tricks"
---

```{r}
#| echo: false
#| purl: true
#| warning: false
library(yaml)
variables <- read_yaml("https://raw.githubusercontent.com/michaelweylandt/STA9750/refs/heads/main/_variables.yml")
course_repo  <- variables$course$repo
course_short <- variables$course$short
```

## Miscellaneous `RStudio` and `Git` Advice

1. <a name="visualeditor"></a>
   When you are writing `quarto` documents in `RStudio`, 
   **do not use** the "Visual" editor mode. It will seem
   tempting at first, but it will lead to difficult
   and hard to debug errors down the road. You should
   disable it in your `RStudio` settings (if possible)
   and include `editor: source` in the header of `quarto`
   documents to keep it from opening automatically. 
2. <a name="gitmisc"></a>
   `git` supports a huge variety of programming workflow
   and was originally designed to allow large teams of
   programmers to work together. This flexibility leads
   to a few "sharp edges", however. 
   
   You will have the best experience with `git` in this
   course if: 
   
   1) <a name="commitoften"></a>
      You `commit` early and often. `git` commits are
      essentially free (40 bytes - less than a text!) and
      give you the ability to roll back changes if you get
      into a bad situation. 
      
   2) <a name="singlecomputer"></a>
      You do all your work on a single computer. Keeping
      files and `git` synchronized across multiple machines
      is possible, but trickier. If, for whatever reason,
      you need to use multiple computers, reach out to
      course staff in office hours for advice.
      
      In particular, even though it may be tempting to edit files
      directly through the GitHub web interface, resist! You will
      get into situations where your local copy and the copy on
      GitHub are out of sync and you will have issues pushing
      your work onto GitHub. 
      
      (It is possible to fix this, but you'll have to learn a lot
      more about how `git` works, far beyond what is expected of
      you in this course.)
      
   3) <a name="pushoften"></a>
      You `git push` at the end of every work session. This
      is the stage when things are most likely to go wrong,
      so you want to be `push`ing early and often, giving
      you the best chance to spot and fix an issues early. 
      
   4) <a name="filesystem"></a>
      You work in a "regular" file system. Synchronization
      and back up tools like DropBox or OneDrive perform 
      magic behind the scenes that doesn't always play well
      with `git`. You don't need to use these tools with
      `git` since `git` (used properly and in conjunction
      with `GitHub`) will give you automatic history and 
      backups. 
      
   5) <a name="antivirus"></a>
      You avoid anti-virus issues. If you are on a Windows
      machine, tell your anti-virus software to ignore the
      directory containing your work for this course. Both
      `git` and `quarto` create and delete lots of little
      files as they execute; certain anti-virus tools
      interpret this as evidence of a cyber attack and
      attempt to lock the file system, leading to a variety
      of subtle and hard to explain errors. 
      
   6) <a name="gitadd"></a>
      You are careful about what files you `add` in `git`.
      You should generally only include your `qmd` files and
      the rendered outputs in the `docs` folder. The 
      `build_site` script included in [Mini-Project
      #00](./miniprojects/mini00.html) is helpful, but not
      bullet-proof, here. 

3. <a name="filenames"></a>
   Use the file names and directory structures specified in
   the course instructions. I use a variety of automated
   processing tools to coordinate this course and if you do
   not follow instructions closely, your assignments will
   not be processed. I will typically manually handle your
   submissions, but with a penalty applied to your grade.
   It is better for all parties if you avoid this. 
   
   The helper scripts below can be used to confirm whether
   your submissions are properly formatted. 
4. <a name="nospaces"></a>Avoid using file or directory names with spaces in 
   them. Some software, particularly on Windows computers, does not handle
   spaces in directory or file names properly. 

## Helper Scripts

The following helper scripts can be used to help
automate several students tasks in this course. To use
any of them, run

> `source("https://michael-weylandt.com/STA9750/load_helpers.R")`

and then run the name of the relevant script. 

### Open Mini-Project GitHub Issue

The following functon will create the GitHub issue necessary to submit a 
mini-project: 

```{r}
#| label: mp_submission_create
mp_submission_create <- function(N, github_id){
  library(rvest)
  library(glue)
  library(tidyverse)
  library(httr2)
  library(gh)
  
  if(missing(N)){
    N <- menu(title="Which Mini-Project would you like to submit on GitHub?", 
              choices=c(0, 1, 2, 3, 4))
  }
    
  mp_url <- glue("https://michael-weylandt.com/STA9750/miniprojects/mini0{N}.html")
  
  mp_text <- read_html(mp_url) |> html_element("#submission-text") |> html_text()
  
  if(missing(github_id)){
    github_id <- readline("What is your GitHub ID? ")
  }
  
  title <- glue("{course_short} {github_id} MiniProject #0{N}")
  
  body <- mp_text |> str_replace("<GITHUB_ID>", github_id)
  
  r <- request("https://github.com/") |>
         req_url_path_append(github_id) |>
         req_url_path_append(course_repo) |>
         req_url_path_append("issues/new") |>
         req_url_query(title=title, 
                       body=body) 
  
  browseURL(r$url)
}
```

### Verify Mini-Project Submission {#mp-submission-verify}

The following function will confirm that the submission issue is properly
formatted. 

```{r}
#| label: mp_submission_verify
mp_submission_verify <- function(N, github_id){
  library(rvest)
  library(glue)
  library(tidyverse)
  library(httr2)
  library(gh)
    
  if(missing(N)){
    N <- menu(title="Which Mini-Project would you like to check was properly submitted on GitHub?", 
              choices=c(0, 1, 2, 3, 4))
  }
  
  mp_url <- glue("https://michael-weylandt.com/STA9750/miniprojects/mini0{N}.html")
  
  mp_text <- read_html(mp_url) |> html_element("#submission-text") |> html_text()
  
  if(missing(github_id)){
    github_id <- readline("What is your GitHub ID? ")
  }
  
  title <- glue("{course_short} {github_id} MiniProject #0{N}")
  
  body <- mp_text |> str_replace("<GITHUB_ID>", github_id) |> str_squish()
  
  issues <- gh("/repos/{owner}/{repo}/issues?state=all", 
               owner=github_id, 
               repo=course_repo)
  
  issue_names <- vapply(issues, function(x) str_squish(x$title), "")
  
  name_match <- which(issue_names == title)
  
  if(length(name_match) == 0){
    cat("I could not find a unique issue with the title:\n", 
        "    ", sQuote(title),"\n",
        "The issues I found had the following titles:\n",
        paste(c("", issue_names), collapse="\n - "), "\n",
        "If something on that list looks correct, please check",
        "capitalization and punctuation.\n")
      
    case_insensitive_match <- which(tolower(issue_names) == tolower(title))
        
    if(length(case_insensitive_match) > 0){
      cat("In particular, please check", 
          sQuote(issue_names[case_insensitive_match]), 
          "which appears to be the same, modulo capitalization.\n")
    }
         
    stop("MINIPROJECT NOT SUBMITTED CORRECTLY.")
  } else if(length(name_match) > 1){
    cat("I found multiple issues with the title:\n", 
        "    ", sQuote(title),"\n",
        "Please change the names of issues so that there is only", 
        "issue with the desired name. (Note that closing an issue is",
        "not sufficient.)\n")
          
    stop("MINIPROJECT NOT SUBMITTED CORRECTLY.")
  }
  
  issue <- issues[[name_match]]
  
  issue_num <- issue$number
  issue_url <- issue$html_url
  issue_body <- issue$body |> str_squish()
  
  cat("Identified Issue #", issue_num, "at", issue_url, "as possible candidate.\n")
  
  if(issue$state != "open"){
    cat("Issue does not appear to be in 'open' status. Please",
        "confirm the issue is open and try again.\n")
    stop("MINIPROJECT NOT SUBMITTED CORRECTLY.")
  }
  
  if(issue_body != body){
    cat("Issue does not appear to have the correct body text. This is not",
        "necessarily an issue, but you may want to confirm all relevant",
        "information is included.\n")
  }
  
  # This isn't quite general enough, but I think it covers everything
  # we'll see in this course. 
  URL_REGEX <- "http[A-Za-z0-9:/\\-\\.]*"
    
  urls <- str_match_all(paste(issue_body, "\n Thanks!\n", sep=""), 
                        URL_REGEX)[[1]]
  expected_url <- str_extract(body, URL_REGEX)
  
  if(length(urls) > 1){
    cat("The following URLs were found in the issue text:",
        paste(c("", urls), collapse="\n - "), 
        "Only one URL was expected; please remove any extras.\n")
    stop("MINIPROJECT NOT SUBMITTED CORRECTLY.")
  }
 
  submitted_url <- urls[1]
  
  if(submitted_url != expected_url){
    cat("The submitted URL does not match the Mini-Project instructions.\n", 
        "Expected:\n - ", expected_url, "\n", 
        "Submitted:\n - ", submitted_url, "\n", 
        "Please correct any differences and try again.\n")
        stop("MINIPROJECT NOT SUBMITTED CORRECTLY.")
  }
  
  resp <- request(submitted_url) |> 
      req_error(is_error = \(r) FALSE) |>
      req_perform()
  
  if(resp_is_error(resp)){
    cat("Something appears to be incorrect at the URL: ", 
        submitted_url, 
        "\n Please confirm that it is working as expected and try again.")
    browseURL(submitted_url)
    stop("MINIPROJECT NOT SUBMITTED SUCCESSFULLY.")
  }
  
  if(N == 0){
    raw_url <- glue("https://raw.githubusercontent.com/{github_id}/{course_repo}/refs/heads/main/index.qmd")
  } else {
    raw_url <- glue("https://raw.githubusercontent.com/{github_id}/{course_repo}/refs/heads/main/mp0{N}.qmd")
  }
  
  resp_raw <- request(raw_url) |> 
      req_error(is_error = \(r) FALSE) |>
      req_perform()
  
  if(resp_is_error(resp_raw)){
      cat("I cannot find the source qmd document at", raw_url, ".\n",
          "Please confirm it was correctly submitted and try again.\n",
          "This document is needed for automated code quality checks.\n")
      
    stop("MINIPROJECT NOT SUBMITTED SUCCESSFULLY.")
  }
  
  cat("Congratulations! Your mini-project appears to have been submitted correctly!\n")
  invisible(TRUE)
}
```

### Create Peer Feedback Comment

As of January 2025, GitHub does not support pre-filling issue comments via URL
so I am currently unable to pre-load the peer feedback template for you. For now,
you should be able to copy the [template on the mini-project 
overview page](./miniprojects.html#peer-feedback-template). (Note the "copy"
button in the top right corner of the code-formatted block.)

### Verify Peer Feedback Properly Formatted: 

```{r}
#| label: mp_feedback_verify
mp_feedback_verify <- function(N, github_id, peer_id){
  library(rvest)
  library(glue)
  library(tidyverse)
  library(httr2)
  library(gh)
    
  if(missing(N)){
    N <- menu(title="Which Mini-Project's Peer Feedback would you like to check was properly submitted on GitHub?", 
              choices=c(0, 1, 2, 3, 4))
  }
  
  if(missing(github_id)){
    github_id <- readline("What is your GitHub ID? ")
  }

  if(missing(peer_id)){
    peer_id <- readline("What is your Peer's GitHub ID? ")
  }
 
  template_url <- "https://michael-weylandt.com/STA9750/miniprojects.html"
  
  template_text <- read_html(template_url) |> 
    html_element("#peer-feedback-template") |> 
    html_text() |>
    str_trim() |>
    str_replace_all("NN", "(.*)") |>
    str_replace_all("TEXT TEXT TEXT", "(.*)")
  
  title <- glue("{course_short} {github_id} MiniProject #0{N}")

  issues <- gh("/repos/{owner}/{repo}/issues?state=all", 
               owner=peer_id, 
               repo=course_repo)
  
  issue_names <- vapply(issues, function(x) str_squish(x$title), "")
  
  name_match <- which(issue_names == title)
  
  if(length(name_match) != 1){
      cat("I could not find a unique issue with the title:\n", 
          "    ", sQuote(title),"\n",
          "I'm afraid I can't verify whether the peer feedback was submitted properly.\n")
          stop("PEER FEEDBACK NOT VERIFIED.")
  } 
  
  issue <- issues[[name_match]]
  
  issue_url <- issue$html_url
  issue_comment_url <- issue$comment_url
  
  comments <- gh(issue_comment_url)
  
  commenters <- vapply(comments, function(x) x$user$login, "")
  
  comment_num <- which(commenters == github_id)
  
  if(length(comment_num) != 1){
    cat("I cannot identify your comment on the issue at", 
       sQuote(issue_url), ". Please verify that this is the correct link.\n")
       
    if(length(comment_num) == 0){
      cat("You do not appear to have commented on this issue.")
    } else {
      cat("You have left multiple comments on this issue.",
          "Please consolidate your feedback")
    }
    stop("PEER FEEDBACK NOT VERIFIED.")
  }
  
  comment_body <- comment$body
  
  cat("I have found your comment, with the following text:")
  
  cat(comment_body)
  
  comment_body  <- comment_body |> str_squish()
  template_text <- template_text |> str_squish()
  
  matches <- str_match_all(comment_body, template_text)[[1]][-1]
  
  if(anyNA(matches)){
    cat("I couldn't match the template to your comment. Please modify and try again.")
    stop("PEER FEEDBACK NOT VERIFIED.")
  }
  
  cat(glue("Congratulations! Your peer feedback on {peer_id}'s MP #0{N} appears properly formatted. Thank you for your contributions to {course_short}"))
  
  invisible(TRUE)
  
}
```

### Count Words

The following script can be used to count words on a
Quarto-generated web page. 

```{r}
#| label: count_words
count_words <- function(url){
    library(rvest)
    library(stringi)
    
    # Note that this includes code inside an inline block, but omits
    # full-sized code blocks
    read_html(url) |>
        html_elements("main p") |>
        html_text() |>
        stri_count_words() |>
        sum()
}
```

### Lint Submission

The following will automatically run the `lintr` package on a
submitted `qmd` document. 

```{r}
#| label: lint_submission
lint_submission <- function(N, peer_id){
  library(rvest)
  library(glue)
  library(tidyverse)
  library(httr2)
  library(lintr)

  if(missing(N)){
    N <- menu(title="Which Mini-Project submission would you like to lint?", 
              choices=c(1, 2, 3, 4))
  }
    
  if(missing(peer_id)){
    peer_id <- readline("What is your Peer's GitHub ID? ")
  }
    
  raw_url <- glue("https://raw.githubusercontent.com/{peer_id}/{course_repo}/refs/heads/main/mp0{N}.qmd")
  
  cat("Attempting to access qmd source at", raw_url, ".\n")
  
  resp <- request(raw_url) |> 
      req_error(is_error = \(r) FALSE) |>
      req_perform()
  
  if(resp_is_error(resp)){
      cat("I could not access the raw qmd document. Attempting to open in browser...\n")
      browseURL(resp)
      return(FALSE)
  } else {
      cat("I was able to access the raw qmd document. Beginning to lint.\n")
      
      tf <- tempfile(pattern=glue("mp0{N}_{peer_id}_lint_document.qmd"))
      
      writeLines(resp_body_string(resp), tf)
      
      lintr::lint(tf)
  }
}

```
