# Optional Review

## Exercise #01

::: {#q1}

```{r}
#| label: q1-cell
#| echo: true
f <- function(x){
  cat("The vector has", length(x), "elements and is of type", class(x))
}
```

:::

**Q:** Write a function `f` that does the following:

```{r}
#| echo: true
f(c(1, 2, 3))
f(c("a", "b", "c"))
f(1:5)
```

*Hint*: Use the `cat` function to print to screen. 

. . . 

{{< contents q1 >}}

## Exercise #02

::: {#q2}

```{r}
#| label: q2-cell
#| echo: true
is_even <- function(x)  (x %% 2) == 0
```

:::

**Q:** Write a *vectorized* function to tell if numbers are even. 

```{r}
#| echo: true
is_even(3)
is_even(c(3, 4, 5))
is_even(1:10)
```

*Hint*: Use the `%%` operator to get remainders from division

. . . 

{{< contents q2 >}}


## Exercise #03

::: {#q3}

```{r}
#| label: q3-cell
#| echo: true
count_even <- function(x)  sum(is_even(x))
```

:::

**Q:** Write a function to count the even elements of a vector: 

```{r}
#| echo: true
count_even(3)
count_even(c(3, 4, 5))
count_even(1:10)
```

*Hint*: Combine Q2 with the "sum of logical = count" trick. 

. . . 

{{< contents q3 >}}


## Exercise #04


**Q:** The `seq()` function lets us construct sequences. What is the average (`mean`) of the first 23 odd numbers?

*Hint*: Read the documentation for `?seq` before trying this question

. . . 

```{r}
#| echo: true
mean(seq(from=1, by=2, length.out=23))
```



## Exercise #05

**Q:** The *alternating harmonic series* is defined as: 

$$1 - \frac{1}{2} + \frac{1}{3} - \frac{1}{4} + \frac{1}{5} + \dots$$

Show that this series converges to $\ln(2)$ by taking a partial sum
of the first one million elements. 

*Hint*: Use recycling to get the signs right: 

```{r}
#| echo: true
seq(1, 5) * c(1, -1)
```

. . . 

```{r}
#| echo: true
sum(1/seq(1, 1e6) * c(1, -1))
```

Compare to: 

```{r}
#| echo: true
log(2)
```

## Exercise #06

**Q:** Write a function that computes the mean of a vector (Don't use the 
built-in `mean` function)

*Hint*: Use the `sum` and `length` functions

. . . 

```{r}
#| echo: true
my_mean <- function(x){
  sum(x) / length(x)
}

my_mean(8)
my_mean(1:10)
```


## Exercise #07

::: {#q7}

```{r}
#| label: q7-cell
#| echo: true
pos_sqrt <- function(x) sqrt(pmax(0, x))
```

:::

**Q:** Write a vectorized function that returns $\sqrt{x}$ if $x$ is positive and 0 otherwise. 

```{r}
#| echo: true
pos_sqrt(c(-4, -1, 1, 4, 9, -9))
```

*Hint*: The `pmax` ("parallel max") function may be useful here: 

```{r}
#| echo: true
x <- c(1, 3, 5,  7)
y <- c(5, 3, 10, 1)
pmax(x, y)
```

. . . 

{{< contents q7 >}}

Alternative - use the `ifelse` function for *vectorized* conditionals (but this
has a warning because it tries to do both conditions)


## Exercise #08

::: {#q8}

```{r}
#| label: q8-cell
#| echo: true
long_string <- function(x)  x[which.max(nchar(x))]
```

:::

**Q:** Write a function that takes in a vector of characters and returns the longest. 


```{r}
#| echo: true
long_string(c("a", "bc", "def"))
long_string(c("My", "name", "is", "Michael"))
```

*Hint*: The `nchar` and `which.max` functions may be
helpful here. 

. . . 

{{< contents q8 >}}

## Exercise #09

**Q:** Write a function that computes the factorial of a value. (Don't use the built-in factorial function)

*Hint*: Use the `seq` and `prod` functions

. . . 

```{r}
#| echo: true
my_factorial <- function(n){
  return(prod(seq(1, n)))
}

my_factorial(8)
```

## Exercise #10

**Q:** Write a function that computes the factorial of a value using a loop (instead of the `prod` function). 

. . . 

```{r}
#| echo: true
my_factorial <- function(n){
  fact <- 1
  for(x in seq(1, n)){
    fact <- fact * x
  }
  return(fact)
}

my_factorial(8)
```

## Exercise #11

**Q:** Show that: $$\sum_{k=0}^{\infty} \frac{1}{k!} = e$$

Use the built-in `factorial` function since it is vectorized. 

. . . 

```{r}
#| echo: true
sum(1/factorial(seq(0, 10000)))
```

## Exercise #12

**Q:** Write a function that takes a vector and returns the maximum element. (Don't use the built-in `max` function.)

*Hint*: You will need to use a loop and a conditional here. 

*Hint*: We started factorial at 1; start max at `-Inf` (why?)

. . . 

```{r}
#| echo: true
my_max <- function(x){
  max_val <- -Inf
  for(v in x){
    if(v > max_val) max_val <- v
  }
  return(max_val)
}

my_max(c(1, -1, 3, -4, 10, -3))
```
