---
title: "{{< var course.short >}} - Week 3 Update"
author: "Michael Weylandt"
format: revealjs
---

# Today

## Today

- Introduction to Course Project
- Special Presentation
- Course Administration
- `R`: Data Frames, Functions, Packages and Control Flow

# Course Project Overview

## Course Project

- Teams: 3-4 classmates
- Stages: 
  - Proposal (in class presentation)
  - Mid-semester check-in (in class presentation)
  - Final: in class presentation, individual report, summary report
- Structure:
  - Shared "Overarching Question"
  - Individual "Specific Question" (one per teammate)
  
Full description [online](https://michael-weylandt.com/STA9750/project.html)

## Finding Data

- Start early!
- [NYC Open Data](https://opendata.cityofnewyork.us/) is great
  - See also: [FRED](https://fred.stlouisfed.org/), [Federal Open Data](https://data.gov), [Nasa EarthData](https://search.earthdata.nasa.gov), [Kaggle](https://www.kaggle.com/datasets)
  - Ask on Piazza for pointers
  - Lots of data hidden in Wikipedia
- Nothing paid or private without express instructor submission
- Everyone loves _spatial_ data!

## Presentation Hints

- Longest time $\neq$ most important
- Story, story, story! _Why_ are you making these choices? 
- Hourglass Structure
  - Start big
  - Motivate _your_ overarching question
  - Specific questions
  - Tie specific to overarching
  - From overarching back to big motivation
- No less than one figure every other slide

# Special Presentation</br> Jason Amey </br> Newman Library

# Course Administration

## MP#00 Submissions

I now have GitHub IDs for 22 of you. 

- If you are in the 6 who have not sent me your GitHub ID
  **and received my acknowledgement** please do so immediately. 
  
- Will follow up with a few of you for VoE directly

- Great submissions: Resumes, Bios, Flowcharts (creative!), Headshots

## MP#00 Goals

MP#00 is ungraded to: 

i)   Sort out tech trouble
ii)  Show danger of waiting to the last minute
iii) Demonstrate importance of closely following instructions 
Hopefully lessons learned will be helpful in future assignments

## MP#00 Peer Feedback

Due to a constraint of GitHub's API, need to do issue
creation and tagging on _my_ 
`{{< var course.repo >}}` repository.

. . .

Instructions updated for MP#01-#04 already. 

- Helper scripts to be updated as well

For MP#00, I am hand-copying all issues to my repo and
assigning peer feedback there (done by noon tomorrow) 

## MP#00 Peer Feedback

3 evaluators per submission


Goals: 

- Practice commenting on GitHub and reviewing code
  (ungraded assignment)
- Learn tricks to improve your own site

> "Good artists copy; great artists steal." -- Steve Jobs

## MP#01 Extended Deadlines

- Released: today (unchanged)
- Due: **2025-03-05** (extend one week)
- Peer Feedback Assigned: **2025-03-06** (shift one week)
- Peer Feedback Due: **2025-03-12** (shift one week)

MP#01 Peer Feedback is day before Project Proposals

- No Pre-Assignment that week

## {{< var course.short >}} Mini-Project #01

- MP#01 released
- **Start early**
  - Not too hard if everything is working (post-MP#00)
  - Tech support takes time
  
Found on [Course
Site](https://michael-weylandt.com/STA9750/miniprojects/mini01.html)

## Piazza Participation

- Average time to response: 15 minutes
- 154 posts

Thanks to those of you who are helping classmates!

## Upcoming Week

Next Wednesday at 11:45pm: 

- Next Pre-Assignment
- MP#00 Peer Feedback due

# Pre-Assignment #03 FAQs

## FAQ: Vector Index Printout Rules

Default vector printing: 
```{r}
#| echo: true
1:10
```

Each line gets a new index:
```{r}
#| echo: true
sqrt(1:10)
```

More complex objects have alternate print styles:
```{r}
#| echo: true
matrix(1:9, nrow=3, ncol=3)
```

Print width is controlled by `getOption("width")`.

## FAQ: Recycling Rules

Alignment by default: 
```{r}
#| echo: true
x <- 1:3
y <- 4:6
x + y
```

Recycling by default: 
```{r}
#| echo: true
x <- 1
y <- 4:6
x + y
```

Recycle warning when vectors don't fit together cleanly: 
```{r}
#| echo: true
#| warning: true
x <- 1:2
y <- 4:6
x + y
```

## FAQ: Recycling Warning

```{r}
#| echo: true
#| warning: true
x <- 1:2
y <- 4:6
x + y
```

Not a problem _per se_, but often a sign that something has gone wrong.

- scalar + vector is usually safe
- 2 vectors of same size is usually safe
- vectors of different size is usually a programming mistake

## FAQ: Warnings vs Errors

- _Warnings_: heuristics pointing at typical problem
  - Code still executed without a problem
  - Try to fix these unless you're **certain** it's not a problem
- _Errors_: code failed to execute
  - You have to fix these to run your code

## FAQ: Changing built-in functions

Most built-in functions can't / shouldn't be changed. 

Some allow alternate behavior via additional arguments:

```{r}
#| echo: true
log(10) # Default is natural (base e) logarithm
log(10, base=10)
```

If you want different behavior, write your own function: 

```{r}
#| echo: true
cosd <- function(x){
    ## Cosine in degrees
    cos(x * pi / 180)
}
cosd(90)
```

Always try `?name` to see documentation.

## FAQ: Git Workflow

Three key commands: 

- `git add`: add some changes to a 'box'
- `git commit`: seal the 'box'
- `git push`: send the 'box' to GitHub

`Git` pane in `RStudio` shows _uncommited changes_, not files. 

If a file 'vanishes' after a commit, that's good!

# `R`: Data Frames, Functions, Packages and Control Flow

## Data Frames

## Function Calls

## Packages

## Control Flow

# Wrap-Up

## Looking Ahead

Course Project: 

- Start looking for teammates and topics
- No in-person office hours on 2025-02-18 (MW on travel)

{{< include ../advice/professor_recs.qmd >}}

