---
session: "`r session <- 6; session`"
class_date: 
  "`r library(tidyverse); read_csv('key_dates.csv', name_repair='universal') |> filter(Course.Element == 'Class Session', Item.Number==session) |> pull(Date)`"
title: "{{< var course.short >}} </br> Week {{< meta session >}} Update </br> {{< meta class_date >}}"
author: "Michael Weylandt"
format: revealjs
---

## {{< var course.short >}} Mini-Project #01

👏 Thank you to everyone who took part in peer feedback! 👏

I will start processing grades and uploading to Brightspace soon. 
(Moving day for me...)

. . .

I've started reading some submissions - really excellent work! 

- Plots (exciting!), sophisticated and insightful analyses, creative policymaking

. . . 

Hopefully automated checks and personalized emails helpful (new this year)

If you have ideas for other useful automated 'supports' let me know

## {{< var course.short >}} Mini-Project #01

Some popular tips:

- The [`kable`](https://bookdown.org/yihui/rmarkdown-cookbook/kable.html) function
  makes nice tables
- The [`gt`](https://gt.rstudio.com/) package makes very nice tables
- `Quarto` allows "code folding": useful for hiding long boring code blocks

```{r}
#| echo: true
#| code-fold: true
#| code-summary: "Show the code"
1 + 1
```

You can set this globally. You also want to keep `echo: true`.

## {{< var course.short >}} Mini-Project #02

```{r}
#| echo: false
#| message: false
#| warning: false
library(glue)
library(rlang)
library(yaml)
get_mp_title <- function(N){
    mp_file <- glue("miniprojects/mini0{N}.qmd")
    mp_text <- readLines(mp_file, n=50)
    
    header_end <- which(grepl("---", mp_text))[2]
    
    header_info <- yaml.load(readLines(mp_file, n=header_end))
    
    header_info$mp_title %||% "TBD"
}

get_mp_deadline <- function(N){
    read_csv("key_dates.csv") |> 
        filter(`Course Element` == "Mini-Projects", 
               `Item Number`== N, 
               str_detect(Details, "Due"), 
               !str_detect(Details, "Peer")) |>
        mutate(due = paste(Date, "at", Time)) |>
        pull(due)
}


get_pa_deadline <- function(N){
    read_csv("key_dates.csv") |> 
        filter(`Course Element` == "Pre-Assignments", 
               `Item Number`== N, 
               str_detect(Details, "Due"), 
               !str_detect(Details, "Peer")) |>
        mutate(due = paste(Date, "at", Time)) |>
        pull(due)
}
```

[MP#02 in process now](../miniprojcts/mini02.html) - `r get_mp_title(2)`

. . .

**Due `r get_mp_deadline(2)`**

-   GitHub post (used for peer feedback) AND Brightspace

. . .

Pay attention to [the rubric](../miniprojects/mini02.html#rubric)

-   Writing and presentation are about 50% of your grade
-   Evaluated on *rigor* and *thoughtfulness*
-   Use what you learned from MP #01

## Upcoming Mini-Projects

Topics

-   MP#03: `r get_mp_title(3)`
-   MP#04: `r get_mp_title(4)`

## Pre-Assignments

Brightspace - Wednesdays at 11:45

- Reading, typically on course website
- Brightspace auto-grades
  - I have to manually change to completion grading

Next pre-assignment is **`r get_pa_deadline(7)`**

. . . 

I missed a few comments in the previous cycle (sorry!) - trying to catch up 
in the next few days

## Course Support

- Synchronous
  - Office Hours 2x / week
- Asynchronous
  - Piazza ($<30$ minute average response time)

## Upcoming Week

Due Wednesday at 11:45pm:

- Pre-Assignment #07 (Brightspace)
  * Introduction to plotting with `ggplot2`

Expect back:

-   MP#01 consolidated grades
-   Project proposal instructor feedback

## Project Proposal Presentations

[Official Description](https://michael-weylandt.com/STA9750/project.html#project-proposal-presentations)

- **6 minute presentation**
- Key topics:
  - *Animating Question*
  - *Team Roster*
- Also discuss: Possible specific questions, data sources, 
  analytical plan, anticipated challenges

. . . 

Most important: team names! 

Last semester: Rat Pack, Subway Surfers, Going for Gold, EVengers, *etc.*

## After Proposals

100% optional discussion of `dplyr` vs `SQL`

- `SQL` is a very common Data Scientist interview topic so if you're not taking
  a database course, might be useful
  
## Reminder: CUNY Citizenship Now!

- Free and confidental immigration law services provided by CUNY
- Targeted at CUNY students and families, but open to all NYers

To access: 

- [Callback Request Form](https://www.cuny.edu/about/administration/offices/communications-marketing/citizenship-now/about/callback/)
- [Locations throughout NYC](https://www.cuny.edu/about/administration/offices/communications-marketing/citizenship-now/about/#locations)

## Reminder: ISS Workshops

See Brightspace Announcement

# On to the Show!

## Course Project Presentations

| Order | Members           |   | Team | Members             |
|-------|-------------------|---|------|---------------------|
| 1     | VH + SG + DS + DL |   | 6    | VG                  |
| 2     | HZ + JLL + CA     |   | 7    | EM + AK + GMdS + JL |
| 3     | MT + CW           |   | 8    | SJB + JC + SB + ZS  |
| 4     | SD + GO + CFGF    |   | 9    | GS                  |
| 5     | GB + RJ + FS + MH |   | 10   | CWo.                |

