---
session: "`r session <- 8; session`"
last_updated: "`r last_updated <- Sys.Date(); 
                  last_updated |> format('%Y-%m-%d')`"
class_date:
  "`r library(tidyverse);
    class_date <- read_csv('key_dates.csv', name_repair='universal') |>
    filter(Course.Element == 'Class Session',
           Item.Number == session) |>
    pull(Date); class_date |> format('%A %Y-%m-%d')`"
format: 
  revealjs:
    echo: true
---


{{< include _setup.qmd >}}

## {{< var course.short >}} Week {{< meta session >}} 

Today: `r TODAY_TOPIC`

. . . 

::: {.smaller}

- Team Formation ✅
- Project Proposals ✅
- Check-In Presentations  ⬅️
  - Mid-Semester Peer Evaluations
- Final Presentations
- Final Reports
  - Individual Technical Report
  - Group Summary Report
  - Final Peer Evaluations

:::

## After Proposals

Introduction to `SQL` and Databases - 100% Optional

- If you've taken DB courses already, won't add much. 
- SQL is a common *lingua franca* of data science
- Useful in interviews

## Schedule Notes

Upcoming schedule is a bit confusing:

::: {.incremental}

- Today: Mid-Semester Check-In Presentations
- Two Weeks of Spring Break (April 2 and 9)
  - MP#02 due over Spring Break
  - MP#02 Peer Feedback due over Spring Break
  - MP#03 Assigned
- April 16: Data Import Part 1 
- **Tuesday** April 21: Make-Up Class (Advanced Plotting)
- April 23: Data Import Part 2

:::

## Mini-Project #02

[MP#02](../mini/mini02.html) - `r get_mp_title(2)`


**Due `r get_mp_deadline(2)`**

::: {.small}

-   GitHub post (used for peer feedback) AND Brightspace
-   Start early to avoid Git issues

:::

. . . 

Use Piazza for questions over Spring Break (no Office Hours)

## Mini-Project #02

Topic: 

- ATUS Analysis of "People Like You"

. . . 

Skills: 

- Joining tables
- Elementary visualization
- Survey Weights

## Mini-Project #03

[MP#03](../mini/mini03.html) - `r get_mp_title(3)`

**Due `r get_mp_deadline(3)`**

. . . 

Analysis of *US internal migration*

- Who moves between states? 
- Predict populations as of 2030 
- Implications for Congress

## Mini-Project #03

Topics Covered: 

::: {.incremental}

- Using API Wrapper Packages
- Using APIs Directly
- Downloading and Parsing Files
- Cleaning and standardizing inputs
- Complex `group_by` structures

:::

. . . 

**Warning**: This is the *longest* MP. Plan accordingly!

# Check-In Presentations

## Check-In Presentations

Today: Mid-Semester Check-In Presentations ([Official Description](../project.qmd#mid-semester-check-in-presentations))

I'm looking for: 

::: {.incremental .smaller}

- Finalized _overarching question_
- Locking in on _specific questions_
- Evaluation of _relevant_ data sources
  - Anticipated data or statistical challenges
- Engagement with _existing literature_
- 6 minutes

:::

. . . 

Mainly, I want to see that you will be able to _succeed_ 


## Peer Evaluations

After today, you will be able to upload (anonymous) evaluations of your
project teammates: 

```{r}
#| eval: false
source("https://michael-weylandt.com/STA9750/load_helpers.R")
project_peer_evals(cycle="mid")
```

Upload result `bspf` file to Brightspace by `r get_project_due_date("mid-semester-peer")`. 

I will collate and summarize feedback and assign scores. 

## On to the Show!

Presentation Order:

```{r}
#| echo: false
BREAKOUT_TABLE
```

# Wrap-Up

## Next Steps

Now you should have your questions and data finalized 

Time to begin the actual analysis

. . . 

Starting next class (mid-April): getting data into `R`

## Recommended Workflow

If using a shared folder or Git repository: 

- `data`
- `team_member01`
- `team_member02`
- `team_member03`

If you only read (never write) to other people's stuff, no Git issues. 

. . . 

Shared `git` is not required - Dropbox, Google Drive also work

## Final Deliverables

Final Deliverables: 

::: {.incremental}

- Final Presentations: `r get_project_due_date("final-presentation")`
- Individual Technical Report: `r get_project_due_date("individual-report")`
- Group Summary Report: `r get_project_due_date("group-report")`

:::

. . . 

I've moved presentaitons forward a week this semester to give you two
weeks to write final reports. Feedback on this decision welcome!

{{< include ../advice/interview_tips.qmd >}}

## Musical Treat

</br>

{{< video https://www.youtube.com/watch?v=vb-_qqmOcCI width="80%" height="400px">}}


# Enrichment Material: SQL

## SQL

*S*tructured *Q*uery *L*anguage (ISO 9075)

::: {.incremental}

- Standard language for interacting with _Relational Database 
  Management Systems_ (RDMS)
- Divided into three main sub-languages: 
  - Data Manipulation Language (DML)
  - Data Declaration Language (DDL)
  - Data Control Language (DCL)
- *Declarative*, not *Imperative* Programming

:::

## SQL Languages

Data Definition Language (DDL)

::: {.small}

- "Design" language for DB
- What tables are present? What types are the columns? What relationships (keys)
  are enforced? 
  
:::

. . . 

Data Manipulation Language (DML)

::: {.small}

- "User" interaction with DB: queries, adding, changing, deleting data
- `SELECT`, `GROUP BY`, `WHERE`, `HAVING`, `JOIN`, *etc.*
- Roughly coextensive with `dplyr`

:::



. . . 

Data Control Language (DCL)

::: {.small}

- "Security" language for DB
- What permissions do different users have? 

:::

## RDMS

Relational Database Management System (RDMS)

::: {.incremental}

- Data management server for complex environments
- _Relational_ tabular data structures
- Manipulated using SQL
  - Common SQL Standard + _Many_ RDMS-specific extensions
- ACID Guarantees

:::

## DDL 

RDMS has a fixed number of _tables_ (roughly `data.frame`s). DDL specifies

- Column names and type
- Can values be left blank (nullability)?
- Identifiers (Primary Keys)
- Constraints (non-negative, `a < b`, *etc.*)
- Relationships to other tables (Foreign Keys)

## DDL 

Basic command is the `CREATE TABLE` statement.

```{sql}
#| eval: false
CREATE TABLE table_name (
  column_name1 TYPE1 PRIMARY KEY -- <- Guaranteed Unique and not blank
  column_name2 TYPE2 NOT NULL -- <- Can't be left blank
  column_name3 TYPE3 -- <- Can be blank
  column_name4 TYPE4 DEFAULT 0 -- <- Has default value
);
```

`TYPE` can be `INTEGER`, `DECIMAL`, `CHAR`, `VARCHAR`, `DATE`, 
`TIMESTAMP` and more

## DDL 

Our beloved `penguins`: 

```{sql}
#| eval: false
CREATE TABLE penguins (
  species VARCHAR(9) NOT NULL, -- Name of up to 9 characters
  island  VARCHAR(9) NOT NULL, -- Also can't be left blank
  bill_len NUMERIC,
  bill_dep NUMERIC,
  flipper_len INTEGER,
  body_mass INTEGER, 
  sex VARCHAR(6), -- Name of 6 (female) but also allow missing
  year INTEGER NOT NULL
);
```

Compare: 

```{r}
glimpse(penguins)
```

## DDL 

We can add a _constraint_ so the year is 2007-2009: 

```{sql}
#| eval: false
CREATE TABLE penguins (
  species VARCHAR(9) NOT NULL, -- Name of up to 9 characters
  island  VARCHAR(9) NOT NULL, -- Also can't be left blank
  bill_len NUMERIC,
  bill_dep NUMERIC,
  flipper_len INTEGER,
  body_mass INTEGER, 
  sex VARCHAR(6), -- Name of 6 (female) but also allow missing
  year INTEGER NOT NULL CHECK (year > 2006 AND year < 2010)
);
```

## DDL 

Constraints for categorical variables (`R` `factor`s) can be tricky. 
Create a _normalized_ set of tables: 

::: {.smaller}

| Sex ID | Name   | 
|--------|--------|
| 1      | Female |
| 2      | Male   |

| Penguin ID | Sex ID | Weight |
|------------|--------|--------|
|       1    |     1  |   4000 |
|       2    |     1  |   3500 |
|       3    |     2  |   5000 |

:::

. . . 

We want to make sure each `Sex ID` in table 2 matches a row of the first table

## DDL 

Constraints for categorical variables (`R` `factor`s) can be tricky. 
Create a _normalized_ set of tables: 

```{sql}
#| eval: false
CREATE TABLE sex (
  sex_id INTEGER PRIMARY KEY, 
  sex_name VARCHAR(6) NOT NULL
)

CREATE TABLE penguins (
  species VARCHAR(9) NOT NULL, -- Name of up to 9 characters
  island  VARCHAR(9) NOT NULL, -- Also can't be left blank
  bill_len NUMERIC,
  bill_dep NUMERIC,
  flipper_len INTEGER,
  body_mass INTEGER, 
  sex_id INTEGER, 
  year INTEGER NOT NULL CHECK (year > 2006 AND year < 2010), 
  FOREIGN KEY (sex_id) REFERENCES sex(sex_id)
);
```

## DDL Normalization

Normalization: 

1) Increases integrity via constraints
2) Reduces redundancy (need to change value in 1 place only)
3) Slight performance benefits

. . . 

More `JOIN` work in `SQL` than in `dplyr`

. . . 

Proper table design is a science and an art

## DML 

**Data Manipulation Language** is for DB users. 

- Close parallels to `dplyr`
- Compare: 

```{r}
#| eval: false
penguins
```

*vs.*

```{sql}
#| eval: false
SELECT * FROM penguins; 
```

`SELECT *` means _all columns_ which matches `R`'s default

## DML 

**Data Manipulation Language** is for DB users. 

- Close parallels to `dplyr`
- Compare: 

```{r}
#| eval: false
penguins |> select(species, island, bill_dep)
```

*vs.*

```{sql}
#| eval: false
SELECT species, island, bill_dep FROM penguins; 
```

## DML 

**Data Manipulation Language** is for DB users. 

- Close parallels to `dplyr`
- Compare: 

```{r}
#| eval: false
penguins |> select(species, island, bill_dep) |> filter(bill_dep > 20)
```

*vs.*

```{sql}
#| eval: false
SELECT species, island, bill_dep FROM penguins WHERE bill_dep > 20; 
```

## DML 

Note that

```{r}
#| eval: false
penguins |> filter(bill_dep > 20) |> select(species, island, bill_dep)
```

and 

```{r}
#| eval: false
penguins |> select(species, island, bill_dep) |> filter(bill_dep > 20)
```

are both 

```{sql}
#| eval: false
SELECT species, island, bill_dep FROM penguins WHERE bill_dep > 20; 
```

. . . 

Why? 

. . . 

They *commute* -- same result in either order

## DML 

- `R` is an *imperative* paradigm - you say _what to do_
- `SQL` is a *declarative* paradigm - you say _what you want_ 

. . . 

Declarative paradigms let the server figure out the _best_ way to compute, but
are more restrictive

. . . 

For a well-defined paradigm like SQL, _much_ work has gone into query optimization

## DML 

- Compare: 

```{r}
#| eval: false
penguins |> group_by(species) |> summarize(mean_mass = mean(body_mass, na.rm=TRUE))
```

*vs.*

```{sql}
#| eval: false
SELECT species, AVERAGE(body_mass) FROM penguins GROUP BY species;
```

`summarize()` is implicit in `SELECT` clause when we have a `GROUP BY`

## DML 

But:

```{r}
#| eval: false
penguins |> 
  group_by(species) |> 
  summarize(mean_mass = mean(body_mass, na.rm=TRUE)) |>
  filter(mean_mass > 4000)
```

is 

```{sql}
#| eval: false
SELECT species, AVERAGE(body_mass) AS avg_mass FROM penguins 
  GROUP BY species HAVING avg_mass > 4000;
```

. . . 

Why not `WHERE` again? 

## DML 

In `SQL`:

  - `WHERE` happens before a `GROUP BY`; and 
  - `HAVING` happens after a `GROUP BY`
  
. . . 
  
In `R`: 

  - Ordering of `summarize()` and `filter` controls execution order
  
. . . 

Because `SQL` is _declarative_, we need different words to encode filtering at different stages (pre/post-grouping)
  
## ACID Guarantees

Most RDMSs provide a set of four `ACID` guarantees: 

::: {.incremental}

- **Atomicity**: Changes occur all-or-nothing
  - *E.g.*: Add charge to customer's account + schedule item for shipping
- **Consistency**: DDL Constraints are always valid
  - *E.g.*: Can't create a user without having an email on file
- **Isolation**: Two different users don't "step on each other's toes"
  - *E.g.*: If I'm updating prices while you are checking out, you get your original price
- **Durability**: When a change is complete, it is guaranteed to be saved to disk

:::

## Getting Started with RDMS

Easiest RDMS to use is `SQLite`

- Most ubiquitous DB in the world (by a long-shot)
- Already installed on your computer (many times over!)
- Access directly through `sqlite3` or `sqlite3.exe` 
- Access from `R` with `RSQLite` package

. . . 

[Excellent documentation](https://sqlite.org/lang.html)

## dbplyr

Can use `dbplyr` (extra `b`!) to put `dplyr`-UX on DB

*E.g.*

```{r}
library(dbplyr)
my_db <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
copy_to(my_db, penguins) # Copy penguins data into a temporary SQLite DB
```


## dbplyr

```{r}
penguins_db <- tbl(my_db, "penguins") # Connect to my_db and load penguins table
penguins_db
```


## dbplyr

```{r}
penguins_db |> 
  group_by(species) |> 
  summarize(mean_mass = mean(body_mass, na.rm=TRUE)) |> 
  show_query()
```

Extra step - `show_query` to see SQL or `collect()` to run in DB:

```{r}
penguins_db |> 
  group_by(species) |> 
  summarize(mean_mass = mean(body_mass, na.rm=TRUE)) |> 
  collect()
```

## dbplyr

::: {.smaller}

```{r}
penguins_db |> 
  filter(body_mass > 4000) |>
  group_by(species) |> 
  summarize(mean_mass = mean(body_mass, na.rm=TRUE)) |> 
  show_query()
```
Compared to: 

```{r}
penguins_db |> 
  group_by(species) |> 
  summarize(mean_mass = mean(body_mass, na.rm=TRUE)) |> 
  filter(mean_mass > 4000) |>
  show_query()
```

Not the best SQL, but RDMS can make quick work of it

:::

## dbplyr {.scrollable}

A more substantial example: 

::: {.smaller}
```{r}
library(dbplyr)
baseball_db <- lahman_sqlite()
```
:::

## dbplyr {.scrollable}

A more substantial example: 

::: {.smaller}

```{r}
#| eval: false
library(dbplyr)
baseball_db <- lahman_sqlite()
```

```{r}
#| echo: false
#| eval: true
cat(system2("sqlite3", c(baseball_db@dbname, ".schema"), stdout=TRUE), sep="\n")
```
:::



## dbplyr {.scrollable}

List the 5 modern-era managers with the highest winning percentages. (Define
modern-era as managed over 1000 games post 1920.)

::: {.smaller}

```{r}
managers <- tbl(baseball_db, "Managers")
people   <- tbl(baseball_db, "People")

managers |> 
    filter(yearID >= 1920) |> 
    group_by(playerID) |> 
    summarize(W = sum(W, na.rm=TRUE),
              L = sum(L, na.rm=TRUE)) |> 
    mutate(G = W + L, 
           record = W / (G + 0.0)) |>  # + 0.0 forces to work with decimals, not ints
    filter(G > 1000) |> 
    inner_join(people, 
               join_by(playerID == playerID)) |> 
    slice_max(record, n=5) |> 
    mutate(name = paste(nameGiven, nameLast)) |> 
    select(name, W, L, G, record) |>
    show_query()
```

:::

## dbplyr {.scrollable}

Add `|> explain()` to see the DB internal _query plan_:

::: {.smaller}

```{r}
managers <- tbl(baseball_db, "Managers")
people   <- tbl(baseball_db, "People")

managers |> 
    filter(yearID >= 1920) |> 
    group_by(playerID) |> 
    summarize(W = sum(W, na.rm=TRUE),
              L = sum(L, na.rm=TRUE)) |> 
    mutate(G = W + L, 
           record = W / (G + 0.0)) |>  # + 0.0 forces to work with decimals, not ints
    filter(G > 1000) |> 
    inner_join(people, 
               join_by(playerID == playerID)) |> 
    slice_max(record, n=5) |> 
    mutate(name = paste(nameGiven, nameLast)) |> 
    select(name, W, L, G, record) |>
    explain() # 
```

:::

Beyond this class to interpret, but in general use of `INDEX` yields 
good performance. Take DB classes for _much more_


## dbplyr {.scrollable}

List the 5 modern-era managers with the highest winning percentages. (Define
modern-era as managed over 1000 games post 1920.)

::: {.smaller}

```{r}
managers <- tbl(baseball_db, "Managers")
people   <- tbl(baseball_db, "People")

managers |> 
    filter(yearID >= 1920) |> 
    group_by(playerID) |> 
    summarize(W = sum(W, na.rm=TRUE),
              L = sum(L, na.rm=TRUE)) |> 
    mutate(G = W + L, 
           record = W / (G + 0.0)) |>  # + 0.0 forces to work with decimals, not ints
    filter(G > 1000) |> 
    inner_join(people, 
               join_by(playerID == playerID)) |> 
    slice_max(record, n=5) |> 
    mutate(name = paste(nameGiven, nameLast)) |> 
    select(name, W, L, G, record) |>
    collect()
```

:::
