---
session: "`r session <- 1; session`"
class_date:
  "`r library(tidyverse); 
    tuesday_date <- read_csv('key_dates_tuesday.csv', name_repair='universal') |>
    filter(Course.Element == 'Class Session', 
           Item.Number == session) |>
    pull(Date) |> format('%a %Y-%m-%d'); 
    thursday_date <- read_csv('key_dates_thursday.csv', name_repair='universal') |>
    filter(Course.Element == 'Class Session', 
           Item.Number == session) |>
    pull(Date) |> format('%a %Y-%m-%d'); 
    paste(c(tuesday_date, thursday_date), collapse=' <br>')`"
---

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
DATES   <- read_csv("key_dates.csv")
TODAY   <- DATES |> filter(`Course Element` == "Class Session", 
                           `Item Number` == session)
TODAY_TUESDAY  <- TODAY |> filter(Section == "Tuesday")  |> pull(Date)
TODAY_THURSDAY <- TODAY |> filter(Section == "Thursday") |> pull(Date)

PROJ_RELEASE <- DATES |> 
    filter(Details == "Course Project Description Released") |>
    pull(Date)

ROSTERS <- DATES |> 
    filter(Details == "Team Roster Submission") |>
    pull(Date)
```
    

## {{< var course.short >}} Week {{< meta session >}} 

Today: 

- Tuesday Section: `r TODAY_TUESDAY`
- Thursday Section: `r TODAY_THURSDAY`

. . . 

Weekly Course Updates: 

- Brief updates and reminders about course logistics
- Syllabus and Brightspace are binding
  - If something is left out of here, it still happens!
  
. . . 

Today: Introduction to {{< var course.short >}}



# {{< var course.short >}}: {{< var course.title >}} 

## Course Overview

{{< var course.short >}} is: 

::: incremental

- [Getting s#!t done]{color="#FF0000"}
- Establishing a professional data analyst portfolio
- Increasing your effectiveness in other courses
- Building your networks with faculty and classmates

:::

## Learning Objectives

Formal learning objectives can be found [here](https://michael-weylandt.com/STA9750/objectives.html)

. . . 

Key aims: 

::: incremental

- Deal with **real** data, no matter how messy and unhelpful
- Engage with important technologies and learn to quickly adopt new tools
- Communicate to both technical and non-technical audiences
- Perform *substantial* and *thoughtful* analyses

::: 

## Course Website

**All course information** can be found [on the course website](https://michael-weylandt.com/STA9750/)

{{< html https://michael-weylandt.com/STA9750/ full-screen-link="false" height="300px">}}

No authentication required - mobile friendly

Source code [on my GitHub](https://github.com/michaelweylandt/STA9750)

## Course Website

Look online for: 

- [Syllabus](https://michael-weylandt.com/STA9750/syllabus.html)
- [Weekly Pre-Assignments](https://michael-weylandt.com/STA9750/preassignments.html)
- [Mini-Projects](https://michael-weylandt.com/STA9750/miniprojects.html)
- [In-Class Activities](https://michael-weylandt.com/STA9750/labs.html)
- [Course Project](https://michael-weylandt.com/STA9750/project.html)
- [Additional Policies and Resources](https://michael-weylandt.com/STA9750/resources.html)

## Instructor

Me: Michael Weylandt

. . . 

Assistant (i.e., new) Professor in the [Department of Info. Systems and Statistics](https://zicklin.baruch.cuny.edu/faculty-research/academic-departments/information-technology-statistics/)

. . . 

Previously: 

::: incremental

- Researcher at Sandia National Labs (US Government Lab)
- Postdoc with US Intelligence Community 
- Ph.D. in Statistics at Rice University (Houston, TX)
- "Quant" at Morgan Stanley and a (defunct) hedge fund

::: 

. . . 

See [my website](https://michael-weylandt.com) for more on me and my research

## Course Schedule

::: {.fragment}
- Lecture / Lab sessions
  - Tuesdays / Thursdays on Zoom (here!) at 6:05pm
  - 14 weeks (with two holidays)
:::

::: {.fragment}
- Office Hours
  - Before class (on days we have class) at 5:00pm
  - *Different Zoom Link*
:::

::: {.fragment}
- Asynchronous
  - Piazza course discussion platform
:::

## Course Schedule

Major Topics: 

::: incremental
- Web Communication and Literate Programming Technologies
- Introduction to `R`
- Tidy Data Manipulation (SQL-type operations)
- Plotting and Data Visualization
- Data Import, Web Scraping, and Text Cleaning
- Statistical Analysis
:::

## Grading

- 24% Pre-Assignments
- 36% Mini-Projects
- 40% Course Project

. . . 

Extra credit for contribution to course materials or 'above and beyond' on Piazza

. . . 

Final grades curved to match ZSB grading guidelines

## Course Policies

See syllabus for fine print: 

::: incremental

- Regrading: Must request within 48 hours; total regrade
- Late work: Not without prior permission or DoS letter *ex post*
- Grace period: two days on mini-project initial submission 
- External Resources / AI: Free to use for *coding* only
- Absences: No excuse required, but must attend on presentation days
- Accomodations: ADA (SDS) or Religious (direct with instructor)

:::

## Pre-Assignments

Pre-Assignments: 

::: incremental

- Weekly reading *before class* followed by a short quiz
- Introduce new material + helps me know where folks are confused
- Completion grading
- Starts next week 
- Every week except three presentation weeks
::: 

## Mini-Projects

Traditional homework assignments - "mini-projects"

::: incremental

- Guided real-data analysis using course technologies
- Increasingly ambitious over the semester
- Practice *real* data analysis
- Communication and coding
- Professional portfolio
- One ungraded 'set-up' + four graded

:::

More later in the course

## Course Project

In lieu of exams, semester-long group project

::: incremental

- Three presentations + two reports
  - Project Proposal
  - Mid-Semester Check-In
  - Final Presentation
  - Individual Final Report
  - Group Final Report
  
:::

More later in the course

## Workload Expectations

Per federal and state requirements, 9 hours weekly = 6 hours outside of class

::: incremental

- 10 hours of Pre-Assignments
- 10 hours of Post-Class Review
- 35 hours of Mini-Projects (homework)
- 35 hours of Course Project

:::

Note that course project = sum of all mini-projects - be ambitious!

## Care Resources

See [Care Resources for Students](https://michael-weylandt.com/STA9750/resources.html#care-resources-for-students1) for helpful resources

- Mental health support
- Physical health / medical care
- Food Security
- Financial Security
- Immigration Support

# Any questions?

# Getting Started with `R` and `RStudio`

## Lab 01

Open [Lab #01](https://michael-weylandt.com/STA9750/labs/lab01.html) and follow
the instructions to get started with `R` and `RStudio`

. . . 

</br>

Random assignment to Zoom breakout rooms

I will visit rooms to provide support

</br>

. . . 

Call everyone back at 8:30pm

## Mini-Project #00

You can now start thinking about [Mini-Project #00](https://michael-weylandt.com/STA9750/miniprojects/mini00.html)

::: incremental

- Create an account on GitHub + GitHub student developer pack
- Setting up your own website using `quarto` (like my course site)
  - Future Mini-Projects submitted here
- 'Road test' to make sure everything is working correctly
- Register on course discussion board (Piazza)
- Verification of Enrollment

:::

```{r}
#| echo: false
mp0_due_date <- DATES |> 
    filter(`Course Element` == "Mini-Projects", 
           `Item Number` == 0, 
           str_detect(Details, "Due"), 
           !str_detect(Details, "Feedback")) |> 
    pull(Date)
```

Due `r mp0_due_date`

# Wrap-Up

## Course Introduction

- All materials on course website
- Review of course structure and key policies
- More details about mini-projects and course-project to follow

## Getting Started with `R`

- Install `R`, `RStudio`, `git`, and `quarto`
- Run a basic bit of `R` code to confirm things work well

## Next Week

Writing documents using `Markdown` and `Quarto`

- First Pre-Assignment
  - Read online
  - Submit quiz on Brightspace

{{< include ../advice/office_hours.qmd >}}

## Musical Treat

</br></br>

::: {.center}

{{< video https://www.youtube.com/watch?v=RHu0ALxqUIo width="800" height="400">}}

:::
