---
session: "`r session <- 8; session`"
last_updated: "`r last_updated <- Sys.Date(); 
                  last_updated |> format('%Y-%m-%d')`"
class_date:
  "`r library(tidyverse);
    class_date <- read_csv('key_dates.csv', name_repair='universal') |>
    filter(Course.Element == 'Class Session',
           Item.Number == session) |>
    pull(Date); class_date |> format('%A %Y-%m-%d')`"
format: 
  revealjs:
    echo: true
---


{{< include _setup.qmd >}}

## {{< var course.short >}} Week {{< meta session >}} 

Today: `r TODAY_TOPIC`

. . . 

::: {.smaller}

- Team Formation ✅
- Project Proposals ✅
- Check-In Presentations  ⬅️
  - Mid-Semester Peer Evaluations
- Final Presentations
- Final Reports
  - Individual Technical Report
  - Group Summary Report
  - Final Peer Evaluations

:::

## After Proposals

Introduction to `SQL` and Databases - 100% Optional

- If you've taken DB courses already, won't add much. 
- SQL is a common *lingua franca* of data science
- Useful in interviews

## Schedule Notes

Upcoming schedule is a bit confusing:

::: {.incremental}

- Today: Mid-Semester Check-In Presentations
- Two Weeks of Spring Break (April 2 and 9)
  - MP#02 due over Spring Break
  - MP#02 Peer Feedback due over Spring Break
  - MP#03 Assigned
- April 16: Data Import Part 1 
- **Tuesday** April 21: Make-Up Class (Advanced Plotting)
- April 23: Data Import Part 2

:::

## Mini-Project #02

[MP#02](../mini/mini02.html) - `r get_mp_title(2)`


**Due `r get_mp_deadline(2)`**

::: {.small}

-   GitHub post (used for peer feedback) AND Brightspace
-   Start early to avoid Git issues

:::

. . . 

Use Piazza for questions over Spring Break (no Office Hours)

## Mini-Project #02

Topic: 

- ATUS Analysis of "People Like You"

. . . 

Skills: 

- Joining tables
- Elementary visualization
- Survey Weights

## Mini-Project #03

[MP#03](../mini/mini03.html) - `r get_mp_title(3)`

**Due `r get_mp_deadline(3)`**

. . . 

Analysis of *US internal migration*

- Who moves between states? 
- Predict populations as of 2030 
- Implications for Congress

## Mini-Project #03

Topics Covered: 

::: {.incremental}

- Using API Wrapper Packages
- Using APIs Directly
- Downloading and Parsing Files
- Cleaning and standardizing inputs
- Complex `group_by` structures

:::

. . . 

**Warning**: This is the *longest* MP. Plan accordingly!

# Check-In Presentations

## Check-In Presentations

Today: Mid-Semester Check-In Presentations ([Official Description](../project.qmd#mid-semester-check-in-presentations))

I'm looking for: 

::: {.incremental .smaller}

- Finalized _overarching question_
- Locking in on _specific questions_
- Evaluation of _relevant_ data sources
  - Anticipated data or statistical challenges
- Engagement with _existing literature_
- 6 minutes

:::

. . . 

Mainly, I want to see that you will be able to _succeed_ 


## Peer Evaluations

After today, you will be able to upload (anonymous) evaluations of your
project teammates: 

```{r}
#| eval: false
source("https://michael-weylandt.com/STA9750/load_helpers.R")
project_peer_evals(cycle="mid")
```

Upload result `bspf` file to Brightspace by `r get_project_due_date("mid-semester-peer")`. 

I will collate and summarize feedback and assign scores. 

## On to the Show!

Presentation Order:

```{r}
#| echo: false
BREAKOUT_TABLE
```

# Wrap-Up

## Next Steps

Now you should have your questions and data finalized 

Time to begin the actual analysis

. . . 

Starting next class (mid-April): getting data into `R`

## Recommended Workflow

If using a shared folder or Git repository: 

- `data`
- `team_member01`
- `team_member02`
- `team_member03`

If you only read (never write) to other people's stuff, no Git issues. 

. . . 

Shared `git` is not required - Dropbox, Google Drive also work

## Final Deliverables

Final Deliverables: 

::: {.incremental}

- Final Presentations: `r get_project_due_date("final-presentation")`
- Individual Technical Report: `r get_project_due_date("individual-report")`
- Group Summary Report: `r get_project_due_date("group-report")`

:::

. . . 

I've moved presentaitons forward a week this semester to give you two
weeks to write final reports. Feedback on this decision welcome!

{{< include _end.qmd >}}
