---
title: "{{< var course.short >}} - Week 3 Update"
author: "Michael Weylandt"
format: revealjs
---

## {{< var course.short >}} Mini-Project #00

- MP#00 submitted
  - A few of you didn't submit; I'll follow up directly for VoE

- MP#00 peer feedback assignments released (check GitHub)
  - Give some feedback to your peers
  - Get ideas for improving your own site

## {{< var course.short >}} Mini-Project #01

- MP#01 released
- **Start early**
  - Not too hard if everything is working (post-MP#00)
  - Tech support takes time

## Graduate Teaching Assistant (GTA)

- Charles Ramirez
- Twice Weekly Office Hours (Zoom)
  - Tuesdays 4-5pm
  - Fridays 12-1pm
- Will also help coordinate peer feedback (GitHub), Piazza responses, *etc.*
- Excellent resource for course project advice!

## Piazza Participation

- Average time to response <9 hours
- 209 posts

Thanks to those of you who are helping classmates!

## Course Project

- 1 team already registered with me!
- Piazza discussions helping to coordinate other teams

## Upcoming Week

Next Wednesday at 11:45pm: 

- Next Pre-Assignment
- MP#00 Peer Feedback due

# Pre-Assignment #03 FAQs

## FAQ: Vector Index Printout Rules

Default vector printing: 
```{r}
#| echo: true
1:10
```

Each line gets a new index:
```{r}
#| echo: true
sqrt(1:10)
```

More complex objects have alternate print styles:
```{r}
#| echo: true
matrix(1:9, nrow=3, ncol=3)
```

Print width is controlled by `getOption("width")`.

## FAQ: Recycling Rules

Alignment by default: 
```{r}
#| echo: true
x <- 1:3
y <- 4:6
x + y
```

Recycling by default: 
```{r}
#| echo: true
x <- 1
y <- 4:6
x + y
```

Recycle warning when vectors don't fit together cleanly: 
```{r}
#| echo: true
#| warning: true
x <- 1:2
y <- 4:6
x + y
```

## FAQ: Recycling Warning

```{r}
#| echo: true
#| warning: true
x <- 1:2
y <- 4:6
x + y
```

Not a problem _per se_, but often a sign that something has gone wrong.

- scalar + vector is usually safe
- 2 vectors of same size is usually safe
- vectors of different size is usually a programming mistake

## FAQ: Warnings vs Errors

- _Warnings_: heuristics pointing at typical problem
  - Code still executed without a problem
  - Try to fix these unless you're **certain** it's not a problem
- _Errors_: code failed to execute
  - You have to fix these to run your code

## FAQ: Changing built-in functions

Most built-in functions can't / shouldn't be changed. 

Some allow alternate behavior via additional arguments:

```{r}
#| echo: true
log(10) # Default is natural (base e) logarithm
log(10, base=10)
```

If you want different behavior, write your own function: 

```{r}
#| echo: true
cosd <- function(x){
    ## Cosine in degrees
    cos(x * pi / 180)
}
cosd(90)
```

Always try `?name` to see documentation.

## FAQ: Git Workflow

Three key commands: 

- `git add`: add some changes to a 'box'
- `git commit`: seal the 'box'
- `git push`: send the 'box' to GitHub

`Git` pane in `RStudio` shows _uncommited changes_, not files. 

If a file 'vanishes' after a commit, that's good!
