---
session: "`r session <- 6; session`"
last_updated: "`r last_updated <- Sys.Date(); 
                  last_updated |> format('%Y-%m-%d')`"
class_date:
  "`r library(tidyverse);
    class_date <- read_csv('key_dates.csv', name_repair='universal') |>
    filter(Course.Element == 'Class Session',
           Item.Number == session) |>
    pull(Date); class_date |> format('%A %Y-%m-%d')`"
format: 
  revealjs: 
    echo: true
---

{{< include _setup.qmd >}}

## {{< var course.short >}} Week {{< meta session >}} 

Today: `r TODAY_TOPIC`

# Today

## Today

- Mini-Projects
- Review
- Diving Deeper into Multi-Table Verbs
- PA#05 FAQs
- Wrap-Up
  - Life Tip of the Day
  
## MP#01 - Deadline Soon

[Mini-Project #01](../mini/mini01.qmd) due on `r get_mp_deadline(01)`

. . . 

::: {.small}

Topics Covered: 

- Reproducible Research Tooling
- Packages and Functions 
- Basic Data Manipulation
- Tables for Data Display

:::

. . . 

Make sure to complete _all_ instructor-provided 
EDA questions and writing prompts

**Be reasonable, justify, and document your analysis.** 

- Do your answers pass the sniff test?  
- At least one sentence of text per short question is a good baseline

## MP#01 - Peer Feedback

Assigned on GitHub - due on `r get_mp_pf_deadline(1)`

- $\approx 4$ feedbacks each 
- Take this seriously: 20\% of this assignment is "meta-review"
- Goal: _rigorous_ _constructive_ critique

Use [helper function](https://michael-weylandt.com/STA9750/tips.html) 
to perform peer feedback assigned to you. Ask on Piazza if still having
trouble.

## MP#01 - Peer Feedback

Submissions may not map perfectly to rubric - use your best judgement

. . . 

Be _generous_ but serious: 

::: {.small}

- Goal is improvement, so "everything is great, no comments" is unhelpful
  - Nothing is completely right nor completely wrong
- Remember, *meta-review* (instructor scores of your feedback) to follow

:::

. . .

Learn from this! What can _you_ adapt for MP#02? 

## MP#01 - Peer Feedback {.scrollable}

Example of poor feedback: 

::: {.smallest}

```
Comments
Website looks really good, nice and clean!

Written Communication
Excellent and straight to the point.

Project Skeleton
Solid skeleton, well-organized .

Formatting & Display
Nicely formatted and balanced across the website.

Code Quality
Code runs like Forrest Gump in a Slump.

Data Preparation
Very great. 

Extra Credit
Added graphes for even better understanding.
```

:::

- Superficial comments / no sign of actually reading work

Red Flag: Repeated verbatim on several posts

Reminder: Poor feedback $\neq$ poor work.

## MP#01 - Peer Feedback {.scrollable}

Example of medium feedback: 

::: {.smallest}

```
## Comments

Love the visuals for Press Releases.

### Written Communication

A short summary or a description of this project would be great to add.

### Project Skeleton

Code completes all instructor-provided tasks correctly.

### Formatting & Display

Tables have well-formatted column names; caption would be great.

### Code Quality

Code is clear and well written.

### Data Preparation

Automatic (10/10). Out of scope for this mini-project.

### Extra Credit

I found all the Press Releases very interesting. The visuals were a great touch.
```

:::

- Gave actionable suggestions
- Directionally correct, but a bit vague

## MP#01 - Peer Feedback {.scrollable}

Example of great feedback: 

::: {.smallest}

```
### Written Communication

Overall, your writing is clear and easy to follow. I noticed a few small typos, 
but nothing major — using the built-in spell check in RStudio should catch those quickly. Everything else looks solid, and I didn’t have any major concerns based
on my review.

### Project Skeleton

All tasks were completed satisfactorily. 

### Formatting & Display

Overall, your tables and figures are well-organized and clear. There is one 
table in the "Data" section with column titles that could be formatted a little
more cleanly for easier reading. I also noticed a few small typos in some of the
table captions, but they should be easy to fix.

### Code Quality

The code quality is generally good, but there are a few minor linter issues. The
comments could be more frequent and clearer in some places. For example, the
comment `#checking dupolicate so that github will not block it` should be 
corrected to "duplicate," and the explanation could be clarified to avoid
confusion–what duplicate are you referring to, and how does it block GitHub? 
It might confuse someone else reading the code.

Also, I would recommend moving all library imports to the top of the script, as that is typically considered standard practice.

### Data Preparation

The data preparation looks solid overall. I like how you handle missing files 
and JSON parsing failures–it demonstrates strong defensive programming.

### Extra Credit

Additional two points for using quarto's video support!
```

:::

- Positive tone
- Detailed comments
- Noted issues & gave suggestions on how to fix
- Noted unclear sections for improvement


## MP#01 - Peer Feedback

Lack of prior experience _is not a hinderance_ here: 

- If something is unclear to you, that's a problem!
- Nothing required super-complex code, so anything overly complex probably could
  have been a simpler way (except for some above-and-beyond stuff)
- You don't have to be definitive in comments - impressions and questions are
  just as helpful.


## MP FAQs

Q: Why doesn't my site look the same on GitHub as it does on my laptop?

. . . 

A: Missing `css` and `js` files. Need to _upload_ everything in `docs`. 

. . . 

New helper function: 

```{r}
#| eval: false
#| echo: true
source("https://michael-weylandt.com/STA9750/load_helpers.R")
mp_submission_ready()
```

Tries to make sure all 'secondary' files are present. 

## MP FAQs

Q: Why doesn't my submission have the right URL? 

. . . 

A: _Most_ computers have case-sensitive file names, Mac doesn't, and Windows is
iffy. 

. . . 

New helper function: 

```{r}
#| eval: false
#| echo: true
source("https://michael-weylandt.com/STA9750/load_helpers.R")
mp_start()
```

Generates a very basic `qmd` file with the correct file name and some relevant
context. 


## Mini-Project #02

[MP#02 released](../mini/mini02.html) - `r get_mp_title(2)`

. . .

**Due`r get_mp_deadline(2)`**

- GitHub post (used for automated checks) AND Brightspace
- `r get_mp_deadline(2, as_d=TRUE) - class_date` days:
  don't wait until the very end

. . . 

Topics covered: 

- Combining Data from Different Sources
- Data Visualization
- Using Survey Weights for Microdata Analysis

## Upcoming Mini-Projects

Tentative Topics

-   MP#03: State-to-State Moving - `r get_mp_deadline(3)`
-   MP#04: Olympics Results - `r get_mp_deadline(4)`


# Review from Last Week

## Single-Table Verbs

`dplyr` single-table verbs

::: {.incremental}

- `select`, `filter`: Selecting rows and columns
- `rename`, `mutate`: Changing rows and columns
- `summarize`, `group_by`: Combining multiple rows
- `arrange`, `slice_min/max`: Re-ordering

:::

. . . 

[Lab #05 Review Activity](../labs/lab05.qmd#review)

## Breakout Rooms

```{r}
#| echo: false
BREAKOUT_TABLE
```

# Multi-Table Verbs

## Multi-Table Analysis

*Multiple Tables*:

- More insights than from a single table
- Maintain 'tidy' structure throughout

Will create new (compound) rows: 

- Dangers: *drops* and *(over) duplication*

## Primary Keys

::: {.small}

_Keys_ are unique identifiers for individual records

- Primary (one column) or compound (multiple columns together) 

:::

. . . 

::: {.small}

The history of corporate IT is largely one of (failed) primary keys

- Finance: Tickers, Tickers + Exchange, Tickers + Share Class, CUSIP, ISIN,
  SEDOL, ... 
  
:::
  
. . . 

::: {.small}

Meaningful true keys are vanishingly rare - cherish them when you find them

Often 'unique enough' for an analysis

`dplyr::group_by() + dplyr::count()` is helpful here

:::

## Joins

Joins combine tables _by identity_ - not simple 'stacking' 

Specify a _join key_ - ideally this is an actual key, but doesn't have to be

In `dplyr`, we use the `join_by` function: 

```{r}
#| eval: false
#| echo: true
dplyr::join_by(table_1_name == table_2_name)
```

Here `table_1_name` and `table_2_name` are column names from two tables

. . . 

Join rows where these values are equal (advanced joins possible)

## Inner and Outer Joins

When tables are perfectly matched, not an issue: 

```{r}
#| echo: false
library(dplyr)
cunys <- tribble(~college, ~campus_borough, 
             "CCNY", "Manhattan",
             "Baruch", "Manhattan", 
             "CSI", "Staten Island",
             "York", "Queens")

routes <- tribble(~borough_name, ~bus_code,
                  "Manhattan", "M",
                  "Staten Island", "S", 
                  "Queens", "Q")
```

```{r}
cunys
```

```{r}
routes
```

## Inner and Outer Joins

When tables are perfectly matched, not an issue: 

```{r}
inner_join(cunys, routes, join_by(campus_borough == borough_name))
```

Default to `inner` but irrelevant

Note automatic repetition of `"M"` row

## Inner and Outer Joins

How to handle 'unaligned' values?

```{r}
#| echo: true
cunys <- tribble(~college, ~campus_borough, 
                 "CCNY", "Manhattan",
                 "Baruch", "Manhattan", 
                 "CSI", "Staten Island",
                 "York", "Queens", 
                 "Medgar Evers", "Brooklyn")

inner_join(cunys, routes, join_by(campus_borough == borough_name))
```

. . . 

[MEC vanished!]{color="red"}

## Inner and Outer Joins

```{r}
#| echo: true
left_join(cunys, routes, join_by(campus_borough == borough_name))
```

MEC stays, but no bus code - `NA` value

. . . 

::: {.smaller}

- `inner_join` - Keep only matches
- `left_join` - Keep all rows in left (first) table even w/o matches
- `right_join` - Keep all rows in right (second) table even w/o matches
- `full_join` - Keep all rows from both tables, even w/o matches

:::

`left_` and `right_` are types of 'outer' joins

## Pivoting

The `pivot_*` functions change the **shape** of data

::: {.incremental}

- Values are not created or destroyed, just moved around
- **wider** data sets are formed by forming multiple rows into columns
- **longer** data sets are splitting columns from the same row into new rows

:::

. . . 

These functions come from the `tidyr` package - not `dplyr`

```{r}
library(tidyr) # included in library(tidyverse)
```

## Pivoting

Untidy example from last week: 

```{r}
#| echo: false
library(tibble)
BARUCH_UNTIDY <- tribble(
    ~Semester, ~Course,      ~Number, ~Type,
       "Fall", "Accounting",     200, "Enrollment",
       "Fall", "Accounting",     250, "Cap",
       "Fall", "Law",        100, "Enrollment",
       "Fall", "Law",        125, "Cap",
       "Fall", "Statistics", 200, "Enrollment",
       "Fall", "Statistics", 200, "Cap",
       "Spring", "Accounting",     300, "Enrollment",
       "Spring", "Accounting",     350, "Cap",
       "Spring", "Law",        50, "Enrollment",
       "Spring", "Law",        100, "Cap",
       "Spring", "Statistics", 400, "Enrollment",
       "Spring", "Statistics", 400, "Cap"
)
BARUCH_UNTIDY
```

## Pivoting

This data was untidy because it split a single unit (course) across multiple rows

`pivot_wider` to get to the right format

. . . 

```{r}
#| echo: true
pivot_wider(BARUCH_UNTIDY, names_from=Type, values_from=Number)
```

## Pivots

`pivot_` changes the shape of a data set. Purposes: 

- Get ready for presentation
- Prep for a join
- Combine rows before looking at 'cross-row' structure

## Pivots

Which penguin species has the largest between-sex mass difference? 

```{r}
#| echo: true
library(tidyr)
avg_mass_tbl <- penguins |> drop_na() |> 
    group_by(sex, species) |> 
    summarize(avg_mass = mean(body_mass), .groups="drop")
    # .groups="drop" is equivalent to |> ungroup()
avg_mass_tbl
```

## Pivots

We want data that is _wider_ (or at least not longer) than our current data: 

| species   | male_avg | female_avg | 
|-----------|----------|-------------
| Adelie    |...|...|
| Chinstrap |...|...|
| Gentoo    |...|...|


## Pivots

For the penguins: 

```{r}
#| echo: true
pivot_wider(avg_mass_tbl, 
            id_cols = species, 
            names_from=sex, 
            values_from=avg_mass)
```

. . . 

::: {.small}

Put it all together:

```{r}
#| echo: true
pivot_wider(avg_mass_tbl, 
            id_cols = species, 
            names_from=sex, 
            values_from=avg_mass) |>
    mutate(sex_diff = male - female) |>
    slice_max(sex_diff)
```

:::

## Pivots

`pivot_wider` Arguments:

::: {.small}

- `id_cols`: kept as 'keys' for new table
- `names_from`: existing column 'spread' to create new
                columns names
- `values_from`: values in new table

:::

. . . 

`pivot_longer`: 

::: {.small}

- 'Inverse' operation
- Spread one row + multiple columns => one col + multiple rows

:::

. . . 

::: {.small}

`pivot_wider` and `pivot_longer` have many additional arguments for dealing with
repeats / missing values. The help page (+ experimenting) is your friend

:::

## Legos of Data Analysis

These functions are like Legos: 

- Simple individually
- Combine for complex structures

## Legos of Data Analysis

**Q: How many distinct flights left NYC in 2013?**

```{r}
library(dplyr)
library(nycflights13)
flights |> 
    n_distinct()
```

. . . 

Not quite what we wanted...

## Legos of Data Analysis

**Q: How many distinct flights left NYC in 2013?**

. . . 

How many unique combinations of `carrier` + `flight` (*e.g.*, United 101)?

```{r}
flights |>
    select(carrier, flight) |>
    n_distinct()
```

## Legos of Data Analysis

**Q: How many distinct flights left NYC in 2013?**

::: {.small}

💡 Did airlines re-use flight numbers for different destinations? 

```{r}
flights |>
    distinct(carrier, flight, dest) |>
    # Find reuse of number across different destinations
    # Shorthand for group_by + summarize(n = n())
    count(carrier, flight) 
```

:::

. . . 

Seems so!

## Legos of Data Analysis {.scrollable}

**Q: How many distinct flights left NYC in 2013?**

Find examples of re-use: 

```{r}
flights |>
    distinct(carrier, flight, dest) |>
    count(carrier, flight) |> 
    slice_max(n)
```

## Legos of Data Analysis {.scrollable}

**Q: How many distinct flights left NYC in 2013?**

Finding most re-used flight number: 

```{r}
flights |>
    distinct(carrier, flight, dest) |>
    count(carrier, flight) |> 
    slice_max(n) |> 
    left_join(flights, join_by(carrier == carrier, flight == flight)) |>
    pull(dest) |> # pull out column as vector
    table() # frequency table
```

## Legos of Data Analysis {.scrollable}

**Q: How many distinct flights left NYC in 2013?**

Finding most re-used flight number: 

```{r}
flights |>
    distinct(carrier, flight, dest) |>
    count(carrier, flight) |> 
    slice_max(n)
```

## Legos of Data Analysis {.scrollable}

**Q: How many distinct flights left NYC in 2013?**

Seeing where our most reused number went:

```{r}
flights |>
    distinct(carrier, flight, dest) |>
    count(carrier, flight) |> 
    slice_max(n) |>
    inner_join(flights, join_by(carrier == carrier, flight == flight)) |>
    count(flight, carrier, dest)
```

## Legos of Data Analysis {.scrollable}

**Q: How many distinct flights left NYC in 2013?**

Additional join to get airport information + formatting:

```{r}
flights |>
    distinct(carrier, flight, dest) |>
    count(carrier, flight) |> 
    slice_max(n) |>
    inner_join(flights, join_by(carrier == carrier, flight == flight)) |>
    count(flight, carrier, dest) |> 
    inner_join(airports, join_by(dest == faa)) |>
    select(name, n, carrier, flight) |>
    arrange(desc(n)) |>
    rename(`Destination Airport` = name, 
           `Number of Times Flown` = n, 
           `Carrier Code` = carrier, 
           `Flight Number`= flight)
```


## Legos of Data Analysis {.scrollable}

**Q: How many distinct flights left NYC in 2013?**

::: {.small}

Extra join to match to `airlines` as well: 

```{r}
head(airlines, 3)
```

:::

. . . 

Also has a column named `name` - need to disambiguate!

## Legos of Data Analysis {.scrollable}

**Q: How many distinct flights left NYC in 2013?**

Additional join to get airport information + formatting:

```{r}
flights |>
    distinct(carrier, flight, dest) |>
    count(carrier, flight) |> 
    slice_max(n) |>
    inner_join(flights, join_by(carrier == carrier, flight == flight)) |>
    count(flight, carrier, dest) |>  
    inner_join(airports, join_by(dest == faa)) |>
    select(name, n, carrier, flight) |>
    rename(dest_name = name) |> 
    inner_join(airlines, join_by(carrier == carrier)) |> 
    arrange(desc(n)) |>
    select(-carrier) |> 
    rename(`Destination Airport` = dest_name, 
           `Number of Times Flown` = n, 
           `Carrier` = name, 
           `Flight Number`= flight)
```

## Legos of Data Analysis

Question: What does this do I can't do in `Excel`? 

. . . 

*Technically*, nothing. All programming languages of sufficient complexity are
equally powerful (Turing equivalence). 

</br>

. . . 

In actuality, quite a lot: 

- `filter` allows more complex filtering than clicking on values
- `group_by` + `summarize` extend array formulas
- `*_join` provides more complex matching than `VLOOKUP`
- `pivot_*` provide general formulation of pivot tables

+ everything else you can do in `R`. 

Ability to *script* minimizes "hard-coding" of names and values. 

## But truthfully

```{r}
fortunes::fortune(59)
```

. . . 

```{r}
fortunes::fortune(222)
```


# Diving Deeper into Multi-Table Verbs

## Diving deeper Into Joins

Data Set: [`nycflights13`](https://nycflights13.tidyverse.org/)

Exercises: [Lab #05](../labs/lab05.html)

## Additional dplyr Tricks

::: {.incremental}

- Ranking functions
  - `row_number`, `min_rank`, `dense_rank`: differ in ties
  - Use with `desc()` to flip ordering
  - `cum_dist`, `percent_rank`: compute quantiles
- Cumulative Statistics
  - `cummean`, `cummax`, `cummin`, ...
  
:::

# Multi-Table FAQs

## Subqueries

> [W]ill we be learning how to perform joins within a subquery?

You don't need subqueries in `R` since it's an imperative language. 
Just create a new variable to represent the result of the subquery
and use that in the next command. 

```
SELECT first_name, last_name
FROM collectors
WHERE id IN (
    SELECT collector_id
    FROM sales
);
```

```
collector_ids <- sales |> pull(collector_id)
collectors |> filter(id %in% collector_ids) |> select(first_name, last_name)
```

## Data Integrity

> [H]ow can we ensure that the information [resulting from a join]
> is accurate and not repeated? 

1. If you have a true unique ID, you're usually safe
2. Pay attention to all warnings
3. Manually examine the result of any joins

## Performance

> Will joining large data sets [...] affect performance?

Somewhat - larger data sets are always slower. 

Bigger danger is "bad joins" creating huge data automatically. 

Note that `R` is less "smart" than `SQL`, so won't optimize
execution order for you automatically.

## dplyr joins vs SQL joins

> What is the difference between `dplyr` and `SQL` joins?

Not too much - biggest difference is no `INDEX` or `FOREIGN KEY`
in `R` so less guarantees of data integrity.

## When to use anti_join()?

Rare: looking for _unmatched_ rows. 

- Useful to find data integrity issues or 'implicit' missingness. 
- I use an `anti_join` to find students who haven't submitted an assignment.

## many-to-many Warning

Tricky to address, but fortunately pretty rare. 

- `SQL` explicitly forbids many-to-many
- Usually a sign that a "key" isn't really unique
  - Check for duplicates in `x` and `y` tables
  - Can occur with "fancy" joins (rolling, inequality)
- Add additional join variables to break "duplication"

## How to Check Efficiency? 

No automatic way. Some rules of thumb: 

- Don't create large tables just to filter down
  - filter before join when possible
- `full_outer` join is a bit dangerous
- `cross_join` is _rarely_ the right answer

## tidyr vs dplyr

> Is `tidyr` more efficient than `dplyr`? 

Nope - different packages from the same developers. 

Designed to work together elegantly. 

## Rare Joins

> What are `cross_join`, filter joins, and `nest_join`?

- `cross_join`: dangerous. 
  - Creates "all pairs" of rows. Useful for 'design' problems 
- filter joins (`anti_`, `semi_`): 
  - Hunting down quietly missing data.
  - Filtering to sub-samples
- `nest_join`: beyond this course. 
  - `left_join` with extra structure to output.

# Wrap-Up

## Review

Multi-Table `dplyr`: 

- `inner_join` and `left_join`
- `join_by` specifications
- `pivot_longer` and `pivot_wider` to get data into optimal formats (`tidyr`)

. . . 

Additional `dplyr`: 

- Ranking, cumulative, and shift functions


## Orientation

- Communicating Results (`quarto`)  ✅
- `R` Basics  ✅
- Data Manipulation in `R`  ✅
- Data Visualization in `R`  ⬅️
- Getting Data into `R`
- Statistical Modeling in `R`

{{< include ../advice/barnette.qmd >}}

{{< include _end.qmd >}}
