---
session: "`r session <- 6; session`"
last_updated: "`r last_updated <- Sys.Date(); 
                  last_updated |> format('%Y-%m-%d')`"
class_date:
  "`r library(tidyverse);
    class_date <- read_csv('key_dates.csv', name_repair='universal') |>
    filter(Course.Element == 'Class Session',
           Item.Number == session) |>
    pull(Date); class_date |> format('%A %Y-%m-%d')`"
format: 
  revealjs: 
    echo: true
---

{{< include _setup.qmd >}}

## {{< var course.short >}} Week {{< meta session >}} 

Today: `r TODAY_TOPIC`

# Today

## Today

- Mini-Projects
- Review
- Multi-Table Verbs
- Pivots and Data Reshaping
- Wrap-Up
  
## MP#01 - Deadline Soon

[Mini-Project #01](../mini/mini01.qmd) due on `r get_mp_deadline(01)`

. . . 

::: {.small}

Topics Covered: 

- Reproducible Research Tooling
- Packages and Functions 
- Basic Data Manipulation
- Tables for Data Display

:::

. . . 

Make sure to complete _all_ instructor-provided 
EDA questions and writing prompts

**Be reasonable, justify, and document your analysis.** 

- Do your answers pass the sniff test?  
- At least one sentence of text per short question is a good baseline

## MP#01 - What to Write

In MP#01, you are not just submitting code, but a full documented analysis.

- Code intermixed with text throughout
- But what to write? 

. . . 

Always explain *why* you are taking particular steps; sometimes (high-level) 
_what_; rarely _how_

- Clean code should (generally) be _self-documenting_

---

Unnecessary: 

````{md}
Next, we rename columns in the data set: 

```{{r}}
data_set <- data_set |> rename(...)
```
````

. . . 

Your target audience is a competent working data scientist, so explaining that
`rename()` _renames_ things is unhelpful. 

. . . 

::: {.smaller}

To be fair, there are some times in life where gnarly non-self-documenting code
is required, but you shouldn't be encountering those in this class

:::

---

Better: 


````{md}
Next, we update column names to include SI units and to more closely reflect the 
terminology of Gorman *et al* (2014):

```{{r}}
better_penguins <- penguins |> 
  rename(culmen_depth_mm = bill_dep, 
         ...)
```
````

. . . 

This explains _what_ you are doing at a level _beyond the code_ (not just 
renaming, but _aligning_ with other data sources). 

## EDA Functions

In MP#01, I ask you to run `glimpse()` and `skim()` on your data. These functions
give summaries of the data set (column types, means, ranges, *etc.*)

. . . 

These _EDA_ functions help you get oriented with new data; normally, you wouldn't
document them, but I'm asking you to run and discuss them (in MP#01 only)

--- 

Example: 

````{md}
After an initial survey of our data, 

```{{r}}
glimpse(penguins)
```

we note that this data has a mix of categorial and numeric columns, as well as 
one column (`year`) that should transformed from numeric to factor (categorical)
for modeling (ANOVA) purposes.
````

## Documentation Sandwich

Documentation sandwich: 

- _Before_: _Why_ are you doing something? 

- _After_: _What_ does this tell you? _What_ does it imply about next steps? 

. . . 

The latter is particularly important with tables and figures. You have a _story_
to tell your audience. Don't just dump a bunch of numbers and expect them to 
figure it out

. . . 

Not always necessary to have both. Use your best judgment 

---

To assess whether automated penguin sexing is possible, we need to compare 
typical male and female body masses across our three species of interest: 

```{r}
#| echo: false
#| message: false
#| warning: false
library(tidyverse)
library(gt)
penguins |> 
    select(species, sex, body_mass) |> 
    drop_na() |> 
    group_by(species, sex) |> 
    summarize(body_mass_kg_mean = mean(body_mass / 1000), 
              body_mass_kg_sd = sd(body_mass / 1000), 
              .groups="drop") |>
    pivot_wider(names_from="sex", 
                values_from=starts_with("body_mass_kg")) |>
    gt(id="penguins_tbl1") |>
    fmt_number(
        columns=starts_with("body_mass"), 
        decimals=2
    ) |> 
    cols_merge(
        columns=c(body_mass_kg_mean_male, body_mass_kg_sd_male), 
        pattern = "{1} ± {2}"
    ) |>
    cols_merge(
        columns=c(body_mass_kg_mean_female, body_mass_kg_sd_female), 
        pattern = "{1} ± {2}"
    ) |>
    cols_label(
        body_mass_kg_mean_male="Body Mass (Male)",
        body_mass_kg_mean_female="Body Mass (Female)",
        species="Species"
    ) |>
    tab_header("Inter-Sex Differences in Penguin Body Mass by Species") |>
    tab_source_note(md("Data originally published in by K.B. Gorman, T.D. Williams, 
                       and W. R. Fraser in 'Ecological Sexual Dimorphism and 
                       Environmental Variability within a Community of Antarctic 
                       Penguins (Genus *Pyogscelis*). *PLoS One* 9(3): e90081.
                       <https://doi.org/10.1371/journal.pone.0090081>. Later 
                       popularized via the `R` package
                       [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/)"))
```

We note that typical inter-sex differences range from 1.25 (Chinstrap) to 2.6 
(Gentoo) times the intra-sex variability, indicating that a simple sex classifier 
based on sex and body mass is likely to perform well for Gentoo penguins, but
struggle for Chinstraps. 


--- 

Examining the same data graphically, we see that it is 
indeed difficult to distinguish Chinstrap penguins by body mass alone, as the 
significant overlap in the central panel below confirms. 

```{r}
#| echo: false
#| fig-align: "center"
penguins |> 
    select(sex, species, body_mass) |> 
    drop_na() |> 
    mutate(body_mass = body_mass / 1000, 
           sex=str_to_title(sex)) |> 
    ggplot(aes(x=body_mass, fill=sex)) + 
    geom_density(color=NA, alpha=0.6) + 
    facet_wrap(~species) + 
    theme_bw() + 
    scale_fill_brewer(name="Sex", palette = 2, type="qual") + 
    xlab("Body Mass (kg)") + 
    theme(legend.position="bottom") + 
    ylab(NULL) + 
    theme(axis.text.y = element_blank()) + 
    ggtitle("Distribution of Measured Body Mass by Sex and Species") + 
    labs(subtitle="Significant distributional overlap for Chinstrap penguins suggest a body-mass based sex classifier will have poor performance", 
         caption="Penguins data originally collected by Gorman et al.")
```

# Review 

## Single-Table Verbs

`dplyr` single-table verbs

::: {.incremental}

- `select`, `filter`: Selecting rows and columns
- `rename`, `mutate`: Changing rows and columns
- `summarize`, `group_by`: Combining multiple rows
- `arrange`, `slice_min/max`: Re-ordering

:::

. . . 

[Lab #05 Review Activity](../labs/lab05.qmd#review)

{{< include _chunks/breakout.qmd >}}

# Multi-Table Verbs

## Multi-Table Analysis

*Multiple Tables*:

- More insights than from a single table
- Maintain 'tidy' structure throughout

Will create new (compound) rows: 

- Dangers: *drops* and *(over) duplication*

## Primary Keys

_Keys_ are unique identifiers for individual records

- Primary (one column) or compound (multiple columns together) 


. . . 


The history of corporate IT is largely one of (failed) primary keys

- Finance: Tickers, Tickers + Exchange, Tickers + Share Class, CUSIP, ISIN,
  SEDOL, ... 
  
  
. . . 

Meaningful true keys are vanishingly rare - cherish them when you find them

Often 'unique enough' for an analysis

`dplyr::group_by() + dplyr::count()` is helpful here


---

What _primary keys_ and _compound keys_ are used to identify you? 

. . . 

::: {.incremental}

- CUNY EmplID
- Name
- Name + Birthday
- Name + Birthday + ???
- National ID Number (SSN?)

:::

## Joins

Joins combine tables _by identity_ - not simple 'stacking' 

Specify a _join key_ - ideally this is an actual key, but doesn't have to be

In `dplyr`, we use the `join_by` function: 

```{r}
#| eval: false
#| echo: true
dplyr::join_by(table_1_name == table_2_name)
```

Here `table_1_name` and `table_2_name` are column names from two tables

. . . 

Join rows where these values are equal (advanced joins possible)

## Inner and Outer Joins

When tables are perfectly matched, not an issue: 

```{r}
#| echo: false
library(dplyr)
cunys <- tribble(~college, ~campus_borough, 
             "CCNY", "Manhattan",
             "Baruch", "Manhattan", 
             "CSI", "Staten Island",
             "York", "Queens")

routes <- tribble(~borough_name, ~bus_code,
                  "Manhattan", "M",
                  "Staten Island", "S", 
                  "Queens", "Q")
```

```{r}
cunys
```

```{r}
routes
```

## Inner and Outer Joins

When tables are perfectly matched, not an issue: 

```{r}
inner_join(cunys, routes, join_by(campus_borough == borough_name))
```

Default to `inner` but irrelevant

Note automatic repetition of `"M"` row

## Inner and Outer Joins

How to handle 'unaligned' values?

```{r}
#| echo: true
cunys <- tribble(~college, ~campus_borough, 
                 "CCNY", "Manhattan",
                 "Baruch", "Manhattan", 
                 "CSI", "Staten Island",
                 "York", "Queens", 
                 "Medgar Evers", "Brooklyn")

inner_join(cunys, routes, join_by(campus_borough == borough_name))
```

. . . 

[MEC vanished!]{color="red"}

## Inner and Outer Joins

```{r}
#| echo: true
left_join(cunys, routes, join_by(campus_borough == borough_name))
```

MEC stays, but no bus code - `NA` value

. . . 

::: {.smaller}

- `inner_join` - Keep only matches
- `left_join` - Keep all rows in left (first) table even w/o matches
- `right_join` - Keep all rows in right (second) table even w/o matches
- `full_join` - Keep all rows from both tables, even w/o matches

:::

`left_` and `right_` are types of 'outer' joins

## Legos of Data Analysis {.scrollable}

These functions are like Legos: 

- Simple individually
- Combine for complex structures

---

**Q: How many distinct flights left NYC in 2013?**

```{r}
library(dplyr)
library(nycflights13)
flights |> 
    n_distinct()
```

. . . 

Not quite what we wanted...

---

**Q: How many distinct flights left NYC in 2013?**

. . . 

How many unique combinations of `carrier` + `flight` (*e.g.*, United 101)?

```{r}
flights |>
    select(carrier, flight) |>
    n_distinct()
```

---

**Q: How many distinct flights left NYC in 2013?**

::: {.small}

💡 Did airlines re-use flight numbers for different destinations? 

```{r}
flights |>
    distinct(carrier, flight, dest) |>
    # Find reuse of number across different destinations
    # Shorthand for group_by + summarize(n = n())
    count(carrier, flight) 
```

:::

. . . 

Seems so!

BTW: This is [bad and dangerous](https://viewfromthewing.com/airlines-are-running-out-of-flight-numbers-and-they-dont-know-what-to-do-about-it/)! 

---

**Q: How many distinct flights left NYC in 2013?**

Find examples of re-use: 

```{r}
flights |>
    distinct(carrier, flight, dest) |>
    count(carrier, flight) |> 
    slice_max(n)
```

---

**Q: How many distinct flights left NYC in 2013?**

Finding most re-used flight number: 

```{r}
flights |>
    distinct(carrier, flight, dest) |>
    count(carrier, flight) |> 
    slice_max(n) |> 
    left_join(flights, join_by(carrier == carrier, flight == flight)) |>
    pull(dest) |> # pull out column as vector
    table() # frequency table
```

---

**Q: How many distinct flights left NYC in 2013?**

Finding most re-used flight number: 

```{r}
flights |>
    distinct(carrier, flight, dest) |>
    count(carrier, flight) |> 
    slice_max(n)
```

---

**Q: How many distinct flights left NYC in 2013?**

Seeing where our most reused number went:

```{r}
flights |>
    distinct(carrier, flight, dest) |>
    count(carrier, flight) |> 
    slice_max(n) |>
    inner_join(flights, join_by(carrier == carrier, flight == flight)) |>
    count(flight, carrier, dest)
```

---

**Q: How many distinct flights left NYC in 2013?**

Additional join to get airport information + formatting:

```{r}
flights |>
    distinct(carrier, flight, dest) |>
    count(carrier, flight) |> 
    slice_max(n) |>
    inner_join(flights, join_by(carrier == carrier, flight == flight)) |>
    count(flight, carrier, dest) |> 
    inner_join(airports, join_by(dest == faa)) |>
    select(name, n, carrier, flight) |>
    arrange(desc(n)) |>
    rename(`Destination Airport` = name, 
           `Number of Times Flown` = n, 
           `Carrier Code` = carrier, 
           `Flight Number`= flight)
```


---

**Q: How many distinct flights left NYC in 2013?**

::: {.small}

Extra join to match to `airlines` as well: 

```{r}
head(airlines, 3)
```

:::

. . . 

Also has a column named `name` - need to disambiguate!

---

**Q: How many distinct flights left NYC in 2013?**

Additional join to get airport information + formatting:

```{r}
flights |>
    distinct(carrier, flight, dest) |>
    count(carrier, flight) |> 
    slice_max(n) |>
    inner_join(flights, join_by(carrier == carrier, flight == flight)) |>
    count(flight, carrier, dest) |>  
    inner_join(airports, join_by(dest == faa)) |>
    select(name, n, carrier, flight) |>
    rename(dest_name = name) |> 
    inner_join(airlines, join_by(carrier == carrier)) |> 
    arrange(desc(n)) |>
    select(-carrier) |> 
    rename(`Destination Airport` = dest_name, 
           `Number of Times Flown` = n, 
           `Carrier` = name, 
           `Flight Number`= flight)
```


## Diving deeper Into Joins

Data Set: [`nycflights13`](https://nycflights13.tidyverse.org/)

Return to your breakout rooms and continue [Lab #05](../labs/lab05.html)

# Pivoting

## Pivoting

The `pivot_*` functions change the **shape** of data

::: {.incremental}

- Values are not created or destroyed, just moved around
- **wider** data sets are formed by forming multiple rows into columns
- **longer** data sets are splitting columns from the same row into new rows

:::

. . . 

These functions come from the `tidyr` package - not `dplyr`

```{r}
library(tidyr) # included in library(tidyverse)
```

## Pivoting

Untidy example from last week: 

```{r}
#| echo: false
library(tibble)
BARUCH_UNTIDY <- tribble(
    ~Semester, ~Course,      ~Number, ~Type,
       "Fall", "Accounting",     200, "Enrollment",
       "Fall", "Accounting",     250, "Cap",
       "Fall", "Law",        100, "Enrollment",
       "Fall", "Law",        125, "Cap",
       "Fall", "Statistics", 200, "Enrollment",
       "Fall", "Statistics", 200, "Cap",
       "Spring", "Accounting",     300, "Enrollment",
       "Spring", "Accounting",     350, "Cap",
       "Spring", "Law",        50, "Enrollment",
       "Spring", "Law",        100, "Cap",
       "Spring", "Statistics", 400, "Enrollment",
       "Spring", "Statistics", 400, "Cap"
)
BARUCH_UNTIDY
```

## Pivoting

This data was untidy because it split a single unit (course) across multiple rows

`pivot_wider` to get to the right format

. . . 

```{r}
#| echo: true
pivot_wider(BARUCH_UNTIDY, names_from=Type, values_from=Number)
```

## Pivots

`pivot_` changes the shape of a data set. Purposes: 

- Get ready for presentation
- Prep for a join
- Combine rows before looking at 'cross-row' structure

## Pivots

Which penguin species has the largest between-sex mass difference? 

```{r}
#| echo: true
library(tidyr)
avg_mass_tbl <- penguins |> drop_na() |> 
    group_by(sex, species) |> 
    summarize(avg_mass = mean(body_mass), .groups="drop")
    # .groups="drop" is equivalent to |> ungroup()
avg_mass_tbl
```

## Pivots

We want data that is _wider_ (or at least not longer) than our current data: 

| species   | male_avg | female_avg | 
|-----------|----------|-------------
| Adelie    |...|...|
| Chinstrap |...|...|
| Gentoo    |...|...|


## Pivots

For the penguins: 

```{r}
#| echo: true
pivot_wider(avg_mass_tbl, 
            id_cols = species, 
            names_from=sex, 
            values_from=avg_mass)
```

. . . 

::: {.small}

Put it all together:

```{r}
#| echo: true
pivot_wider(avg_mass_tbl, 
            id_cols = species, 
            names_from=sex, 
            values_from=avg_mass) |>
    mutate(sex_diff = male - female) |>
    slice_max(sex_diff)
```

:::

## Pivots

`pivot_wider` Arguments:

::: {.small}

- `id_cols`: kept as 'keys' for new table
- `names_from`: existing column 'spread' to create new
                columns names
- `values_from`: values in new table

:::

. . . 

`pivot_longer`: 

::: {.small}

- 'Inverse' operation
- Spread one row + multiple columns => one col + multiple rows

:::

. . . 

::: {.small}

`pivot_wider` and `pivot_longer` have many additional arguments for dealing with
repeats / missing values. The help page (+ experimenting) is your friend

:::


## Legos of Data Analysis

Question: What does this do I can't do in `Excel`? 

. . . 

*Technically*, nothing. All programming languages of sufficient complexity are
equally powerful (Turing equivalence). 

</br>

. . . 

In actuality, quite a lot: 

- `filter` allows more complex filtering than clicking on values
- `group_by` + `summarize` extend array formulas
- `*_join` provides more complex matching than `VLOOKUP`
- `pivot_*` provide general formulation of pivot tables

+ everything else you can do in `R`. 

Ability to *script* minimizes "hard-coding" of names and values. 

## But truthfully

```{r}
fortunes::fortune(59)
```

. . . 

```{r}
fortunes::fortune(222)
```



## Additional dplyr Tricks

::: {.incremental}

- Ranking functions
  - `row_number`, `min_rank`, `dense_rank`: differ in ties
  - Use with `desc()` to flip ordering
  - `cum_dist`, `percent_rank`: compute quantiles
- Cumulative Statistics
  - `cummean`, `cummax`, `cummin`, ...
  
:::

# FAQs

## Subqueries

> [W]ill we be learning how to perform joins within a subquery?

You don't need subqueries in `R` since it's an imperative language. 
Just create a new variable to represent the result of the subquery
and use that in the next command. 

```
SELECT first_name, last_name
FROM collectors
WHERE id IN (
    SELECT collector_id
    FROM sales
);
```

```
collector_ids <- sales |> pull(collector_id)
collectors |> filter(id %in% collector_ids) |> select(first_name, last_name)
```

## Data Integrity

> [H]ow can we ensure that the information [resulting from a join]
> is accurate and not repeated? 

1. If you have a true unique ID, you're usually safe
2. Pay attention to all warnings
3. Manually examine the result of any joins

## Performance

> Will joining large data sets [...] affect performance?

Somewhat - larger data sets are always slower. 

Bigger danger is "bad joins" creating huge data automatically. 

Note that `R` is less "smart" than `SQL`, so won't optimize
execution order for you automatically.

## dplyr joins vs SQL joins

> What is the difference between `dplyr` and `SQL` joins?

Not too much - biggest difference is no `INDEX` or `FOREIGN KEY`
in `R` so less guarantees of data integrity.

## When to use anti_join()?

Rare: looking for _unmatched_ rows. 

- Useful to find data integrity issues or 'implicit' missingness. 
- I use an `anti_join` to find students who haven't submitted an assignment.

## many-to-many Warning

Tricky to address, but fortunately pretty rare. 

- `SQL` explicitly forbids many-to-many
- Usually a sign that a "key" isn't really unique
  - Check for duplicates in `x` and `y` tables
  - Can occur with "fancy" joins (rolling, inequality)
- Add additional join variables to break "duplication"

## How to Check Efficiency? 

No automatic way. Some rules of thumb: 

- Don't create large tables just to filter down
  - filter before join when possible
- `full_outer` join is a bit dangerous
- `cross_join` is _rarely_ the right answer

## tidyr vs dplyr

> Is `tidyr` more efficient than `dplyr`? 

Nope - different packages from the same developers. 

Designed to work together elegantly. 

## Rare Joins

> What are `cross_join`, filter joins, and `nest_join`?

- `cross_join`: dangerous. 
  - Creates "all pairs" of rows. Useful for 'design' problems 
- filter joins (`anti_`, `semi_`): 
  - Hunting down quietly missing data.
  - Filtering to sub-samples
- `nest_join`: beyond this course. 
  - `left_join` with extra structure to output.

# Wrap-Up

## Review

Multi-Table `dplyr`: 

- `inner_join` and `left_join`
- `join_by` specifications
- `pivot_longer` and `pivot_wider` to get data into optimal formats (`tidyr`)

. . . 

Additional `dplyr`: 

- Ranking, cumulative, and shift functions

## Orientation

::: {.small}

- Communicating Results (`quarto`)  ✅
- `R` Basics  ✅
- Data Manipulation in `R` ⬅️
  - Single-Table Data Manipulation ✅
  - Multiple-Table Data Manipulation ⬅️
- Data Visualization in `R`
- Getting Data into `R` 
- Statistical Modeling in `R`

:::


## Next Time

::: {.small}

- Communicating Results (`quarto`)  ✅
- `R` Basics  ✅
- Data Manipulation in `R` ✅
- Data Visualization in `R` ⬅️ 
- Getting Data into `R` 
- Statistical Modeling in `R`

:::


. . . 

**Data Visualization** 

- Grammar of Graphics
- `ggplot2`
- EDA Plots vs Publication-Ready Plots


{{< include _end.qmd >}}
