---
title: "{{< var course.short >}} - Week 12"
author: "Michael Weylandt"
format: revealjs
echo: true
---

## {{< var course.short >}} Mini-Project #04

[MP#04 online now](https://michael-weylandt.com/STA9750/miniprojects/mini04.html)

- Due **2024-12-04** ($\approx$ 3 weeks - 2 remaining)
- Topic: financial modeling
  - Comparison of two retirement plans
  - Historical data + Monte Carlo ("bootstrapping") 
- Format: 
  - *Decision Analytics* - Play the role of financial advisor
  - GitHub post **AND Brightspace submission**
  
## Week 12 Pre-Assignment 

Due at midnight tonight - take a moment to do it now 
if you haven't already!

# Going Forward

## Pre-Assignments

Brightspace - Wednesdays at 11:45

-   Reading, typically on course website
-   Brightspace auto-grades
    -   I have to manually change to completion grading

Next (and final!) pre-assignment is **December 4th**

. . .

Thank you for FAQs and (honest) team feedback. Keep it coming!

## Next Semester Topics

- NYC Open Data
- Sports (Baseball?)
- Spotify / Music
- Healthcare / Pharmaceutical (might be tricky...)
- Video Games
- Quant Finance / Time Series?
- Baruch Demographics
- Job Market
- Real Estate

## Grading

Returned: 

- MP#02 grade
- MP#02 meta-review grade
- Videos now on Vocat

We owe you: 

- Mid-Term Check-In Feedback

## Grading - *Ex Post* Adjustments

FYI: At the end of the course, I curve individual peer grades. 

*Example:* If grader $X$ is on average, 5 points lower, I re-center all their grades, raising the gradees by an average of 1.25. 

Try to be consistent over the semester so I can calibrate this correctly. 

## GitHub Notifications

Make sure you check GitHub notifications, via email or at
[https://github.com/notifications](https://github.com/notifications)
to make sure you get all peer feedback assignments. 

- I tag you *in other folks' repo* when you are **supposed to review**

- People tagged *in your repo* are **evaluating you**

## Course Support

- Synchronous
  - Office Hours 4x / week
    - MW Office Hours on **Monday** + **Thursday**
    - CR **Tuesday** + **Friday**
    - No OH during Thanksgiving break
- Asynchronous
  - Piazza ($38$ minute average response time)
  
**Change**: MW Thursday Zoom OH now **4:00pm** to **5:00pm**

## Upcoming

Nov 27 - Thanksgiving Holiday (No Class on Nov 28)

- Check-In Peer Feedback (Vocat)

## Comments

Bad - Trivial:

```{r}
#| eval: false
# Set x to 3
x <- 3
```

Bad - Opaque: 

```{r}
#| eval: false
# Follow instructions
x <- 3
```

Bad - Redundant / Explaining Code
```{r}
#| eval: false
# Fit a model
mod <- model(x, y)

# Build a query
query_build() |> query_add() |> query_formulate()
```

## Comments

Good - Purpose ("Business Logic"):

```{r}
#| eval: false
# Regulation XX.YY requires us to apply a risk multiplier to output
# As of 2024-11-01, CFO says risk multiplier is 3
# Cf. Email to risk group, subject "New Risk Multiplier"
RISK_MULTIPLIER <- 3
```

Good - Higher Level Structure (Example from [`googledrive` package](https://github.com/tidyverse/googledrive/blob/main/R/utils-io.R#L18)): 

```{r}
#| eval: false
# https://github.com/gaborcsardi/rencfaq#with-base-r
write_utf8 <- function(text, path = NULL) {
  # sometimes we use writeLines() basically to print something for a snapshot
  if (is.null(path)) {
    return(base::writeLines(text))
  }

  # step 1: ensure our text is utf8 encoded
  utf8 <- enc2utf8(text)
  upath <- enc2utf8(path)

  # step 2: create a connection with 'native' encoding
  # this signals to R that translation before writing
  # to the connection should be skipped
  con <- file(upath, open = "w+", encoding = "native.enc")
  withr::defer(close(con))

  # step 3: write to the connection with 'useBytes = TRUE',
  # telling R to skip translation to the native encoding
  base::writeLines(utf8, con = con, useBytes = TRUE)
}
```

## Comments

More Advice on [StackOverflow](https://stackoverflow.blog/2021/12/23/best-practices-for-writing-code-comments/)

# Pre-Assignment #12 FAQs

## FAQ: Unicode Resources

- Unicode Tables: [unicodeplus.com/](https://unicodeplus.com)
- Taco Emoji [History](https://www.huffpost.com/entry/taco-bell-emoji-petition_n_6416798)
- Taco Emoji [Controversy](https://www.vice.com/en/article/the-brief-history-of-the-taco-emoji-now-has-a-happy-ending/)

## FAQ: Regular Expression Tools

- Testing Regular Expressions Interactively: [regex101.com/](https://regex101.com/)
- Alternative [regexr.com/](https://regexr.com/)
- Automated Regular Expression Builder: [regex-generator](https://regex-generator.olafneumann.org) 
- AI Regexp Builder: [hregexgo.com/](https://www.regexgo.com/)


## FAQ: Substrings and String Splitting

```{r}
fruits <- c("apples and oranges and pears and bananas", 
            "pineapples and mangos and guavas")

stringr::str_split(fruits, " and ")
```

```{r}
stringr::str_split_fixed(fruits, "and", n=2)
```

## Sub-Strings / Splitting

```{r}
x <- "Baruch College, CUNY"
stringr::str_sub(x, end=6) # Includes endpoints
```

```{r}
stringr::str_sub(x, start=-4) # Count from end
```

```{r}
x <- c("Baruch College, CUNY", "Brooklyn College, CUNY")
stringr::str_sub(x, end=-7) # Drop last _6_
```

## FAQ: Start and End Anchors

> When to use the `^` and `$` anchors? 

Start and end of a _line_. 

- Very useful for _structured_ text (computer log outputs)
- In data analysis, a bit less useful
  - Applied to output of `str_split`

## FAQ: Exclusion + Detection

```{r}
x <- c("10 blue fish", "three wet goats")
stringr::str_detect(x, "[^0123456789]")
```

`str_detect` has a `negate` option:

```{r}
stringr::str_detect(x, "[0-9]", negate=TRUE)
```


## FAQ: `str_detect` vs `str_match` vs `str_extract`

- `str_detect` is there a 'fit'?
- `str_extract` extract the _whole_ 'fit'
- `str_match` extract specific groups

```{r}
x <- "Baruch College, CUNY is a wonderful place to work!"
stringr::str_detect(x, "(.*), CUNY")
stringr::str_extract(x, "(.*), CUNY")
stringr::str_match(x, "(.*), CUNY")
```

## FAQ: Subset Selection + Indexing

`str_match(group=)` is useful for complex data extraction. 

```{r}
x <- c("Michael Weylandt teaches STA9750", "KRR teaches STA9891")
pattern <- c("(.*) teaches (.*)")
stringr::str_extract(x, pattern, group=1)
stringr::str_extract(x, pattern, group=2)
```
(Not sure what negatives do here...)

Also allows named groups: 

```{r}
x <- c("Michael Weylandt teaches STA9750 on Thursday", "KRR teaches STA9891 on Wednesday")
pattern <- c("(?<instructor>.*) teaches (?<course>.*) on (?<weekday>.*)")
stringr::str_match(x, pattern) |> as.data.frame()
```

## FAQ: Homoglyphs

```{r}
x <- c("Η", "H")
tolower(x)
```

Why?

. . . 

```{r}
uni_info <- Vectorize(function(x) Unicode::u_char_name(utf8ToInt(x)), "x")
uni_info(x)
```

. . . 

Particularly nasty with dashes - lean on `[[:punct::]]` where possible.

```{r}
x <- c("Em Dash —", "En Dash –", "Hyphen ‐")
stringr::str_remove(x, "[[:punct:]]") # Works
stringr::str_remove(x, "-")  # Keyboard minus = Fail
```

## FAQ: `?` Symbol (Quantifiers)

Quantifiers (multiple matches): 

- `.{a, b}`: anywhere from `a` to `b` copies (inclusive)
- `.{, b}`: no more than `b` copies
- `.{a,}`: at least `a` copies
- `.?`: zero-or-one, same as `.{0,1}`
- `.*`: zero-or-more, same as `.{0,}`
- `.+`: one-or-more, same as `{1,}`

## FAQ: `stringr` vs `grep` / `grepl`

Ultimately the same functionality, but `stringr` has a more consistent interface. 

Conversion table [online](https://stringr.tidyverse.org/articles/from-base.html)

## FAQ: Working Columnwise

All `stringr` functions work well in `dplyr` pipelines ("vectorized"): 

```{r}
library(dplyr); library(stringr)
df <- data.frame(lower_letters = letters)
df |> mutate(upper_letters = str_to_upper(lower_letters))
```

## FAQ: How to Convert to UTF-8

If you know the source encoding: 

```{r}
#| eval: false
inconv(STRING, from="latin1", to="UTF-8")
```

If you don't know the source, .... 

# Today

## Agenda

- Unicode Discussion
- Regex Discussion
- Regex Exercises
- Completion of Cocktail Exercise
- *Time Permitting*: More Scraping
- *Time Permitting*: Statistical Inference

## Breakout Rooms

| Order| Team           |  | Order| Team            |
|------|----------------|--|------|-----------------|
| 1    | Rat Pack       |  | 6    | Ca\$h VZ        |
| 2    | Subway Surfers |  | 7    | Listing Legends |
| 3    | Chart Toppers  |  | 8    | TDSSG           |
| 4    | Metro Mindset  |  | 9    | Broker T's      |
| 5    | Apple Watch    |  | 10   | EVengers        |
