---
session: "`r session <- 4; session`"
last_updated: "`r last_updated <- Sys.Date(); 
                  last_updated |> format('%Y-%m-%d')`"
class_date:
  "`r library(tidyverse);
    class_date <- read_csv('key_dates.csv', name_repair='universal') |>
    filter(Course.Element == 'Class Session',
           Item.Number == session) |>
    pull(Date); class_date |> format('%A %Y-%m-%d')`"
format: revealjs
---

{{< include _setup.qmd >}}

## Course Project Proposals

Today: Course project proposal presentations ([Official Description](../project.qmd#project-proposal-presentations))

- **6 minute presentation**
- Key topics:
  - *Animating Question*
  - *Team Roster*
- Also discuss: Possible specific questions, data sources, 
  analytical plan, anticipated challenges

. . . 

Most important: team names! 

Previous: Rat Pack, Subway Surfers, Going for Gold, *etc.*

## After Proposals

Extra Review of `R` Basics - 100% Optional

. . . 

- Variables, Vectors, Types, Control Flow


## Mini-Project #00

MP#00 due last Friday - Course Infrastructure Set-Up

::: {.incremental}

- Setup RStudio Project
- Create GitHub Account and Repo
- Connect Local Computer to GitHub
- Deploy GitHub Pages

:::

. . . 

I enjoyed getting to know (a bit more about) you!

## Mini-Project #00 

[Mini-Project #00](../mini/mini00.qmd) peer feedback assigned

- Due `r get_mp_pf_deadline(0)`
- 3 peer feedback assignments per student
  - Give 3 comments, get 3 comments
  
. . . 

[Peer Feedback Instructions](../miniprojects.qmd#peer-feedback)
can be found online

## Mini-Project #00

Interactive script automates this process: 

```{r}
#| eval: false
#| echo: true
source("https://michael-weylandt.com/STA9750/load_helpers.R")
mp_pf_perform(0, github="YOUR_GITHUB_ID")
```

Will request the secret code I gave you *via* Piazza when you completed MP#00

. . . 

Will ask you a series of questions and save your 
responses in a specifically formatted file

## MP#00 Peer Feedback

Ungraded assignment, so simple feedback: 

- One strength, one weakness, one suggestion for improvement (3x)

After completion, upload `bspf` file to Brightspace (for privacy / confidentiality)

## MP PF Cycle

Aims of Mini-Project Peer Feedback:

::: {.small}

- Learn to _read_ and _evaluate_ code
- In analysis, rarely _right_ and _wrong_; 
  definitely _better_ and _worse_
- Learn tricks to improve your own site

:::

> "Good artists copy; great artists steal." -- Steve Jobs

. . . 

*Most* coding is reading - *most* reading is reading _your own old code_

## Course Support

Asynchronous Support: Piazza

- All registered now in Piazza!
- **Sub-1 Hour** average time to response

. . . 

Synchronous Support: Office Hours

- Wednesdays (in person) and Thursdays (Zoom) at 5pm

## Pre-Assignments

. . .
  
`r (session + 1) |> sprintf("Pre-Assignment #%02d", x= _)` - Due `r get_pa_deadline(session+1)`

::: {.small}

  - Day before class at 6:00pm
  - Available on course website + Brightspace after 9pm
  - Unlimited re-tries so make sure you get 30/30!
  
:::

# Quick Review

## Values

Basic "things" in `R` ("scalars"): 

::: {.incremental}

- Numeric values: `10`, `3.14`, `0.0002`, `1.234e5`, `3 + 4i`
  - `R` distinguishes between `integer` and `numeric` / `double` but you don't need to
- Character values: `"Baruch"`, `"Michael Weylandt"`, 
  - Arbitrary length - matched quotes (double or single)
  - Mix quotes to put quotes inside a string: `"He said to me: 'Code is great!'"`
- Logical values: `TRUE`, `FALSE` (no quotes)

:::

## Values

Use the `class()` function to see types:

```{r}
#| echo: true
class(3)
```

. . . 

```{r}
#| echo: true
class(3.14159)
```

. . . 

```{r}
#| echo: true
class("I love R!")
```

## Variables and Assignment

We can assign a name to a value: 

```{r}
#| echo: true
x <- 3
```

. . . 

Now anywhere we use `x`, the value `3` automatically is introduced:

```{r}
#| echo: true
x^2
```

## Variables and Assignment

Variable names must be: 

- One word (no spaces)
- All alpha, numeric, or underscores
- Start with a letter

Avoid special "reserved words": `if`, `else`, `for`, *etc*

## Variables and Assignment

Assignment (`<-`) is the _last_ operation, so can use to save results for later use

```{r}
#| echo: true
five_factorial <- 5 * 4 * 3 * 2 * 1
five_factorial
```

. . . 

Then

```{r}
#| echo: true
six_factorial <- 6 * five_factorial
six_factorial
```

## Vectors

An ordered collection of the same type is called a *vector*:

- Create with `c` ("concatenate"): 

```{r}
#| echo: true
x <- c(1, 2, 3)
```

. . . 

```{r}
#| echo: true
class(x)
```

. . . 

```{r}
#| echo: true
length(x)
```

## Vectors

Vectors are everywhere in `R`: 

- The "single" values we saw earlier are just vectors of length `1`

```{r}
#| echo: true
length(3)
```

## Vectors

Access specific elements of a vector with `[]`: 

```{r}
#| echo: true
x <- c("a", "b", "c")
x[2]
```

. . . 

Or give a vector of indices: 

```{r}
#| echo: true
x[c(2, 1, 3)]
```

. . . 

```{r}
#| echo: true
x[c(2, 1, 2, 1, 3, 3)]
```

## Vectors

Negative indices _drop_: 

```{r}
#| echo: true
x[-2]
```

. . . 

This is different from Python: not count backwards!

## Functions

Functions take input ("arguments") and produce results and side-effects: 

```{r}
#| echo: true
x <- c(1, 4, 9, 16)
sqrt(x)
```

. . . 

`x` is an input to the function `sqrt`

`sqrt` doesn't "see" the name `x`; it sees the vector `1, 4, 9, 16`

. . . 

::: {.small}

Not 100% true, but close enough!

:::

## Functions

Some examples: 

- Print to screen with some formatting: `print`
- "Pure" print to screen (no formatting): `cat`
- Combine strings: `paste`
- Math: `sin`, `sqrt`
- Load a package: `library`

## Vectorization

Where possible, functions are *vectorized*: 

```{r}
#| echo: true
x <- c(1, 2, 3)
y <- c(4, 5, 6)

x * y
```

. . . 

Operations occur "in parallel" on matched elements

## Control Flow

Two useful operations for small code snippets: 

```{r}
#| eval: false
#| echo: true
if(condition){
  do_if_true
} else { # Optional - can omit this 'side'
  do_if_false
}
```

. . . 

This is a *conditional* operator: 

::: {.incremental .small}

- Used to run code _sometimes_ 
  - Download a missing file, but not a file already present
  - Throw an error if something bad happens
  - Turn off parts of code with (`if(FALSE){}`)
- _Everything_ between `{` and `}` is handled
- `else` branch is optional
::: 

## Compound Conditions

Can do 'compound' or nested `if/else`: 

```{r}
#| echo: true
#| eval: false
if(x > 10){
  cat("x is very positive.")
} else if(x > 0) {
  cat("x is a little positive.\n")
} else {
  cat("x is negative.\n")
}
```

. . . 



## Control Flow

Two useful operations for small code snippets: 

```{r}
#| eval: false
#| echo: true
for(element in vector){
  process_one_at_a_time(element)
}
```

. . . 

Goes through `vector` taking out one `element` at a time: 

```{r}
#| echo: true
nums <- c(1, 2, 3, 4, 5)
for(n in nums){
  cat(n, "squared is", n^2, "\n")
}
```

*Don't focus on these too much - better alternatives coming soon!*

## Control Flow

By default, the last line of a function is the returned value:

```{r}
#| echo: true
my_absolute_value <- function(x){
  if(x > 0){
    x
  } else {
    -x
  }
}

my_absolute_value(-3)
my_absolute_value(3)
```

## Control Flow

Can override with `return` statement - instantly returns and 'exits' function:

```{r}
#| echo: true
say_hello <- function(name, scream=FALSE, quiet=FALSE){
  text <- paste("Hello", name)
  
  if(scream){
    text <- paste(toupper(text), "!!!")
  }
  
  if(quiet){return(text)} # Stop here if quiet and skip print
  
  print(text)
}
```

. . . 

::: {.smaller}

```{r}
#| echo: true
say_hello("Michael")
```

```{r}
#| echo: true
say_hello("Michael", scream=TRUE)
```

```{r}
#| echo: true
say_hello("Michael", quiet=TRUE)
```

:::

## Control Flow

Overly complicated $\text{sign}(x)$ function:

```{r}
#| echo: true
sign <- function(x){
  if(x > 0){
    1
  } else {
    if(x < 0){
      -1
    } else {
      0
    }
  }
}
```

## Control Flow

Somewhat better $\text{sign}(x)$ function:

```{r}
#| echo: true
sign <- function(x){
  if(x > 0){
    1
  } else if(x < 0){
      -1
  } else {
      0
  }
}
```

## Control Flow

Decent $\text{sign}(x)$ function:

```{r}
#| echo: true
sign <- function(x){
  if(x > 0) return(1) # Use 'return' to stop function here
  if(x < 0) return(-1)
  return(0)
}
```

. . . 

Would be even better to vectorize

## Optional Review

Programming exercises to practice these concepts

# Proposal Presentations

## Presentation Order

```{r}
#| echo: false
PRESENTATION_TABLE
```

# Wrap-Up

## Orientation

- Communicating Results (`quarto`)  ✅
- `R` Basics  ✅
- Data Manipulation in `R`
- Data Visualization in `R`
- Getting Data into `R` 
- Statistical Modeling in `R`


## Next Time

**Data Frames**: 

- Organizing several 'connected' vectors into a table
- Table operations with `dplyr`

## Upcoming Work

Upcoming work from [course calendar](../index.qmd#calendar)

::: {.incremental}

- Weekly Pre-Assignment
- [Mini-Project #00](../mini/mini00.qmd) peer feedback
- [Mini-Project #01](../mini/mini01.qmd) due on `r get_mp_deadline(1)`

:::

{{< include ../advice/amazon.qmd >}}

{{< include _end.qmd >}}

# Optional Review

## Exercise #01

::: {#q1}

```{r}
#| label: q1-cell
#| echo: true
f <- function(x){
  cat("The vector has", length(x), "elements and is of type", class(x))
}
```

:::

**Q:** Write a function `f` that does the following:

```{r}
#| echo: true
f(c(1, 2, 3))
f(c("a", "b", "c"))
f(1:5)
```

*Hint*: Use the `cat` function to print to screen. 

. . . 

{{< contents q1 >}}

## Exercise #02

::: {#q2}

```{r}
#| label: q2-cell
#| echo: true
is_even <- function(x)  (x %% 2) == 0
```

:::

**Q:** Write a *vectorized* function to tell if numbers are even. 

```{r}
#| echo: true
is_even(3)
is_even(c(3, 4, 5))
is_even(1:10)
```

*Hint*: Use the `%%` operator to get remainders from division

. . . 

{{< contents q2 >}}


## Exercise #03

::: {#q3}

```{r}
#| label: q3-cell
#| echo: true
count_even <- function(x)  sum(is_even(x))
```

:::

**Q:** Write a function to count the even elements of a vector: 

```{r}
#| echo: true
count_even(3)
count_even(c(3, 4, 5))
count_even(1:10)
```

*Hint*: Combine Q2 with the "sum of logical = count" trick. 

. . . 

{{< contents q3 >}}


## Exercise #04


**Q:** The `seq()` function lets us construct sequences. What is the average (`mean`) of the first 23 odd numbers?

*Hint*: Read the documentation for `?seq` before trying this question

. . . 

```{r}
#| echo: true
mean(seq(from=1, by=2, length.out=23))
```



## Exercise #05

**Q:** The *alternating harmonic series* is defined as: 

$$1 - \frac{1}{2} + \frac{1}{3} - \frac{1}{4} + \frac{1}{5} + \dots$$

Show that this series converges to $\ln(2)$ by taking a partial sum
of the first one million elements. 

*Hint*: Use recycling to get the signs right: 

```{r}
#| echo: true
seq(1, 5) * c(1, -1)
```

. . . 

```{r}
#| echo: true
sum(1/seq(1, 1e6) * c(1, -1))
```

Compare to: 

```{r}
#| echo: true
log(2)
```

## Exercise #06

**Q:** Write a function that computes the mean of a vector (Don't use the 
built-in `mean` function)

*Hint*: Use the `sum` and `length` functions

. . . 

```{r}
#| echo: true
my_mean <- function(x){
  sum(x) / length(x)
}

my_mean(8)
my_mean(1:10)
```


## Exercise #07

::: {#q7}

```{r}
#| label: q7-cell
#| echo: true
pos_sqrt <- function(x) sqrt(pmax(0, x))
```

:::

**Q:** Write a vectorized function that returns $\sqrt{x}$ if $x$ is positive and 0 otherwise. 

```{r}
#| echo: true
pos_sqrt(c(-4, -1, 1, 4, 9, -9))
```

*Hint*: The `pmax` ("parallel max") function may be useful here: 

```{r}
#| echo: true
x <- c(1, 3, 5,  7)
y <- c(5, 3, 10, 1)
pmax(x, y)
```

. . . 

{{< contents q7 >}}

Alternative - use the `ifelse` function for *vectorized* conditionals (but this
has a warning because it tries to do both conditions)


## Exercise #08

::: {#q8}

```{r}
#| label: q8-cell
#| echo: true
long_string <- function(x)  x[which.max(nchar(x))]
```

:::

**Q:** Write a function that takes in a vector of characters and returns the longest. 


```{r}
#| echo: true
long_string(c("a", "bc", "def"))
long_string(c("My", "name", "is", "Michael"))
```

*Hint*: The `nchar` and `which.max` functions may be
helpful here. 

. . . 

{{< contents q8 >}}

## Exercise #09

**Q:** Write a function that computes the factorial of a value. (Don't use the built-in factorial function)

*Hint*: Use the `seq` and `prod` functions

. . . 

```{r}
#| echo: true
my_factorial <- function(n){
  return(prod(seq(1, n)))
}

my_factorial(8)
```

## Exercise #10

**Q:** Write a function that computes the factorial of a value using a loop (instead of the `prod` function). 

. . . 

```{r}
#| echo: true
my_factorial <- function(n){
  fact <- 1
  for(x in seq(1, n)){
    fact <- fact * x
  }
  return(fact)
}

my_factorial(8)
```

## Exercise #11

**Q:** Show that: $$\sum_{k=0}^{\infty} \frac{1}{k!} = e$$

Use the built-in `factorial` function since it is vectorized. 

. . . 

```{r}
#| echo: true
sum(1/factorial(seq(0, 10000)))
```

## Exercise #12

**Q:** Write a function that takes a vector and returns the maximum element. (Don't use the built-in `max` function.)

*Hint*: You will need to use a loop and a conditional here. 

*Hint*: We started factorial at 1; start max at `-Inf` (why?)

. . . 

```{r}
#| echo: true
my_max <- function(x){
  max_val <- -Inf
  for(v in x){
    if(v > max_val) max_val <- v
  }
  return(max_val)
}

my_max(c(1, -1, 3, -4, 10, -3))
```
