---
session: "`r session <- 1; session`"
last_updated: "`r last_updated <- Sys.Date(); 
                  last_updated |> format('%Y-%m-%d')`"
class_date:
  "`r library(tidyverse);
    class_date <- read_csv('key_dates.csv', name_repair='universal') |>
    filter(Course.Element == 'Class Session',
           Item.Number == session) |>
    pull(Date); 
    class_date |> format('%A %Y-%m-%d')`"
format: revealjs
---

{{< include _setup.qmd >}}

## {{< var course.short >}} Week {{< meta session >}} 

Weekly Course Updates: 

- Brief updates and reminders about course logistics
- Syllabus and Brightspace are binding
  - If something is left out of here, it still happens!
  
. . . 

Today: Introduction to {{< var course.short >}}

# {{< var course.short >}}: {{< var course.title >}} 

## Course Overview

{{< var course.short >}} is: 

::: incremental

- Establishing a professional data analyst portfolio
- Increasing your effectiveness in other courses
- Building your networks with faculty and classmates
- [Getting s#!t done]{color="#FF0000"}

:::

## Learning Objectives

Formal learning objectives can be found [here](../objectives.qmd)

. . . 

Key aims: 

::: incremental

- Deal with **real** data, no matter how messy and unhelpful
- Engage with important technologies and learn to quickly adopt new tools
- Communicate to both technical and non-technical audiences
- Perform *substantial* and *thoughtful* analyses

::: 

## Course Website

**All course information** can be found [on the course website](../index.qmd)

{{< html https://michael-weylandt.com/STA9750/ full-screen-link="false" height="300px">}}

No authentication required - mobile friendly

Source code [on my GitHub](https://github.com/michaelweylandt/STA9750)

## Course Website

Look online for: 

- [Syllabus](../syllabus.qmd)
- [Weekly Pre-Assignments](../preassignments.qmd)
- [Mini-Projects](../miniprojects.qmd)
- [In-Class Activities](../labs.qmd)
- [Course Project](../project.qmd)
- [Additional Policies and Resources](../resources.qmd)

## Instructor

Me: Michael Weylandt

. . . 

Assistant (i.e., new-ish) Professor in the [Department of Info. Systems and Statistics](https://zicklin.baruch.cuny.edu/faculty-research/academic-departments/information-technology-statistics/)

. . . 

Previously: 

::: {.incremental .small}

- Researcher at Sandia National Labs (US Government Lab)
- Postdoc with US Intelligence Community 
- Ph.D. in Statistics at Rice University (Houston, TX)
- "Quant" at Morgan Stanley and a (defunct) hedge fund

::: 

. . . 

See [my website](https://michael-weylandt.com) for more on me and my research

## Course Schedule

```{r}
#| message: false
#| echo: false
holidays <- DATES |> 
    filter(`Course Element` == "No Class") |> 
    pull(Date) |> 
    unique() |>
    sort() |> 
    format("%b %d") |> 
    paste(collapse=" + ")

presentations <- DATES |> 
    filter(str_detect(Details, "Presentation")) |> 
    pull(Date) |> 
    unique() |> 
    sort() |>
    format("%b %d") |> 
    paste(collapse=" + ")
```

Lecture / Lab sessions:

::: {.incremental}

  - Thursdays on Zoom (here!) at 6:05pm
    - Rescheduled class on *Tuesday* on 2026-04-21 ("Conversion Day")
  - 14 weeks. Holidays on `r holidays`
  - No penalty for missing class. 
    - Slides and in-class activities online
    - Lectures generally not posted 
    
:::
  
## Course Schedule
  
Presentation days: 

  - Student Presentations on `r presentations`
  - First two can be one per team, so only **required** day is 
    final presentation
  - Order will be randomly selected by instructor
    - I expect you to be able to attend *all* of class on presentation day
    
## Getting Help

Online Office Hours:

  - Before class (on days we have class) at 5:00pm
  - *Different Zoom Link*

. . . 

In-person Office Hours: 

- Wednesdays at 5pm (on days we have class)

. . . 

Asynchronous

  - Piazza course discussion platform

## DataCamp

This semester, I am providing free access to DataCamp courses

- **100% optional**
- Additional exercises and review of class material
- Access through sign-up link on Brightspace

. . . 

::: {.small}

Restricted to `@baruchmail.cuny.edu` emails - Piazza message if you have a 
different CUNY email and I'll special-case you

:::

## Course Schedule

Major Topics: 

::: {.incremental}

- Web Communication and Literate Programming Technologies - 2 weeks
- Introduction to `R` - 1 week
- Tidy Data Manipulation (SQL-type operations) - 2 weeks
- Plotting and Data Visualization - 2 weeks
- Data Import, Web Scraping, and Text Cleaning - 3 weeks
- Statistical Analysis - 1 week

:::

## Grading

- 24% Pre-Assignments
- 30% Mini-Projects
- 46% Course Project

. . . 

Extra credit for contribution to course materials or 'above and beyond' on Piazza

. . . 

Final grades curved to match ZSB grading guidelines

## Course Policies

See syllabus for fine print: 

::: {.incremental .small}

- Regrading: Must request within 48 hours; total regrade of assignment
- Late work: Not without prior permission or DoS letter *ex post*
- Grace period: two days on mini-project initial submission, none otherwise
- Grade Curve: One at the end of the semester
- External Resources / AI: Free to use for *coding* only, not writing
- Absences: No excuse required, but must attend on presentation days
- Accommodations: ADA (SDS) or Religious (direct with instructor)

:::

## Pre-Assignments

Pre-Assignments: 

::: incremental

- Weekly reading *before class* followed by a short quiz
- Introduce new material + helps me know where folks are confused
- Starts next week 
- Every week except three presentation weeks

::: 

## Pre-Assignments

Pre-Assignments: 

::: incremental

- 24% of Overall Grade
- Mix of new material and review questions
- Lowest 2 dropped (best 8 of 10)
- Take as many times as needed to get 100% (30/30)
- Brightspace configuration is fragile, so Piazza if issues

:::

## Mini-Projects

Traditional homework assignments - "mini-projects"

::: incremental

- **Guided** real-data analysis using course technologies
- Increasingly ambitious over the semester
- Practice *real* data analysis
- Communication and coding
- Professional portfolio
- One ungraded 'set-up' ("MP#00") + best 3 of 4 graded

:::

More later in the course

## Mini-Projects

```{r}
#| echo: false
#| message: false
mp_00_span <- get_mp_deadline(0, as_date_obj=TRUE) - 
              class_date

mp_01_span <- get_mp_deadline(1, as_date_obj=TRUE) - 
              get_mp_deadline(0, as_date_obj=TRUE)

mp_02_span <- get_mp_deadline(2, as_date_obj=TRUE) - 
              get_mp_deadline(1, as_date_obj=TRUE)

mp_03_span <- get_mp_deadline(3, as_date_obj=TRUE) - 
              get_mp_deadline(2, as_date_obj=TRUE)

mp_04_span <- get_mp_deadline(4, as_date_obj=TRUE) - 
              get_mp_deadline(3, as_date_obj=TRUE)
```

Due Dates: 

::: {.incremental}

- MP#00: `r get_mp_deadline(0)` (`r sprintf("%02g days from today", mp_00_span)`)
- MP#01: `r get_mp_deadline(1)` (`r sprintf("%02g days", mp_01_span)`)
- MP#02: `r get_mp_deadline(2)` (`r sprintf("%02g days", mp_02_span)`)
- MP#03: `r get_mp_deadline(3)` (`r sprintf("%02g days", mp_03_span)`)
- MP#04: `r get_mp_deadline(4)` (`r sprintf("%02g days", mp_04_span)`)

:::

These take time - "due date" $\neq$ "do date"

## Mini-Projects

- MP#00: [`r get_mp_title(0)`](../mini/mini00.qmd)
- MP#01: [`r get_mp_title(1)`](../mini/mini01.qmd)
- MP#02: [`r get_mp_title(2)`](../mini/mini02.qmd)
- MP#03: [`r get_mp_title(3)`](../mini/mini03.qmd)
- MP#04: [`r get_mp_title(4)`](../mini/mini04.qmd)

Suggestions welcome - see Archives for previous semesters

## Mini-Projects

100 points per (graded) MP: 

- 80 points for submission 
- 20 points for *peer review*

. . . 

Important to learn to **read** and **write** code

- Single-blind peer review process
- Grade from instructor on quality of comments

. . . 

You *can't * do peer review if you don't submit (auto 0)

. . . 

Best 3 of 4 (total score out of 100) used


## Mini-Projects

A mix of 'warm-up' activities + final deliverable analysis: 

. . . 

Each project is graded on an 8 part rubric: 

::: {.incremental .smaller}

- Project Skeleton
- Written Communication
- Tables & Presentation
- Data Visualization
- Exploratory Data Analysis
- Code Quality
- Data Preparation
- Analysis and Findings

:::

. . . 

::: {.small}

For earlier projects, automatic 10/10 (if you do the bare minimum) 
on some elements

:::

## Course Project

In lieu of exams, semester-long group project

::: {.incremental}

- Three presentations + two reports
  - Project Proposal
  - Mid-Semester Check-In
  - Final Presentation
  - Individual Final Report
  - Group Final Report
  
:::

More [online](../project.qmd)

## Course Project

46% of grade: 

::: {.incremental}

- Blend of individual (20%) and group (26%) grades
- Timeline moved up a bit to give more end of semester breathing room
- Start looking for teammates: form team by `r ROSTERS |> format("%b %d")`
  - Proposal presentation and team contract due `r PROPOSALS |> format("%b %d")`
- First two presentations are mainly about making sure project is well-posed
  - Analysis due in final presentation + individual reports
  
:::

## Workload Expectations

Per federal and state requirements, 9 hours weekly = 6 hours outside of class

::: incremental

- 10 hours of Pre-Assignments
- 5 hours of Post-Class Review
- 45 hours of Mini-Projects (homework)
- 30 hours of Course Project

:::

Note that course project = sum of all mini-projects - be ambitious!

## Care Resources

See [Care Resources for Students](../resources.html#care-resources-for-students1)
for helpful resources:

- Mental health support
- Physical health / medical care
- Food Security
- Financial Security
- Immigration Support

# Any questions?

## Syllabus Quiz

"Syllabus quiz": 

  - Due `r DATES |> filter(str_detect(Details, "Syllabus")) |> 
           pull(Date) |> format("%a %b %d")`
  - Brightspace
  - Unlimited attempts
  - Unlocks all subsequent assignments so do early

# Getting Started with `R` and `RStudio`

## Lab 01

Open [Lab #01](https://michael-weylandt.com/STA9750/labs/lab01.html) and follow
the instructions to get started with `R` and `RStudio`

. . . 

</br>

Random assignment to Zoom breakout rooms

I will visit rooms to provide support

</br>

. . . 

Call everyone back at 8:30pm

## Mini-Project #00

You can now start thinking about [Mini-Project #00](../mini/mini00.qmd)

::: incremental

- Create an account on GitHub + GitHub student developer pack
- Setting up your own website using `quarto` (like my course site)
  - Future Mini-Projects submitted here
- 'Road test' to make sure everything is working correctly
- Register on course discussion board (Piazza)

:::

Due `r get_mp_deadline(0)`

# Wrap-Up

## Course Agenda

- Communicating Results (`quarto`) ⬅️
- `R` Basics 
- Data Manipulation in `R` 
- Data Visualization in `R` 
- Getting Data into `R` 
- Statistical Modeling in `R`

## Course Introduction

- All materials on course website
- Review of course structure and key policies
- More details about mini-projects and course project to follow

## Getting Started with `R`

- Install `R`, `RStudio`, `git`, and `quarto`
- Run a basic bit of `R` code to confirm things work well

## Next Week

**Syllabus Quiz**  on Brightspace

. . . 

Writing documents using `Markdown` and `Quarto`

- First Pre-Assignment due `r get_pa_deadline(next_session)`
  1. [Online reading](../pre/pa02.qmd)
  2. Submit quiz on Brightspace
  
{{< include ../advice/office_hours.qmd >}}

## Musical Treat

</br></br>

::: {.center}

{{< video https://www.youtube.com/watch?v=RHu0ALxqUIo width="800" height="400">}}

:::
