---
title: "{{< var course.short >}} - Week 7"
author: "Michael Weylandt"
format: revealjs
---

## {{< var course.short >}} Mini-Project #01

Grades returned this afternoon.

Review [regrade policy](https://michael-weylandt.com/STA9750/syllabus.html#regrading-policy) and [late work policy](https://michael-weylandt.com/STA9750/resources.html#late-work-policy) if you have questions.

## {{< var course.short >}} Mini-Project #02

[MP#02](../miniprojcts/mini02.html) - Hollywood Movies -- **Due October 23rd**

-   GitHub post (used for peer feedback) AND Brightspace
-   Start early to avoid Git issues

. . .

Pay attention to [the rubric](../miniprojects/mini02.html#rubric)

-   Writing and presentation are about 50% of your grade
-   Evaluated on *rigor* and *thoughtfulness*
-   Use what you learned from MP#01
-   Pre-processed data now available as well

## Upcoming Mini-Projects

Tentative Topics

-   MP#03: Political Analysis
-   MP#04: Retirement Forecasting

## Course Project

Proposal feedback *soon* (need to do 2 more before releasing)

## Pre-Assignments

Brightspace - Wednesdays at 11:45

-   Reading, typically on course website
-   Brightspace auto-grades
    -   I have to manually change to completion grading

Next pre-assignment is **October 23rd**

. . .

Thank you for FAQs and (honest) team feedback. Keep it coming!

## Course Support

-   Synchronous
    -   Office Hours 4x / week
        -   MW Office Hours on **Monday** + **Thursday** for rest of semester
        -   CR **Tuesday** + **Friday**
        -   No OH during Thanksgiving break
-   Asynchronous
    -   Piazza ($<30$ minute average response time)

## Upcoming Week

Due Wednesday at 11:45pm:

-   Pre-Assignment #08 (Brightspace)
    -   Advanced plotting with `ggplot2`
-   MP #02 on GitHub AND Brightspace

## Additional Resources

-   C. Wilke. [*Fundamentals of Data Visualization*](https://clauswilke.com/dataviz/)
-   K. Healy. [*Data Visualization*](https://socviz.co/)
-   H. Wickham [*`ggplot2`: Elegant Visualizations for Data Analysis*](https://ggplot2-book.org/)

. . .

-   B. Yu and R. Barter [*Veridical Data Science*](https://vdsbook.com/)

# Pre-Assignment #07 FAQs

## FAQ: `ggplot2` vs `Tableau`

-   Tableau
    -   \$\$\$
    -   IT department automatically integrates with data sources
    -   Easy, if it does what you want
-   `ggplot2`
    -   Free
    -   Can use arbitrary data sources, with effort
    -   Flexible / customizable

## FAQ: `ggplot2` vs `matplotlib`

-   `ggplot2`
    -   *Data* visualizations
    -   Enforces "good practice" via `gg`
-   `matplotlib`
    -   *Scientific* visualizations
    -   More flexible for good or for ill
    -   Inspired by `Matlab` plotting

Closest Python analogue to `ggplot2` is `seaborn`

## FAQ: Why use `+` instead of `|>`

-   `ggplot2` is older than `|>`
-   Per H. Wickham: if `ggplot3` ever gets made, will use `|>`
-   Unlikely to change: too much code depends on it

## FAQ: Performance

> I tried an interactive plot with $n=132,000$ points, but it brought my computer to a halt. *\[Ed. Paraphrased\]*

That's a lot of plots!!

`ggplot2` is itself pretty fast, but it depends on (possibly slow) graphics backends

- Different file types implement graphics differently.
- You should also think about overplotting / pre-processing

## FAQ: Overplotting

Large data sets can lead to _overplotting_:

  - Points "on top of" each other
  - Can also occur with "designed" experiments / rounded data
  
Ways to address: 

- `geom_jitter`
- `geom_hex`
  
## FAQ: Overplotting

Jitter: add a bit of random noise so points don't step on each other

```{r}
#| echo: true
#| warning: false
library(ggplot2); library(patchwork)
p <- ggplot(mpg, aes(cyl, hwy))
p1 <- p + geom_point() + ggtitle("geom_point")
p2 <- p + geom_jitter() + ggtitle("geom_jitter")
p1 + p2
```

## FAQ: Hexagonal Binning

Little "heatmaps" of counts. Hexagons to avoid weird rounding artifacts

```{r}
#| echo: true
library(ggplot2); library(patchwork)
p <- ggplot(diamonds, aes(carat, price))
p1 <- p + geom_point() + ggtitle("geom_point")
p2 <- p + geom_hex() + ggtitle("geom_hex")
p1 + p2
```


## FAQ: Inside vs. Outside `aes()`

`aes` maps *data to values*. Outside of `aes`, set constant value

```{r}
#| echo: true
library(ggplot2); library(palmerpenguins)
ggplot(penguins, 
       aes(x=bill_length_mm, y=bill_depth_mm, color=species))+ geom_point()
```

## FAQ: Inside vs. Outside `aes()`

`aes` maps *data to values*. Outside of `aes`, set constant value

```{r}
#| echo: true
library(ggplot2); library(palmerpenguins)
ggplot(penguins, 
       aes(x=bill_length_mm, y=bill_depth_mm))+ geom_point(color="blue")
```

## FAQ: Global vs `geom_` specific `aes()`

-   Elements set in `ggplot()` apply to entire plot
-   Elements set in specific `geom` apply there only
    -   Override globals

```{r}
#| echo: true
library(ggplot2); library(palmerpenguins)
ggplot(penguins, 
       aes(x=bill_length_mm, y=bill_depth_mm, color=species))+
    geom_smooth() + 
    geom_point(color="blue")
```

## FAQ: How to choose plot types

Two "modes"

- Exploratory data analysis. Quick, rapid iteration, for your eyes only
  - Let the data tell you a story
  - Low pre-processing: scatter plots, lines, histograms
- "Publication quality". Polished, 
  - You tell the reader a story
  - More processing, more modeling: trends, line segments, ribbons

## FAQ: Color Palettes

Three types of color palettes: 

- Sequential: ordered from 0 to "high"
  - Example: rain forecast in different areas
- Diverging: ordered from -X to +X with meaningful 0 in the middle
  - Example: political leaning
- Qualitative: no ordering

. . . 

When mapping quantitative variables to palettes (sequential/diverging), two approaches: 

- Binned: $[0, 1)$ light green, $[1, 3)$ medium green; $[3, 5]$ dark green
- Continuous

## FAQ: Color Palettes

```{r}
#| echo: true
library(ggplot2); library(palmerpenguins)
ggplot(penguins, aes(x=bill_length_mm, y=bill_depth_mm, color=body_mass_g)) + 
    geom_point() + theme_bw() + 
    scale_color_distiller(type="seq") # Continuous
```

## FAQ: Color Palettes

```{r}
#| echo: true
library(ggplot2); library(palmerpenguins)
ggplot(penguins, aes(x=bill_length_mm, y=bill_depth_mm, color=body_mass_g)) + 
    geom_point() + theme_bw() + 
    scale_color_fermenter(type="seq") # Binned
```

## FAQ: Color Palettes

```{r}
#| echo: true
library(ggplot2); library(palmerpenguins)
ggplot(penguins, aes(x=bill_length_mm, y=bill_depth_mm, color=body_mass_g)) + 
    geom_point() + theme_bw() + 
    scale_color_fermenter(type="seq") # Binned + Sequential
```

## FAQ: Color Palettes

```{r}
#| echo: true
library(ggplot2); library(palmerpenguins)
ggplot(penguins, aes(x=bill_length_mm, y=bill_depth_mm, color=body_mass_g)) + 
    geom_point() + theme_bw() +
    scale_color_fermenter(type="qual") # Binned + Qualitative
```

## FAQ: Color Palettes

```{r}
#| echo: true
library(ggplot2); library(palmerpenguins)
ggplot(penguins, aes(x=bill_length_mm, y=bill_depth_mm, color=body_mass_g)) + 
    geom_point() + theme_bw() + 
    scale_color_fermenter(type="div") # Binned + Diverging
```

## FAQ: How to "hard-code" colors

```{r}
#| echo: true
library(dplyr)
data <- data.frame(x = rnorm(5), 
                   y = rnorm(5), 
                   group = c("a", "a", "b", "b", "b"))

data |> 
    group_by(group) |>
    mutate(n_count = n()) |>
    ungroup() |>
    mutate(color = ifelse(n_count == max(n_count), "red", "black")) |>
    ggplot(aes(x=x, y=y, shape=group, color=color)) + 
    geom_point() + 
    scale_color_identity()
```


## FAQ: How to Customize Themes

Built-in themes + `ggthemes` package: 

```{r}
#| echo: true
#| warning: false
library(ggplot2); library(ggthemes); 
library(palmerpenguins); library(ggpmisc)
p <- ggplot(penguins, 
       aes(x=flipper_length_mm, 
           y=body_mass_g, 
           color=species)) + 
    geom_point() + 
    stat_poly_line(se=FALSE, 
                   color="black") +
    stat_poly_eq() + 
    xlab("Flipper Length (mm)") + 
    ylab("Body Mass (g)") + 
    scale_color_brewer(type="qual", 
                       palette=2, 
                       name="Species") + 
    facet_wrap(~species)
```

## FAQ: Themes

Default theme (`ggplot2::theme_grey()`):

```{r}
p
```

## FAQ: Themes

Black and White theme (`ggplot2::theme_bw()`):

```{r}
p + theme_bw()
```

## FAQ: Themes

Minimal theme (`ggplot2::theme_minimal()`):

```{r}
p + theme_minimal()
```

## FAQ: Themes

Light theme (`ggplot2::theme_light()`):

```{r}
p + theme_light()
```

## FAQ: Themes

Dark theme (`ggplot2::theme_dark()`):

```{r}
p + theme_dark()
```

## FAQ: Themes

Excel theme (`ggthemes::theme_excel()`):

```{r}
p + theme_excel()
```

## FAQ: Themes

Google Docs theme (`ggthemes::theme_gdocs()`):

```{r}
p + theme_gdocs()
```

## FAQ: Themes

*The Economist* theme (`ggthemes::theme_economist()`):

```{r}
p + theme_economist()
```

## FAQ: Themes

*The Economist* theme (`ggthemes::theme_economist()`):

```{r}
p + theme_economist()
```

## FAQ: Themes

*Solarized* theme (`ggthemes::theme_solarized()`):

```{r}
p + theme_solarized()
```

## FAQ: Themes

*Solarized2* theme (`ggthemes::theme_solarized_2()`):

```{r}
p + theme_solarized_2()
```


## FAQ: Themes

Stata theme (`ggthemes::theme_stata()`):

```{r}
p + theme_stata()
```

## FAQ: Themes

Tufte theme (`ggthemes::theme_tufte()`):

```{r}
p + theme_tufte()
```

## FAQ: Themes

*Wall Street Journal* theme (`ggthemes::theme_wsj()`):

```{r}
p + theme_wsj()
```

## FAQ: Themes

Many more online:

- [ThemePark](https://github.com/MatthewBJane/ThemePark) for movie themes

## FAQ: Order of Layers

Order of layers _technically_ matters, but the effect is small

```{r}
#| echo: true
p1 <- ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm)) +
        geom_point(color="black") + 
        geom_smooth(color="blue", method="lm") + ggtitle("Line on points")
p2 <- ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm)) +
        geom_smooth(color="blue", method="lm") + 
        geom_point(color="black") + ggtitle("Points on line")
p1 + p2
```

## FAQ: Order of layers

Order matters more with theme. Adding a `theme_*()` will override any
`theme()` customization you did: 

```{r}
#| echo: true
p1 <- p + theme_bw() + theme(legend.position="bottom")
p2 <- p + theme(legend.position="bottom") + theme_bw() 
p1 + p2
```

## FAQ: `stat_poly_{line,eq}` vs `geom_smooth`

By default `geom_smooth` fits a _generalized additive model_ (GAM)

`ggpmisc::stat_poly_{line,eq}` fit linear models, so they can expose more machinery. 

. . . 

What is a GAM? Take 9890 with me (Spring, Tuesdays at 6) to find out!

## FAQ: Titles and Captions

```{r}
#| echo: true
ggplot() + 
    labs(title="Title", subtitle="Subtitle", caption="Caption",
         tag="Tag", alt="Alt-Text", alt_insight="Alt-Insight")
```

`+ggtitle("text")` is just shorthand for `+labs(title="text")`

## FAQ: Relative Importance of Aesthetics

Perceptually: 

- Location > Color > Size > Shape

Humans are better at: 

- Length > Area > Volume

## FAQ: When to Use Facets?

Facets are `group_by` for plots. Useful for

- Distinguishing intra- vs inter-group trends
- Avoiding overplotting

## FAQ: Simpson's Paradox

```{r}
#| code-fold: true
#| warning: false
#| label: "Create Data"
library(datasauRus)
library(dplyr)
simpsons_paradox <- datasauRus::simpsons_paradox %>%
  filter(dataset == "simpson_2") %>%
  mutate(group = cut(x + y, c(0, 55, 80, 120, 145, 200), 
       labels = LETTERS[1:5])) %>%
  select(- dataset)
```

```{r}
ggplot(simpsons_paradox, aes(x=x, y=y)) + 
    geom_point() + geom_smooth(method="lm")
```

## FAQ: Simpson's Paradox

```{r}
ggplot(simpsons_paradox, aes(x=x, y=y, color=group)) + 
    geom_point() + geom_smooth(method="lm") + facet_grid(~group)
```

## FAQ: Twin Axes Plots

> How can I implement a dual (twin) axis plot in `ggplot2`?

Disfavored. But if you must ...

[sec.axis](https://r-graph-gallery.com/line-chart-dual-Y-axis-ggplot2.html)

Doesn't allow *arbitrary* secondary axes; allows *transformed* axes (e.g., Celsius and Fahrenheit)


## FAQ: Embedding images in `ggplot`

See the `ggimage` or `ggflags` package for images as "points":

```{r}
#| echo: true
#devtools::install_github("jimjam-slam/ggflags"); 
library(ggflags)
d <- data.frame(x=rnorm(50), y=rnorm(50), 
                country=sample(c("ar","fr", "nz", "gb", "es", "ca"), 50, TRUE), 
                stringsAsFactors = FALSE)
ggplot(d, aes(x=x, y=y, country=country, size=x)) + 
  geom_flag() + 
  scale_country()
```

## FAQ: Embedding Images

See `cowplot::draw_image()` for image background: 

```{r}
#| echo: true
library(cowplot)
p <- ggplot(iris, aes(x = Sepal.Length, fill = Species)) +
  geom_density(alpha = 0.7) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  theme_half_open(12)

logo_file <- system.file("extdata", "logo.png", package = "cowplot")
ggdraw() +
  draw_image(
    logo_file, scale = .7
  ) +
  draw_plot(p)
```

# Today

## Diving Deeper with `ggplot2`

Data Sets: 

- [`diamonds`](https://ggplot2.tidyverse.org/reference/diamonds.html) from the `ggplot2` package
- `cdiac` from the `CVXR` package
- `gapminder` from the `gapminder` package

You need to install `CVXR` and `gapminder` now.

Exercise: [Lab #07](../labs/lab07.html)

## Diving Deeper with `ggplot2`: Learning Goals


Today: 

- Fluency with basic `geom`s
- Animation

Next Week: 

- Spatial data
- Interactive graphics
- Dashboards (time allowing)

## Breakout Rooms

| Room | Team           |  | Room | Team            |
|------|----------------|--|------|-----------------|
| 1    | Rat Pack       |  | 6    | Ca\$h VZ        |
| 2    | Subway Surfers |  | 7    | Listing Legends |
| 3    | Chart Toppers  |  | 8    | TDSSG           |
| 4    | Metro Mindset  |  | 9    | Broker T's      |
| 5    | Apple Watch    |  | 10   | EVengers        |
