---
session: "`r session <- 10; session`"
last_updated: "`r last_updated <- Sys.Date(); 
                  last_updated |> format('%Y-%m-%d')`"
class_date:
  "`r library(tidyverse);
    class_date <- read_csv('key_dates.csv', name_repair='universal') |>
    filter(Course.Element == 'Class Session',
           Item.Number == session) |>
    pull(Date); class_date |> format('%A %Y-%m-%d')`"
format: 
  revealjs:
    echo: true
course_repo: "`r yaml::read_yaml('_variables.yml')$course$repo`"
---

{{< include _setup.qmd >}}

## {{< var course.short >}} Week {{< meta session >}} 

Today: `r TODAY_TOPIC`

. . . 


::: {.smaller}

- Communicating Results (`quarto`)  ✅
- `R` Basics  ✅
- Data Manipulation in `R`  ✅
- Data Visualization in `R`  ⬅️
  - Static Plots  ✅
  - Interactivity, Maps, Animated Plots  ⬅️
- Getting Data into `R`
- Statistical Modeling in `R`

:::

# Today

## Today

- Course Administration
- Review Exercise
- Advanced `ggplot2`
  - Spatial Data (Maps)
  - Animation
  - Interactivity
- Wrap-Up
  - Life Tip of the Day

# Administrative Business

## {{< var course.short >}} Mini-Projects

- Mini-Project #01 ✅ (`r get_mp_deadline(1)`)
- Mini-Project #02 ✅ (`r get_mp_deadline(2)`)
- Mini-Project #03 (`r get_mp_deadline(3)`)
  - Submission ⬅️
  - Peer Feedback
- Mini-Project #04 (`r get_mp_deadline(4)`)
  
## Mini-Project #03

[MP#03](../mini/mini03.html) - `r get_mp_title(3)`

**Due `r get_mp_deadline(3)`**

. . . 

Topics covered: 

::: {.incremental .smaller}

- Data Import
  - Package Usage
  - Static Files
  - API Calls
- Advanced Data Manipulation
- Spatial Data Visualization (Optional)
  
:::

## Future Mini-Projects

[MP#04](../mini/mini04.html) - `r get_mp_title(4)`

**Due `r get_mp_deadline(4)`**

. . . 

Topics covered: 

::: {.incremental .smaller}

- Data Import
  - HTTP Requests
  - HTML Scraping (Tabular)
  - Text Processing
- $t$-tests
- Putting Everything Together

:::


## Course Support

- Synchronous: MW Office Hours 2x / week: 
  - **Wednesdays 5pm** in-person
  - **Thursdays 5pm** on Zoom
- Asynchronous: Piazza

## Course Project

[Course Project](../project.html) should be your *main focus* 
for rest of course

- But you still need to do mini-projects and pre-assignments(!)

## Course Project

Final submissions: 

::: {.incremental}

- Detailed rubrics for group report + individual report
  - [Final Group Report](../project.html#rubric-group-report)
  - [Final Individual Report](../project.html#rubric-individual-report)
  
:::


. . . 

Notes added on _integration_ of SQs and strategies for estimating
causal effects. Ask questions! I'm happy to fill these out further.

# Review Exercise

## SSA Registered Births

Time series data of babies born and registered with the 
Social Security Administration (previously in [Lab #05](../labs/lab05.qmd))

. . . 

```{r}
#| echo: false

if(!file.exists("births.csv")){
    download.file("https://raw.githubusercontent.com/michaelweylandt/STA9750/main/births.csv", 
                  destfile="births.csv")
}
library(readr)
library(dplyr)

births <- read_csv("births.csv")
glimpse(births)
```

## SSA Registered Births

Practice visualization: 

1. Are there months with more births?
2. Are there days of the week with more births? 
3. Is there a long-term trend in births?

See [Lab #08](`r url_full`#review) for details.


## Breakout Rooms {.scrollable}

```{r}
#| echo: false
BREAKOUT_TABLE
```

Review activities from [today's lab](`r url_full`#review)


# Advanced `ggplot2`

## Spatial Data

Maps are more interesting than you think!

- The world isn't flat!

. . . 

This can get *intense*, but we will get by with simple features (`sf`)

## Map Projections

{{< video https://www.youtube.com/watch?v=UPaca-SrvEkhttps://www.youtube.com/watch?v=vb-_qqmOcCI width="80%" height="400px">}}

## WGS 84

[WGS 84](https://en.wikipedia.org/wiki/World_Geodetic_System#WGS_84) is a 
robust and widely-used way of creating maps from 3D coordinates. 

- Good default
- Convert existing `sf` using `sf::st_transform(4326)`

. . . 

Not universal!

{{< video https://www.youtube.com/watch?v=L9Di-UVC-_4 width="40%" height="200px" >}}

## sf Package

The `sf` package provides tools for dealing with spatial data: 

- `st_read` for reading `geojson` or `shp` files
- `st_join` for combining spatial data
- Integrates with `ggplot2::geom_sf()`  for plotting

## sf Objects

`sf` object - `data.frame` with additional _geometry_ information 

```{r}
#| echo: true
library(sf); library(tidyverse)
nc <- st_read(system.file("shape/nc.shp", package="sf"))
glimpse(nc)
```

## Chloropleths

A map whose fill depends on the value of interest is called a _chloropleth_:

::: {.small}

```{r}
ggplot(nc, aes(fill=BIR79 / BIR74 - 1, geometry=geometry)) + 
  geom_sf() + 
  guides(fill="none") + 
  ggtitle("% Change in Number of Children Born Per Year 1974-1979")
```

:::

. . . 

Integrates well with `ggplot2`: 

- `geom_sf` knows how to handle geometries
- other features all work as expected (fill, animation, legends, *etc.*)

## Breakout Exercises

Two sets of spatial exercises: 

- Chloropleths
- Cartograms

[Spatial Data in Lab #08](`r url_full`#maps)

# Animation

## Animated Graphics

`gganimate` is the *de facto* standard for animated `ggplot2`: 

- Make a bunch of `png` files
- Combine into a `gif`

. . . 

Most commonly: `+transition_time(VARIABLE)`

NB: Unlike `facet_wrap`, no `~` before `VARIABLE`

## transition_*() functions

Animation is implemented via a new plot element: `transition_*()`: 

```{r}
#| code-line-numbers: "4"
library(gganimate)
ggplot(births, aes(x=day_of_year, y=births)) +
  geom_line() + 
  transition_time(year)
```

## transition_*() functions

Add a title to indicate _time_ being shown: 

```{r}
#| code-line-numbers: "4"
library(gganimate)
ggplot(births, aes(x=day_of_year, y=births)) +
  geom_line() + 
  transition_time(year) + ggtitle("Year: {frame_time}")
```


## transition_*() functions

`gganimate` "tweens" (interpolates) between frames for
smoothness - works best for _continuous_ time data. 

It's a bit weird here and we might prefer a more discrete
interpolation:

```{r}
#| code-line-numbers: "4"
library(gganimate)
ggplot(births, aes(x=day_of_year, y=births)) +
  geom_line() + 
  transition_states(year) + ggtitle("Year: {closest_state}")
```

## Easing

Use the `ease_` functions (or `enter_()`/`exit_*()` separately) for finer control: 

```{r}
ggplot(births, aes(x=day_of_year, y=births, group=year)) +
  geom_line() + transition_states(year, transition_length = 4, state_length=1) + 
  ggtitle("Year: {closest_state}") + enter_grow() + exit_fade()
```

## Grouping

The `group`  variable is used to determine whether something is permanent over time. 

```{r}
penguins |> drop_na() |>
  ggplot(aes(x = bill_dep, y = body_mass)) +
  geom_point() + ggtitle("{frame_time}") + 
  transition_time(year)
```

. . . 

**Bad!** Doesn't make sense to _interpolate_ penguins

## Grouping

```{r}
penguins |> drop_na() |>
  ggplot(aes(x = bill_dep, y = body_mass, group=year)) +
  geom_point() + ggtitle("{floor(frame_time)}") + 
  transition_time(year)
```

See [`gganimate` Getting Started
Documentation](https://gganimate.com/articles/gganimate.html) 
for more

## Breakout Exercises

Animation exercises: 

- `gapminder` data
- How has the world gotten healthier and richer over the past century?

[Animated Visualizations in Lab #08](`r url_full`#animated)

## Animated Graphics

When it works, `gganimate` is great

- PITA when external software is busted

. . . 

Alternatives: 

- Facet plots
- Split facets over pages and scroll quickly: `ggforce::facet_wrap_paginate()`
- Interactivity with autoplay

. . . 

Caution: think about how transitions should be structured

# Interactivity

## Interactivity 

The `ggiraph` package provides for _interactive_ `ggplot2` 

- 

## Alternatives

The `plotly` package provides a `ggplotly()` function which 
works similarly. 

::: {.small .incremental}

Advantages: 

- Has a more sophisticated "time change" interactivity

Disadvantages

- Doesn't support all of `ggplot2`

:::

## Alternatives

Example of `ggplotly` - `frame` gives rise to a time slider:

```{r}
#| code-line-numbers: "2|4"
library(plotly)
G <- ggplot(births, aes(x=day_of_year, y=births, frame=year)) +
  geom_point() + geom_line() + theme_bw()
ggplotly(G)
```

# Shiny


# Plotting FAQs

## ggplot2 & Pie Charts

Why do Pie Charts have a bad reputation?

- Use of _area_ and _angle_ over _length_: less accurate perception
- Depends on `fill` to convey category - limited categories

. . . 

But primarily - "insider smugness"  and hate of `Excel`


## ggplot2 Plot Type Choice

For me: 

- Exploratory mode: 
  - Simple: line, scatter, bar, frequency
- Publication mode:
  - Very context specific
  - "Simplicity" / "Elegance" depend on audience expectations

## ggplot2 Font Sizing

Theme machinery!

```{r}
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + 
    geom_point() + theme(axis.text = element_text(size=24))
```

## Overplotting / ScatterBlobs

Student asked about "scatterblobs" - typo(?) but I love it!

. . . 

- Density based plotting: hexbins, histograms, rugplots
- Data reduction: summarization or sub-sampling

## Optimizing ggplot2 Performance

Active project of `ggplot2` team - not much you can do

. . . 

*Practical advice*: 

- Plot less (see previous slide)
- Take advantage of [`quarto` Caching](https://quarto.org/docs/computations/caching.html)

## ggplot2 Beyond Scatter and Line

Some favorite semi-advanced plot types: 

- Violin plots: combination of boxplot and histogram
- Ridgelines
- Beeswarms

_Deep_ rabbit hole

## ggplot2 + High-Dimensional Data

High-dimensional data: measure many variables per observation ("wide")

High-dimensional data is _hard_ to visualize

. . . 

Approaches: 

- Pair plots for "moderate" HDD
- PCA (or similar dimension reduction. Take 9890!)

## Custom ggplot2 Theme

```{r}
my_theme <- theme_bw() + theme(panel.background = element_rect(fill = 'lightblue'))
ggplot(iris, aes(x=Sepal.Length, y = Sepal.Width)) + geom_point() + my_theme
```

Advanced: 

- `theme_set()` - change `ggplot2` defaults
- `.Rprofile` - set code to run every time you start `R`

## ggplot2 - When Not to Use 

`ggplot2` is designed to make good statistical graphics. Sub-par for: 

- Very advanced interactivity (unless combined with `shiny` or custom JS)
- _Really_ big data
- Hardcore customization / "infographics"

## `git` WTF

Reference: [Happy Git with R](https://happygitwithr.com/)

```{r}
#| echo: false
fs::dir_ls("slides10_files/figure-revealjs/") |>
  fs::file_copy("docs/slides/slides10_files/figure-revealjs/", 
                overwrite=TRUE)
```

# Wrap-Up

## Review

Advanced `ggplot2`: 

- `ggplot2` as a platform for powerful extensions
- Spatial data: `sf`
- Animation: `gganimate`
- Interactivity: `ggiraph`, `shiny`

# Wrap-Up

## Orientation

- Communicating Results (`quarto`)  ✅
- `R` Basics  ✅
- Data Manipulation in `R`  ✅
- Data Visualization in `R`  ✅
- Getting Data into `R` ⬅️
  - Files and APIs ✅
  - Web Scraping **[Thursday]**
  - Cleaning and Processing Text
- Statistical Modeling in `R`

<!-- TODO: LIFE TIP? -->

{{< include _end.qmd >}}
