
$$\newcommand{\P}{\mathbb{P}} \newcommand{\E}{\mathbb{E}} \newcommand{\R}{\mathbb{R}}$$

```{r}
#| echo: false
#| message: false
#| warning: false
set.seed(9750) # For reproducibility
library <- function(pkg){
    options(repos = c(CRAN = "https://cloud.r-project.org"))
    
    pkg <- as.character(substitute(pkg))
        
    if(!require(pkg, character.only=TRUE, quietly=TRUE)) 
        install.packages(pkg)
        
    stopifnot(require(pkg, character.only=TRUE, quietly=TRUE))
}

library(tidyverse)
DATES <- readr::read_csv("key_dates.csv") |>
    rename(element=`Course Element`,
           item=`Item Number`) |>
    mutate(dt = case_when(is.na(Time) ~ as.character(Date),
                          TRUE ~ paste(Date, Time)))
```

```{r}
#| echo: false
#| warning: false
library(glue)
library(rmarkdown)
mp_id <- as.integer(rmarkdown::metadata$mp_num)
mp <-  DATES |> filter(element == "Mini-Projects", 
                       item == mp_id)
mp_file <- list(qmd=glue("`mp{sprintf('%02g', mp_id)}.qmd`"), 
                html=glue("`mp{sprintf('%02g', mp_id)}.html`"),
                html_long=glue("`docs/mp{sprintf('%02g', mp_id)}.html`"))
```

```{r}
#| echo: false
#| message: false
library(yaml)
get_mp_yaml <- function(N){
    fname <- glue("mini/mini0{N}.qmd")
    lines <- readLines(fname, n=30)
    lines_header <- which(lines == "---")
    lines_start <- min(lines_header) + 1
    lines_end <- max(lines_header) - 1
    
    lines_range <- seq(lines_start, lines_end)
    
    lines_yaml <- lines[lines_range]
    
    yaml.load(lines_yaml)
}

get_mp_skills <- function(N){
    get_mp_yaml(N)$mp_skills
}

get_mp_rhetoric <- function(N){
    get_mp_yaml(N)$mp_rhetoric
}


get_mp_application <- function(N){
    get_mp_yaml(N)$mp_application
}
```

### Due Dates

-   Released to Students: `{r} mp |> filter(str_detect(Details, "Mini-Project Released")) |> pull(dt)`
-   **Initial Submission: `{r} mp |> filter(str_detect(Details, "Mini-Project Due")) |> pull(dt)` on GitHub and Brightspace**
-   **Peer Feedback:**
    -   Peer Feedback Assigned: `{r} mp |> filter(str_detect(Details, "Mini-Project Peer Feedback Assigned")) |> pull(dt)` on GitHub
    -   **Peer Feedback Due: `{r} mp |> filter(str_detect(Details, "Mini-Project Peer Feedback Due")) |> pull(dt)` on GitHub**
    
*Estimated Time to Complete: 13-15 Hours*

*Estimated Time for Peer Feedback: 1 Hour*

------------------------------------------------------------------------
