---
mp_num: "03"
mp_title: "TBD"
mp_skills: "TBD"
mp_application: "TBD"
mp_rhetoric: "TBD"
mp_max_ec: 4
filters: 
 - list-table
---

```{r}
#| echo: false
#| warning: false
#| message: false
library(tidyverse)
library(dplyr)
DATES <- readr::read_csv("key_dates.csv") |>
    rename(element=`Course Element`,
           item=`Item Number`) |>
    mutate(dt = case_when(is.na(Time) ~ as.character(Date),
                          TRUE ~ paste(Date, Time)))
```

```{r}
#| echo: false
#| warning: false
library(glue)
mp_id <- as.integer(rmarkdown::metadata$mp_num)
mp_id_s <- glue("MP #0{mp_id}")
mp <-  DATES |> filter(element == "Mini-Projects", 
                       item == mp_id)
mp_file <- list(qmd=glue("`mp0{mp_id}.qmd`"), 
                html=glue("`mp0{mp_id}.html`"),
                html_long=glue("`docs/mp0{mp_id}.html`"))
```

### Due Dates

-   Released to Students: `{r} mp |> filter(str_detect(Details, "Mini-Project Released")) |> pull(dt)`
-   **Initial Submission: `{r} mp |> filter(str_detect(Details, "Mini-Project Due")) |> pull(dt)` on GitHub and Brightspace**
-   **Peer Feedback:**
    -   Peer Feedback Assigned: `{r} mp |> filter(str_detect(Details, "Mini-Project Peer Feedback Assigned")) |> pull(dt)` on GitHub
    -   **Peer Feedback Due: `{r} mp |> filter(str_detect(Details, "Mini-Project Peer Feedback Due")) |> pull(dt)` on GitHub**

*Estimated Time to Complete: 13-15 Hours*

*Estimated Time for Peer Feedback: 1 Hour*


------------------------------------------------------------------------

## Introduction

Welcome to Mini-Project #{{< meta mp_num >}}! TBD

### Student Responsibilities


Recall our basic analytic workflow and table of student responsibilities:

{{< include _responsibilities.qmd >}}

In this project, I am no longer providing code to download and read the 
necessary data files. The data files I have selected for this mini-project are 
relatively easy to work with and should not provide a significant challenge, 
particularly after our in-class discussion of Data Import. 

### Rubric


{{< var course.short >}} Mini-Projects are evaluated using *peer grading* with 
*meta-review* by the course GTAs. The following basic rubric will be used for 
all mini-projects:

{{< include _rubric.qmd >}}

At this point, you are responsible for the 'Data Preparation' portion of the
project, but I am still providing a set of basic EDA activities. Accordingly,
reports completing all tasks described under [Data Integration and
Exploration](#eda) below should receive an automatic 10/10 for the 
'Exploratory Data Analysis' rubric element. 


Taken together, you are only really responsible for these portions of the
rubric in this assignment: 

- Written Communication
- Project Skeleton
- Tables & Document Presentation
- Data Visualization
- Code Quality
- Data Preparation
- Analysis and Findings

Reports completing all key steps outlined below essentially start with
10 free points. 

{{< include _submission.qmd >}}

{{< include _submission.qmd >}}

## Mini-Project #{{< meta mp_num >}}: {{< meta mp_title >}}

### Data Acquisition

TBD


### Data Cleaning and Preparation 

TBD

### Data Integration and Initial Exploration {#eda}

TBD

### Final Deliverable: TBD

TBD

{{< include _aidisclosure.qmd >}}


{{< include _echeader.qmd >}}

TBD

{{< include _footer.qmd >}}

