---
mp_num: "03"
mp_title: "Who Goes There? US Internal Migration and Implications for Congressional Reapportionment"
mp_skills: "Data Import, API Usage, File Parsing"
mp_application: "Census Data, Migration Flows"
mp_rhetoric: "Targeted Advertising, Political Action"
mp_max_ec: 10
filters: 
 - list-table
aliases: 
  - "/miniprojects/mini03.html"
---

{{< include _mpsetup.qmd >}}


## Introduction

Welcome to Mini-Project #{{< meta mp_num >}}! In this mini-project, we will
explore patterns of *internal migration* within the United States with an eye
towards forecasting the size of state congressional delegations for the 2032 
midterm elections. Our analysis will rely on 
[Migration Flows](https://www.census.gov/topics/population/migration/guidance/migration-flows.html)
data inferred from the 
[American Community Survey (ACS)](https://en.wikipedia.org/wiki/American_Community_Survey), 
the long-running premier national survey conducted by the US Census 
Bureau.[^atus] Unlike the decadal census, which occurs once every ten years, the
ACS is constantly collecting new data, giving a up-to-date picture of the 
American public. ACS data is released across a range of granularities, from
national averages to small census-block level estimates. High-level estimates, 
*e.g.*, those for the country as a whole or individual states, are published on
a year-by-year basis, in a data product often called "ACS-1" since they are
derived from samples taken in a one-year window. In order to preserve respondent
privacy and ensure statistical reliability at finer scales, ACS estimates for 
smaller regions are published based on a rolling five-year window: these 
estimates are known as "ACS-5".[^lookback] Larger regions are also available 
from ACS-5 to avoid alignment issues, but these are a bit less useful to us. 

[^lookback]: Specifically ACS-5 estimates are computed over a retrospective 
(looking backwards) five year window: *e.g.*, the 2022 ACS-5 results are 
obtained by combining survey responses from 2018 to 2022. 

For this project, we will primarily make use of state-level ACS-1 flows for our
demographic projections, but we will use metro-level ACS-5 to identify the 
primary endpoints of those flows. We will combine these with other ACS data
via the `tidycensus` package to (roughly) predict state populations in 2030. 
These flows are estimated by asking respondents where they lived
in the prior year, providing detailed data on population movement within the 
United States.  

[^atus]: The American Time Use Survey (ATUS) from [Mini-Project #02](./mini02.qmd)
is a downstream product from the ACS.

### Student Responsibilities

Recall our basic analytic workflow and table of student responsibilities:

{{< include _responsibilities.qmd >}}

In this project, I am no longer providing code to download and read the 
necessary data files. The data files I have selected for this mini-project are 
relatively easy to work with and should not provide a significant challenge, 
particularly after our in-class discussion of Data Import. 

### Rubric

{{< include _rubric.qmd >}}

At this point, you are responsible for the 'Data Preparation' portion of the
project, but I am still providing a set of basic EDA activities. Accordingly,
reports completing all tasks described under [Data Integration and
Exploration](#eda) below should receive an automatic 10/10 for the 
'Exploratory Data Analysis' rubric element. 

Taken together, you are only really responsible for these portions of the
rubric in this assignment: 

- Written Communication
- Project Skeleton
- Tables & Document Presentation
- Data Visualization
- Code Quality
- Data Preparation
- Analysis and Findings

Reports completing all key steps outlined below essentially start with
10 free points. 

{{< include _submission.qmd >}}

## Mini-Project #{{< meta mp_num >}}: {{< meta mp_title >}}

::: {.callout-caution title="Longest Mini-Project of the Semester"}

**NB**: This is the *longest* mini-project of the semester.
[Mini-Project #04](./mini04.qmd) is significantly shorter to give you sufficient
time to complete your [course project](../project.qmd) successfully. Unlike
the previous two mini-projects, you are responsible for data import and cleaning
in this project. Though this isn't glamorous, this is the aspect of data analysis
that typically requires the most time and effort and you should plan accordingly.

:::

### Data Acquisition

When downloading data from a web-based resource, it is important to be a 
*polite* and *responsible* user of that resource. Specifically, we want to
avoid excessively and repeatedly re-downloading a file if it is not changing.
Running a public data source is costly and downloading a file 1,000 times when
you only need it once is simply abusing the good will of others. 

We can ensure responsible usage by adopting a practice of *file caching*:

- Create a "data directory" where data files relevant to the project can 
  be stored
- When attempting to load a data resource, first see if the relevant file is
  in the data directory:
  - If the file is in the data directory, simply read and return the contents of
    that file to the user
  - If the file is not in the data directory, download it into the data 
    directory and save a permanent copy for later use before reading and 
    returning the contents of the file to the user. 
    
This process of "saving" an intermediate result (data file) to avoid an expensive
step (downloading) is known as *caching*. Here, in addition to reducing load on
the data server, caching will also make your analysis faster (as you can avoid 
download times) and allow you to work even when you don't have reliable internet
access. 

Review the data download code I provided you in [MP#01](./mini01.qmd##data-acquisition) 
and [MP#02](./mini02.qmd##data-acquisition) to see how I implemented caching 
there. You should adopt similar patterns for this assignment. 

We will use three different data sources for this assignment: 

- ACS-1 State-to-State Migration Flows
- ACS-5 Metro-to-Metro Migration Flows
- ACS-1 Baseline State Population Data

The last of these will actually be the simplest, so we will start there. 

#### State Population - Package Usage

In order to make our predictions about how many residents each state will have
in 2030, we need to know, as a baseline, how many residents that state has
*presently*.[^presently] Additionally, we want to know how many children are
being born in that state. While fertility is a complex and deeply-studied topic
in demography, we will use a *very simple* model below that holds the natural
growth rate constant across the whole country: that is, we assume families have
(on average) the same number of kids in each state, life expectancies are the
same in every state, and the age breakdown is the same in every state. See
[Extra Credit #03](#ec03) below for opportunities to build a more 
realistic analysis.

Census data is very high-quality, but accessing it can be somewhat tricky The
census reports data in the form of "tables" which have systematic, but difficult
to interpret identifiers. We can access these using the `get_acs()` function 
from the `tidycensus` package, which works like this: 

```{r}
#| message: false
#| warning: false
library(tidycensus)
library(gt)

ny_income <- get_acs(
    c(median_household_income="B19013_001"), 
    state="NY", 
    geography="county", 
    year=2018, 
    survey="acs1"
)
```

In this case, `B19013` is the table "Median Household Income in the Past 
12 Months (in YYYY Inflation-Adjusted Dollars)" and the suffix "001" indicates
that we want the first row of the table. Here, we asked for results from the
2018 ACS-1 county-level estimates for New York State. After
some basic formatting, we get the following results, which can be verified
against the [Census Data Explorer
website](https://data.census.gov/table?q=B19013&g=040XX00US36$0500000&y=2018):

```{r}
#| echo: false
ht <- function(x, n=3){
    rbind(head(x, n) |> mutate(.range="high"), 
          tail(x, n) |> mutate(.range="low"))
}
ny_income |> 
    select(-moe) |>
    pivot_wider(id_cols=c(GEOID, NAME), 
                names_from=variable, 
                values_from=estimate) |>
    arrange(desc(median_household_income)) |>
    ht(5) |>
    gt(id="tbl_income") |>
    cols_hide(".range") |> 
    tab_row_group(label="Lowest Median Household Income Counties", rows= .range == "low") |>
    tab_row_group(label="Highest Median Household Income Counties", rows= .range == "high") |>
    cols_label(median_household_income="Median Household Income") |>
    fmt_currency(median_household_income, 
                 currency="USD", 
                 decimals=0)
```

Here, the `GEOID`, also known as the FIPS code, is a unique identifier used for
different administrative units with the US (so states, counties, cities, *etc*
all have distinct `GEOID`s).  Note that I'm only showing the top and bottom 5 
counties here for brevity.

Note also the use of a _named vector_ as the first argument to `get_acs()` lets
us control the name the result is given in `R`. `get_acs()` has two additional
features that will be useful: 

- the ability to automatically download shapefile information associated with
  each result; 
- the ability to cache results automatically. 

You will want to use both of these. 

::: {.callout-tip title="Task 1: Data Acquisition - State Population and Birth Information"}

Use the `get_acs()` function from the `tidycensus` package to download 
state-level estimates of total population (`B01003_001`) for the years 2015 to
2024. Additionally, use `get_acs()` to automatically download state shape 
information and to appropriately cache your results locally. 

Read the help page for `get_acs()` to identify the caching and shapefile download
arguments. 

You should not need to provide a census API key at this step. If you receive an
**error** (not just a warning or message) indicating that an API key is required, 
please contact the instructor.

Once you have made your request, construct a tidy data frame with the following 
columns:

1) Total population
2) State Name or Abbreviation
3) Year

:::

[^presently]: As of the time of writing this assignment, the latest ACS data are
2024, so interpret words like "current" or "present" to refer to 2024. If 2025
data are available when you are working on this project, please use those 
instead to make more accurate forecasts. 

#### State-to-State Migration Flows - File Download

Since our primary focus is in predicting future state populations, we will
principally work with state-to-state migration flows. The Census Bureau provides
Excel files with these data
[here](https://www.census.gov/data/tables/time-series/demo/geographic-mobility/state-to-state-migration.html). 
While convenient for manual use, this site actually has several of the challenges
associated with complex file-based data exchange: 

- The URLs of the individual files are not consistent
- The formatting of files changes over time

Unfortunately, there is rarely a 'simple' way of dealing with these problems and
we will instead need to write code to handle different cases separately. When
dealing with scenarios like this, a good first step is to decide what you want
your final result to look like and to work backwards from there. While census
provides us with quite a bit, the actual data we need from this data set is
reasonably straightforward: 

| state_current | state_1y | population | year_current | year_1y |
|---------------|----------|------------|--------------|---------|
| AL            | AL       | 4,960,560  | 2024         | 2023    |
| AL            | AK       | 387        | 2024         | 2023    |
| AL            | AL       | 4,920,737  | 2023         | 2022    |
| AL            | AK       | 780        | 2023         | 2022    |

Look at the 
[2024](https://www2.census.gov/programs-surveys/demo/tables/geographic-mobility/2024/state-to-state-migration/State_to_State_Migration_Table_2024_T13.xlsx)
and 
[2023](https://www2.census.gov/programs-surveys/demo/tables/geographic-mobility/2023/state-to-state-migration/State_to_State_Migration_Table_2023_T13.xlsx) 
files and confirm for yourself how I computed these numbers. 

Earlier files in this series have different formats and, while they could be
helpful for our analysis, we are only going to use 2023 and 2024 to keep the
scope of this project tractable. 

::: {.callout-tip title="Task 2: Data Acquisition - State-to-State Flows 2024"}

Write a function `state_to_state_migration_2024()` which extracts state flows
from the 2024 migration file and formats them as shown above. Your function
should, at a minimum, 

1) Create the directory `data/mp03` if it does not already exist. 
   (See previous MPs for examples of how this is done.)
   
2) Check if the 2024 migration file is present in the `data/mp03` directory
   and, if it is not, download it using the `download.file` function. 

3) Read the file into `R` using the `read_excel()` function from the `readxl` 
   package. (This package has two reader functions-`read_xls()` and 
   `read_xlsx()`-and the wrapper function, `read_excel()` which will determine
   which is appropriate for a given file. Unless you have a specific reason not 
   to use it, `read_excel()` should be your default import function.)
   
   You actually will need to use this function twice, varying the `sheet` 
   argument:
   
   i) One time, set `sheet="Table"` to get the actual state-to-state flows
      from the first sheet of the `xlsx` file. 
   ii) On a second time, use `sheet="Supplemental - Current Res"` so you can
       get the "stationary" population. 
       
   In these, you may also find the `skip` function or the `range` argument to 
   import only those rows and columns that will be useful to your analysis. 
   
   *Hint*: Often, Excel's "merged cells" are used as headers in complex data 
   like this, but these are not easily brought into `R`. It is sometimes
   useful to change the `col_names` argument to provide headers yourself 
   rather than trying to read them from the data set. 
   
4) Manipulate the data into an appropriate format: 

   i) For this data, the `"Table"` sheet requires fairly little work, but you 
   will want to remove the "non-moving" rows as these have `X` values instead of 
   actual numbers. 
   
   Additionally, there are some data irregularities in the `Table` sheet you 
   will need to handle. For state-to-state pairs with very little migration, 
   a value of `"N"` is given to indicate that there are not enough values to
   preserve privacy. Use the `replace_values()` function from the `dplyr` 
   package[^new] to replace `"N"` with `"0"` and then use the `as.integer()` 
   function to ensure you have numeric values. 
   
   Finally, add the two year columns and make sure your data is in the format
   given above. 
   
   ii) The `Supplemental` sheet only requires a bit more work. As before, use
   `sheet`, `skip`, and `col_names` to get only what you need. A useful trick
   here is to use `select()` to rename columns by position: *e.g.*, 
   
   ```{r}
   library(dplyr)
   penguins |> select(Species = 1, bill_length=3) |> head(5)
   ```
   
   Add together two columns to get a "stayed within the state" total and 
   manipulate everything else to ensure your data is in the format 
   prescribed above.
   
5)  Use either the `bind_rows()` or `list_rbind()` function to combine the 
    results from each sheet together into a single 2,500 
    ($=50 \times 50$)[^fifty] row data frame.  Note  that these functions combine
    data sets rowwise (*i.e.*, stacks them vertically) so you will need to make 
    sure your data has consistent column names and types before calling it. 
   
    Note that these two functions do essentially the same thing, but differ in
    the expect format of the input: 
   
    ```{r}
    library(dplyr)
    penguins_list <- list(penguins, penguins, penguins)
    all.equal(list_rbind(penguins_list),
              bind_rows(penguins, penguins, penguins))
    ```
    
    So `list_rbind()` assumes its inputs are all in a single list - as you might
    have from, *e.g.*, the output of the `purrr::map()` function - while 
    `bind_rows()` takes its inputs as separate arguments. 
:::

[^fifty]: This data includes `Abroad` or `Foreign` country as a "one year ago"
destination, so you have $50 * (50 + 1 - 1) = 2500$ total rows, not 
$50 * (50 - 1) = 2450$. 

This analysis will be a bit tricky to figure out, but it shouldn't require that
much code when you get all of the pieces in place. I did this in about 40 lines
of reasonably clean code, so if you find yourself using more than 75, step back
and reconsider. 

As you are cleaning data like this, you should _constantly_ be printing output
and comparing it to what you see when you look at the "raw" spreadsheets (keeping
a copy open in Excel). This data is not quite small enough that you can check 
each cell individually, but you can definitely "spot check" data as you work. 
   
[^new]: This function is new in `dplyr` 1.2.0, released in Februrary 2026. If you
do not have this function, restart `R` and run `update.packages()` to ensure
you have the latest versions. 

Once you have read in the 2024 data file, it is time to read the 2023 data file.
This is a bit trickier, but we can make it work. 

::: {.callout-tip title="Task 3: Data Acquisition - State-to-State Flows 2023"}

Write a function `state_to_state_migration_2023()` which extracts state flows
from the 2023 migration file and formats them as shown above. Your function
should, at a minimum, 

1)  Create the directory `data/mp03` if it does not already exist. 
    (See previous MPs for examples of how this is done.)
   
2)  Check if the 2023 migration file is present in the `data/mp03` directory
    and, if it is not, download it using the `download.file()` function. 

3)  Read the file into `R` using the `read_excel()` function from the `readxl` 
    package. In this scenario, you will not need to supply the `sheet`
    argument to `read_excel()` since everything we need is on the `"Table"` sheet. 
   
    Given the complexity of this data, I recommend using the `range` argument
    to get _only_ those cells you need. I recommend using two different calls
    to `read_excel()` to get this right: 
   
    i) First, read only the first 6 columns of the sheet and add together two 
    columns to get the population that stayed within the state. Format your results
    into the five columns specified above. 
   
    ii) Secondly, read in the bulk (columns J to DG) of the table to get the 
    actual flows that we want to extract. The formatting of this sheet makes it
    difficult to get the column names, so it will be useful to provide the
    `col_names` argument directly. Thankfully, `R` gives us a built-in data vector
    with the names of each state (`state.name`) that can be used here, like this: 
   
    ```{r}
    #| eval: false
    library(readxl)
    states_with_dc <- c(state.name[1:8], 
                        "District of Columbia", 
                        state.name[9:50])
                       
    read_excel(..., # Fill in these dots
               col_names=vctrs::vec_interleave(states_with_dc, NA)) |>
        select(any_of(states_with_dc)) |>
        mutate(state_to=states_with_dc) 
    ```
                    
    After this second import, you will want to `pivot_longer()` this data into a 
    format where each row is a pair of states. Finally, prepare this data for use
    by: 
   
    i) Dropping any rows with DC
    ii) Dropping any "within" rows (*e.g.*, (Texas, Texas)) since that value is
        N/A and we get that data elsewhere. 
    iii) Converting the numeric column to integer. 
   
4)  Combine the results of your two imports using `bind_rows()` or `list_rbind()`
    and confirm that you have a 2,500 row 5 column data set as before.[^fifty_two]

:::

[^fifty_two]: For consistency with the prior table, use the total residents
previously living abroad (sum of Puerto Rico, U.S. Island Area, and Foreign
Country) from column DH as your `abroad` factor. 

::: {.callout-tip title="Task 4: Data Acquisition - Combined State-to-State Flows"}

Write a function `state_to_state_migration` which invokes the two functions
you wrote above and uses `bind_rows()` or `list_rbind()` to create a 5,000 row 
table of state to state migration over the previous two years. Confirm that your
results match the example rows shown above. 

:::

::: {.callout-caution title="Do Not Use `tidycensus` for Migration Flows"}

It is possible to get migration flows data using the `tidycensus` package: see
documentation [here](https://walker-data.com/tidycensus/articles/other-datasets.html#migration-flows). 

**Do not use `tidycensus` to access this data.** You must use generic tools
like the `download.file()` function or the `httr2` (not `httr`) package. 

The point of this portion is to practice getting data for which an easy wrapper
is not available. Use of `tidycensus` to access migration data will receive an
automatic 0 for the "Data Preparation" portion of this mini-project and 
significant penalization on the "Project Skeleton" portion. 

You may however use `tidycensus` to cross-check your results: *i.e.*, to 
see if the data you acquire directly matches values given by `tidycensus`.

:::

#### Metro-to-Metro Migration Flows - API Usage

While we can complete our analysis using the state-level data acquired above, 
understanding movement between metro areas (cities) will be useful in gaining a
deeper understanding of migration patterns (*e.g.*, many people move from 
other states to NYS, but almost all of those move to NYC and suburbs, not to 
Albany). For this step, we will use a census API rather than reading formatted
data files. As you will see, the API is typically much easier to parse in `R` 
than working directly with data files. 

The latest metro-to-metro migration flows available from the API are those
from the 2016-2020 ACS-5 window, and can be found
[here](https://api.census.gov/data/2020/acs/flows.html). Review that link and
look at the example API calls.  

::: {.callout-tip title="Task 5: Data Acquisition - Metro-to-Metro Migration Flows"}

Write a function `metro_to_metro_migration` to get city-level migration flows
from the Census API. Your function should, at a minimum, 

1)  Create the directory `data/mp03` if it does not already exist. 
    (See previous MPs for examples of how this is done.)
   
2)  Check if the 2023 migration file is present in the `data/mp03` directory
    and, if it is not, make an API request and save the resulting JSON file using
    the `download.file()` function. 
   
    Determine the URL of the API request using the example API calls linked above.
    Your request should request data at the 
    [metropolitan/micropolitan statistical
    area](https://en.wikipedia.org/wiki/Metropolitan_statistical_area) level for
    the following fields: 
   
    - MOVEDIN: The number of people who moved **into** `FULL1` **from** `FULL2`
    - MOVEDOUT The number of people who moved **into** `FULL2` **from** `FULL1`
    - FULL1_NAME: An MSA Name
    - FULL2_NAME: An MSA Name
   
3)  Read the saved JSON file using the `read_json()` function from the `jsonlite`
    library. 

4)  The census API response isn't formatted quite properly, so `read_json()` will
    interpret the first row of the response as data instead of column names. This
    code snippet can be used to move the first row into column names and do some
    basic type clean-up: 
   
    ```{r}
    #| message: false
    library(tidyverse)
    library(jsonlite)
    move_row_to_colnames <- function(X, row = 1){
        X <- as_tibble(X)
        Xrow <- X[row,]
        
        X <- X[-row,]
        
        colnames(X) <- Xrow
        
        readr::type_convert(X)
    }
    ```
    
    Feel free to use it in your data preparation. 
5)  Simplify column names and drop unneeded columns. 
6)  Use the following mutate command to pull out the (principal) state
    for a metro area. For metro areas contained within a single state, *e.g.*, 
    Los Angeles/Long Beach, this will just pull that one state (`CA`); for metro
    areas spread across multiple states (*e.g.*, "New York-Newark-Jersey City,
    NY-NJ-PA Metro Area") this will pull the first which I'm assuming is the
    largest: 
    
    ```{r}
    pull_state_from_metro <- function(metro_name){
      str_extract(metro_name, ".*, (\\S{2})[-[:alpha:]]* Metro Area", group=1)
    }
    ```
    
    You can use it as follows: 
    
    ```{r}
    example_metros <- c(
      "Los Angeles-Long Beach-Anaheim, CA Metro Area", 
      "Dallas-Fort Worth-Arlington, TX Metro Area", 
      "Riverside-San Bernardino-Ontario, CA Metro Area", 
      "Atlanta-Sandy Springs-Alpharetta, GA Metro Area",
      "Chicago-Naperville-Elgin, IL-IN-WI Metro Area", 
      "New York-Newark-Jersey City, NY-NJ-PA Metro Area", 
      "Outside Metro Area within U.S. or Puerto Rico", 
      "Houston-The Woodlands-Sugar Land, TX Metro Area", 
      "Washington-Arlington-Alexandria, DC-VA-MD-WV Metro Area", 
      "San Jose-Sunnyvale-Santa Clara, CA Metro Area"
    )
    
    data.frame(metro = example_metros) |> 
      mutate(metro_state = pull_state_from_metro(metro)) |> 
      gt(id="tbl_metro_states") |> 
      cols_label(metro="Metropolitan Statistical Area", 
                 metro_state="Primary State")
    ```
:::

::: {.callout-caution title="Do Not Use `tidycensus` for Migration Flows"}

It is possible to get migration flows data using the `tidycensus` package: see
documentation [here](https://walker-data.com/tidycensus/articles/other-datasets.html#migration-flows). 

**Do not use `tidycensus` to access this data.** You must use generic tools
like the `download.file()` function or the `httr2` (not `httr`) package. 

The point of this portion is to practice getting data for which an easy wrapper
is not available. Use of `tidycensus` to access migration data will receive an
automatic 0 for the "Data Preparation" portion of this mini-project and 
significant penalization on the "Project Skeleton" portion. 

You may however use `tidycensus` to cross-check your results: *i.e.*, to 
see if the data you acquire directly matches values given by `tidycensus`.

:::

### Data Cleaning and Preparation 

You should have completed your data cleaning as part of the 
import steps above. While it is generally good to "spot check" data for accuracy,
in this case we can trust the Census Bureau and move immediately to EDA, only 
performing additional checks if issues are identified. 

### Data Integration and Initial Exploration {#eda}

::: {.callout-tip title="Task 6: Exploratory Data Analysis"}

Answer the following questions to perform your Exploratory Data Analysis of the
various data sets used in this project. For each question, answer using inline
values, a table, or a graphic as you feel is most effective. Note that you
will need to determine which data are best suited to answer each question: some
questions might require multiple data sets to answer fully. 

01. Which states have had the highest net population growth rates over the 
    past decade? 
02. If you meet someone who moved to New York State in the last year, what are
    the most likely states they moved from? If one of your friends announces they
    are moving out of the state, where are they most likely to move? 
03. If you meet someone who moved to New York City in the last year, what are
    the most likely metro areas they moved from? If one of your friends announces
    they are moving out of the city, where are they most likely to move?

04. Which states have had the highest amounts of in-migration and 
    out-migration? Furthermore, which states have had the highest amount of
    net migration? 
   
    To get total migration in a direction, add up the total migration in that
    direction: *e.g.*, the total in-migration to Alabama is the sum of people
    who live in Alabama in 2024 but lived elsewhere in 2023 (this is a sum over
    49 other states). 
   
    Net migration is simply total in-migration minus total out-migration.
   
05. Which metro areas have had the highest amounts of total in-migration and 
    out-migration? Furthermore, which metro areas have had the highest amount of
    net migration? 
   
06. Which states have the highest fraction of residents who lived there last year?
    (That is, which states have the lowest proportional total in-migration?)
   
07. Which state has the highest fraction of its population growth attributable
    to net internal migration? That is, determine the total population growth of 
    each state and see what fraction of it corresponds to net migration (as 
    opposed to deaths, births, or international migration).
   
These final questions are specific to the state you are focusing on (see final
deliverable below for more). 

08. Which state had the highest amount of migration into your state? Of the
    people who moved out of your state, what was the most popular destination
    state? 
   
09. What is the largest metropolitan area in your state? (You can answer this
    just using background knowledge; you do not have to use Census data here.)
    
    Which metro area, not located in your state, has the highest amount of 
    migration into your state's largest metro? To which metro area outside of
    your state do the most residents of your largest metro area move? 

10. Are there any metro areas that have a particularly high connection to your
    state's largest metro? (*E.g.*, NYCers were over 15% of people moving
    into the Miami Metro area.)

:::

### Final Deliverable: Ensuring State Political Power in 2032

You are a policy analyst working for the governor of a state (possibly,
but not necessarily, New York State). In light of recent news stories around
redistricting, the governor has asked you to look ahead to the 2030 decennial 
census and the corresponding 2032 apportionment and redistricting cycle. 

Under the US Constitution (Article 1, Section 2, Clause 3), seats in the House 
of Representatives are allocated to the states on the basis of state population.
In particular, 

> Representatives and direct Taxes shall be apportioned among the several States 
  which may be included within this Union, according to their respective Numbers
  [...]. The actual Enumeration shall be made [...] within every subsequent Term 
  of ten Years, in such Manner as they shall by Law direct. 
> The Number of Representatives shall not exceed one for every thirty Thousand,
  but each State shall have at Least one Representative.
  
As such, after each decennial (every 10 year) census, the Congress had to pass
a new bill determining exactly how many seats each state was granted.[^apport]
Because this process became quite contentious, Congress passed the *Permanent
Apportionment Act of 1929* which fixed the size of the House of Representatives
at 435.[^size] Given this fixed cap, states whose population is growing faster
than others are predicted to gain more congressional seats, while states whose
population is growing more slowly (or even shrinking) are expected to lose seats.
Accordingly, estimates of state populations as of 2030 are quite politically
interesting.[^timing]

You have been tasked with i) estimating your state's population in 2030; ii)
estimating your state's congressional apportionment as of 2032, based on this
population; and iii) helping to design a marketing campaign to encourage more
migration to your state.

[^size]: An interesting, but somewhat obscure, idea for increasing the
representativeness of the House of Representative (and, by extension, the
electoral college) is to increase this limit from 435 to something much higher.
See *e.g.*, {{< video https://www.youtube.com/watch?v=KhQGHY44XPM >}} from the
political left or {{< video https://www.youtube.com/watch?v=EpYPQtwXUCQ >}} from
the political right. You can explore this possibility in more detail in 
[Extra Credit #01](#ec01) below. 

[^timing]: Due to the timing of the census, the federal election two years after
a census is typically the first to use this updated apportionment. (So, *e.g.*, 
results of the 2000 census were first used to determine parameters of the 2002
election.)

[^apport]: This is the process of (re-)apportionment, or determining the total
number of seats given to each state. After apportionment, that state determines
how those seats are distributed within its boundaries ("redistricting"), via
state legislative action, independent commission, judicial decree, *etc.* 

Let's begin by constructing estimates of 2030 populations. We can use a very
simple population forecast model: 

$$P_{i}^{(t+1)} = P_i^{(t)} (1 + g_i^{(t)}) + \sum_{j} m_{j \to i} \sqrt{P^{(t)}_iP^{(t)}_j}$$

where 

- $P_i^{(t)}$ is the population of region $i$ in year $t$
- $P_i^{(t+1)}$ is the population of region $i$ in year $t+1$
- $g_i$ is the natural population growth rate (births and deaths) within 
  region $i$
- $m_{i \to j}$ is the migration rate (as a percent of population) from region
  $i$ to region $j$
  
This model is rather simple - notably, it ignores age and cohort effects - but
it's good enough for us to get started. To simplify this further, we're going to
assumes that the natural population growth rate is constant across states
($g_i = \overline{g}$ for all $i$). While this expression looks a bit complex,
it is linear in both of the unknown parameters $g$ and $m_{i \to j}$. We can 
estimate this in two steps: 

1)  Compute $\overline{g}$ for the country as a whole as: 
    $$\overline{g} = \frac{P^{(t+1)} - M^{(t)}}{P^{(t)}} - 1 = \frac{P^{(t+1)} - P^{(t)} - M^{(t)}}{P^{(t)}}$$
    where $M^{(t)}$ is the total net migration (all sources) into region the
    country between year $t$ and year $t+1$. 
    
2)  Compute $m_{i \to j}$ for each state pair as

So, for example, if we have three regions (A, B, C) with the following 
population data: 

```{r}
#| echo: false
library(gt)
pop_example <- tribble(
    ~region, ~pop_y2, ~pop_y1_a, ~pop_y1_b, ~pop_y1_c, ~pop_y1_d, 
        "A",    1100,      1000,         0,         0,        50,
        "B",    1500,        25,      1250,         0,        20,
        "C",    2000,        25,         0,      1975,        50
) |>
  mutate(pop_y1 = rowSums(across(starts_with("pop_y1"))))

pop_example |> 
    gt(id="tbl_pop_example",
       rowname_col="region") |>
    cols_label(region = "Region", 
               pop_y2 = "Total Population (Year 2)", 
               pop_y1_a = "Previously in Region A",
               pop_y1_b = "Previously in Region B",
               pop_y1_c = "Previously in Region C", 
               pop_y1_d = "Previously Living Abroad", 
               pop_y1 = "Total Population (Year 1)") |>
    cols_move("pop_y1", "pop_y2") |> 
    tab_spanner("Location in Year 1", 
                columns=starts_with("pop_y1_")) |> 
    grand_summary_rows(fns = list("Total (Nationwide)" ~ sum(.)))
```

From these, we compute the following key statistics: 

- Total Population in Year 2: $P^{(2)} = 4600$
- Total Population in Year 1: $P^{(1)} = 4275$
- In-Migration: $120$[^outmigration]

[^outmigration]: You might ask about *out-migration* (leaving the country) in
this data. As a general rule, we do not get out-migration from US Census data 
because out-migrants are, by definition, no longer within the US. For our 
purposes, we are simply treating "out-migration" as a part of the death effect
in the natural growth rate. After all, what is death but moving to the great
census tract in the sky? (*Cf.*, my childhood dog who moved to a nice 
farm upstate.)

From these, we see that 

$$\overline{g} = \frac{4600 - 4275 - 120}{4275} \approx 4.8\%$$

so this population is growing at a *very* healthy clip. (For reference, the 
actual US growth rate has never surpassed 1\% in the past 25 years. See
[this Census news 
release](https://www.census.gov/library/stories/2024/12/population-estimates.html).)

Next, we want to determine the migration factors $m_{j \to i}$ for each pair
of regions: 

$$m_{j \to i} = M_{j \to i}^{(t)} / \sqrt{P_i^{(t)}P_j^{(t)}}$$

This model assumes that the number of people who move from region $j$ to region
$i$ depends on both the population in region $j$ and region $i$, so as $j$ 
becomes larger it has more out-migration and as $i$ becomes bigger, it has more
in-migration. This is not terribly realistic, but it will do for our purposes. 

Applying this calculation to our data, we get the following migration factors: 

- $m_{B \to A} = m_{C \to A} = m_{C \to B} =  m_{B \to C} = 0$
- $m_{A \to B} = 25 / \sqrt{1000 * 25} = 0.16$
- $m_{A \to C} = 25 / \sqrt{1000 * 25} = 0.16$

From here, we can uses these to compute population projections for Year 3: 

$$P_{A}^{(3)} = P_{A}^{(2)}(1 + \overline{g}) + \sum_{j \in \{B, C\}}m_{j \to A}\sqrt{P_{A}^{(t)}P_j^{(2)}} = 1100 * (1 + 0.048) + \sum_{j \in \{B, C\}} 0 * \sqrt{1100 * P_j^{(2)}} = 1153$$


$$P_{B}^{(3)} = P_{B}^{(2)}(1 + \overline{g}) + \sum_{j \in \{A, C\}}m_{j \to B}\sqrt{P_{B}^{(t)}P_j^{(2)}} = 1500 * (1 + 0.048) + 0.16 * \sqrt{1500 * 1100} = 1778$$


$$P_{C}^{(3)} = P_{C}^{(2)}(1 + \overline{g}) + \sum_{j \in \{A, B\}}m_{j \to C}\sqrt{P_{C}^{(t)}P_j^{(2)}} = 2000 * (1 + 0.048) + 0.16 * \sqrt{2000 * 1100} = 2333$$
If we wanted to predict further ahead to Year 4, we would simply repeat this 
calculation using $P^{(3)}$ as inputs, holding $\overline{g}$ and $m_{j \to i}$
constant. 

::: {.callout-tip title="Task 7: Population Projections"}

Fit the above model to the US and use it to make population predictions for
2030. Because we have two data sets (migration flows 2023 and 2024), let's fit 
our parameters twice and average them to improve the accuracy of our model. 

1) First, for 2024, use the migration flows data to see how many people were 
   living outside the US in the previous year (2023) to determine $M^{(2023)}$.
2) Using the national population $\overline{P}^{(2023)}$ and 
   $\overline{P}^{(2024)}$, as well as $M^{(2023)}$, determine 
   $\overline{g}^{(2023)}$. 
3) Using the migration flows data and the per-state populations in 2023 and 2024, 
   estimate $m_{j \to i}^{(2023)}$ for 2,450 ($=50 \times (50 - 1)$) state pairs.
4) Repeat Steps 1-3 for 2022 to 2023. 
5) Get parameter estimates by averaging over each of the previous two years, 
   *e.g.*, $\overline{g} = \frac{\overline{g}^{(2023)} + \overline{g}^{(2024)}}{2}$.
6) Using your two year estimates, predict per-state populations for 2025, 2026, 
   2027, 2028, 2029, and finally 2030. 
   
A word of warning: you will compute approximately 7,650 numbers in this
calculation. ($2,451 * 3$ parameters and 300 future populations.) Even though
none of these are too hard to compute, that is still quite a lot of bookkeeping. 
Think *carefully* about how you want to organize your calculations. You will
want to use a combination of `group_by` calculations and `join`s to make this
work. If you find yourself defining scalar variables outside of data frames, you
are almost certainly doing down a difficult path. 

:::

At this point, you have (estimates) for the 2030 population of each state. We 
will use these to estimate congressional apportionments for the 2032 election. 
In the US, this is done using the [Huntington-Hill method](https://en.wikipedia.org/wiki/Huntington%E2%80%93Hill_method). The 
Wikipedia entry is reasonably clear, but I will also demonstrate how this was done
for the 2020 reapportionment, *a.k.a*, the apportionment currently in effect.

```{r}
#| warning: false
#| message: false
library(fs)
library(tidyverse)
library(gt)
library(readxl)

# CA has 52 districts, so if we put in a hard max of 100 we're safe
MAX_DISTRICTS <- 100 
# Fixed per Reapportionment Act of 1929
N_DISTRICTS <- 435

data_mp03 <- fs::path("data", "mp03")
data_file <- fs::path(data_mp03, "apportionment-2020-tableA.xlsx")
if(!fs::dir_exists(data_mp03)) fs::dir_create(data_mp03, recurse=TRUE)
    
if(!fs::file_exists(data_file)){
    download.file("https://www2.census.gov/programs-surveys/decennial/2020/data/apportionment/apportionment-2020-tableA.xlsx", 
                  destfile=data_file)
}    

data_file |> 
    read_excel(skip=4) |> 
    select(state=1, 
           population=2) |> 
    filter_out(state %in%  c("District of Columbia", "U.S. Total")) |> 
    expand_grid(cd=seq(0, MAX_DISTRICTS)) |> 
    mutate(hh_den = sqrt(cd * (cd+1)),
           population=as.integer(population),
           hh_weight = population / hh_den) |> 
    slice_max(hh_weight, n=N_DISTRICTS) |> 
    group_by(state) |> 
    summarize(population = first(population), 
              n_districts = n()) |> 
    mutate(pop_per_district = population / n_districts) |> 
    arrange(desc(n_districts), desc(population), state) |> 
    gt(id="tbl_2020_apportionment",
       rowname_col="state") |> 
    cols_label(state=md("State"), 
               population="Apportionment Population", 
               n_districts="Number of Congressional Districts", 
               pop_per_district="Approximate Population per District") |> 
    fmt_integer(c(population, pop_per_district)) |> 
    data_color(columns=pop_per_district, 
               palette="PuOr") |>
    grand_summary_rows(columns=c(population, n_districts), 
                       fn=list("Total" ~ sum(.)), 
                       fmt = ~ fmt_integer(.)) |> 
    tab_header("Congressional Apportionments from the 2020 Census") |> 
    tab_source_note(md("Apportionments computed using the 
[Huntington-Hill Method](https://en.wikipedia.org/wiki/Huntington%E2%80%93Hill_method).
Data from the US Census Bureau 
[Reapportionment Table A](https://www2.census.gov/programs-surveys/decennial/2020/data/apportionment/apportionment-2020-tableA.xlsx)
(Resident Population + Assigned Overseas Population).")) |> 
    tab_footnote(
        locations=cells_column_labels(pop_per_district),
        "Colors indicate population per congressional representative, ranging from
almost 1 million residents per representative (DE, purple) to just over 540 thousand 
per representative (Montana, brown)."
    )
```

Note that this analysis uses the so-called *Apportionment Population* which
includes 'actual' residents as well as assigned overseas populations (*e.g.*, 
mapping active duty military back to their home states). You do not need to 
follow this subtlety in your analysis. 

::: {.callout-tip title="Task 8: Reapportionment"}

Apply the Huntington-Hill method to your population forecasts in order to 
estimate the number of congressional districts your state will have *as of* 
the 2030 redistricting cycle. 

:::

We are now finally ready for our final deliverable. 

::: {.callout-tip title="Task 9: Developing an Advertising Strategy to Induce Internal Migration and Increase State Political Power"}

Your boss, the governor, wants to increase the number of congressional
seats your state has following the next census. To do so, you have been 
charged with developing an advertising strategy to convince more people
to move to your state. Using your analysis up to this point (and any additional
analysis you find helpful), develop a proposal for a new state advertising 
strategy. This strategy should: 

i) Target common sources of migration *into* your state and its major metro
   areas in an attempt to increase *in-migration*. 
ii) Target common destinations of migration *out of* your state and its major
    metro areas in an attempt to convince people to move back. 
    
For additional value, you might want to target metro areas in states that are
near the low end of the population per resident statistic. If their population
drops much lower, they might loose a congressional seat and you might hope to 
take it. *E.g.*, in the 2020 apportionment shown above, Montana has just an 
apportionment population just over 1 million and two congressional seats. If 
just 10,000 Montanans were to move to New York, NY would gain an additional seat
(going to 27) and Montana would have only one seat.[^ny_gain]

To help 'size' your advertising campaign - and to get legislative support to pay
for it - estimate how much you would internal migration you would need to net
an additional congressional representative. (You do not need to do this 
analytically, rough numbers will suffice.) 

Finally, write up a proposal for your advertisements including: 

i) Where you want to advertise (see, at least, the two target metros above)
ii) How much migration you are hoping to induce
iii) A new slogan or marketing pitch for your state.[^fl]

:::

[^ny_gain]: In 2020, NY could have actually gained an extra seat with just a few
hundred Montanans, but that seat would have come from Minnesota, not Montana. 
The apportionment algorithm is not entirely intuitive under small perturbations.

[^fl]: *E.g.*, if you were trying to draw New Yorkers who retired to Florida 
back into the state, a slogan of "At least we keep our gators in the sewers"
might be appropriate. 

{{< include _aidisclosure.qmd >}}

{{< include _echeader.qmd >}}

### Extra Credit Opportunity #01: Expanding the House {#ec01}

The (relatively) small size of the US House of Representatives is often cited
as a cause of many systemic problems in US politics. For up to two points, 
write a brief 3 paragraph argument for upsizing the House focusing on i) 
equilibration of population per representative across states; and ii) the
additional (relative) power gained by your state in a larger House. You may
assume that the basic contours of the system remain unchanged (minimum one
representative per state, the Huntington-Hill allocation) and that only the 
number 435 is up for debate. 

### Extra Credit Opportunity #02: Spatial Visualizations {#ec02}

When analyzing data like this, *maps* and other data visualizations can be 
very helpful. For up to three points, add additional visualizations to your
analysis. You should include:

1) two maps
2) one chord diagram.[^help]

The `geometry=TRUE` argument to `tidycensus` functions will be *very* helpful 
to you here. 

If you omit the chord diagram, but have at least two maps, you can only get up
to two points here. 

[^help]: The `ggraph` package, especially the linear layout with `circular=TRUE`
might be helpful, as well as [this page](https://www.data-to-viz.com/graph/chord.html)
from the R Graph Gallery. 

### Extra Credit Opportunity #03: More Realistic Growth Models {#ec03}

The population growth model described above is rather limited and assumes
a lot of temporal and cross-sectional (between state) homogeneity. For up to
four points, build a more realistic growth model that takes additional
demographic data into account. 

For one point, fit a different growth factor $g$ for each state instead of a
national growth factor. 

For up to two more points beyond that, further modify your model to take advantage
of additional census variables. There are many ways that you might choose to 
modify this model, but some possibilities include: 

i) Forecasting on smaller levels than the state level (*e.g.*, county or after
   dividing into a rural/urban split)
ii) Using demographic variables (especially age) to forecast how growth and
    death rates will change. (Younger populations have more children.)
    
For one more point, use a longer history to improve estimation of long-run 
growth parameters. This may involve parsing more data files. 

### Extra Credit Opportunity #04: Model Validation {#ec04}

Finally, for up to four points, evaluate the performance of your population 
forecast model _retrospectively_. *E.g.*, if you had used the same model in 
2020, how well would it have predicted populations in 2022 to 2025? There 
are many ways you can go with this, but at a minimum, you should 

i) Make at least one additional set of predictions based on historical data
ii) Compare the accuracy of those predictions with the realized values on a 
    one-year, two-year, and three-year horizon. 
iii) Use that error estimate to put some sort of margin of error on your 2030
     population predictions. 

{{< include _footer.qmd >}}

