---
mp_num: "01"
mp_title: "What We Watched: Exploring Netflix Viewing Data"
mp_skills: "Data Summarization, Exploratory Analysis, Table Formatting"
mp_application: "Media, Internet"
mp_rhetoric: "Strategy Memo"
mp_max_ec: 4
aliases: 
  - "/miniprojects/mini01.html"
---

{{< include _mpsetup.qmd >}}

## Welcome to {{< var course.short >}} Mini Projects!

In the {{< var course.short >}} Mini-Projects, you will perform basic data
analyses intended to model best practices for your course final project.
(Note, however, that these are *mini*-projects; your final course project is
expected to be far more extensive than any single MP.)

## Introduction

The global streaming giant Netflix needs no introduction. Following Netflix's
original pivot-to-streaming (2010-2011) and move into original content (2013
*House of Cards*), Netflix has defined a new type of media company that 
leverages an enormous library of licensed and original content in a bid to 
capture as much viewing time as possible. For years, Netflix's actual viewership
data was a closely guarded secret, but in recent years, two factors have forced
Netflix to be more transparent: 

- The rise of an _advertising_ tier: advertisers want to know how many people
  are likely to actually see their ads and to know which shows drive the most
  eyeballs
- The development of profit participation for creators: historically, Hollywood
  allowed creatives to own some of the equity in their products, so more
  successful productions would lead to higher pay on the backend. Netflix 
  resisted this model for a long time-preferring to provide larger 'up-front'
  payments to talent-but recent pressure from the creative community has forced
  Netflix to relent. In order to ensure that creators receive fair and accurate
  backend compensation, Netflix has had to share data with the larger Hollywood
  community. 
  
To make this data available, Netflix has published a "What We Watched" report
every six months starting in 2023.[^top10] These WWW reports aggregate total
global viewership for almost the entire Netflix catalog over six month periods.[^small] Netflix has argued that these statistics show the depth and
success of their programming, but they have also recently announced their
intent to [release data less frequently and not at the same time as financial
statements](https://deadline.com/2026/07/netflix-viewership-data-reports-change-to-annual-1236982322/)
so you can judge for yourself. 

In this mini-project, you will explore the WWW data to get a sense of what types
of programming captures the bulk of Netflix watch time. Along the way, you will
use your `dplyr` data manipulation skills and the `gt` package to provide 
sophisticated well-formatted insights. Your findings will be summarized in a 
"strategy memo" for Netflix senior leadership, in which you advise them on how
to most effectively spend their 2027 budget to maximize leadership. 

[^top10]: The WWW reports give _global_ data for _all_ of Netflix. Netflix also
releases country-specific Top 10 data. See
<https://michael-weylandt.com/STA9750/archive/AY-2025-FALL/miniprojects/mini01.html>
for a prior mini-project using this data. 

[^small]: Shows with less than 100,000 hours of watch time are omitted and 
viewership is founded to the nearest 100,000, so this data doesn't capture
small differences in unpopular shows. The 'winners' are still quite clear in
this data. 

### Student Responsibilities

For purposes of MPs, we are dividing the basic data analytic workflow into
several major stages:

{{< include _responsibilities.qmd >}}

In early stages of the course, such as this MP, I will 'scaffold' much of the
analysis for you, leaving only those stages we have discussed in class for you
to fill in. As the course progresses, the mini-projects will be more
self-directed and results less standardized.

### Rubric

{{< include _rubric.qmd >}}

At this early point, you are not responsible for all elements of this rubric. 
In particular, all submissions will receive an automatic 10/10 for Data 
Visualization as this is outside the scope of this mini-project. Furthermore, 
because I am providing code to download the data, load it into
`R`, and prepare it for analysis, all reports submitted using my code will 
receive an automatic 10/10 for the 'Data Preparation' element of the rubric. 
Finally, reports completing all tasks described under [Data Integration and
Exploration](#eda) below should receive a 10/10 for the 'Exploratory 
Data Analysis' rubric element. 

Taken together, you are only really responsible for these portions of the
rubric: 

- Written Communication
- Project Skeleton
- Tables & Document Presentation
- Code Quality
- Analysis and Findings

Reports completing all key steps outlined below essentially start with
30 free points. Note however that you must submit this assignment to get these 
points: no points are given for submitting nothing, but it's worth submitting
*something* even if you do not have a fully complete and polished assignment.

<div id="mp_rubric_auto" style="display: none;">Data Visualization;Exploratory Data Analysis;Data Preparation</div>

::: {.callout-warning title="Writing Requirements"}

Note that you are evaluated *on writing and communication* in these 
Mini-Projects. You are required to write a report in the prescribed style, 
culminating in an Op-Ed. A submission that performs the instructor-specified 
tasks, but does not write and give appropriate context and commentary will 
score very poorly  on the relevant rubric elements. 

In particular, if a submission does not include a clearly delineated Op-Ed and
only answers the instructor prompts in narrative text, peer evaluators should
judge  it to have "Good" quality Written Communication (at best) as key findings
are  not conveyed appropriately. 

Quarto's [code 
folding](https://quarto.org/docs/output-formats/html-code.html#folding-code) 
functionality is useful for "hiding" code so that it doesn't break the flow of
your writing. 

You can also make use of [Quarto's `contents`
shortcode](https://quarto.org/docs/authoring/contents.html) to present
code and findings in an order other than how the code should be executed.
This is particularly useful if you want to include a figure or table in an 
"Executive Summary" at the top of your submission.

:::

{{< include _submission.qmd >}}

## Mini-Project #{{< meta mp_num >}}: {{< meta mp_title >}}

### Data Acquisition

The following code can be used to acquire data from Netflix's corporate
relations website. Specifically, once run, this code will download and aggregate
the previous seven WWW reports (first-half 2023 `2023H1` to first-half 2026
`2026H1`). Netflix provides slightly different data from each report, so this
function standardizes on a small set of key features including: the 
'watch window', the title of the program, the runtime, the total hours viewed, 
the original release date of the program, whether the program is available in 
all countries in which Netflix is available, and whether the program is
pre-recorded or live. 


To be efficient, this function will save a copy of the downloaded 
data in a folder called `data/mp01` and use that copy to avoid re-downloading a
file if it is already present on your computer. This will make your code faster
to run and will avoid putting unnecessary stress on the Netflix servers. 

```{r}
#| message: false
#' Acquire Netflix What We Watched Reports
#' 
#' This function will acquire and standardize all data for MP#01
#' from Netflix. Because Netflix hosts the data files behind a CDN, we
#' are accessing the release pages and automatically extracting links to the
#' relevant files (with CDN hash). 
#' 
#' After import, files are stacked vertically and some basic cleaning is
#' performed. 
load_netflix_data <- function(){
  # Note to students: Make sure you have installed the tidyverse and fs packages
  #                   before beginning this assignment. If you aren't sure,
  #                   simply run `library(tidyverse)` at your R Console and, if
  #                   you don't see an error message, `tidyverse` has been 
  #                   installed. (Same for `fs`) If you still need to install
  #                   either package, use the `install.packages` function as 
  #                   demonstrated in class.
  library(tidyverse)
  library(rvest)
  library(readxl)
  library(fs)
  
  data_dir <- path("data", "mp01")
    
  if(!dir_exists(data_dir)){
    dir_create(data_dir, showWarnings=FALSE, recurse=TRUE)
  }
  
  NETFLIX_URLS <- list(
    `2023H1` = "https://about.netflix.com/news/what-we-watched-a-netflix-engagement-report",
    `2023H2` = "https://about.netflix.com/news/what-we-watched-the-second-half-of-2023",
    `2024H1` = "https://about.netflix.com/news/what-we-watched-the-first-half-of-2024",
    `2024H2` = "https://about.netflix.com/news/what-we-watched-the-second-half-of-2024",
    `2025H1` = "https://about.netflix.com/news/what-we-watched-the-first-half-of-2025",
    `2025H2` = "https://about.netflix.com/news/what-we-watched-the-second-half-of-2025",
    `2026H1` = "https://about.netflix.com/news/what-we-watched-the-first-half-of-2026"
  )
  
  download_if_needed <- function(url, local){
    if(!file_exists(local)){
      download.file(url, local, quiet = TRUE, mode = "wb")
    }
    
    return(local)
  }
    
  NETFLIX_URLS |> 
    map(read_html) |> 
    map(html_element, xpath = "//a[contains(@href, '.xlsx')]") |> 
    map_chr(html_attr, "href") |> 
    enframe("period", "url") |> 
   mutate(path = path(data_dir, paste0(period, ".xlsx")), 
          path = map2(url, path, download_if_needed), 
          data = map(path, read_excel, skip=3, na="*"), 
          data = map(data, \(x) x |> mutate(`Release Date` = as.Date(`Release Date`)))) |>
    unnest(data) |> 
    select(-url, -path) |> 
    mutate(year=str_sub(period, 1, 4), 
           half=str_sub(period, start=6), 
           period_start = make_date(year, if_else(half==1, 1, 7), 1), 
           period_end = make_date(year, if_else(half==1, 7, 12), if_else(half==1, 30, 31))) |> 
    filter(!is.na(`Hours Viewed`)) |> 
    mutate(live_event = !is.na(Runtime) & str_detect(Runtime, "\\d+:\\d+\\*"), 
           clean_time = str_remove(Runtime, fixed("*")),
           Runtime = 60 * hour(hm(clean_time)) + minute(hm(clean_time))) |>
    select(-clean_time) |>
    mutate(Runtime = coalesce(Runtime, mean(Runtime, na.rm=TRUE)), 
           .by=Title, 
           Title = stringi::stri_trans_general(Title, "Latin-ASCII"),
           Title = str_remove(Title, " // .*"), 
           Title = str_remove(Title, "[^[:ascii:]].*"), 
           Title = str_trim(Title)) |>
    filter_out(Title == "") |>
    select(year, 
           period, 
           period_start, 
           period_end, 
           Title, 
           -half, 
           Runtime, 
           Views, 
           `Hours Viewed`, 
           `Release Date`,
           `Available Globally?`, 
           live_event)
}

NETFLIX_DATA <- load_netflix_data()
```

This creates a data frame called `NETFLIX_DATA` that has `r NROW(NETFLIX_DATA)` 
rows and `r NCOL(NETFLIX_DATA)` columns in your local environment. This data 
will be used for the remainder of this mini-project. 

::: {.callout-tip title="Task 1: Data Acquisition"}

Using the code above, acquire the Netflix "What We Watched" data. Copy the code 
into your Quarto document and make sure it runs successfully. 

:::

::: {.callout-caution title="Do Not `git add` Data Files"}

Make sure that `git` is set to ignore data files, such as the one created above.
Check the `git` pane in `RStudio` and make sure that the files in `data/mp01`
do not appear. (If you set up your `.gitignore` file correctly in
[MP#00](./mini00.html), it should already be ignored.) If it is appearing, you
may need to edit your `.gitignore` file.

Removing a large data file from `git` is possible, but difficult. Don't get
into a bad state!

:::

### Data Cleaning and Preparation

The Netflix "What We Watched" reports provide a high-level view of Netflix's
internal viewership data. While still interesting for our analysis, this data
has several limitations, you keep in mind throughout this Mini-Project. I note
some of them below in the description of each column. 

The `NETFLIX_DATA` from above has several columns: 

- `year`, `period`, `period_start`, and `period_end`. These reflect the period
  over which the viewership data was aggregated, *e.g.*, `2024H1` refers to the
  first half of 2024 (January 1, 2024 to June 30, 2024). I have created the
  `period_start` and `period_end` columns as dates which may be more useful for
  filtering and comparing against the published release dates. 
- `Title`. The title of the show or movie, roughly standardized.[^ascii] Note 
  that Netflix is a bit inconsistent in how it represents different seasons
  of TV shows; we have `Seinfeld: Season 1` and `Seinfeld: Season 3`, but
  `Stranger Things` and `Stranger Things 3`.
- `Runtime`. The run time in minutes of the show or movie (aggregated across all
   episodes for a show). It is not clear why certain catalog entries do not
   have a published run time (*e.g.*, Season 1 of Suits has no run time, but 
   later seasons do). Here, we can only work with what Netflix gives us.
- `Hours Viewed`. The global aggregate view time (in hours) during the period
   for that title. Note that this seems to be rounded to the nearest 100,000 
   hours and low viewership shows (those with less than 50,000 hours of view time)
   are omitted. This makes it hard to compare the popularity of two not-so
   popular shows, but we can still make some sense of major titles.
-  `Views`. The approximate number of views, again rounded. Note that these are
   "implied views" (computed by dividing Hours Viewed by Runtime) that assume
   all viewers watched a title in its entirety. This measure of `Views` cannot
   distinguish 600 people watching 10 minutes of a show and turning it off from
   one person watching the entirety of a 100 hour documentary.[^views]
- `Release Date`: The date on which a title was released. This is inconsistently
  available and should be used with caution. 
- `Available Globally?`: Was this title available in all Netflix territories 
  (typical for Netflix Originals) or only in certain territories (typical for
  content made available under license from a different creator).
- `live_event`: Was the title a live sporting event (a new area of focus for 
  Netflix)?
   
[^ascii]: To keep this analysis simple, I have restricted the data set to shows
with titles that are fully-ASCII (roughly, English language characters). We will
discuss the challenges associated with non-English characters later in the 
semester. For now, just be aware that your results are restricted to shows with
an English title, so if a show is popular in a country with a non-ASCII alphabet
(*e.g.*, Japan or Korea) but was never released with an English title, it may
be excluded from our analysis. 

[^views]: Netflix defines "Views" differently in other contexts. At one point,
a user was said to have "viewed" a show if they watched two or more minutes of
it. This, predictably, lead Netflix to debut the "most viewed show ever" with
surprising regularity. 


Before we can begin to look more closely at this data, we have a little bit more
clean-up we need to complete to get this data in its most usable form. 


One requirement for data to be easily manipulated is easy column names. Since we
use base column names in many `tidyverse` functions, including all of the `dplyr`
functions we have covered to this point, we want to ensure that our column names
are valid `R` variable names and easily typed. This means that we want: 

- Only numbers and simple letters, no punctuation or accents
- Starting with a letter
- All lower-case
- If multi-word, separated by underscores (`_`) not spaces

So a column name like `Flipper Length (mm)` will be hard to work with, while one
like `flipper_length_mm` will be fine.[^var_len] Compare the column names of
the `penguins` data set, which have been "tidied up" for easy use, to those of 
the `penguins_raw` which reflect the original data source. 

To get nice tidy names, we will use the `rename()` function from the `dplyr` 
package. The `rename()` function works similarly to the `mutate` function, but 
we can _only_ rename and cannot perform transformations. This might seem like
a limitation, but it makes the point of our code clearer. 

The `rename` function works like this: 

```{r}
#| eval: false
library(tidyverse)
iris |>
  rename(petal_length = Petal.Length, 
         petal_width  = Petal.Width, 
         ...)
```

Each argument is a pair of the form `NEW_NAME = OLD_NAME` and, by adding more
names in the `...`, we can perform even more renamings at once. If the old names
are sufficiently nasty that they are not even valid (but ugly) names, we need
to surround them with backticks (`), such as: 

```{r}
#| eval: false
library(tidyverse)
iris |>
  rename(petal_length = `Petal.Length`, # Note backticks around old names
         petal_width  = `Petal.Width`, 
         ...)
```

Anything placed between two backticks will be interpreted as a name. 

[^var_len]: Don't worry too much about the length of variable names: 
tab-completion is a wonderful thing and code is read more than written)


::: {.callout-tip title="Task 2: Fix Column Titles"}

Modify the code you copied in Task 1 to ensure that we have nice "tidy" column
names satisfying the four rules listed above. 

You can do this by either: 

i) Adding a `rename()` step to the end of the pipeline in my 
   `acquire_netflix_data()` function. (Recommended)
ii) Adding a `rename()` line after my `acquire_netflix_data()` function. 
    (If you do this, make sure to assign your modified data back into a 
    variable.)

:::

Additionally, we want to make sure our data is all reasonably typed. I have done
most of the clean-up work for you and only two columns need to be modified. 
Because you are actually changing the values from one type to another, we have
to actually use the `mutate()` function, and not `rename()`, but this idiom will
work well: 

```{r}
#| eval: false
data |>
  mutate(column = some_function(column))
```

Here, the new column is created by calling a function (`some_function`) but then
assigned back to the old name, having the effect of changing the column's values
in place. 

::: {.callout-tip title="Task 3: Fix Column Types"}

To proceed, we need to fix the `year` column and the `Available Globally?` 
column (which will have a slightly different name after Task 2). 

The `year` column is rather easy. Right now, it is _string_ values ("2023", 
"2024", *etc.*) but we want to convert these to numbers so that they can be used
as such. This is a simple "like-for-like" change that can be done with the
`as.integer()` (or `as.numeric()`) coercion function. 

The global availability column requires a bit more care At the end of my code,
it consists of the string values `"Yes"` and `"No"`, but we would really like
to use logical `TRUE` / `FALSE` values. Unfortunately, the `as.logical()` 
function does not do what we want here: 

```{r}
as.logical("Yes") # We want TRUE
```

Instead, we want to use a _vectorized_ check to see whether the value is
`"Yes"`, in which case we will convert to `TRUE`, or something else that will
be converted to `FALSE`. 

Complete the following code snippet to perform these two conversions: 

```{r}
#| eval: false
|> mutate(
  ... = as.integer(...), 
  ... = ... == ... # Recall that double equals (==) checks for equality
)
```

As in the previous task, either insert your code into the end of my 
`acquire_netflix_data()` function or call it on the output of my code.

:::

### Initial Data Exploration {#eda}

Before moving to our final analysis, we will do a bit of 
_Exploratory Data Analysis_ (EDA). EDA serves many purposes in data science-- 
quality control, hypothesis generation, outlier identification, *etc.*--but 
perhaps the most important is simply knowing what information can be found in a 
novel data set. Now that our data is imported and cleaned, it's almost 
time to start our EDA. Before we do EDA, however, we should pause briefly to
think about how we want to display data. 

#### Displaying Data in Tables

While we could continue investigating our data using `R`'s basic print-outs, 
this is a good time to introduce the [`gt` package](https://gt.rstudio.com/) 
which can be used to create complex tables natively in `R`.

What follows here is a very brief introduction to the `gt` package. **You 
do not need to copy this into your submission - it is provided only as 
background.** You will use the `gt` package to format your answers to the 
next few tasks. Click on the blue bar below to read the introduction to `gt`.

::: {.callout-note collapse="true" title="Making Attractive Tables in R"}

For this introduction, I am going to use the `penguins` data, but the same
functions can be applied to any data that you are analyzing. 


Let's first look at what a "basic" or "raw" display of a data frame gives us: 

```{r}
penguins
```

This has *several* problems: 

1. We are showing *way* too much data. A reader will not be able to easily find
   meaningful trends or patterns in a big data set like this. 
   
   As a general rule, you should rarely have more than 10-15 rows in a table;
   even then, you will still want to guide your reader to the point of the table.
2. The column names are rather ugly. Some, like `species` are not too bad, though
   it would still be better if they were capitalized. Others, like `bill_len`
   are pretty terrible: `bill_len` is not an English word, the underscore
   exists only to separate two words "in code" (recall `R`'s restrictions on 
   variable names), and the unit isn't clear. In this case, a column name like
   `Bill Length (mm)` would be far preferable.[^names]
   
[^names]: I know this is the opposite of the advice given about names earlier. 
The "best practices" for code are not the "best practices" for final 
presentations.
   
3. The row numbers are essentially pointless and just take up space, adding no
   value. Any content that is not adding value is simply distracting the reader
   from the content that has value. 
   
4. The "point" of the table is unclear. What is a reader supposed to get from
   this? As a data analyst - doing work on behalf of a reader who may not be a
   data analyst - you have a responsibility to clearly convey the "story" of
   your findings and this does not do so. 
   
   I may want to use this data to show that Gentoo penguins are, on average, 
   heavier than the other two species in this data set, but this is far from 
   clear. 
5. It's just a bit ugly. 

Good table design requires us to take on the mindset of the reader. Tools like
`gt` can help pretty things up, but you still have to think about *what* you 
want to display. Well-formatted garbage is still garbage. 

To start improving this table, let's do the calculations for our reader instead
of expecting them to do it all manually: 

```{r}
library(tidyverse)
penguins |> 
    group_by(species) |> 
    summarize(n_species = n(), 
              avg_body_mass = mean(body_mass, na.rm=TRUE)) |>
    arrange(desc(avg_body_mass))
```

We're definitely not done - but here the "point" of the table is clear, at least
if we also put some text surround it. 

To improve this further, we can also pass this smaller summary data frame to
the `gt` function from the package of the same name: 

```{r}
#| warning: false
library(gt)
penguins |> 
    group_by(species) |> 
    summarize(n_species = n(), 
              avg_body_mass = mean(body_mass, na.rm=TRUE)) |>
    arrange(desc(avg_body_mass)) |>
    gt()
```

Note here that `gt` recognizes we are rendering an HTML page and produces a 
"real" HTML table here. If you were to copy and paste the table above into 
"table" software, *e.g.* Google Sheets or Microsoft Excel, it would be properly 
and automatically handled. For us, the table is the end-point, but it's a nice 
courtesy to your reader who may want to use your results in their own 
presentations. 

The `gt` package provides *many* functions for tweaking and improving the 
appearance of a table. You will almost always want to, at a minimum, use these 
for: 

- Ordering and (re-)naming columns
- Adding titles and footers
- Formatting values

Let's to through these one at a time. Firstly, we want to rename and reorder
the columns. This can be done in pure `dplyr` with the `select` and `rename`
columns, but we'll show the `gt` way here: 

```{r}
library(gt)
penguins |> 
    group_by(species) |> 
    summarize(n_species = n(), 
              avg_body_mass = mean(body_mass, na.rm=TRUE)) |>
    arrange(desc(avg_body_mass)) |>
    gt() |>
    cols_move_to_end(n_species) |>
    cols_label(species="Species", 
               avg_body_mass = "Avg. Body Mass (g)", 
               n_species = "Number of Penguins in Sample") 
```

Here, we used the `cols_move_to_end()` function to move the `n_species` column
to the end (no surprise!). In other contexts, we might want to use the 
`cols_move_to_start()` function to move a column to the leftmost side of a table
or `cols_move()` to put a column in the middle of the table. 

The `cols_label()` function essentially serves as a renaming operation: the 
left side of each parentheses is the old column name in the table and the right 
side gives the new name. (Note, a bit confusingly, that this is the *reverse*
of `dplyr::rename()`.) While we can just pass a basic string here, we can also
use the `md()` function to pass _Markdown_ which lets us do some custom formatting: 

```{r}
penguins |> 
    group_by(species) |> 
    summarize(n_species = n(), 
              avg_body_mass = mean(body_mass, na.rm=TRUE)) |>
    arrange(desc(avg_body_mass)) |>
    gt() |>
    cols_move_to_end(n_species) |>
    cols_label(species=md("**Species**"), 
               avg_body_mass = md("Avg. Body Mass (*g*)"), 
               n_species = md("*Number of Penguins in Sample*"))
```

Here, we could use **boldface** and *italics* for certain text using standard
Markdown syntax. 

Next, we can add a table title and subtitle to make the content and point
of this table clear to our reader: 

```{r}
penguins |> 
    group_by(species) |> 
    summarize(n_species = n(), 
              avg_body_mass = mean(body_mass, na.rm=TRUE)) |>
    arrange(desc(avg_body_mass)) |>
    gt() |>
    cols_move_to_end(n_species) |>
    cols_label(species=md("**Species**"), 
               avg_body_mass = md("Avg. Body Mass (*g*)"), 
               n_species = md("*Number of Penguins in Sample*")) |>
    tab_header(title="Average Body Mass of Three Penguin Species", 
               subtitle="Gentoo penguins are the largest in the study")
```

While we can and should describe our analysis in more detail in the main text,
I like this pattern of having the super-simple one-liner present directly in
the table. This also makes it convenient to clip the table (or a screenshot
thereof) for use in other documents and presentations. 

Next, we should always note the source of the data used to get our results. 
In this case, the original penguins data comes from 
[this article](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0090081) 
so we can cite that in our work Note the use of Markdown (`md()`) to let us include 
a link to the original source within our table: 

```{r}
penguins |> 
    group_by(species) |> 
    summarize(n_species = n(), 
              avg_body_mass = mean(body_mass, na.rm=TRUE)) |>
    arrange(desc(avg_body_mass)) |>
    gt() |>
    cols_move_to_end(n_species) |>
    cols_label(species=md("**Species**"), 
               avg_body_mass = md("Avg. Body Mass (*g*)"), 
               n_species = md("*Number of Penguins in Sample*")) |>
    tab_header(title="Average Body Mass of Three Penguin Species", 
               subtitle="Gentoo penguins are the largest in the study") |>
    tab_source_note(md("Data originally published in by K.B. Gorman, T.D. Williams, 
                       and W. R. Fraser in 'Ecological Sexual Dimorphism and 
                       Environmental Variability within a Community of Antarctic 
                       Penguins (Genus *Pyogscelis*). *PLoS One* 9(3): e90081.
                       <https://doi.org/10.1371/journal.pone.0090081>. Later 
                       popularized via the `R` package
                       [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/)"))
```

A small, but lovely, quality of life feature here is the fact that `md()` will
automatically re-align the text to the fit the dimensions of the rendered
table. This lets us put new lines within our citation text so that our code 
doesn't exceed the 80 characters-per-line guideline. 

Finally, we want to make sure the number values are formatted appropriately. 
For these numbers, plain formatting really isn't much of a problem, but for
very large or small numbers, we might want to use scientific notation; for dates,
we might want to control the formatting; *etc.*. This is done with the `fmt_*`
family of functions. 

Each `fmt_` function takes one or more column names and applies a formatting
transformation to that column. The specifics of the formatting can be controlled
with additional optional arguments. For example, if we want to round the average
weight to the nearest gram, we would use the `fmt_number()` function with the
argument `deicmals=0`:

```{r}
penguins |> 
    group_by(species) |> 
    summarize(n_species = n(), 
              avg_body_mass = mean(body_mass, na.rm=TRUE)) |>
    arrange(desc(avg_body_mass)) |>
    gt() |>
    cols_move_to_end(n_species) |>
    cols_label(species=md("**Species**"), 
               avg_body_mass = md("Avg. Body Mass (*g*)"), 
               n_species = md("*Number of Penguins in Sample*")) |>
    tab_header(title="Average Body Mass of Three Penguin Species", 
               subtitle="Gentoo penguins are the largest in the study") |>
    tab_source_note(md("Data originally published in by K.B. Gorman, T.D. Williams, 
                       and W. R. Fraser in 'Ecological Sexual Dimorphism and 
                       Environmental Variability within a Community of Antarctic 
                       Penguins (Genus *Pyogscelis*). *PLoS One* 9(3): e90081.
                       <https://doi.org/10.1371/journal.pone.0090081>. Later 
                       popularized via the `R` package
                       [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/)")) |>
    fmt_number(avg_body_mass, decimals=0)
```

In this case, it may be more appropriate to display the body mass in *kilograms*
and we can do so semi-automatically with the `format_number_si()` formatter: 

```{r}
penguins |> 
    group_by(species) |> 
    summarize(n_species = n(), 
              avg_body_mass = mean(body_mass, na.rm=TRUE)) |>
    arrange(desc(avg_body_mass)) |>
    gt() |>
    cols_move_to_end(n_species) |>
    cols_label(species=md("**Species**"), 
               avg_body_mass = md("Avg. Body Mass "), 
               n_species = md("*Number of Penguins in Sample*")) |>
    tab_header(title="Average Body Mass of Three Penguin Species", 
               subtitle="Gentoo penguins are the largest in the study") |>
    tab_source_note(md("Data originally published in by K.B. Gorman, T.D. Williams, 
                       and W. R. Fraser in 'Ecological Sexual Dimorphism and 
                       Environmental Variability within a Community of Antarctic 
                       Penguins (Genus *Pyogscelis*). *PLoS One* 9(3): e90081.
                       <https://doi.org/10.1371/journal.pone.0090081>. Later 
                       popularized via the `R` package
                       [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/)")) |>
    fmt_number_si(avg_body_mass, 
                  decimals=2, 
                  unit = "g")
```

Note that, because `fmt_number_si()` automatically includes the unit and transforms
it to the most natural scale (here `kg`), we can remove the unit from the column
name. `gt` has many more advanced options that can be used for further 
customization; refer to the package documentation for more details or ask on the
course discussion board.

:::

#### Exploratory Analysis

When faced with a new data set, it is tempting to look only at the first few
rows to get a sense of the data: `R` does this by default. In practice, I 
recommend viewing a _random_ selection of rows instead. This won't guarantee you
find any issues, but it increases the probability of finding issues in 
older parts of a data set. The `slice_sample()` function can be used for this. 

While EDA can be an extensive activity on its own, at an *absolute minimum*,
I recommend you always do at least two basic checks: 

- Ensure that you know what `R` *thinks* your data is. You might see a value like
  `"2025-12-01"` and think that `R` is reading it as a date value, but `R`
  might instead interpret it as a string value.[^excel] Then, when you try to do
  "date things" you will run into trouble. To avoid this, you often need to 
  explicitly convert object types in `R`, but knowing when to convert requires
  you to first know what type `R` thinks something is. You have several
  options for determining object type in `R`, but when working with a data 
  frame, I'd recommend the `glimpse()` function from the `dplyr` package: *e.g.*
  
  ```{r}
  glimpse(penguins)
  ```
  
  After getting the basic dimensions of this data frame, `glimpse()` will print
  a line summary of each column giving its name, type, and the first few values
  in the table. 
  
  In this case, since I did most of the data prep for you, all columns should be
  of the correct type after you complete Task 3, but this is
  a quick and easy check. If there are issues with your data types, it's better
  to catch them early than to have silent and hard to identify errors further
  down the line. ("Fail fast" is *great* advice in any programming exercise.)
  
- Take a quick look at some basic (univariate) summary statistics for each
  column. There are several functions for this in base `R`: *e.g.*, 
  
  ```{r}
  summary(penguins)
  ```
  
  But I actually prefer the `skim()` function from the `skimr` package as it
  provides a bit more type-specific information and has more handsome formatting:
  
  ```{r}
  #| warning: false
  library(skimr)
  skim(penguins)
  ```
  
  You'll see that this gives some nice overview about data structure and types,
  identifies grouping structure (if any is present), and gives summaries
  appropriate to the type of each column (here `factors`, *i.e.*, categorical 
  variables and `numeric`). I like that this summary gives means, standard 
  deviations, minima (`p0`), maxima (`p100`), medians (`p50`), and a cute little
  histogram of each variable. 
  
  Note that this type of summary only reveals *univariate* structure: if there
  are interesting multi-dimensional outliers or weird correlation patterns, they
  won't appear here. 
  
[^excel]: In one famous (and slightly tragic) example, Microsoft Excel silently
misinterpreted the names of various genes as numeric values and changed them
from (what it thought was) scientific notation to (what it thought was) standard
numeric formatting. This wound up ruining several important scientific studies. 
Always check your data types! The original study identifying this problem can be
found [here](https://link.springer.com/article/10.1186/s13059-016-1044-7) and
a popular news summary is 
[here](https://qz.com/768334/years-of-genomics-research-is-riddled-with-errors-thanks-to-a-bunch-of-botched-excel-spreadsheets). 

::: {.callout-tip title="Task 4: Initial EDA Pass"}

Perform the two checks described above on the Netflix data you 
loaded and processed above. **Include the code to perform the checks, 
the output of the checks, and describe what you see.**

In general, you shouldn't include this in analysis reports, but we're making
an exception here to practice a good habit. I would advise *against* including 
code in your *future* mini-project or course project reports, but these steps
are easy and very valuable, so I recommend you make them part of your standard
EDA workflow. (Remember, not everything you do as part of EDA needs to be 
included in your final submission, especially outside of a classroom setting. 
Make sure you are maximizing signal-to-noise for your intended readers.)

:::

We are finally ready to begin some EDA. Analysts organize their EDA in a variety
of ways, but one of my favorites is to think of a variety of interesting 
questions relevant to the data set and to attempt to answer them. When the 
answers don't match my intuition, I know I've found somewhere I want to dig 
deeper. For the first few mini-projects, I will provide these exploratory 
questions. Later in the course, particularly after we have discussed the role of
plotting and graphics in EDA, you will have opportunities to organize your EDA 
in other fashions. 

::: {.callout-tip title="Task 5: Exploratory Questions - Inline Values"}

Using `dplyr` tools, answer the following six questions. 
Each of these questions can be answered with one or two scalar values. Use 
Quarto's [inline code](https://quarto.org/docs/computations/inline-code.html)
functionality to place the values in a sentence; that is, you should answer in
complete sentences, written as normal text with inline code for computed values.

1.  What date range is covered by this data set? 
2.  How many different shows and movies are covered in this data set? 

    (Don't worry about combining seasons for now: you can treat "Stranger Things
    2" and "Stranger Things 3" as different shows.)
3.  How many total hours have been spent watching shows from the *Bridgerton* 
    universe? 
   
    *Hint:* The `str_detect()` function can be used to test whether a string 
    (like a name) contains a substring:[^re] *e.g.*, 

    ```{r}
# True because "CUNY" (2nd arg) is in the longer string (1st arg)
str_detect("CUNY Bernard M. Baruch College", "CUNY")

# False because "Hunter" is nowhere in the first argument
str_detect("CUNY Bernard M. Baruch College", "Hunter")
    ```

    As with other functions in `R`, `str_detect()` is vectorized, making it easy
    to use inside of other functions, such as `filter()`.

    ```{r}
    names <- c("City College-Miami", "CUNY City College", "CUNY Hunter College")
    str_detect(names, "CUNY")
    ```

[^re]: `str_detect()` can actually be used to perform significantly more complex
string analysis than simple "does it contain this subset of letters" but we
won't cover that sort of string processing for a [few more weeks](../pre/pa10.qmd). 

4.  How many total hours have been spent watching the five seasons of 
    *Stranger Things*? For this question, we only want the "real" *Stranger Things*
    shows and not the companion behind-the-scenes show ("Beyond Stranger Things"), 
   the animated spin-off ("Tales from '85"), or the background music 
    ("Stranger Things: Fireplace"). 
   
    To answer this question, you will need to use a combination of `filter()`
    and `filter_out()` statements to sequentially apply multiple `str_detect()`
    steps. 
   
    *Hint*: Recall that you generally shouldn't combine multiple steps in a single
    `filter_out()` as the results can be a bit unintuitive.
5.  What percentage of total watch time was spent on the Top 10 titles of 2026H1?
6.  What fraction of Netflix's catalog is globally available? (Answer this both
    by title and weighted by hours viewed)

As you go through these questions, you may find it useful to create new 
variables in the data to avoid repeated lengthy calculations, *e.g.*, 
`is_stranger_things`. Do this is you find it useful, but it is not strictly
required. 

Where appropriate, the `scales` package can be used to format numbers 
attractively: *e.g.*, 

```{r}
#| message: false
library(scales)
comma(12345.6789)
comma(12345.6789, accuracy=0.01)
dollar(12345.6789)
percent(0.1234)
percent(0.1234, accuracy=0.01)
```

:::

The next set of questions have slightly more complex answers and should be
answered with a table, formatted using the techniques described above. Longer
term, you will prefer plots to tables as they are a bit easier to interpret
(humans being visual creatures) but these are the types of questions you might
use to create plots as well. 

::: {.callout-tip title="Task 6: Exploratory Questions - Table Answers"}

Using `dplyr` tools, answer the following questions. Each of these questions can
be answered with a table of just a few rows. Use the `gt` package, as introduced
above, to present your results in an attractive 'publication-quality' format, 
not just a "raw" `R` output.

1. What was the top show (by total hours viewed) in each period and what 
   fraction of total viewership did it capture? 


2.  How has Netflix's catalog changed in size over the periods for which we 
    have questions? You should consider both the number of titles and the 
    total runtime of content. 

3.  Which shows have had the most consistently high viewership across all 
    periods? These shows are likely to be 'staples' that viewers fall back
    on as 'comfort food', *e.g.*, re-runs of *Friends*. 
    
    *Hint*: To answer this question, you might want to compute the average
    veiwership, the minimum viewership, and the standard deviation of viewership
    across different periods and develop a filter using some combination of
    these quantities.

4. Within each period, what fraction of time was spent on 'shows' vs movies?
   
   Unfortunately our data does not clearly delineate shows from movies, so you
   will need to come up with a heuristic for doing so. I'd recommend some 
   combination of: 
   
   - Length: anything over 4-ish hours is almost certainly a TV show, not a 
     movie
   - "Season": if a title has "Season" in it, it is almost certainly a TV show
   
   You might think of other filters or thresholds that make sense and can
   incorporate them into your analysis. 
   
   This is a problem that commonly arises in real data analysis: the data is
   _not quite_ what we might want to answer a question, but we could definitely
   construct the feature manually if we were willing to put in the effort. In
   this case, we have to come up with an _approximate_ way of implementing our
   feature that (roughly) captures our human intuition.[^llm]
   
   [^llm]: This is an interesting place where LLMs can _augment_ data analysts.
   It would be pretty easy - but a bit beyond this assignment - to query your
   favorite LLM to classify these titles as TV shows or movies. See
   <https://tidydesign.substack.com/i/214337800/structured-data> for a 
   discussion of these ideas. 

5. How does the performance of live events compare to pre-taped content? Does
   the average live event draw more viewership than an average piece of 
   pre-taped content? How about a 'top tier' live event vs 'top tier' pre-taped
   content?
   
There are **many ways** you might choose to approach these questions, and there
isn't a single correct answer. Make sure to _clearly explain_ any analytical 
choices you make. 
    

:::


### Final Deliverable: Platform Strategy Memo

At this point, you have acquired your data, cleaned and prepared it, and
performed your EDA. Now, you are ready to get to work on the final deliverable
of your analysis. Everything that comes before this is important, but typically
less visible to your final customer. 

::: {.callout-tip title="Task 7: Corporate Strategy Memo"}

Write a brief (no more than 750 words) _corporate strategy memo_ addressed to
Netflix leadership. Your memo should help inform Netflix's original content
strategy for 2027 and beyond. 

You can use the calculations performed above - or something similar - to include
specific _numerical_ evidence in support of your recommendations. For example,
you should not say "live events are an increasingly large share of viewership" and should instead prefer a more specific claim like "live events have risen from X%
to Y% of total watch time."

Your memo should touch on **at least 2** of the following possible topics: 

- Globally available *vs.* regionally licensed content
- TV shows *vs.* movies
- 'Big hits' *vs.* the long-tail of smaller titles
- Shorter 'bite sized' *vs.* longer content 
- 'Burstiness' of viewership: do titles have a _rush_ of viewership immediately 
  after release (to partially answer this, you can use the period end and release
  date columns to get a "days available" value; combine this with a focus on
  titles that release near a period end)
- Trends over time: identify time changes in the data, don't simply aggregate
  2023 to the present. 
  
You may also address other structure or trends in the data you find interesting.

:::

You may include additional tables or even visualizations if you want, 
but these do not replace the requirement to write the core strategy memo. Your
memo should stand "alone" and not be mixed in with your code. Place the code
necessary to perform the relevant supporting calculations in a separate section
and use inline code chunks to include results of your analysis in the text of 
your strategy memo. Deliverables that hard-code calculated values will be
penalized. 

{{< include _aidisclosure.qmd >}}


{{< include _echeader.qmd >}}

### Gini Analysis (Up to 2 Points)

When analyzing data with a 'long tail', it is often interesting to look at
_inequality_ in the data. We know that viewership data will be 'top-heavy' and
that some shows will be major hits with meaningful cultural footprints, while
others will be watched by only a small number of viewers. To wit, 

{{< video https://www.youtube.com/watch?v=lqRQ5Y6OYi4 >}}

To put some numbers of this phenomenon, we can adapt the idea of a 
[Gini coefficient](https://en.wikipedia.org/wiki/Gini_coefficient) from
economics. Wikipedia gives several useful formulas for the Gini coefficient. 
For example, suppose a country has only 5 citizens, and their incomes are as
follows: 

| Citizen | Income |
|---------|--------|
| Alice   |  $20K  |
| Barbara | $200K  |
| Carol   |   $0K  |
| Debbie  |  $50K  |
| Emily   |  $40K  |

The Gini coefficient of this distribution can be computed as: 

$$G = \frac{2}{n} \frac{\sum_{i=1}^n i y_{(i)}}{\sum_{i=1}^n y_{(i)}} - \frac{n+1}{n}$$

where $y_{(i)}$ are the _ordered_ incomes ($y_{(1)} = 0 \leq y_{(2)} = 20 \leq y_{(3)} = 40 \leq y_{(4)} = 50 \leq y_{(5)} = 200$). This gives a Gini 
coefficient of around 55.5% for this country.  (Make sure you can replicate
this number.)

Up to two points of extra credit will be given to submissions that include
a Gini-coefficient analysis in their strategy memo. Submissions must (briefly)
define the Gini coefficient in non-technical language. When computing the 
Gini coefficient, make sure to consider whether it makes sense to look at
raw view time or some sort of length-normalized viewership metric. 

### Data Visualization (2 point)

Inclusion of a well-formatted visual element to accompany your memo may earn up
to 2 points of extra credit. 

{{< include _footer.qmd >}}
