---
mp_num: "01"
mp_title: "Assessing the Impact of SFFA on Campus Diversity One-Year Later"
mp_skills: "Data Summarization, Exploratory Analysis, Table Formatting"
mp_application: "Official Statistics, Demography"
mp_rhetoric: "Op-Ed Writing"
mp_max_ec: 6
---

{{< include _mpsetup.qmd >}}

## Welcome to {{< var course.short >}} Mini Projects!

In the {{< var course.short >}} Mini-Projects, you will perform basic data
analyses intended to model best practices for your course final project.
(Note, however, that these are *mini*-projects; your final course project is
expected to be far more extensive than any single MP.)

## Introduction

On June 29, 2023, The US Supreme Court handed down its decision in the closely
watched case of [*Students for Fair Admissions v. 
Harvard*](https://en.wikipedia.org/wiki/Students_for_Fair_Admissions_v._Harvard)
("SFFA"). In *SFFA*, the court found that the admissions programs at Harvard
and at the University of North Carolina violated the Equal Protection Clause
of the Fourteenth Amendment to the US Constitution and engaged in impermissible
race-conscious admissions practices. While the court had explicitly disallowed
race-conscious practices in other contexts, Harvard argued that the admissions
process served a compelling governmental interest - educational benefits from a 
diverse student body - that had been recognized by the court as [recently as 
2016](https://en.wikipedia.org/wiki/Fisher_v._University_of_Texas_(2016)). 

While the *SFFA* case touched on many aspects of law, one important factual
question was the practical impact of the programs in question: were they 
essential to maintaining a diverse student body or were they merely a slight
"thumb on the scale" that had only a minor impact?

Now that the first set of post-*SFFA* admissions data has been released,[^year] 
you are going to attempt to analyze the extent to which post-*SFFA* admissions 
differ from pre-*SFFA* practice. You will use admissions data distributed via 
*IPEDS*, the Integrated Postsecondary Education Data System, managed by the 
National Center for Education Statistics within the Federal Department of 
Education. 

After some initial analysis, you will write a brief (750) word 
[Op-Ed](https://en.wikipedia.org/wiki/Op-ed) from the perspective of a college
president describing the impact of *SFFA* on the demographics of admitted 
students at your college and at colleges across the country. 

In this mini-project, you will: 

1) Practice use of `dplyr` for analysis of tabular data
2) Practice use of `quarto` and Reproducible Research Tools for Effective
   Communication of Data Analysis Results
3) Begin your professional data science portfolio. 

[^year]: Following the decision in June 2023, universities were first applying
post-*SFFA* admissions policies for the undergraduate class entering in Fall 
2024. Due to reporting lags, these data were released in late 2025 and are 
currently the latest available. As data from more admissions cycles are
released in IPEDS, the impact (or not) of *SFFA* will become clearer. 

### Student Responsibilities

For purposes of MPs, we are dividing the basic data analytic workflow into
several major stages:

{{< include _responsibilities.qmd >}}

In early stages of the course, such as this MP, I will 'scaffold' much of the
analysis for you, leaving only those stages we have discussed in class for you
to fill in. As the course progresses, the mini-projects will be more
self-directed and results less standardized.

### Rubric

{{< var course.short >}} Mini-Projects are evaluated using *peer grading* with 
*meta-review* by the course staff. The following basic rubric will be used for 
all mini-projects:

{{< include _rubric.qmd >}}

At this early point, you are not responsible for all elements of this rubric. 
In particular, all submissions will receive an automatic 10/10 for Data 
Visualization as this is outside the scope of this mini-project. Furthermore, 
because I am providing code to download the data, load it into
`R`, and prepare it for analysis, all reports submitted using my code will 
receive an automatic 10/10 for the 'Data Preparation' element of the rubric. 
Finally, reports completing all tasks described under [Data Integration and
Exploration](#eda) below should receive a 10/10 for the 'Exploratory Data Analysis'
rubric element. 

Taken together, you are only really responsible for these portions of the
rubric: 

- Written Communication
- Project Skeleton
- Tables & Document Presentation
- Code Quality
- Analysis and Findings

Reports completing all key steps outlined below essentially start with
30 free points. 


::: {.callout-warning title="Writing Requirements"}

Note that you are evaluated*on writing and communication in these 
Mini-Projects. You are required to write a report in the prescribed style, 
culminating in an Op-Ed. A submission that performs the instructor-specified 
tasks, but does not write and give appropriate context and commentary will 
score very poorly  on the relevant rubric elements. 

In particular, if a submission does not include a clearly delineated Op-Ed and
only answers the instructor prompts in narrative text, peer evaluators should
judge  it to have "Good" quality Written Communication (at best) as key findings
are  not conveyed appropriately. 

Quarto's [code 
folding](https://quarto.org/docs/output-formats/html-code.html#folding-code) 
functionality is useful for "hiding" code so that it doesn't break the flow of
your writing. 

You can also make use of [Quarto's `contents`
shortcode](https://quarto.org/docs/authoring/contents.html) to present
code and findings in an order other than how the code should be executed.
This is particularly useful if you want to include a figure or table in an 
"Executive Summary" at the top of your submission.

:::

{{< include _submission.qmd >}}

## Mini-Project #{{< meta mp_num >}}: {{< meta mp_title >}}

### Data Acquisition

The following code can be used to acquire data from
[IPEDS](https://nces.ed.gov/ipeds/datacenter/DataFiles.aspx). Specifically, 
once run, this code will download the fall enrollment ("EF") and institution
description ("HD") files for the previous 15 years. To be efficient, this
function will save a copy of the downloaded data in a folder called
`data/mp01` and use that copy to avoid re-downloading a file if it is already
present on your computer. This will make your code faster to run and will avoid
putting unnecessary stress on the IPEDS servers. 

```{r}
#| message: false
#' Acquire IPEDS Data for MP#01
#' 
#' This function will acquire and standardize all data for MP#01
#' from IPEDS (https://nces.ed.gov/ipeds/datacenter/DataFiles.aspx)
#' 
#' We're starting in 2010 as the data seems to be reasonably complete 
#' after that point. 
acquire_ipeds_data <- function(start_year=2010, end_year=2024){
    library(tidyverse)
    library(glue)
    
    data_dir <- file.path("data", "mp01")
    
    if(!dir.exists(data_dir)){
        dir.create(data_dir, showWarnings=FALSE, recursive=TRUE)
    }
    
    YEARS <- seq(start_year, end_year)
    
    EFA_ALL <- map(YEARS, function(yy){
        if(yy <= 2022){
            ef_url <- glue("https://nces.ed.gov/ipeds/datacenter/data/EF{yy}A.zip")
            
        } else {
            ef_url <- glue("https://nces.ed.gov/ipeds/data-generator?year={yy}&tableName=EF{yy}A&HasRV=0&type=csv")
        }
        
        ef_file <- file.path(data_dir, glue("ef{yy}a.csv.zip"))
        
        if(!file.exists(ef_file)){
            message(glue("Downloading Enrollment Data for {yy} from {ef_url}"))
            download.file(ef_url, destfile = ef_file, quiet=TRUE)    
        }
        
        read_csv(ef_file, 
                 show_col_types=FALSE) |>
            mutate(year = yy, 
                   # American Indian or Alaskan Native
                   enrollment_m_aian = EFAIANM, 
                   enrollment_f_aian = EFAIANW, 
                   # Asian
                   enrollment_m_asia = EFASIAM, 
                   enrollment_f_asia = EFASIAW, 
                   # Black or African-American, 
                   enrollment_m_bkaa = EFBKAAM, 
                   enrollment_f_bkaa = EFBKAAW, 
                   # Hispanic 
                   enrollment_m_hisp = EFHISPM, 
                   enrollment_f_hisp = EFHISPW, 
                   # Native Hawaiian or Other Pacific Islander 
                   enrollment_m_nhpi = EFNHPIM, 
                   enrollment_f_nhpi = EFNHPIW, 
                   # White
                   enrollment_m_whit = EFWHITM, 
                   enrollment_f_whit = EFWHITW, 
                   # Two or More Races
                   enrollment_m_2mor = EF2MORM, 
                   enrollment_f_2mor = EF2MORW, 
                   # Unknown / Undisclosed Race
                   enrollment_m_unkn = EFUNKNM, 
                   enrollment_f_unkn = EFUNKNW, 
                   # US Non-Resident
                   enrollment_m_nral = EFNRALM, 
                   enrollment_f_nral = EFNRALW, 
            ) |> filter(
                (EFALEVEL %in% c(2, 12)) | (LINE %in% c(1, 15))
                # Per 2024 Data Dictionary, 
                # - EFALEVEL 2 = undergrad 
                # - EFALELVE 12 = grad
                # - Line 1 = first year first time full-time undergrad
                # - Line 15 = first year first time part-time undergrad
            ) |> mutate(level = case_when(
                       EFALEVEL == 2 ~ "all undergrad", 
                       EFALEVEL == 12 ~ "all graduate",
                       LINE %in% c(1, 15) ~ "first year undergrad"
                   )
            ) |>
            select(institution_id = UNITID, 
                   year, 
                   level,
                   starts_with("enrollment_")) |>
            group_by(institution_id, 
                     year, 
                     level) |>
            summarize(across(starts_with("enrollment_"), sum), 
                      .groups = "drop")
        
    }) |> bind_rows()
    
    DESC_ALL <- map(YEARS, function(yy){
        if(yy <= 2022){
            hd_url <- glue("https://nces.ed.gov/ipeds/datacenter/data/HD{yy}.zip")
            
        } else {
            hd_url <- glue("https://nces.ed.gov/ipeds/data-generator?year={yy}&tableName=HD{yy}&HasRV=0&type=csv")
        }
        
        hd_file <- file.path(data_dir, glue("hd{yy}.csv.zip"))
        
        if(!file.exists(hd_file)){
            message(glue("Downloading Institutional Descriptions for {yy} from {hd_url}"))
            download.file(hd_url, destfile = hd_file, quiet=TRUE)    
        }
        
        suppressWarnings(
            read_csv(hd_file, 
                 show_col_types=FALSE, 
                 locale=locale(encoding=if_else(yy==2024, "utf-8", "windows-1252"))) |>
            mutate(year = yy, 
                   INSTNM) |> 
            select(institution_id = UNITID, 
                   institution_name = INSTNM, 
                   state = STABBR, 
                   year)
        )
        
    }) |> bind_rows()
    
    inner_join(EFA_ALL, 
               DESC_ALL, 
               join_by(institution_id == institution_id, 
                       year == year))
}

IPEDS <- acquire_ipeds_data()
```

This creates a data frame called `IPEDS` that has `r NROW(IPEDS)` rows and
`r NCOL(IPEDS)` columns in your local environment. This data will be used for
the remainder of this mini-project. 

::: {.callout-tip title="Task 1: Data Acquisition"}

Using the code above, acquire the IPEDS data. Copy the code into
your Quarto document and make sure it runs successfully. 

:::

::: {.callout-caution title="Do Not `git add` Data Files"}

Make sure that `git` is set to ignore data files, such as the one created above.
Check the `git` pane in `RStudio` and make sure that the files in `data/mp01`
do not appear. (If you set up your `.gitignore` file correctly in
[MP#00](./mini00.html), it should already be ignored.) If it is appearing, you
may need to edit your `.gitignore` file.

Removing a large data file from `git` is possible, but difficult. Don't get
into a bad state!

:::

### Data Cleaning and Preparation

IPEDS provides *many* different pieces of information, so I have provided you
a subset of interesting variables. These include: 

- A unique institutional ID code (`institution_id`)
- The name of the institution (`insitution_name`)
- The state in which the institution's principal campus is located
- The year of reporting. Note that, for enrollment data, these are the enrollment
  as of the *Fall* semester, so, *e.g.*, Year 2010 is for the first semester 
  of the 2010-2011 academic year. 
- The "level" of students being considered: 
  - `"all graduate"` refers to all graduate students on campus (both full-time
     and part-time)
  - `"all undergraduate"` refers to all undergraduate students on campus (both
    full-time and part-time)
  - `"first year undergrad"` refers to "stereotypical" first year college 
    students. These are students (either full-time or part-time) who are in 
    their first year of study during their first undergraduate enrollment (*i.e.*,
    excluding students who took time off and re-enrolled, students who 
    transferred from another institution, and students who previously studied at
    a different institution earning an associates' degree or similar credential)
- Number of enrolled students in various demographic groups: these 18 variables
  are formatted as: `enrollment_X_YYYY`, where
  
  - `X` is `m` or `f` for male or female students
  - `YYYY` is a racial/ethnic group identifier:
    - `aian`: American Indian or Alaskan Native
    - `asia`: Asian-American
    - `bkaa`: Black or African-American
    - `hisp`: Hispanic
    - `nhpi`: Native Hawaiian or Pacific Islander
    - `whit`: White
    - `2mor`: Two or More Races
    - `unkn`: Unknown / Unreported
    - `nral`: Not a US Resident before enrollment ("Non-Resident Alien")
    
  Note that these are somewhat non-standard identifiers and do not match other
  official sources. (*E.g.*, the US Census considers Hispanic ancestry a 
  separate axis so it is possible to have Non-Hispanic White, Hispanic Black, 
  *etc.*) We use these as given since they are what IPEDS provides. 
  
  Together these give *18* enrollment categories that, when summed, give the
  total enrollment at an institution. 
  
Before proceeding, we will create two additional variables for our analyses: 

- `is_cuny`: A Boolean (True/False) variable indicating whether an institution
  is part of the CUNY system. 
- `is_calpublic`: A variable indicating whether an institution is a public
  college or university of the state of California
  
::: {.callout-tip title="Task 2: Identify CUNY Schools"}

Create a new column `is_cuny` in your IPEDS data. For our
purposes, we will assume that CUNY schools all have "CUNY" in
their institution name. 

The `str_detect` function can be used to test whether a string (like a name)
contains a substring:[^re] *e.g.*, 

```{r}
# True because "CUNY" (2nd arg) is in the longer string (1st arg)
str_detect("CUNY Bernard M. Baruch College", "CUNY")

# False because "Hunter" is nowhere in the first argument
str_detect("CUNY Bernard M. Baruch College", "Hunter")
```

As with other functions in `R`, `str_detect` is vectorized, making it easy
to use inside of other functions. 

```{r}
names <- c("City College-Miami", "CUNY City College", "CUNY Hunter College")
str_detect(names, "CUNY")
```

Use this function, in conjunction with a `mutate` function, to create a new
column called `is_cuny` inside the IPEDS data. Make sure to assign your mutated
data frame so that you can use it later. 

Your code should look something like: 

```{r}
#| eval: false
IPEDS <- IPEDS |>
    mutate(is_cuny = ...)
```

where the `...` is some code using the string `"CUNY"`, the existing column 
names in `IPEDS` and the `str_detect` function. 

:::

[^re]: `str_detect` can actually be used to perform significantly more complex
string analysis than simple "does it contain this subset of letters" but we
won't cover that sort of string processing for a [few more weeks](../pre/pa10.qmd). 

The public colleges and universities of the state of California are an 
interesting test case. Persuant to 
[California Proposition 209 (1996)](https://en.wikipedia.org/wiki/1996_California_Proposition_209),
public institutions of higher education in the state were specifically not
allowed to implement any sort of Affirmative Action program, so the ending of
Affirmative Action would theoretically have no impact on their admissions 
practices. We will investigate this possibility below. For now, we need to 
create an additional variable which identifies the California public 
institutions in our data set. 

Unlike the CUNY system, which combines community colleges, senior (four-year)
colleges, and specialized graduate institutions into a single system, California
has *three* separate public systems: 

- the [*Univeristy of California* System](https://en.wikipedia.org/wiki/University_of_California)
- the [*California State* System](https://en.wikipedia.org/wiki/California_State_University) and
- the [*California Community Colleges*](https://en.wikipedia.org/wiki/California_Community_Colleges) system. 

Because of this, we will need to do a bit more work to create our `is_calpublic`
variable. Because the Community Colleges are open-enrollment, we can ignore them
from our analysis and focus exclusively on the UC system and the Cal State 
system. 

::: {.callout-tip title="Task 3: Identify UC and Cal State Schools"}

Add a new variable `is_calpublic` to the `IPEDS` data that is `TRUE` for 
institutions that are part of the University of California System or the 
California State system. 

To do this, you will need to use `str_detect` (see the previous task) and one
of the Boolean logical operators (`!`, `&`, `|`). 

Once you have created your variable, use a combination of `dplyr` functions to
get a list of the institutions identified by your new variable. You may wish to
compare this against the Wikipedia articles linked above to ensure your results
are accurate. 

*Note:* If you know what a *regular expression* is, you know that you can use
only a single `str_detect` call here and avoid the use of a Boolean operator. 
Do **not** do this - you must use some Boolean logic to get full credit for this
task. If you do not know what a regular expression is, you can disregard this
note. 

:::

### Initial Data Exploration {#eda}

Before moving to our final analysis, we will do a bit of 
_Exploratory Data Analysis_ (EDA). EDA serves many purposes in data science-- 
quality control, hypothesis generation, outlier identification, *etc.*--but 
perhaps the most important is simply knowing what information can be found in a 
novel data set. Now that our data is imported and cleaned, it's almost 
time to start our EDA. Before we do EDA, however, we should pause briefly to
consider how we want to display data. 

#### Displaying Data in Tables

While we could continue investigating our data using `R`'s basic print-outs, 
this is a good time to introduce the [`gt` package](https://gt.rstudio.com/) 
which can be used to create complex tables natively in `R`.

What follows here is a very brief introduction to the `gt` package. You 
do not need to copy this into your submission - it is provided only as background. 
You will use the `gt` package to format your answers to the next few tasks. 

For this introduction, I am going to use the `penguins` data, but the same
functions can be applied to the IPEDS data that you are analyzing. 

Let's first look at what a "basic" or "raw" display of a data frame gives us: 

```{r}
penguins
```

This has *several* problems: 

1. We are showing *way* too much data. A reader will not be able to easily find
   meaningful trends or patterns in a big data set like this. 
   
   As a general rule, you should rarely have more than 10-15 rows in a table;
   even then, you will still want to guide your reader to the point of the table.
2. The column names are rather ugly. Some, like `species` are not too bad, though
   it would still be better if they were capitalized. Others, like `bill_len`
   are pretty terrible: `bill_len` is not an English word, the underscore
   exists only to separate two words "in code" (recall `R`'s restrictions on 
   variable names), and the unit isn't clear. In this case, a column name like
   `Bill Length (mm)` would be far preferable. 
   
3. The row numbers are essentially pointless and just take up space, adding no
   value. Any content that is not adding value is simply distracting the reader
   from the content that has value. 
4. The "point" of the table is unclear. What is a reader supposed to get from
   this? As a data analyst - doing work on behalf of a reader who may not be a
   data analyst - you have a responsibility to clearly convey the "story" of
   your findings and this does not do so. 
   
   I may want to use this data to show that Gentoo penguins are, on average, 
   heavier than the other two species in this data set, but this is far from 
   clear. 
5. It's just a bit ugly. 

Good table design requires us to take on the mindset of the reader. Tools like
`gt` can help pretty things up, but you still have to think about *what* you 
want to display. Well-formatted garbage is still garbage. 

To start improving this table, let's do the calculations for our reader instead
of expecting them to do it all manually: 

```{r}
library(tidyverse)
penguins |> 
    group_by(species) |> 
    summarize(n_species = n(), 
              avg_body_mass = mean(body_mass, na.rm=TRUE)) |>
    arrange(desc(avg_body_mass))
```

We're definitely not done - but here the "point" of the table is clear, at least
if we also put some text surround it. 

To improve this further, we can also pass this smaller summary data frame to
the `gt` function from the package of the same name: 

```{r}
#| warning: false
library(gt)
penguins |> 
    group_by(species) |> 
    summarize(n_species = n(), 
              avg_body_mass = mean(body_mass, na.rm=TRUE)) |>
    arrange(desc(avg_body_mass)) |>
    gt()
```

Note here that `gt` recognizes we are rendering an HTML page and produces a 
"real" HTML table here. If you were to copy and paste the table above into 
"table" software, *e.g.* Google Sheets or Microsoft Excel, it would be properly 
and automatically handled. For us, the table is the end-point, but it's a nice 
courtesy to your reader who may want to use your results in their own 
presentations. 

The `gt` package provides *many* functions for tweaking and improving the 
appearance of a table. You will almost always want to, at a minimum, use these 
for: 

- Ordering and (re-)naming columns
- Adding titles and footers
- Formatting values

Let's to through these one at a time. Firstly, we want to rename and reorder
the columns. This can be done in pure `dplyr` with the `select` and `rename`
columns, but we'll show the `gt` way here: 

```{r}
library(gt)
penguins |> 
    group_by(species) |> 
    summarize(n_species = n(), 
              avg_body_mass = mean(body_mass, na.rm=TRUE)) |>
    arrange(desc(avg_body_mass)) |>
    gt() |>
    cols_move_to_end(n_species) |>
    cols_label(species="Species", 
               avg_body_mass = "Avg. Body Mass (g)", 
               n_species = "Number of Penguins in Sample") 
```

Here, we used the `cols_move_to_end` function to move the `n_species` column
to the end (no surprise!). In other contexts, we might want to use the 
`cols_move_to_start` function to move a column to the leftmost side of a table
or `cols_move` to put a column in the middle of the table. 

The `cols_label` function essentially serves as a renaming operation: the 
left side of each parentheses is the old column name in the table and the right 
side gives the new name. (Note, a bit confusingly, that this is the *reverse*
of `dplyr::rename`.) While we can just pass a basic string here, we can also
use the `md` function to pass _Markdown_ which lets us do some custom formatting: 

```{r}
penguins |> 
    group_by(species) |> 
    summarize(n_species = n(), 
              avg_body_mass = mean(body_mass, na.rm=TRUE)) |>
    arrange(desc(avg_body_mass)) |>
    gt() |>
    cols_move_to_end(n_species) |>
    cols_label(species=md("**Species**"), 
               avg_body_mass = md("Avg. Body Mass (*g*)"), 
               n_species = md("*Number of Penguins in Sample*"))
```

Here, we could use **boldface** and *italics* for certain text using standard
Markdown syntax. 

Next, we can add a table title and subtitle to make the content and point
of this table clear to our reader: 

```{r}
penguins |> 
    group_by(species) |> 
    summarize(n_species = n(), 
              avg_body_mass = mean(body_mass, na.rm=TRUE)) |>
    arrange(desc(avg_body_mass)) |>
    gt() |>
    cols_move_to_end(n_species) |>
    cols_label(species=md("**Species**"), 
               avg_body_mass = md("Avg. Body Mass (*g*)"), 
               n_species = md("*Number of Penguins in Sample*")) |>
    tab_header(title="Average Body Mass of Three Penguin Species", 
               subtitle="Gentoo penguins are the largest in the study")
```

While we can and should describe our analysis in more detail in the main text,
I like this pattern of having the super-simple one-liner present directly in
the table. This also makes it convenient to clip the table (or a screenshot
thereof) for use in other documents and presentations. 

Next, we should always note the source of the data used to get our results. 
In this case, the original penguins data comes from 
[this article](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0090081) 
so we can cite that in our work Note the use of Markdown (`md()`) to let us include 
a link to the original source within our table: 

```{r}
penguins |> 
    group_by(species) |> 
    summarize(n_species = n(), 
              avg_body_mass = mean(body_mass, na.rm=TRUE)) |>
    arrange(desc(avg_body_mass)) |>
    gt() |>
    cols_move_to_end(n_species) |>
    cols_label(species=md("**Species**"), 
               avg_body_mass = md("Avg. Body Mass (*g*)"), 
               n_species = md("*Number of Penguins in Sample*")) |>
    tab_header(title="Average Body Mass of Three Penguin Species", 
               subtitle="Gentoo penguins are the largest in the study") |>
    tab_source_note(md("Data originally published in by K.B. Gorman, T.D. Williams, 
                       and W. R. Fraser in 'Ecological Sexual Dimorphism and 
                       Environmental Variability within a Community of Antarctic 
                       Penguins (Genus *Pyogscelis*). *PLoS One* 9(3): e90081.
                       <https://doi.org/10.1371/journal.pone.0090081>. Later 
                       popularized via the `R` package
                       [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/)"))
```

A small, but lovely, quality of life feature here is the fact that `md` will
automatically re-align the text to the fit the dimensions of the rendered
table. This lets us put new lines within our citation text so that our code 
doesn't exceed the 80 characters-per-line guideline. 

Finally, we want to make sure the number values are formatted appropriately. 
For these numbers, plain formatting really isn't much of a problem, but for
very large or small numbers, we might want to use scientific notation; for dates,
we might want to control the formatting; *etc.*. This is done with the `fmt_*`
family of functions. 

Each `fmt_` function takes one or more column names and applies a formatting
transformation to that column. The specifics of the formatting can be controlled
with additional optional arguments. For example, if we want to round the average
weight to the nearest gram, we would use the `fmt_number` function with the
argument `deicmals=0`:

```{r}
penguins |> 
    group_by(species) |> 
    summarize(n_species = n(), 
              avg_body_mass = mean(body_mass, na.rm=TRUE)) |>
    arrange(desc(avg_body_mass)) |>
    gt() |>
    cols_move_to_end(n_species) |>
    cols_label(species=md("**Species**"), 
               avg_body_mass = md("Avg. Body Mass (*g*)"), 
               n_species = md("*Number of Penguins in Sample*")) |>
    tab_header(title="Average Body Mass of Three Penguin Species", 
               subtitle="Gentoo penguins are the largest in the study") |>
    tab_source_note(md("Data originally published in by K.B. Gorman, T.D. Williams, 
                       and W. R. Fraser in 'Ecological Sexual Dimorphism and 
                       Environmental Variability within a Community of Antarctic 
                       Penguins (Genus *Pyogscelis*). *PLoS One* 9(3): e90081.
                       <https://doi.org/10.1371/journal.pone.0090081>. Later 
                       popularized via the `R` package
                       [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/)")) |>
    fmt_number(avg_body_mass, decimals=0)
```

In this case, it may be more appropriate to display the body mass in *kilograms*
and we can do so semi-automatically with the `format_number_si` formatter: 

```{r}
penguins |> 
    group_by(species) |> 
    summarize(n_species = n(), 
              avg_body_mass = mean(body_mass, na.rm=TRUE)) |>
    arrange(desc(avg_body_mass)) |>
    gt() |>
    cols_move_to_end(n_species) |>
    cols_label(species=md("**Species**"), 
               avg_body_mass = md("Avg. Body Mass "), 
               n_species = md("*Number of Penguins in Sample*")) |>
    tab_header(title="Average Body Mass of Three Penguin Species", 
               subtitle="Gentoo penguins are the largest in the study") |>
    tab_source_note(md("Data originally published in by K.B. Gorman, T.D. Williams, 
                       and W. R. Fraser in 'Ecological Sexual Dimorphism and 
                       Environmental Variability within a Community of Antarctic 
                       Penguins (Genus *Pyogscelis*). *PLoS One* 9(3): e90081.
                       <https://doi.org/10.1371/journal.pone.0090081>. Later 
                       popularized via the `R` package
                       [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/)")) |>
    fmt_number_si(avg_body_mass, 
                  decimals=2, 
                  unit = "g")
```

Note that, because `fmt_number_si` automatically includes the unit and transforms
it to the most natural scale (here `kg`), we can remove the unit from the column
name. `gt` has many more advanced options that can be used for further 
customization; refer to the package documentation for more details or ask on the
course discussion board.

#### Exploratory Analysis

When faced with a new data set, it is tempting to look only at the first few
rows to get a sense of the data: `R` does this by default. In practice, I 
recommend viewing a _random_ selection of rows instead. This won't guarantee you
find any issues, but it increases the probability of finding issues in 
older parts of a data set. The `slice_sample` function can be used for this. 

While EDA can be an extensive activity on its own, at an *absolute minimum*,
I recommend you always do at least two basic checks: 

- Ensure that you know what `R` *thinks* your data is. You might see a value like
  `"2025-12-01"` and think that `R` is reading it as a date value, but `R`
  might instead by interpreting it as a string value.[^excel] You have several
  options for this in `R`, but when working with a data frame, I'd recommend the
  `glimpse` function: *e.g.*
  
  ```{r}
  glimpse(penguins)
  ```
  
  After getting the basic dimensions of this data frame, `glimpse()` will print
  a line summary of each column giving its name, type, and the first few values
  in the table. 
  
  In this case, since I prepared the data for you, all columns are of the 
  correct type (mostly numeric, with a few character columns for institution name
  and state name, and the two Boolean columns you created earlier), but this is
  a quick and easy check. If there are issues with your data types, it's better
  to catch them early than to have silent and hard to identify errors further
  down the line. ("Fail fast" is *great* advice in any programming exercise.)
  
- Take a quick look at some basic (univariate) summary statistics for each
  column. There are several functions for this in base `R`: *e.g.*, 
  
  ```{r}
  summary(penguins)
  ```
  
  But I actually like the `skim` function from the `skimr` package: 
  
  ```{r}
  #| warning: false
  library(skimr)
  skim(penguins)
  ```
  
  You'll see that this gives some nice overview about data structure and types,
  identifies grouping structure (if any is present), and gives summaries
  appropriate to the type of each column (here `factors`, *i.e.*, categorical 
  variables and `numeric`). I like that this summary gives means, standard 
  deviations, minima (`p0`), maxima (`p100`), medians (`p50`), and a cute little
  histogram of each variable. 
  
  Note that this type of summary only reveals *univariate* structure: if there
  are interesting multi-dimensional outliers or weird correlation patterns, they
  won't appear here. 
  
[^excel]: In one famous (and slightly tragic) example, Microsoft Excel silently
misinterpreted the names of various genes as numeric values and changed them
from (what it thought was) scientific notation to (what it thought was) standard
numeric formatting. This wound up ruining several important scientific studies. 
Always check your data types! The original study identifying this problem can be
found [here](https://link.springer.com/article/10.1186/s13059-016-1044-7) and
a popular news summary is 
[here](https://qz.com/768334/years-of-genomics-research-is-riddled-with-errors-thanks-to-a-bunch-of-botched-excel-spreadsheets). 

::: {.callout-tip title="Task 4: Initial EDA Pass"}

Perform the two checks described above on the IPEDS data you 
loaded and processed above. Include the code to perform the checks, 
the output of the checks, and describe what you see. 

In general, you shouldn't include this in analysis reports, but we're making
an exception here to practice a good habit. I won't ask you to include this
analysis in future mini-projects or your course project reports, but these
checks are easy and very valuable, so I recommend you make them part of all
analyses. 

:::

We are now ready to begin some EDA. Analysts organize their EDA in a variety of
ways, but one of my favorites is to think of a variety of interesting questions
and to answer them. When the answers don't match my intuition, I know I've found
somewhere I want to dig deeper. For the first few mini-projects, I will provide
these exploratory questions. Later in the course, particularly after we have
discussed the role of plotting and graphics in EDA, you will have opportunities
to organize your EDA in other fashions. 

::: {.callout-tip title="Task 5: Exploratory Questions - Inline Values"}

Using `dplyr` tools, answer the following questions:

1. How many distinct institutions appear in this data set? 
2. How many graduate students were enrolled at Baruch in 2024? 

   *Hint:* Use the `str_detect` function discussed above inside a `filter`
   command to identify Baruch. 
3. How many _total_ students were enrolled at Baruch in 2024? 

   *Hint:* Make sure to avoid double-counting first-year undergraduates. 
4. Which institution had the highest number of enrolled female students in 2019?

   Report at least both the institution and the total number of female students. 
5. Which institution with over 1000 total students *admitted* the highest 
   *proportion* of Native Hawaiian or Pacific Islander (`nhpi`) *first-year*
   undergraduates in 2024? 
   
   Report at least both the institution and the fraction of relevant students. 

As you go through these questions, you may find it useful to create new 
variables in the `IPEDS` data to avoid repeated lengthy calculations, *e.g.*,
a new variable for the total number of enrolled students. 

Each of these questions can be answered with one or two scalar values. Use 
Quarto's [inline code](https://quarto.org/docs/computations/inline-code.html)
functionality to place the values in a sentence; that is, you should answer in
complete sentences, written as normal text with inline code for computed values.

Where appropriate, the `scales` package can be used to format numbers 
attractively: *e.g.*, 

```{r}
#| message: false
library(scales)
comma(12345.6789)
comma(12345.6789, accuracy=0.01)
dollar(12345.6789)
percent(0.1234)
percent(0.1234, accuracy=0.01)
```

:::

The next set of questions have slightly more complex answers and should be
answered with a table, formatted using the techniques described above. Longer
term, you will prefer plots to tables as they are a bit easier to interpret
(humans being visual creatures) but these are the types of questions you might
use to create plots as well. 

::: {.callout-tip title="Task 6: Exploratory Questions - Table Answers"}

Using `dplyr` tools, answer the following questions:

1. Which 5 states had the highest number of graduate students across all 
   institutions located in that state? 

2. In 2024, how many first year undergraduate students were enrolled at CUNY 
   colleges and which colleges did they attend? Report both absolute enrollment 
   numbers and percent of total first-year undergraduates? 
   
   *Hint*: The `fmt_percent` and `fmt_number` functions will be useful here.
3. How has Baruch's total undergraduate enrollment changed over the study 
   period? Report both enrollment numbers and percent change year-over-year. 
   
   *Hint*: The `lag` function will be helpful here. 
4. At what 5 institutions did the *fraction of white students* decrease the most
   over the period from 2010 to 2020? 
   
   *Hint*: You may want to pre-filter by total enrollment to make sure you are
   not only reporting very small institutions in your analysis, as these are
   more prone to large fluctuations.
5. In which 3 states did the *fraction of female undergraduates* increase the
   most over the period from 2010 to 2024?
   
Each of these questions can be answered with a table of just a few rows. Use the
`gt` package, as introduced above, to present your results in an attractive
'publication-quality' format, not just a "raw" `R` output.

:::


### Final Deliverable: College Newpaper Op-Ed

At this point, you have acquired your data, cleaned and prepared it, and
performed your EDA. Now, you are ready to get to work on the final deliverable
of your analysis. Everything that comes before this is important, but typically
less visible to your final customer. 


::: {.callout-tip title="Task 7: Op-Ed"}

Write a brief (no more than 750 words) Op-Ed from the perspective of a college 
president to be published in the campus newspaper. You can be the president of 
any institution you want, so long as it has a meaningful undergraduate population, 
(*i.e.*, not stand-alone graduate schools like the CUNY School of Law), so
have some fun with choosing your persona. 

Your Op-Ed should include (at a minimum) the following information: 

- The definition of diversity you are using
- Some brief statistics about the size and make-up of the student body at your
  institution
- A year-over-year comparison of the entering first-year first-time 
  undergraduate class. 
- A year-over-year comparison of the demographics of the entire student body
- A discussion of long-term diversity trends at your institution
- A comparison of changes at your institution to one-or-more of the California
  public institutions. 
  
  (Think of these as a 'critical value' in a statistical
  test for a change: if your change is smaller than a California change, when
  the California schools shouldn't have had to change their policy at all, your
  change is likely just noise.)

:::

You may include optionally additional tables or even visualizations if you want, 
but these do not replace the requirement to write an Op-Ed. Your Op-Ed should 
stand  "alone" and not be mixed in with your code. Place the code necessary to
perform the op-ed calculations in a separate section and use inline code chunks
to include results of your analysis in the text of your op-ed. Op-Eds that 
hard-code calculated values will be penalized. 

For purposes of this exercise, you can measure "diversity" as the fraction of
non-White/non-Asian-American students in your undergraduate class. If you want
to use a more sophisticated metric, see the first extra credit opportunity below.
It is up to you whether you want to consider gender diversity or not in your
analysis: if you don't want to consider gender diversity, simply sum 
corresponding `m` and `f` columns within each racial group.

{{< include _aidisclosure.qmd >}}


{{< include _echeader.qmd >}}

### Entropy Analysis (Up to 2 Points)

Diversity of a population is a difficult quantity to measure. While simple 
statistics ("percent female" or "percent underrepresented minority") are often
used, they suffer from various challenges in an increasingly diverse world.
(For example, NYC does not have a racial majority, so what does it mean to be
an underrepresented minority in the context of city politics?) History and 
social context can guide the choice of diversity metric, but for this extra 
credit opportunity, you can use a *statistical* measure of diversity known as
[entropy](https://en.wikipedia.org/wiki/Entropy_(information_theory)). 

Op-Eds that use the concept of entropy may be awarded up to two points of
extra credit. Specifically, entropy must be used to answer all of the questions
specified in Task 7 above. 

Entropy of a distribution measures how hard it is to predict. Consider two 
scenarios: 

i) An urn filled with 99% red balls and 1% green balls; 
ii) An urn filled with an equal mix of red and green balls. 

If you are asked repeatedly to guess the color of the next ball drawn from an urn,
knowing only the baseline mixture, your predictions for the first urn will be 
correct 99% of the time (assuming you use the obvious "always guess red" strategy),
while no strategy can be correct more than 50% of the time for the second urn. 

Entropy formalizes this concept as follows: given a (discrete) random variable 
$X$ taking values in a set $\mathcal{X}$, each with probability $p(\cdot)$. The
entropy of $X$ is given by: 

$$H(X) = -\sum_{x \in \mathcal{X}: p(x) > 0} p(x) \log p(x)$$

The sum is taken over all outcomes $x$ with non-zero probability. For our two
urns above: 

$$\begin{align*}
H(\text{Urn 1}) = -\left(0.99 * \log(0.99) + 0.01 * \log(0.01)\right) = 0.056 \\
H(\text{Urn 2}) = -\left(0.50 * \log(0.50) + 0.50 * \log(0.50)\right) = 0.693 \\
\end{align*}$$

This tells us that Urn 2 is quite a bit more random than Urn 1.[^units] Entropy
is particularly helpful in our context as it applies naturally to *categorial*
quantities like race where numerical measures of randomness like variance don't
naturally fit. 

This definition can be extended straightforwardly to multi-category quantities. 
Consider these demographics of Bronx and Queens counties in the 2020 census: 

```{r}
#| echo: false
demos <- tribble(
    ~County, ~`Asian`, ~`Black`, ~`Hispanic`, ~`White`, ~`Other`,
    "Bronx",   0.0460,   0.2848,      0.5476,   0.0888,  0.0336,
    "Queens",  0.2730,   0.1585,      0.2276,   0.2284,  0.0625
) 

demos |> 
    gt() |> 
    cols_label(County = "County / Borough", 
               Other  = "All Other") |>
    tab_source_note(md("Percentages from relevant 
                       [Queens](https://en.wikipedia.org/wiki/Demographics_of_Queens#Race_and_ethnicity) 
                       and 
                       [Bronx](https://en.wikipedia.org/wiki/Demographics_of_the_Bronx#Race_and_ethnicity) 
                       Wikipedia articles")) |>
    tab_footnote(md("Asian, Black, and White percentages correspond to census 
                 estimates of Asian (only, non-hispanic), *etc.* Hispanic 
                 percentage corresponds to census estimates of Hispanic (any 
                 race), and All Other was calculated to values would sum to 
                 100%.")) |>
    tab_header("Demographics of Two NYC Boroughs", 
               "Data from 2020 Census") |>
    fmt_percent(-County)
```
If you repeat the entropy calculation here (with five terms in the sum), you 
find that the Bronx has an entropy of approximately 1.16 while Queens has an 
entropy of 1.49, indicating a more diverse population, even though the Bronx has
a higher proportion of Black and Hispanic residents.  

*Hint*: When computing entropy, it is useful to take $0 * \log(0) = 0$ so that
impossible outcomes are automatically discarded. As such, code like this is 
useful to avoid `NA` issues: 

```{r}
bronx_demos <- c(0.046, 0.2848, 0.5476, 0.0888, 0.0336)
-sum(bronx_demos * log(bronx_demos + 1e-10))
```

The small 'nugget' term (`1e-10`) has no real impact on most probabilities, but
prevents issues arising from $\log(0)$ being undefined. 

[^units]: The units of entropy are somewhat tricky to understand. For us, 
it suffices to know that a larger entropy means more randomness. 

### Advanced `dplyr` Programming (Up to 3 Points)

For up to three extra credit points, add the following question to your Task 6 
EDA: In 2024, which 3 institutions of at least 1000 undergraduates had student 
populations that were 'most representative' of the US undergraduate population
as a whole?

Extra credit will be determined on the basis of how accurately and how
efficiently the KL divergence is calculated, in addition to how well it is 
explained and presented in table format. 
   
The *Kullback–Leibler divergence* or KL divergence can be used to 
measure the difference between two different probability distributions. For
this problem, you can use the enrollment counts at each institution to define
a probability distribution (*e.g.*, $\mathbb{P}(\text{White Male}) = 20\%$,
$\mathbb{P}(\text{White Female}) = 22\%$, *etc.*) for that campus and compare
it to the national undergraduate population. 
   
The following example may help: suppose I have categorized my personal 
library into three categories of books: i) fiction; ii) non-fiction; and
iii) textbooks. Furthermore, my books are unorganized and spread randomly
across 5 shelves as follows: 
   
```{r}
#| echo: false
books <- tribble(
    ~shelf, ~fiction, ~nonfiction, ~textbook, 
    1,       15,          20,         5,
    2,       20,          10,        10,
    3,       40,          20,        20,
    4,       10,          10,        30,
    5,       15,          20,        20
)
   
books_base_counts <- books |> 
    select(-shelf) |>
    summarize(across(everything(), sum)) 

books_base_probs <- books_base_counts |> 
    mutate(total = rowSums(across(everything())), 
           across(everything(), \(x) x/total)) |> 
    select(-total)

shelf_probs <- books |> 
    group_by(shelf) |>
    mutate(total = rowSums(across(everything())), 
           # No need to `-shelf` here as grouping variables aren't included in `everything()` 
           across(everything(), \(x) x/total)) |>
    select(-total)
     
shelf1_probs <- shelf_probs |> filter(shelf == 1)
     
# KL can be computed in two ways
#
# This is simpler, but requires some tools not yet introduced (joins and pivots)
kl <- inner_join(
    pivot_longer(shelf_probs, 
                 -shelf, 
                 names_to="category",
                 values_to="shelf_prob"), 
    pivot_longer(books_base_probs, 
                 everything(), 
                 values_to="collection_prob", 
                 names_to="category"), 
    join_by(category)) |>
    summarize(kl=sum(collection_prob * log(collection_prob / shelf_prob)))

# This avoids joins and pivots, but is pretty fragile and requires 
# rather advanced use of the across() idiom
kl_alt <- books |> 
    mutate(shelf_total = rowSums(across(-shelf)), 
           # Count total # of books per genre
           across(-contains("shelf"), 
                  \(x) sum(x), 
                  .names="{.col}_collection"), 
           # Compute total # of books total, to get collection probabilities
           collection = rowSums(across(contains("collection"))), 
           # Compute collection-wide probabilities
           across(-c(contains("shelf"), 
                     contains("collection")), 
                  \(x) x / shelf_total, 
                  .names="{.col}_p"),
           # Compute per shelf probabilities of each genre
           across(ends_with("_collection"), 
                  \(x) x/collection, 
                  .names="{.col}_p")
    )|>
    # We can now work with probabilities only
    select(shelf, ends_with("_p")) |>
    mutate(across(contains("collection"), 
                  # This is some nasty magic to let us access the population (collection)
                  # and shelf probabilities jointly
                  # cur_column() will give a string like "fiction_collection_p", so
                  # removing "_collection" leaves us with the string "fiction_p". 
                  # Passing this to get() gives access to the shelf-level `fiction_p`
                  # variable that we can then use to get the relevant term of the KL sum
                  \(x) x * log(x / get(str_remove(cur_column(), "_collection"))), 
                  .names="{.col}_kl"), 
           # Sum across individual KL terms for overall KL
           kl=rowSums(across(ends_with("_kl")))) |>
    select(shelf, kl)

# Check two approaches give the same result
stopifnot(all.equal(kl, kl_alt))

   
books |> 
    gt() |> 
    cols_label_with(fn=str_to_title)
```
   
Comparing these shelves is a bit tricky since there aren't the same 
number of books on each shelf and a purely 'numbers-based' comparison
might treat Shelf 2 as more representative than Shelf 3 simply because it
has fewer books. The KL divergence things of these as probabilities (*e.g.*,
Shelf 2 is 50% fiction) and can be used to compare them to the overall 
probabilities of the whole collection. 
   
Adding up the columns, I have `r books_base_counts$fiction` fiction books, 
`r books_base_counts$nonfiction` non-fiction books, and 
`r books_base_counts$textbook` textbooks,  so my collection is about
`r percent(books_base_probs$fiction)` fiction, 
`r percent(books_base_probs$nonfiction)` non-fiction, and 
`r percent(books_base_probs$textbook)` textbooks. 
   
We use these probabilities to compare each row against the overall 
population using the KL divergence formula: 
 
$$\mathcal{D}(P \| Q) = \sum_{x} P(x) \log \frac{P(x)}{Q(x)}$$
   
where the sum $x$ is taken over all categories and $P, Q$ are different 
probability distributions. Here $P$ is the "baseline" (correct) distribution and
$Q$ is the approximation. So the KL divergence between Shelf 1, which is about
`r percent(shelf1_probs$fiction)` fiction, 
`r percent(shelf1_probs$nonfiction)` non-fiction, and 
`r percent(shelf1_probs$textbook)` textbooks, and the collection as a whole
is given by
   
$$
   `r books_base_probs$fiction` \log \frac{`r books_base_probs$fiction`}{`r shelf1_probs$fiction`} + 
   `r books_base_probs$nonfiction` \log \frac{`r books_base_probs$nonfiction`}{`r shelf1_probs$nonfiction`} + 
   `r books_base_probs$textbook` \log \frac{`r books_base_probs$textbook`}{`r shelf1_probs$textbook`} \approx
   `r kl |> filter(shelf == 1) |> pull(kl)`
$$
   
We can repeat this analysis across all 5 shelves to get the following KL table
   
```{r}
#| echo: false
kl |> 
    gt() |>
    cols_label(shelf="Shelf", 
              kl="KL Divergence from Entire Collection") |>
    fmt_number(kl, decimals=4)
    
```
   
From which we see that Shelf `r kl |> slice_min(kl) |> pull(shelf)` is 
"most representative" of my whole collection. Note also that Shelves 2
and 3 have the same KL Divergence because they are (proportionally) the same.

*Hint*: Note that this is a *significantly* more complex analysis than we have
performed so far and requires some ideas that won't be covered until Week 6 and
some column selection functionality that will not be covered in lecture. 
Make sure to access the source code for this assignment and use it as a guide
for completing this extra credit task.

### Data Visualization (1 point)

Inclusion of a well-formatted visual element to accompany your op-ed will get
one extra credit point. 

{{< include _footer.qmd >}}
