---
mp_num: "02"
mp_title: "TBD"
mp_skills: "TBD"
mp_application: "TBD"
mp_rhetoric: "TBD"
mp_max_ec: 5
filters:
    - list-table
---

```{r}
#| echo: false
#| message: false
#| warning: false
library(tidyverse)
DATES <- readr::read_csv("key_dates.csv") |>
    rename(element=`Course Element`,
           item=`Item Number`) |>
    mutate(dt = case_when(is.na(Time) ~ as.character(Date),
                          TRUE ~ paste(Date, Time)))
```

```{r}
#| echo: false
#| warning: false
library(glue)
mp_id <- as.integer(rmarkdown::metadata$mp_num)
mp <-  DATES |> filter(element == "Mini-Projects", 
                       item == mp_id)
mp_file <- list(qmd=glue("`mp{sprintf('%02g', mp_id)}.qmd`"), 
                html=glue("`mp{sprintf('%02g', mp_id)}.html`"),
                html_long=glue("`docs/mp{sprintf('%02g', mp_id)}.html`"))
```

### Due Dates

-   Released to Students: `{r} mp |> filter(str_detect(Details, "Mini-Project Released")) |> pull(dt)`
-   **Initial Submission: `{r} mp |> filter(str_detect(Details, "Mini-Project Due")) |> pull(dt)` on GitHub and Brightspace**
-   **Peer Feedback:**
    -   Peer Feedback Assigned: `{r} mp |> filter(str_detect(Details, "Mini-Project Peer Feedback Assigned")) |> pull(dt)` on GitHub
    -   **Peer Feedback Due: `{r} mp |> filter(str_detect(Details, "Mini-Project Peer Feedback Due")) |> pull(dt)` on GitHub**


*Estimated Time to Complete: 13-15 Hours*

*Estimated Time for Peer Feedback: 1 Hour*


------------------------------------------------------------------------

## Introduction

Welcome to Mini-Project #{{< meta mp_num >}}! TBD

In this mini-project, you will: 

1) Practice use of `dplyr` for combining and manipulating data from distinct
   sources
2) Practice use of `ggplot2` to create compelling statistical visualizations. 
   Note that this assignment only requires "basic" `ggplot2` visualizations
   like scatterplots or line graphs; [Mini-Project #03](mini03.html) will require
   more complex visualization. 
3) Practice summarizing complex analytical pipelines for non-technical audiences.

As with [Mini-Project #01](mini01.html), a major portion of your grade is based on
the _communication and writing_ elements of this mini-project, so make sure to
budget sufficient time for the writing process. Code without context has little
practical value - you are always writing and conveying your analysis to a specific
target audience. 

### Student Responsibilities

Recall our basic analytic workflow and table of student responsibilities:

{{< include _responsibilities.qmd >}}

In this mini-project, you are mainly responsible for data alignment and basic 
statistical analyses. While not the main focus of this mini-project, you are 
also expected to provide basic data visualizations to support your findings; 
the grading rubric, below, emphasizes the `dplyr` tools used in this project, 
but reports without any visualization will be penalized severely. Note that data
visualization will play a larger role in Mini-Project #03.

As before, I will provide code to automatically download and read in the data 
used for this project. Also note that, as compared with Mini-Project #01, I am
providing less 'scaffolding' for tabular display and EDA: students are more 
responsible for directing their own analyses.

### Rubric

{{< var course.short >}} Mini-Projects are evaluated using *peer grading* with 
*meta-review* by the course staff. The following basic rubric will be used for 
all mini-projects:

{{< include _rubric.qmd >}}

At this point, you are responsible for (elementary) data visualization and will
be graded on that aspect of the rubric, as well as the sections on which you were
assessed in [the previous mini-project](./mini01.qmd).

Note that, because I am still providing code to download the data, load it into
`R`, and prepare it for analysis, all reports submitted using my code will 
receive an automatic 10/10 for the 'Data Preparation' element of the rubric. 
Additionally, reports completing all tasks described under [Data Integration and
Exploration](#eda) below should receive a 10/10 for the 'Exploratory Data Analysis'
rubric element. 

Taken together, you are only really responsible for these portions of the
rubric in this assignment: 

- Written Communication
- Project Skeleton
- Tables & Document Presentation
- Data Visualization
- Code Quality
- Analysis and Findings

Reports completing all key steps outlined below essentially start with
20 free points. 

{{< include _submission.qmd >}}


## Mini-Project #{{< meta mp_num >}}: {{< meta mp_title >}}

### Data Acquisition

TBD


::: {.callout-tip title="Task 1: Data Acquisition"}

Using the code above, acquire the TBD. Copy the code into
your Quarto document and make sure it runs successfully. 

:::

::: {.callout-caution title="Do Not `git add` Data Files"}

Make sure that `git` is set to ignore data files, such as the one created above.
Check the `git` pane in `RStudio` and make sure that TBD
does not appear. (If you set up your `.gitignore` file correctly in
[MP#00](./mini00.html), it should already be ignored.) If it is appearing, you
may need to edit your `.gitignore` file.

Removing a large data file from `git` is possible, but difficult. Don't get
into a bad state!

:::

### Data Cleaning and Preparation 

TBD

### Data Integration and Initial Exploration {#eda}

TBD

### Final Deliverable: TBD

TBD

    
{{< include _aidisclosure.qmd >}}


{{< include _echeader.qmd >}}

TBD

{{< include _footer.qmd >}}

