---
mp_num: "04"
mp_title: "Going for the Gold"
mp_skills: "Web Scraping, Data Cleaning, Forecasting"
mp_application: "Sports Analytics"
mp_rhetoric: "Popular Media, Fundraising Pitch"
mp_max_ec: 7
filters: 
 - list-table
aliases: 
  - "/miniprojects/mini04.html"
---


{{< include _mpsetup.qmd >}}

## Introduction

Welcome to Mini-Project #{{< meta mp_num >}}!

This project will bring together all of the skills you have
developed in this course: 

- Accessing data from the web using HTTP requests (`httr2`)
- Extracting data from HTML and cleaning it (`rvest`, `stringr`)
- Joining together data from different tables (multi-table `dplyr`)
- "Pivoting" and preparing data for analysis (single-table `dplyr`)
- Exploring and visualizing data (`ggplot2`)
- Statistical analysis and inference (`infer` or `stats`)
- Communicating your findings using web-based reproducible research tools (Quarto)

You will use these skills to deliver a very early prediction of Team USA's 
medal haul from the upcoming *2028 Los Angeles Summer Olympics*. 

Note that, compared to previous mini-projects, the scope of this project
is a bit smaller: in light of this, and the more advanced skills you have
spent the past 3 months developing, this mini-project should be the least
demanding of the course. (It is still likely to be the most *difficult* simply
because you are responsible for more steps of the data analysis than any prior 
project.) At this point in the course, you should be spending the majority of 
your out-of-class hours on your [Course Project](../project.html). 

## The Final Mini-Project

This mini-project completes our whirlwind tour of several different
forms of data-driven writing: 

- MP01: `r get_mp_rhetoric(1)`
- MP02: `r get_mp_rhetoric(2)`
- MP03: `r get_mp_rhetoric(3)`
- MP04: `r get_mp_rhetoric(4)`

Within these projects, you have developed and showcased a diverse array
of skills: 

- MP01: `r get_mp_skills(1)`
- MP02: `r get_mp_skills(2)`
- MP03: `r get_mp_skills(3)`
- MP04: `r get_mp_skills(4)`

The skills you have developed are powerful and flexible and can be used across
a variety of application areas: 

- MP01: `r get_mp_application(1)`
- MP02: `r get_mp_application(2)`
- MP03: `r get_mp_application(3)`
- MP04: `r get_mp_application(4)`

There are, of course, many other ways that data can be used to generate and
communicate insights and many other topics where your skills can be applied, but
hopefully this "hit parade" has given you a sense of just how widely your 
new skills can take you. In each of these domains, you have seen how sophisticated
and thoughtful analysis - not simply computing a single mean or regression 
coefficient - has allowed deeper understanding of the complexities of real world
problems. You have learned to think beyond the simple binary "correct/incorrect"
of an analysis and to see that even superficially simple questions - 
"is this true?", "what is the average?", "is there a correlation?" - can lead
down deep and rewarding rabbit holes of complexity. But those rabbit holes 
are where your real power as a data analyst can be found.  

The tools of quantitative analysis and communication you have
developed in this course can be used in essentially infinite contexts--
we have only scratched the surface--and I'm excited to see what you do
in the remainder of this course, in your remaining time at Baruch, and
in your future careers. 

### Student Responsibilities

Recall our basic analytic workflow and table of student responsibilities:

{{< include _responsibilities.qmd >}}

In this mini-project, you are in charge of the whole pipeline, from data 
acquisition to statistical inference. The rubric below evaluates your work 
on all aspects of this project. 

### Rubric

{{< include _rubric.qmd >}}

For this mini-project, students are responsible for **all elements** of the 
analysis and will be evaluated on all rubric elements. 

{{< include _submission.qmd >}}

## Mini-Project #{{< meta mp_num >}}: {{< meta mp_title >}}

With the 2026 Winter Olympics complete, the world is now preparing for the 
2028 Summer Olympics, to be held in Los Angeles, California.[^la_prep] At the
same time, sports betting--and indeed, gambling on essentially *everything*--seems
to be an ever-more common society phenomenon.[^gambling] As such, serious money
is likely to be wagered on Team USA's performance in the 2028 Olympics. Your 
task in this final mini-project is to build a (relatively simple) model to
predict the total amount of gold medals Team USA is likely to win in 2028. 

Your model will need to consider *at least* three factors. For extra credit, you
will be able to consider other factors to build a (hopefully!) more accurate
forecast; [see below]{#extra_credit} for more. Your model needs to incorporate
three primary effects:

A)  The baseline performance of Team USA. As a large and rich nation that spans
    several different different climates (sun and snow), the United States have 
    long performed well in both winter and summer olympics. 
B)  The Host Nation effect. Most analyses suggest that the host nation 
    overperforms relative to its performance in the Olympiad immediately before
    or afterwards. The reasons for this are not entirely clear: they seem to be
    some combination of: 
    
    i) the psychological benefits of having more supporters in the stands (the
       "home turf" effect); 
    ii) the benefit of receiving automatic qualification to *every* event. The
        host nation is basically guaranteed to have at least one athelete 
        qualify in every event. (The specifics vary by sport.) Given that these
        "automatic qualifiers" have to compete against athletes who qualify
        based on their strong performance in non-Olympic international competition,
        the impact of this factor is likely small, but there are lots of small
        events and every extra entry increases the probability of a 'fluke' 
        victory. 
    iii) the pre-competition enthusiasm for the games leading to more national
         support and interest in Olympic competition. 
         
    (Note that [some authors](https://doi.org/10.1038/s41598-022-27259-8) 
    dispute whether a host nation benefit exists at all.)
    
C)  The new sport effect. The host nation is able to introduce up to six new 
    sports to the Olympics. Given that these have to be sports that are not 
    already included in the "standard" Olympic schedule, these sports are often
    (relatively) niche events that are particularly popular in the host nation. 
    As such, the host nation is often expected to perform well in these sports. 
    
    In the 2020 Olympics (actually held in 2021), Japan introduced several 
    sports, including Karate in which Japanese atheletes won five of twelve 
    possible gold medals and baseball/softball in which the Japanese won both
    gold medals. The 2024 Paris Olympics introduced the (seemingly short-lived) 
    Olympic Breaking competition to [mixed
    reviews](https://www.youtube.com/watch?v=T8q1d4LZb1Y). The 2028 
    Olympics are (re-)introducing baseball/softball, cricket, flag football, 
    lacrosse, and squash. While the US is not known as a cricket powerhouse, it 
    is expected to be quite strong in the other sports. 
    
To determine the medal impact of these three factors, you will perform a series
of statistical analyses intended to estimate confidence intervals for each
effect. Before we can do so, however, it's time to acquire some data. 

[^la_prep]: The degree to which LA is itself preparing for the 2028 Olympics is
[rather more complicated](https://en.wikipedia.org/wiki/Twenty-eight_by_%2728).

[^gambling]: For the legal and societal preconditions of this trend, see [this
article](https://bclawreview.bc.edu/articles/10.70167/DDAX4275) from the 
Boston College Law Review. 

### Data Acquisition

The [Olympedia](https://www.olympedia.org/) website contains detailed records
of all modern Olympic games. You will use *web-scraping* techniques to extract
the following information: 

- Modern Olympic Games: for both modern[^modern] summer and winter Olympics, 
  extract the following from <https://www.olympedia.org/editions>: 
  
  - The year in which the Olympics were held
  - the Olympiad ID for that games (*e.g.*, Chamonix 1924 was the 1st Winter 
    Olympiad while Paris 1924 was the 8th Summer Olympiad)
    
    Note that Olympiads are traditionally written using Roman numerals. The
    following `R` snippet may be useful in translating these back to more
    standard (decimal) integers. 
    
    ```{r}
    olympiad_ids <- c("I", "III", "IV", "V", "VI", "XII", "XXV")
    roman_to_integer <- function(x){
        setNames(as.integer(as.roman(olympiad_ids)), x)  
    }
    roman_to_integer(olympiad_ids)
    ```
    
  - The Olympedia internal ID for that games; Olympedia refers to these as
    "editions". 
  - The host nation for that games. Do this by parsing the URL associated
    with the flag in the `Country` column. *E.g.*, the Greek flag used to
    identify the 1896 Olympics is loaded from the URL
    `https://olympedia-flags.s3.eu-central-1.amazonaws.com/GRE.png`; by parsing
    the three letter country code (`GRE`), we can identify this as Greece. 
    
    Note that, since you want to extract some information from the URLs in the
    table, not just the text of the table, you will need to use more than just
    `rvest::html_table()`.
- For each games, identify the table of medal disciplines (sports) at that games.
  *E.g*. at the 1896 Paris Olympics, [we can see
  that](https://www.olympedia.org/editions/1) 10 medal disciplines were 
  represented. For each of these, extract a link to the relevant page, *e.g.*,
  <https://www.olympedia.org/editions/1/sports/ATH> for Athletics. 
- For each medal discipline + edition combination, extract the medal table. 

**NB**: Olympedia is a free resource, made available at the expense of a small
group of dedicated volunteers, and we do not want to abuse it. One way to
avoid abusing their servers is to only download pages one; we can save
a copy of the page locally the first time we access it and use that instead of 
accessing the URL repeatedly. The following code will do this for you: 

```{r}
#| message: false
library(rvest)

slow_download <- purrr::slowly(download.file, purrr::rate_delay(pause=5))

read_olympedia <- function(url, ...){
    if(stringr::str_starts(url, "/")) url <- stringr::str_remove(url, "/")
    cache_dir <- fs::path("data", "mp04", "olympedia_cache")
    if(!fs::dir_exists(cache_dir)) fs::dir_create(cache_dir)

    cache_path <- fs::path(cache_dir, 
                           url |> as.character() |> stringr::str_replace_all("/", "_"))
    if(!fs::file_exists(cache_path)){
        src_url <- paste0("https://www.olympedia.org/", url)
        slow_download(src_url, cache_path, mode="wb")
    }
    rvest::read_html(cache_path)
}
```

Use this code to avoid taxing Olympedia's servers. If you access the site too
often or too quickly, you will be automatically blocked for a short period. 
If this happens to you, take a short break and come back to it later. To use
this function, simply pass the *path* part of the URL and the hostname will be
appended automatically; *e.g.*, to read the contents of
`https://www.olympedia.org/editions` into `R`, you would simply run
`read_olympedia("editions")`. Note that this function is designed to be slow
the first time you access a site (`pause=5` indicating that it won't hit
Olympedia more than once every 5 seconds) but it will be faster each subsequent
time since the HTML is now cached on your computer.
    
::: {.callout-tip title="Task 1: Acquire Olympedia Data"}

Using the user provided function above, download all necessary Olympedia data.
Once your have acquired all of the relevant pages and parsed them into a single
data frame, combine all of your results into a single tidy table. With this data 
in hand, you should have everything you need to create a table like this: 

```{r}
#| echo: false
#| message: false
library(gt)
library(necountries)

# Use a random selected subset of rows for presentation purposes
# If you are a student viewing this code, don't use the whole data set, not
# just a subset!

olympic_sample <- 
structure(list(edition_id = c(2L, 3L, 2L, 2L, 2L, 2L, 2L, 1L, 
1L, 1L, 1L, 2L, 1L, 3L, 2L, 2L, 2L, 2L, 3L, 2L, 3L, 1L, 2L, 2L, 
3L), edition_roman = structure(c(2L, 3L, 2L, 2L, 2L, 2L, 2L, 
1L, 1L, 1L, 1L, 2L, 1L, 3L, 2L, 2L, 2L, 2L, 3L, 2L, 3L, 1L, 2L, 
2L, 3L), class = "roman"), edition_url = c("/editions/2", "/editions/3", 
"/editions/2", "/editions/2", "/editions/2", "/editions/2", "/editions/2", 
"/editions/1", "/editions/1", "/editions/1", "/editions/1", "/editions/2", 
"/editions/1", "/editions/3", "/editions/2", "/editions/2", "/editions/2", 
"/editions/2", "/editions/3", "/editions/2", "/editions/3", "/editions/1", 
"/editions/2", "/editions/2", "/editions/3"), year = c(1900L, 
1904L, 1900L, 1900L, 1900L, 1900L, 1900L, 1896L, 1896L, 1896L, 
1896L, 1900L, 1896L, 1904L, 1900L, 1900L, 1900L, 1900L, 1904L, 
1900L, 1904L, 1896L, 1900L, 1900L, 1904L), city = c("Paris", 
"St. Louis", "Paris", "Paris", "Paris", "Paris", "Paris", "Athina", 
"Athina", "Athina", "Athina", "Paris", "Athina", "St. Louis", 
"Paris", "Paris", "Paris", "Paris", "St. Louis", "Paris", "St. Louis", 
"Athina", "Paris", "Paris", "St. Louis"), host_country_iso = c("FRA", 
"USA", "FRA", "FRA", "FRA", "FRA", "FRA", "GRE", "GRE", "GRE", 
"GRE", "FRA", "GRE", "USA", "FRA", "FRA", "FRA", "FRA", "USA", 
"FRA", "USA", "GRE", "FRA", "FRA", "USA"), season = c("Summer", 
"Summer", "Summer", "Summer", "Summer", "Summer", "Summer", "Summer", 
"Summer", "Summer", "Summer", "Summer", "Summer", "Summer", "Summer", 
"Summer", "Summer", "Summer", "Summer", "Summer", "Summer", "Summer", 
"Summer", "Summer", "Summer"), discipline_name = c("Rowing", 
"Golf", "Swimming", "Swimming", "Equestrian Dressage", "Sailing", 
"Cycling Track", "Fencing", "Athletics", "Wrestling", "Athletics", 
"Rugby", "Athletics", "Swimming", "Equestrian Driving", "Athletics", 
"Athletics", "Cycling Track", "Tug-Of-War", "Water Polo", "Artistic Gymnastics", 
"Shooting", "Golf", "Tennis", "Weightlifting"), discipline_url = c("/editions/2/sports/ROW", 
"/editions/3/sports/GLF", "/editions/2/sports/SWM", "/editions/2/sports/SWM", 
"/editions/2/sports/EDR", "/editions/2/sports/SAL", "/editions/2/sports/CTR", 
"/editions/1/sports/FEN", "/editions/1/sports/ATH", "/editions/1/sports/WRE", 
"/editions/1/sports/ATH", "/editions/2/sports/RGB", "/editions/1/sports/ATH", 
"/editions/3/sports/SWM", "/editions/2/sports/EDV", "/editions/2/sports/ATH", 
"/editions/2/sports/ATH", "/editions/2/sports/CTR", "/editions/3/sports/TOW", 
"/editions/2/sports/WPO", "/editions/3/sports/GAR", "/editions/1/sports/SHO", 
"/editions/2/sports/GLF", "/editions/2/sports/TEN", "/editions/3/sports/WLF"
), competitor_country = c("Germany", "Canada", "Denmark", "Great Britain", 
"France", "France", "Germany", "Denmark", "Germany", "Germany", 
"Australia", "Germany", "Great Britain", "Hungary", "Belgium", 
"Hungary", "France", "France", "United States", "Great Britain", 
"Germany", "United States", "United States", "Great Britain", 
"Greece"), competitor_iso = c("GER", "CAN", "DEN", "GBR", "FRA", 
"FRA", "GER", "DEN", "GER", "GER", "AUS", "GER", "GBR", "HUN", 
"BEL", "HUN", "FRA", "FRA", "USA", "GBR", "GER", "USA", "USA", 
"GBR", "GRE"), Gold = c(1L, 1L, 0L, 4L, 1L, 7L, 0L, 0L, 0L, 1L, 
2L, 0L, 0L, 2L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 2L, 2L, 4L, 1L), 
    Silver = c(0L, 0L, 0L, 0L, 1L, 9L, 1L, 0L, 1L, 0L, 0L, 1L, 
    1L, 1L, 0L, 0L, 4L, 2L, 1L, 0L, 1L, 1L, 0L, 1L, 0L), Bronze = c(2L, 
    0L, 1L, 1L, 1L, 10L, 0L, 1L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 
    1L, 2L, 2L, 1L, 0L, 2L, 0L, 0L, 3L, 0L), Total = c(3L, 1L, 
    1L, 5L, 3L, 26L, 1L, 1L, 1L, 1L, 2L, 1L, 2L, 4L, 1L, 2L, 
    7L, 6L, 3L, 1L, 4L, 3L, 2L, 8L, 1L)), row.names = c(NA, -25L
), class = c("tbl_df", "tbl", "data.frame"))

clean_countries <- necountries::countries() |> 
    as_tibble() |> 
    select(iso2, iso3) |> 
    mutate(iso3 = replace_values(iso3, 
                          "GRC" ~ "GRE", 
                          "DEU" ~ "GER", 
                          "DNK" ~ "DEN"))

olympic_sample |> 
    slice_sample(n=5) |> 
    mutate(host_country_iso = replace_values(host_country_iso, 
                                             from=clean_countries$iso3, 
                                             to=clean_countries$iso2), 
           competitor_iso = replace_values(competitor_iso, 
                                             from=clean_countries$iso3, 
                                             to=clean_countries$iso2), 
           edition_url = paste0("https://www.olympedia.org", edition_url), 
           discipline_url = paste0("https://www.olympedia.org", discipline_url)) |>
    arrange(year, desc(season), discipline_name, desc(Total)) |>
    gt(id="tbl_olympic_example") |> 
    cols_hide(c(edition_id)) |>
    fmt_flag(c(competitor_iso, host_country_iso)) |>
    fmt_url(c(edition_url, discipline_url), 
            label="(Details)") |>
    cols_label(edition_roman="Games", 
               competitor_country="Competitor",
               discipline_name="Discipline") |> 
    cols_merge(c(edition_roman, season, year, city, edition_url, host_country_iso), 
               pattern="{2} Games of the {1} Olympiad ({3} {4} {6}) {5}") |>
    cols_merge(c(competitor_country, competitor_iso), 
               pattern="{1} ({2})") |>
    cols_merge(c(discipline_name, discipline_url)) |>
    tab_spanner("Medal Count", 
                c(Gold, Silver, Bronze, Total)) |> 
    tab_source_note("Data from Olympedia. 5 randomly selected rows - not the full data set.") |> 
    tab_header("Olympic Medal Results by Discipline")
```

Note that you do not need to actually make a table like this - just check that
you have everything you would need to make this table. 

*Hint:* See the source code for this Mini-Project on GitHub to see how I 
formatted this table. You do not need to use the same column names I did, but
you are always encouraged to use syntactic (valid `R`) names to make your code
easier to read and write. 

:::

Some general tips: 

- This is a difficult scraping exercise, but don't go overboard. I did this
  whole thing in about 80 lines of clear well-formatted code (not including
  `read_olympedia` above). If you find yourself needing more than 200 lines of
  code to read everything, reconsider your strategy. 
- Make sure to `DRY` your code. Write a function to parse a page and extract the
  relevant data; since the pages are formatted consistently, this will be far
  easier than copy-and-pasting the same code over and over. 
  
  If your function takes in a `URL` and parses the results, but is not vectorized,
  remember that you can use the *functional* idiom: 
  
  ```{r}
  #| eval: false
  library(tidyverse)
  ... |>
    mutate(results = map(url, read_results_page))
  ```
  
  To apply `read_results_page` to each URL in sequence. If the result of
  `read_results_page` is a tidy `data.frame`, the `results` column will be a
  "list column", which is not something we have used in class. Use 
  `tidyr::unnest()` to convert it to a more standard tabular format. 
  So the whole idiom would be
  something like: 
  
  ```{r}
  #| eval: false
  library(tidyverse)
  ... |>
    mutate(results = map(url, read_results_page)) |>
    unnest(results)
  ```
- If you want to pull out a specific column of a table, consider using a CSS
  selector like `tr > td:nth-of-type(3)`: this will pull the third `td` within
  each `tr`. Because each `tr` is a **t**able **r**ow and each `td` is a 
  **t**able **d**atum or cell, this gives the third cell of each row, *i.e.*, the
  third column. 
- For some pages, you might want to get the first table after a section header
  with a specific text. This cannot be done with simple CSS selectors, but 
  an `xpath` selector will suffice here. *E.g.*, 
  
  ```{r}
  #| eval: false
  ... |> 
     html_element(xpath = "//h1[normalize-space() = 'Section Header']/following-sibling::table[3]") |>
     html_table()
  ```
  
  This will look for an `h1` header element whose text reads `Section Header`;
  once it finds it, it will get the third `table` that follows and convert it
  to a table using our trusty `html_table()` function. Modify this example as 
  needed to work with Olympedia. 
  
[^modern]: For this assignment, we are of course only looking at the *modern*
Olympics that resumed in 1896. We will not look at the ancient (pre-fall of 
the Roman Empire) Olympics. Note that, while the Summer Olympics resumed in
1896, the first modern Winter Olympics were only held starting in 1924 and 
the Winter Olympics have only been "off-cycle" from the Summer Olympics starting
with the 1994 Lillehammer games. Additionally, note that certain Olympics were 
not held during WWI and WWII. Finally, we will only focus on "mainline" Summer
and Winter Olympics and will ignore the 1956 Equestrian Olympics and similar
events. 

::: {.callout-caution title="Use Olympedia Only"}

You **must** use Olympedia as your data source for this mini-project. While
Olympic results are found on many websites, including Wikipedia, the point of
this Mini-Project is to practice scraping a *new* website you have not previously
used. (We discuss scraping tables from Wikipedia in class.)

Submissions that acquire data from alternate sources will be penalized 
**heavily.**

:::


::: {.callout-tip title="Task 2: Standardize and Clean Countries"}

Note that Olympedia often uses a three-letter code for each country, but these
do not always correspond with the [three letter ISO country
codes](https://en.wikipedia.org/wiki/ISO_3166-1_alpha-3), *e.g.*, `GRC` vs `GRE`
for Greece. You will need to identify and fix these mismatches (see below). The
`replace_values()` function from (recent versions of) the `dplyr` package will 
be helpful to do these replacements and the `anti_join()` operator will be useful 
in finding where they are needed.

Furthermore, at least one code, `URS` for the Soviet Union (USSR), is no longer
applicable. For simplicity, we are going to replace `URS` with `RUS` (Russia) to
ensure continuity of our analysis, though this choice is not geopolitically
un-fraught.

Given the spatial structure of this data, we will want to augment it with `sf`
data for geospatial visualization. The `countries()` function from the 
`necountries` package returns an `sf` data frame of 199 countries that contains
their two- and three-digit country codes, their name, and `geometry` column that
can be used to visualize them on a map. 

Join your Olympedia data to this object. (This is where you will start running
into pain from the mismatch between Olympedia's three-digit codes and the official
ISO codes.) Design your join to identify country codes not present in the
`countries()` data frame and, where possible, perform 'spot fixes' to 
standardize these onto countries present in the data set. (*E.g.*, the `URS` to
`RUS` standardization described above.)

Once your data is standardized, create at least one geospatial data 
visualization of this data to highlight an interesting fact. 

:::

### Data Integration and Exploration {#eda}

Now that you have acquired your data, it is time to explore it and further 
familiarize yourself with it. For this mini-project, you are on your own for EDA. 
Explore your data, find and note any irregularities (using them to improve your
data import if appropriate) and to get baseline facts relevant to your analysis. 

::: {.callout-tip title="Task 3: Exploratory Data Analysis"}

Perform EDA on your data set. Document your results in (at least) 12 findings. 
These should include at least: 

- one inline value (scalar statistic); 
- one data visualization; and
- one table

for each of the three effects we plan to study: US baseline performance, host 
country effect, new sport effect. You can choose the topic and presentation of
your final three EDA findings however you wish. 

:::

### Statistical Analysis

While Exploratory Data Analysis (EDA) can inspire questions, it remains useful
to perform _formal statistical inference_ to assess whether observed differences
are larger than can be explained by pure randomness. The `infer` package can
be used to neatly integrate $t$-tests and binomial proportion tests into a 
`tidyverse` workflow. 

We will use the `t_test()` and `prop_test()` functions from this package to 
complete our analyses. These can be used as demonstrated below. 

-   Example 01: Suppose we want to test whether Gentoo penguins are heavier than
    Adelie penguins on average. Since we have a numerical response and two different
    groups, this is a _two-sample $t$ test_: 
  
    ```{r}
    library(infer)
    library(tidyverse)
    penguins_ok <- penguins |>
        drop_na() |>
        # Things will work better if we ensure species is a character vector 
        # and not a factor vector
        # 
        # You probably don't need to copy this step since your import is unlikely
        # to 'accidentally' make a factor
        mutate(species = as.character(species)) 
    
    penguins_ok |>
        filter(species != "Chinstrap") |>
        t_test(body_mass ~ species, 
               order = c("Adelie", "Gentoo"))
    ```
  
    From this, we see that Adelie penguins are about 1375 grams _lighter_ on average
    than Gentoos and the $p$-value is far less than 0.01, so the difference is
    unlikely to be solely due to sampling variability. Further more, the 
    `lower_ci` and `upper_ci` columns give us a confidence interval (95% by 
    default, but this can be adjusted by changing the `conf_level` argument to
    `t_test()`.) for the difference in means. 
    
    Here, we perform the test by specifying the `response` (quantity of interest)
    on the left hand side of the `~` and the `explanatory` variable 
    (that which might meaningfully predict a differerence in groups) on the
    right hand side of the `~`. If we swap the order (or don't provide it), the
    sign of the estimated difference may differ, but the actual $p$-value won't
    change for a two-sided test (as done here). 
    
    We can extend this analysis to perform it separately for each of three years
    using a bit of nested data trickery. The mechanics of this are more advanced,
    but you should be able to extend this to your own analysis by just changing
    variable names: 
    
    ```{r}
    penguins_ok |>
    filter(species != "Chinstrap") |>
    group_by(year) |>
    nest() |>
    mutate(p_value_adelie_gentoo = map(data, \(d) t_test(d, 
                                                         body_mass ~ species,
                                                         order=c("Adelie", "Gentoo")))) |>
    unnest(everything())
    ```
    
    Here, we see that our result has columns for the $t$ statistic, 
    the $p$-value, the estimated difference, and many other useful inferential
    quantities. 
    
2.  Example 02: In other contexts, we may have a _categorial_ response. In this 
    case, we should not perform a $t$-test on means, but we instead want to use 
    a binomial proportion test *via* the `prop_test()` function. In the simplest
    case, we might want to examine the fraction of sampled penguins which are
    female: 
    
    ```{r}
    penguins_ok |> 
      mutate(is_female = (sex == "female")) |> 
      prop_test(is_female ~ NULL)
    ```
    
    Note that, here the `statistic` is the test statistic, not the sample
    proportion. Examining the confidence interval, we see that it's quite clear
    the fraction of female penguins is basically 50% (as we would expect). 
    Here, the `~ NULL` of the specification implies that we aren't looking
    at any potential covariates. 
    
    We can extend this to the two-sample setting to see if the frequency
    differs between two groups. For instance, if we want to see whether Adelie 
    penguins are more likely to be female than Gentoos: 
    
    ```{r}
    library(infer)
    library(tidyverse)
    penguins_ok |>
       filter(species != "Chinstrap") |>
       mutate(is_female = (sex == "female")) |> 
       prop_test(is_female ~ species, 
                 order = c("Adelie", "Gentoo"))
    ```
    
    As expected, we see that there is basically no evidence that the fraction of
    female penguins differs by species. Note the key difference between the 
    `t_test()` and the `prop_test()`: in `t_test()` the response is a numeric 
    variable and we are testing the _mean_ of that distribution; in a 
    `prop_test()` the response is a Boolean (`TRUE`/`FALSE`) value and we are 
    testing the probability of observing a `TRUE`.[^prop]
    
    We can also use this with a derived quantity. For instance, if we want to 
    see if Gentoos are more likely to be over 4000 grams than non-Gentoos: 
    
    ```{r}
    penguins |>
        mutate(is_gentoo = (species == "Gentoo"),
               over_4k = body_mass > 4000) |>
        prop_test(over_4k ~ is_gentoo, 
                  order = c("TRUE", "FALSE"))
    ```
    
    Here, we see that there is _strong_ statistical evidence (in the form of a
    tiny $p$-value) that being a Gentoo penguin increases probability of weighing
    over 4000 grams. While this interface does not give us an estimate of 
    
    We can also modify this to be a one-sided test: 
    
    ```{r}
    penguins |>
        mutate(is_gentoo = (species == "Gentoo"),
               over_4k = body_mass > 4000) |>
        prop_test(over_4k ~ is_gentoo, 
                  alternative = "greater",
                  order = c("TRUE", "FALSE"))

    ```
    
    Finally, recall that these values can be pulled out of the test result and
    used in the body of your text using the `pull` function from the `dplyr`
    package.
    
::: {.callout-tip title="Task 4: Statistical Inference"}

Using the tools of the `infer` package, construct confidence intervals for the 
following quantities of interest: 

1. The average number of *gold* medals won by the US at a summer Olympics. 
1. The percent increase in *gold* medals won by the host country vs their
   performance in the prior same-season Olympics. (*E.g.*, make sure to compare
   Russia's performance in the 2014 Sochi Olympics to the 2010 Vancouver Olympics,
   not the 2012 London Olympics.) Here, we use the *percent* increase over the
   prior Olympics to account for the fact that some countries have a higher
   'baseline' medal expectation than others. 
1. The probability that the host country wins a *gold* medal in a new sport 
   (defined as a sport that did not appear in the previous Olympics). 
   
Repeat your analysis for *silver* medals, *bronze* medals, and *total* medals. 

For each of these, you may wish to focus on a relatively more recent subset
of Olympics (say 1960 and later) to ensure your data is relevant to the 2028 
Olympics. This is an instance of a common tension in the analysis of data 
observed at different points in time: you want to focus on recent data points
as they are more similar to the future, but you don't want to focus in too much
and have insufficient data. 

:::

::: {.callout-tip title="Task 5: Combining Confidence Intervals to Forecast US Medal Counts"}

*Combining* confidence intervals is a bit of an art, especially when the
intervals are not independent (as they are not here). We are going to make a 
working assumption that the intervals are independent which will make it a bit
easier to combine things. 

We can forecast Team USA's 2028 Olympic medal count using the following equation:

$$ \text{Medals-2028} \approx \text{USA Baseline} * \text{Host Nation Factor} + 22 * \text{Probability Host Nation Wins} $$

where we used the fact that there will be 22 more medal events in 2028 than there
were in 2024. 

To get a range of predictions for the expression above, we will use a simple
*Monte Carlo* sample. That is, rather than doing the calculations exactly, we
will simply simulate them many times over and use the results of that simulation
to get an interval on the derived quantity. 

For example, if I have two parameters $\alpha, \beta$ with 95% confidence 
intervals $[2, 5]$ and $[-1, 3]$ respectively, I can get a confidence interval
for their product as follows:[^pred_int]

[^pred_int]: You might ask whether we want a *confidence* interval or a 
*prediction* interval here. You are right to do so, but we are omitting this
distinction as it's a far less odious problem than the other statistical sins
we are already committing here. 

```{r}
alpha_ci_lower <- 2
alpha_ci_upper <- 5
beta_ci_lower <- -1
beta_ci_upper <- 3

alpha_ci_mean <- (alpha_ci_lower + alpha_ci_upper) / 2
beta_ci_mean  <- (beta_ci_lower + beta_ci_upper) / 2

alpha_ci_sd <- (alpha_ci_upper - alpha_ci_mean) / qnorm(1 - 0.05/2)
beta_ci_sd  <- (beta_ci_upper - beta_ci_mean) / qnorm(1 - 0.05/2)
```

To wit, we have the confidence interval for $\alpha$ given by: 

```{r}
qnorm(c(0.025, 0.975), mean=alpha_ci_mean, sd=alpha_ci_sd)
```

and similarly for $\beta$. To get a CI for $\alpha\beta$, we then can simply
proceed as: 

```{r}
alpha_samples <- rnorm(1e7, mean=alpha_ci_mean, sd=alpha_ci_sd)
beta_samples  <- rnorm(1e7, mean=beta_ci_mean, sd=beta_ci_sd)

alphabeta_samples <- alpha_samples * beta_samples

alphabeta_interval <- quantile(alphabeta_samples, c(0.025, 0.975))
```

which gives an interval of [`r scales::number(min(alphabeta_interval), 0.01)`, 
`r scales::number(max(alphabeta_interval), 0.01)`]. It's not obvious how we
could have computed this analytically. 

Adapt this approach to get 95% confidence intervals on Team USA's medal counts
in LA 2028. 

:::

[^prop]: Of course, if you haven taken 
[STA 9715](https://michael-weylandt.com/STA9715/), you know that there is
(essentially) only one univariate Boolean distribution (the Bernoulli) and that
the probability of observing a success is equal to the mean of the distribution. 
($X \sim \text{Bernoulli} \implies \E[X] = \P(X = 1)$ and all that). This might 
lead you to ask whether there really is a meaningful difference between a 
$t$-test and a proportion test. These are good questions to ask your STA 
9719 professor.

### Final Deliverable: Team USA Fundraising Fact Sheet

With all of your analyses in hand, it is time to start promoting the LA 2028 
Olympics. Specifically, write a short brief for the LA 2028 Organizing 
Committee forecasting Team USA's performance that the fundraising team can use
to start soliciting donations from potential sponsors. Specifically, you want
to help your fundraising team speak confidently about how well you expect Team USA to perform so that sponsors feel comfortable giving to the cause. (Corporations
always want to pay more to be associated with winners.)


::: {.callout-tip title="Task 6: Write Your Final Deliverable"}

Your forecast report should be no more than two pages in length and should
describe your expected Team USA performance, using all of the work you have done
above. Give your 'three-factor' model, interpret the terms individually, and 
then put them together for your final prediction. Be sure to provide suitable
background for each factor, to write at a level that could be easily interpreted
by a marketing executive at a major corporation, and to include well-formatted
graphics and tables as appropriate. 

:::



{{< include _aidisclosure.qmd >}}


{{< include _echeader.qmd >}}


### Extra Credit Opportunity #01: Relaxing Distributional Assumptions {#ec01}

The `prop_test()` and `t_test()` are classical normal-theory based statistical 
tests. This makes them simple to implement, but at the cost of additional 
assumptions. Assumptions are not necessarily a bad thing - and most statistical
procedures are robust to at least mild violations of assumptions - but for 
two points of extra credit, we can explore ways to relax these assumptions. 

Review the [`infer` documentation](https://infer.tidymodels.org/articles/infer.html)
and implement bootstrap based alternatives to `prop_test()` and `t_test()` in
your analysis pipeline. Compare your parametric and bootstrap based forecasts.
Chapters 7-10 of the [Statistical Inference via Data
Science](https://moderndive.com/v2/index.html) book will give additional useful
background. 

### Extra Credit Opportunity #02: Model Improvement {#ec02}

For up to two points of extra credit, add an additional factor to your forecast
system and compare your forecasts with and without that extra term. Be sure to
do the full work-up (some EDA, a statistical test to get a confidence interval,
inclusion in the forecast model, propagation through to a final confidence
interval). Make sure to justify your additional factor and to explain why you
think it will make your model more accurate. 

### Extra Credit Opportunity #03: Model Validation {#ec03}

Finally, for up to three points, evaluate the performance of your forecast model
_retrospectively_. *E.g.*, if you had used the same model for Japan in 2020, how
well would it have predicted medal outcomes for those Olympics? There 
are many ways you can go with this, but at a minimum, you should 

i) Make at least 10 additional sets of predictions based on historical data
ii) Compare the accuracy of those predictions with the realized values for each
    medal category 
iii) Use that error estimate to put some sort of margin of error on your 2028
     population predictions. 

{{< include _footer.qmd >}}
