---
lab_num: "`r lab <- 8; lab`"
lab_title: "More `ggplot2` - Chloropleths and Cartograms + Tools for Interactive Data Analysis"
engine: knitr
format: live-html
---

{{< include _setup.qmd >}}

{{< include _gradethis_init.qmd >}}

```{webr}
#| echo: false
webr::install(c("tidyverse", "mgcv"), quiet=TRUE)
suppressPackageStartupMessages(library(tidyverse))
```

# Review Practice {#review}

Recall the SSA Birth Count Time Series from 
[earlier in the semester](./lab05.qmd).

You can load this data into `R` using the following code: 


```{webr}
if(!file.exists("births.csv")){
    download.file("https://raw.githubusercontent.com/michaelweylandt/STA9750/main/births.csv", 
                  destfile="births.csv")
}
library(readr)
library(dplyr)

births <- read_csv("births.csv")
glimpse(births)
```

Recall that the columns here are:

- `id` : The day in the entire time series (going up to $\approx 365 * 20$ plus 
   a few for leap days)
- `day_of_year`: the day in the year (1 to 365/366)
- `day_of_week`: the day of the week, coded as an integer
- `day`, `month`, `year`: the parts of the date as we normally think of them
- `births`: the number of births that day. 

We are going to look for structure in this data using our data visualization
tools. 

1. Compute the average number of babies born per day for each year. Is there
   evidence of a long term trend? 

::: {.callout-tip collapse="true" title="Solution"}

```{webr}
births |> 
    group_by(year) |> 
    summarize(avg_births = mean(births)) |> 
    ggplot(aes(x=year, y=avg_births)) + 
        geom_line() +  
        geom_point(size=5) + 
        stat_smooth(se=FALSE) + 
        xlab("Year") + 
        ylab("Average Number of Births per Day") + 
        theme_bw()
```

Yes, following a crash in the early 1970's, the number of births has increased
most years, consistent with a growing national population (more people leading
to more babies *ceteris paribus*).

:::

2. The previous plot may sacrifice too much resolution in taking yearly averages.
   Instead of using yearly averages, use a 90-day moving window to estimate 
   the median number of children born each day.
   
   *Hint: `stats::runmed()` can be used to compute a running median.*
   

::: {.callout-tip collapse="true" title="Solution"}

```{webr}
births |> 
    mutate(med_births = runmed(births, 91)) |> 
    ggplot(aes(x=id, y=med_births)) + 
    geom_line()
```

There are clearly several interesting trends at play in this data, but use of
`id` on the $x$-axis makes this plot a bit harder to comprehend, so let's
put a bit more formatting into it: 

```{webr}
births |>
  mutate(med_births = runmed(births, 91), 
         date = ymd(paste(year, month, day))) |>
  ggplot(aes(x=date)) + 
  geom_point(aes(y=births), alpha=0.05) + 
  geom_line(aes(y=med_births), color="red4") + 
  theme_bw() + 
  xlab("Date") + 
  ylab("Number of Children Born per Day") 
```

Visualizing the individual points, we see quite a lot of variation that is worth
further exploration, but the highest-level trend essentially matches what we
saw above. 

*Note*: If you look closely, you'll see that `runmed` has somewhat 
counterintuitive behavior around the ends. For finer control over moving window
statistics like this, look into the `slider::slide_dbl()` framework. 

:::

3. Use `dplyr` and `ggplot2` to compute and visualize the average number of 
   children born each day of the week. You can use a line graph or a bar graph
   to visualize your results: 
   
::: {.callout-tip collapse="true" title="Solution"}

```{webr}
births |> 
    group_by(day_of_week) |> 
    summarize(avg_births = mean(births)) |> 
    ggplot(aes(x=day_of_week, y=avg_births)) + 
        geom_col(fill="lightblue1", color="blue4", alpha=0.4) + 
        geom_line(col="red4", linewidth=1.5) + 
        geom_point(size=5) + 
        xlab("Day of Week (1 = Monday)") + 
        ylab("Average Number of Births per Day") + 
        theme_bw()
```

Again, we can do a bit better by manually adjusting the labels: 

```{webr}
births |> 
    mutate(date = ymd(paste(year, month, day)), 
           wday = weekdays(date)) |>
    group_by(wday) |> 
    summarize(day_of_week=mean(day_of_week), 
              avg_births = mean(births)) |> 
    arrange(day_of_week) |> 
    mutate(wday = factor(wday, levels=wday, ordered=TRUE)) |> 
    ggplot(aes(x=wday, y=avg_births, group=1)) + 
        geom_col(fill="lightblue1", color="blue4", alpha=0.4) + 
        geom_line(col="red4", linewidth=1.5) + 
        geom_point(size=5) + 
        xlab(NULL) + 
        ylab("Average Number of Births per Day") + 
        theme_bw()
```

:::


4. Use `dplyr` and `ggplot2` to compute and visualize the average number of 
   children born per day for each month. 
   
::: {.callout-tip collapse="true" title="Solution"}

```{webr}
births |> 
    group_by(month) |> 
    summarize(avg_births = mean(births)) |> 
    # Optional: use the built-in month.name to map 
    # 1,2,3 => January, February, March, ...
    mutate(month_name = month.name[month], 
           month_name = factor(month_name, 
                               levels = month.name, 
                               ordered = TRUE)) |>
    ggplot(aes(x=month_name, y=avg_births)) + 
        geom_col(fill="lightblue1", color="blue4", alpha=0.4) + 
        geom_point(size=5) + 
        geom_line(col="red4", linewidth=1.5) +
        xlab(NULL) + 
        ylab("Average Number of Births per Day") + 
        theme_bw()
```

:::

From this analysis, it seems clear that there are actually several different
periodic and trend components to this data set. Extracting them is a fascinating
analysis on its own but beyond this course. See, *e.g.*, 
[this page](https://research.cs.aalto.fi/pml/software/gpstuff/demo_births.shtml)
for details of how you might approach this to produce a plot like:

![GP-Based Decomposition of Births Time Series from Vehtari Group
(Aalto)](https://research.cs.aalto.fi/pml/software/gpstuff/pics/births_pic2.png)

Note that this type of 'multi-seasonal' decomposition is _incredibly_ useful
in business contexts. The [`prophet` package](https://facebook.github.io/prophet/),
an open-source product of the data science teams at Meta is designed for precisely
this sort of analysis. 

# Advanced `ggplot2` Usage

*Instructor's Note: This week, I am providing non-interactive code snippets
and encouraging you to run these examples locally (on your own instance
of `RStudio`). Some of these tools rely on technologies you should have installed
as part of [this week's pre-assignment](../pre/pa08.html), but that are
a bit tricky to perfectly emulate in-browser.*

## Spatial Visualization with `ggplot2` {#maps}

Today, we're going to dive deeper into `ggplot2`, with a focus on
visualizing spatial data. Spatial data can be [quite
complex](https://postgis.net/documentation/manual/), but we can get pretty far
with the **simple features** (`sf`) paradigm. Also, if we focus on relatively 
small regions, we can avoid the complexities that come from Earth not being
planar. 

{{< video https://www.youtube.com/watch?v=UPaca-SrvEk >}}

Key to this approach is a _shapefile_. A _shapefile_, conventionally stored
with a `shp` extension, gives precise coordinates outlining spatial regions. 
By plotting a polygon with those points at its boundaries, we can visualize
spatial regions. 

The `sf` package includes a `shp` file for the counties of North Carolina: 

```{r}
#| warning: false
#| message: false
library(tidyverse)
library(sf)
nc <- read_sf(system.file("shape/nc.shp", package="sf"))
nc
```

We represent this data in "tidy" format, with each row being a county. The
"magic" is in the `geometry` column: 

```{r}
library(dplyr)
nc |> select(NAME, AREA, PERIMETER, geometry)
```

The `geometry` column is of type `MULTIPOLYGON`, essentially a list of GPS
coordinates tracing the corners of each county.

In this activity, we'll mainly just use these for plotting, though it is
possible to do quite sophisticated analyses here. 

`ggplot2`'s `geom_sf` makes it easy to visualize spatial data. It has 
one required aesthetic, fittingly named `geometry`: 

```{r}
#| fig.asp: 0.3
library(ggplot2)
ggplot(nc, aes(geometry = geometry)) + geom_sf()
```

Note that `ggplot2` is doing some work behind the scenes to keep the
aspect ratio reasonable and not "stretching" North Carolina north-south.[^asp]

[^asp]: The _aspect ratio_ of a plot is the ratio of its height to its width. 
While we can change the aspect ratio pretty safely for most plot types, messing
with the aspect ratio of a map leads to some weird visualizations. Unfortunately,
to keep the aspect ratio correct, `ggplot2` and `quarto` work to add some extra
white space around a figure (see example below with default settings).  If you 
want to cut off this white space and have a more professionally-formatted figure,
you will need to adjust the figure width and height using the `fig.width` and 
`fig.height` execution options. Rather than trying to tweak these both, you can
also set `fig.asp` to manually set the aspect ratio. In this case, because NC
runs about 2.5 degrees north-to-south and 8 degrees east-to-west, I have set 
the aspect ratio of most of the figures in this set of notes to 
$0.3 \approx 2.5/8$, which gives a comfortable appearance without too much 
whitespace. 

    ```{r}
ggplot(nc, aes(geometry = geometry)) + 
    geom_sf() + 
    ggtitle("Map of NC with Default Aspect Ratio and Margins")
    ```
    
    Look at all this wasted space!

This is nice enough, if we like maps for maps' sake, but we can do more. The
`BIR74` is the number of children born in that county between July 1, 1974
and June 30, 1978. We can use that to set the _fills_ for each county. 

```{r}
#| fig.asp: 0.3
library(ggplot2)
ggplot(nc, aes(geometry = geometry, 
               fill = BIR74)) + geom_sf()
```

This type of plot is called a **chloropleth** plot, coming from the Greek
for "area-multitude". It is commonly used to visualize _areal_[^areal] geospatial
data - that is quantities associated with spatial regions - as opposed to
point data. 

[^areal]: The word "areal" (area based) is a homonym to the word "arial" (the 
font) or "aerial" (relating to flight or the atmosphere), but don't get them 
confused.

Building a chloropleth plot is not too hard, if you have an appropriate
shape file. Thankfully, governments are in the business of knowing _exactly_
what the boundaries of their territories lie and providing shape files.
Typically, these shape files are distributed in a "zip" format, with other
geodata files that we won't use in this course.

::: {.callout-note title="Adding Points on Chloropleths" collapse=true}

We can add the [City of
Charlotte](https://en.wikipedia.org/wiki/Charlotte,_North_Carolina),
the largest city of North Carolina on this map: 

```{r}
#| fig.asp: 0.3
charlotte <- data.frame(
    x = -80 - 50/60 - 35/60^2, # West = negative
    y = +35 + 13/60 + 38/60^2  # North = positive
)

ggplot() + 
    geom_sf(data=nc, 
            aes(geometry = geometry, 
                fill = BIR74)) + 
    geom_point(data=charlotte, 
               aes(x=x, y=y),
               color="red4")
```

Note that I pass _different_ data sets and `aes` mappings to each layer
here, so I do them inside the individual `geom_`s instead of in the global
`ggplot()` call. 

:::

### Exercises 1 - Geospatial Visualization of MTA Counties {#ex01}

Before completing the following exercises, run the following commands to 
download shape data and census demographics for 12 counties within New York
State that are subject to the Metropolitan Commuter Transportation Mobility
Tax:[^mta_tax]

[^mta_tax]: These are roughly the 12 counties served by the MTA 
(NYC Bus and Subway, MetroNorth, LIRR) and which capture most of the commuting
population of NYC. 

```{r}
#| cache: true
#| warning: false
#| message: false
ACS_TABLES <- tribble(
    ~concept, ~id, 
    "total_population", "B01001_001",
    "total_male_population", "B01001_002",
    "total_female_population", "B01001_026",
    "total_white_population", "B01001A_001", 
    "total_black_population", "B01001B_001",
    "total_american_indian_population", "B01001C_001",
    "total_asian_population", "B01001D_001",
    "total_native_hawaiian_population", "B01001E_001",
    "total_orther_race_population", "B01001F_001",
    "total_two_or_more_population", "B01001G_001",
    "total_hispanic_popoulation", "B01001I_001"
)

library(tidycensus)
options(tigris_use_cache = TRUE)

ACS_DATA <- ACS_TABLES |> 
    pull(id) |> 
    map(\(table_id) get_acs("county", 
                            county=c("Queens", 
                                     "Kings", 
                                     "Bronx", 
                                     "New York", 
                                     "Richmond", 
                                     "Westchester", 
                                     "Rockland", 
                                     "Suffolk", 
                                     "Nassau", 
                                     "Dutchess",
                                     "Putnam",
                                     "Orange"), 
                            state="NY", table_id, year=2020, 
                            geometry=TRUE, cache_table=TRUE, 
                            keep_geo_vars=TRUE)) |>
    bind_rows() |> 
    select(NAME.y, ALAND, AWATER, variable, estimate, geometry) |>
    rename(county = NAME.y, 
           land_area = ALAND, 
           water_area = AWATER, 
           value = estimate) |>
    inner_join(ACS_TABLES, join_by(variable == id)) |> 
    select(-variable) |> 
    pivot_wider(names_from="concept", 
                values_from="value") |>
    mutate(county = str_replace(county, ", New York", ", NY"))

glimpse(ACS_DATA)
```

Here, I have joined together the data in a fashion that should be easy to use
for our visualization exercises.

1. Using `ACS_DATA`, create a simple map of the 11 MTA Commuter counties. 

::: {.callout-tip collapse="true" title="Solution"}

```{r}
#| fig.asp: 0.5
ggplot(ACS_DATA, aes(geometry=geometry)) + geom_sf()
```

:::

2. Using the `fill` aesthetic, create a legend with the name of each county: 

::: {.callout-tip collapse="true" title="Solution"}

```{r}
#| fig.asp: 0.5
ggplot(ACS_DATA, aes(geometry=geometry, fill=county)) + 
  geom_sf()
```

With a bit of polish, we may instead present this as: 

```{r}
#| fig.asp: 0.5
ggplot(ACS_DATA, aes(geometry=geometry, fill=county)) + 
  geom_sf() + theme_bw() + 
  scale_fill_brewer(name="County", type="qual", palette="Paired") + 
  theme(legend.position="bottom")
```

:::

3. Using the `total_population` column, color the counties by population instead.
   
   *Hint: What type of color palette do you want to use here?*
   
::: {.callout-tip collapse="true" title="Solution"}

```{r}
#| fig.asp: 0.5
ggplot(ACS_DATA, aes(geometry=geometry, fill=total_population)) + 
  geom_sf()
```

This can be refined as: 

```{r}
#| fig.asp: 0.5
ggplot(ACS_DATA, aes(geometry=geometry, fill=total_population)) + 
    geom_sf() + 
    theme_bw() + 
    scale_fill_distiller(name="Total Population", 
                         type="seq", 
                         palette="Reds", 
                         direction=1) + # Flip direction of legend
    theme(legend.position="bottom")
```

:::

4. Because NYS loves complex governance structures, the MTA has a 
   21-member board wherein: 
   
   - 5 members are appointed by the Governor of NYS
   - 4 members are appointed by the Mayor of NYC
   - 3 members are appointed by the Executives of Nassau, Suffolk, 
     and Westchester Counties (one per county)
   - 4 members are appointed by the Executives of Dutchess, Orange, Rockland
     and Putnam counties, but they share a single vote
   - The remaining members are appointed by a mixture of customer and labor
     representatives
     
   If we focus on the 11 members appointed by county and city leadership, we
   can add the following column to `ACS_DATA`: 
   
```{r}
ACS_DATA <- ACS_DATA |>
   mutate(effective_mta_vote = case_when(
      county %in% c("Bronx County, NY", 
                    "Kings County, NY", 
                    "Queens County, NY", 
                    "Richmond County, NY", 
                    "New York County, NY") ~ 4/5, 
      county %in% c("Westchester County, NY", 
                    "Suffolk County, NY", 
                    "Nassau County, NY") ~ 1, 
      TRUE ~ 1/4)
   )
```

   Modify your map to color each county by its effective representation on the 
   MTA board. 

::: {.callout-tip collapse="true" title="Solution"}

```{r}
#| fig.asp: 0.5
ggplot(ACS_DATA, aes(geometry=geometry, fill=effective_mta_vote)) + 
    geom_sf() + 
    theme_bw() + 
    scale_fill_distiller(name="Effective MTA Representation", 
                         type="seq", 
                         palette="Blues", 
                         direction=1) + # Flip direction of legend
    theme(legend.position="bottom")
```

:::

5. Under the "one person, one vote"-principle,[^one_vote] we would expect county
   representation on the MTA board to be proportional to the number of residents
   of each county. Divide the effective votes of each county by the total 
   population of that county to get a measure of _representative effectiveness_ 
   and visualize your results. 

[^one_vote]: NYC's baroque governance famously did not adhere to this principle 
until the [Board of Estimate was struck down as a violation of the 14th 
Amendment](https://en.wikipedia.org/wiki/Board_of_Estimate_of_City_of_New_York_v._Morris)
in 1989. Other NYC and NYS structures continue this legacy of irregular 
representation to this day.

::: {.callout-tip collapse="true" title="Solution"}

```{r}
#| fig.asp: 0.5
ACS_DATA |>
    mutate(voter_power = effective_mta_vote / total_population) |>
    ggplot(aes(geometry=geometry, fill=voter_power)) + 
        geom_sf() + 
        theme_bw() + 
        scale_fill_distiller(name="Effective Population-Normalized MTA Representation", 
                             type="seq", 
                             palette="Greens", 
                             direction=1) + # Flip direction of legend
        theme(legend.position="bottom")
```

:::
   
   
6. While this map is interesting, it is more interpretable to instead identify
   which counties have relatively too-much or too-little voting power on the MTA
   board. Compute a "baseline" measure of voter power by dividing the 8 total
   local MTA votes by the total population of the 12-county region. Then divide
   each county's voter power by this baseline to get a sense of its relative
   power. Visualize your result using a suitable color palette. 
   
   *Hint: You may want to subtract 1 from your voter power measure to center 0 
   at "average".*


::: {.callout-tip collapse="true" title="Solution"}

```{r}
#| fig.asp: 0.5
ACS_DATA |>
    mutate(county_voter_power = effective_mta_vote / total_population, 
           baseline_voter_power = sum(effective_mta_vote) / sum(total_population), 
           relative_voter_power = county_voter_power / baseline_voter_power -1) |>
    ggplot(aes(geometry=geometry, fill=relative_voter_power)) + 
        geom_sf() + 
        theme_bw() + 
        scale_fill_fermenter(name="Relative Representation on the MTA Board [Red Regions are Over-Powered]", 
                             type="div", 
                             palette="RdBu", 
                             breaks=c(-3, -1.5, 0, 1.5, 3)) +
        theme(legend.position="bottom", 
              legend.text = element_blank()) # Disable legend ticks
```


:::

Based on this analysis, which county's residents have the most relative power on
on the MTA board? Which have the least? 

Note that this analysis is somewhat limited as it assumes NYC's four votes are
equally split by borough. In practice, the interests of the larger boroughs may
be more strongly represented among mayoral appointees. 

### Limitations and Extensions of Chloropleths

Chloropleth plots are rightly popular but they have a major limitation: 
humans are not uniformly distributed across space! Smaller regions may
have higher populations (*cf.* Connecticut *vs.* Montana) and administrative
regions are not the same size. Because human perception is naturally drawn
to larger regions, we often need to adjust the color schemes used to 
compensate for area effects. This process can be powerful, but it is a bit
difficult to get right. 

For instance, let's see what happens if we create a chloropleth of the
`SID74` data, the number of children who die in North Carolina of 
[SIDS](https://en.wikipedia.org/wiki/SIDS) during our sample period: 

```{r}
#| fig.asp: 0.3
ggplot(nc, aes(geometry = geometry, fill = SID74)) + 
    geom_sf()
```

Here, [Mecklenberg County](https://en.wikipedia.org/wiki/Mecklenburg_County,_North_Carolina)
(Charlotte; bottom-middle of the state) sticks out because it has a large
population, not because SIDS was particularly more common in that county. 
We can modify our plot to show the SIDS rate, rather than raw counts: 

```{r}
#| fig.asp: 0.3
nc |>
    mutate(sids_rate_74 = SID74 / BIR74) |>
    ggplot(aes(geometry = geometry, 
               fill = sids_rate_74)) + 
    geom_sf()
```

We see less variation here, but there is still some. Note that this type
of data -- rate estimation from rare counts -- can be pretty tricky to analyze,
(and thankfully SIDS remains quite rare) but that's not the primary focus of 
this class so we'll leave it here. 

This plot still isn't perfect however: [Sampson
county](https://en.wikipedia.org/wiki/Sampson_County%2C_North_Carolina)
is the largest county in NC by area, but it has a relatively small population. 
That means that the largest area - and hence the place our eyes will most
immediately look - is assigned to a relatively unimportant county. Relatedly, 
[Anson County](https://en.wikipedia.org/wiki/Anson_County%2C_North_Carolina) 
is a county with a small population, but it appears to be a "hot-spot".
This could be true, but it is much more likely a noise-effect resulting from a
small population.

To address these issues, we can use a _cartogram_ which will "adjust" 
the map so that area maps to a relevant quantity. This is quite complex
mathematically, but the `cartogram` package handles it reasonably transparently
for us[^1]: 

```{r}
#| fig.asp: 0.3
library(cartogram)
nc |> 
    mutate(sids_rate_74 = SID74 / BIR74) |>
    st_transform(26916) |> 
    cartogram_cont(weight="BIR74") |> 
    ggplot(aes(geometry=geometry, 
               fill = sids_rate_74)) + 
    geom_sf()
```

Here the `st_transform` function species _what projection_ to equilibrate
area in. I'm not an expert on this, but the internet seems to suggest 26916
is not a bad default.

Looking at this plot, we see that counties are now adjusted to population, 
(or, more precisely, birth count) and that and Anson county-while still high-is
clearly down-weighted in accordance with its population. 

### Exercises 2 - Cartogram Analysis of MTA Counties {#ex02}

1. Modify your plot of the MTA counties so that each county is resized proportional
   to its total population. 
   
::: {.callout-tip collapse="true" title="Solution"}

```{r}
#| fig-asp: 0.5
library(cartogram)
ACS_DATA |>
    st_transform(26916) |> 
    cartogram_cont(weight="total_population") |> 
    ggplot(aes(geometry=geometry, 
               fill = county)) + 
    geom_sf() + 
    scale_fill_brewer(name="County", 
                       type="qual", 
                       palette="Paired") + 
    theme(legend.position="bottom")
```

:::

2. The result of the previous cartogram is a bit janky. Adapt your code to instead
   use a Dorling cartogram (`cartogram_dorling`) which gives a more recognizable 
   cartogram representation in some circumstances. 

::: {.callout-tip collapse="true" title="Solution"}
   
```{r}
#| fig-asp: 0.5
library(cartogram)
ACS_DATA |>
    st_transform(26916) |> 
    cartogram_dorling(weight="total_population") |> 
    ggplot(aes(geometry=geometry, 
               fill = county)) + 
    geom_sf() + 
    scale_fill_brewer(name="County", 
                       type="qual", 
                       palette="Paired") + 
    theme(legend.position="bottom")
``` 
   
:::

3. Apply a Dorling Cartogram to visualize effective population-adjusted MTA 
   board representation of each county. 
   
   Is this graphic more or less interpretable than your previous attempt 
   using true geometries? In particular, how does this visualization balance
   the relative overpowering of Putnam County with its smaller overall
   population?  (That is, the residents of Putnam county may have more voting
   power _per capita_, but if there are very few of them, the net effect of this
   bias might not actually be that large.)

::: {.callout-tip collapse="true" title="Solution"}
```{r}
ACS_DATA |>
    mutate(voter_power = effective_mta_vote / total_population) |>
    st_transform(26916) |> 
    cartogram_dorling(weight="total_population") |> 
    ggplot(aes(geometry=geometry, fill=voter_power)) + 
        geom_sf() + 
        theme_bw() + 
        scale_fill_distiller(name="Effective Population-Normalized MTA Representation", 
                             type="seq", 
                             palette="Greens", 
                             direction=1) + # Flip direction of legend
        theme(legend.position="bottom")
```
:::

### Improving Cartogram Visualizations

As we have seen, creating an attractive cartogram is not entirely trivial and we
can often do better using a visualization that has a bit more of an artistic
touch. One popular alternative - popularized by sites like FiveThirtyEight in
the context of US elections - is a *Hex Map*, which represents political
boundaries as a set of hexagonal tiles. If the hexagonal tiles each represent
an area of roughly equal proportion, we can get to a scenario where area does
indeed match population. Some software attempts to do this automatically (see
the `tilemaps` `R` package or the [`mapshaper`](https://mapshaper.org/) website)
but you will usually get a better result more easily if you can find a hex map 
that has already been developed and published. 

In the context of NYC, the `nychex` (read as `"nyc-hex"` not `"ny-chex"` like 
the snack mix) package provides these at several levels. Note that this
package is **not** available on CRAN and we have to install it from an alternate
source: 

```{r}
#| eval: false
install.packages(
  "nychex",
  repos = c("https://kjhealy.r-universe.dev", "https://cloud.r-project.org")
)
```

Once installed, we can see the various hexmaps and square tiled maps this 
package provides: 

```{r}
#| echo: false
#| cache: true
#| label: nychex_demo
#| warning: false
library(sf)
library(ggplot2)
library(ggiraph)
library(nychex)
library(patchwork)

ggplot(nyc_nta20_hex_sf) +
    geom_sf(aes(fill = boro_name), 
            color = "white", 
            linewidth = 0.3, 
            alpha=0.3) +
    geom_sf_text_interactive(aes(label=nta_abbrev, 
                                 tooltip=nta_name), 
                             size=1.8, 
                             alpha=0) + 
    geom_sf(data=nyc_nta_boros_hex_sf, 
            aes(color=boro_name), 
            fill=NA, 
            linewidth=0.8) + 
    scale_fill_brewer(palette = "Set2") +
    scale_color_brewer(palette="Set2") + 
    labs(fill = "Borough", color="Borough") +
    theme_void() + 
    ggtitle("NYC NTAs - HexMap") -> G1


ggplot(nyc_nta20_sq_sf) +
    geom_sf(aes(fill = boro_name), 
            color = "white", 
            linewidth = 0.3, 
            alpha=0.3) +
    geom_sf_text_interactive(aes(label=nta_abbrev, 
                                 tooltip=nta_name), 
                             size=1.8, 
                             alpha=0) + 
    geom_sf(data=nyc_nta_boros_sq_sf, 
            aes(color=boro_name), 
            fill=NA, 
            linewidth=0.8) + 
    scale_fill_brewer(palette = "Set2") +
    scale_color_brewer(palette="Set2") + 
    labs(fill = "Borough", color="Borough") +
    theme_void() + 
    ggtitle("NYC Square Tiles - HexMap") -> G2


ggplot(nyc_ct20_hex_sf) +
    geom_sf(aes(fill = boro_name), 
            color = "white", 
            linewidth = 0.3, 
            alpha=0.3) +
    geom_sf_text_interactive(aes(label=ct2020, 
                                 tooltip=geoid),
                             alpha=0,
                             size=1.8) + 
    geom_sf(data=nyc_ct_boros_hex_sf,
            aes(color=boro_name), 
            fill=NA, 
            linewidth=0.8) + 
    scale_fill_brewer(palette = "Set2") +
    scale_color_brewer(palette="Set2") + 
    labs(fill = "Borough", color="Borough") +
    theme_void() + 
    ggtitle("NYC Census Tracts - HexMap") -> G3


ggplot(nyc_ct20_sq_sf) +
    geom_sf(aes(fill = boro_name), 
            color = "white", 
            linewidth = 0.3, 
            alpha=0.3) +
    geom_sf_text_interactive(aes(label=ct2020, 
                                 tooltip=geoid), 
                             size=1.8, 
                             alpha=0) + 
    geom_sf(data=nyc_ct_boros_sq_sf, 
            aes(color=boro_name), 
            fill=NA, 
            linewidth=0.8) + 
    scale_fill_brewer(palette = "Set2") +
    scale_color_brewer(palette="Set2") + 
    labs(fill = "Borough", color="Borough") +
    theme_void()  + 
    ggtitle("NYC Census Tracts - Square Tiles") -> G4


G <- G1 + G2 + G3 + G4 + 
    plot_annotation("Comparison of nychex Cartograms", 
                    caption="Mouseover for NTA/CT IDs") + 
    plot_layout(guides="collect", nrow=2) & 
    theme(legend.position="bottom")

Gi <- girafe(ggobj=G, 
             options=list(
                 opts_hover_inv(css="opacity:0.3;"),
                 opts_toolbar(saveaspng=FALSE), 
                 opts_sizing(rescale=TRUE)),
             width_svg=12, 
             height_svg=10
             )

Gi
```

For your use, you will typically want to use either `nyc_nta20_hex_sf` which
provides a hexmap for [NYC's Neighborhood Tabulation Areas](https://www.nyc.gov/content/planning/pages/resources/datasets/neighborhood-tabulation)
or `nyc_ct20_hex_sf` which provides a hexmap for NYC's 2020 Census Tracts. 
Note that NYC's NTAs are formed by aggregating census tracts in a way that roughly
aligns with NYC's more interpretable neighborhood boundaries, so we can get 
NTA-level values from the tract data by suitable aggregation: 

```{r}
#| warning: false
#| message: false
library(tidycensus)
library(nychex)
library(ggplot2)
library(sf)
get_acs(year=2024, 
        geography="tract", 
        variable=c(agg_income="B19025_001", 
                   total_hhs="B19001_001"),
        county=c("New York", 
                 "Richmond", 
                 "Bronx", 
                 "Queens", 
                 "Kings"), 
        state="NY",
        geometry=FALSE) |>
    select(-NAME, 
           -moe) |>
    pivot_wider(names_from=variable, 
                values_from=estimate) |>
    filter(total_hhs > 0) |>
    inner_join(nyc_ct20_hex_sf, 
               join_by(GEOID == geoid)) |>
    # Drop unneeded the CT shapes, since we're going to plot at NTA level
    st_drop_geometry() |> 
    select(total_hhs, 
           agg_income, 
           nta2020) |>
    group_by(nta2020) |>
    summarize(total_hhs = sum(total_hhs), 
              agg_income = sum(agg_income)) |> 
    mutate(hh_income_avg = agg_income / total_hhs) |> 
    inner_join(nyc_nta20_hex_sf, 
               join_by(nta2020 == nta2020)) |> 
    ggplot(aes(geometry=tile_map)) + 
    geom_sf(aes(fill=hh_income_avg), 
            color=NA) + 
    geom_sf(aes(color=boro_name), 
            data=nyc_nta_boros_hex_sf, 
            linewidth=2) + 
    scale_color_brewer(type="qual", 
                       palette=2, 
                       name="Borough") + 
    scale_fill_distiller(type="seq", 
                         palette=2, 
                         direction = 1, 
                         name="Average Household Income", 
                         labels=scales::label_dollar()) + 
    theme_void() + 
    theme(legend.position="bottom", 
          legend.box="vertical") + 
    guides(fill=guide_colorbar(barwidth=unit(10, "cm")))
```

Note that here I used aggregate income and count of households because these
aggregate easily. For other types of statistics (proportions, medians), more
sophisticated aggregation may be required.

[^1]: If you get an error reading "`stat_sf()` requires the following
missing aesthetics: geometry", make sure you include `geometry=geometry`
in your `aes`

## Animated Graphics {#animated}

Animation is a natural and impressive way of adding "pop" to your graphics. 
It's not right for all scenarios, but when it applies, *e.g.* when showing how
something has changed over time, it can be quite magical. 

Some caution before you become too enamored with animated graphics: 

i) Animation isn't well-supported across all platforms. While HTML-based 
   presentations are becoming standard and quite common, you can't always
   assume you will have the option to use them. If you need to embed animation
   inside a PDF, you are at the mercy of the system PDF viewer, which may not be
   entirely reliable.
   
   In general, I recommend designing your presentations to 'fail gracefully', so
   that they still work even if the animation is not displayed properly. 
   Typically, this involves setting things up so the first frame of your 
   animation is a plot showing all of the data; if the animation won't play, you
   still have the ability to show all of the data you intended to show in your
   presentation
ii) Animation is showy, but not great for deep dives into data. Because most 
    animations don't allow control of playback speed, it can be hard to stop
    and focus in on an important frame or two. I tend to use animations only for
    'finished' analyses and prefer interactivity or small-multiples for EDA-type
    work. 
iii) Animation has accessibility limitations that you should generally be 
     aware of. 

### Exercise 3: Animated Graphics {#ex03}

In this section, you will use the `gganimate` extension to `ggplot2` to create 
animated graphics. We will use the famous `gapminder` data set from the 
`gapminder` package. Install the `gganimate`, `gapminder`, `gifski`, and `av`
packages before attempting attempting this set of exercises. 

0. For background, watch Hans Rosling's talk on human prosperity.

{{< video https://www.youtube.com/watch?v=jbkSRLYSojo >}}

1. Load the `gapminder` package and create a static scatter plot of the 
   relationship between GDP and Life Expectancy in the year 1952. Color points
   by continent and use the size aesthetic to represent population.
   
   *Hint: You might want to put some quantities on a log-scale.* 
   
::: {.callout-tip collapse="true" title="Solution"}

```{r}
library(tidyverse)
library(gapminder)
library(gganimate)

gapminder |>
    filter(year == 1952) |> 
    ggplot(aes(y=gdpPercap, x=lifeExp, size=pop, color=continent)) + 
    geom_point() + 
    theme_bw() + 
    ylab("GDP Per Capita (Inflation Adjusted US-Dollars") + 
    xlab("Life Expectancy at Birth (Years)") + 
    theme_bw() + 
    scale_y_log10(labels = scales::dollar) + 
    scale_color_brewer(name="Continent", type="qual", palette=2) + 
    scale_size_continuous(name="Population") + 
    theme(legend.position="bottom") + 
    ggtitle("Global GDP Per Capital and Life Expectancy in 1952")
```

:::
2. There is an outlier country in this data with very high GDP. What is it? 
   Identify and remove it from your data set. 
   
::: {.callout-tip collapse="true" title="Solution"}

```{r}
library(tidyverse)
library(gapminder)

gapminder |>
    filter(year == 1952) |> 
    slice_max(gdpPercap)
```

The outlier is Kuwait. Removing it from our data yields the following: 

```{r}
library(tidyverse)
library(gapminder)

gapminder |>
    filter(year == 1952, 
           country != "Kuwait") |> 
    ggplot(aes(y=gdpPercap, x=lifeExp, size=pop, color=continent)) + 
    geom_point() + 
    theme_bw() + 
    ylab("GDP Per Capita (Inflation Adjusted US-Dollars") + 
    xlab("Life Expectancy at Birth (Years)") + 
    theme_bw() + 
    scale_y_log10(labels = scales::dollar) + 
    scale_color_brewer(name="Continent", type="qual", palette=2) + 
    scale_size_continuous(name="Population") + 
    theme(legend.position="bottom") + 
    ggtitle("Global GDP Per Capital and Life Expectancy in 1952")
```

which is a bit easier to read. 

:::

3. Load the `gganimate` package and use it to create an _animated_ version of 
   this plot. 
   
   A common use for animated plots is showing how data evolve over time and we 
   will do the same here. Specifically, use the `transition_time` function to
   make your plot animated. The `transition_*` functions of `gganimate` work
   similarly to the `facet_*` functionality of base `ggplot2`, but don't require
   a `~` symbol in their specification. 
   
   *Hint: Since we are looking for changes over time, you do not need to `filter`
   down to 1952 any longer.*
   
::: {.callout-tip collapse="true" title="Solution"}

Making this an animated plot only requires us to change two lines: 

```{r}
#| label: gapminder1
#| cache: true
library(tidyverse)
library(gapminder)
library(gganimate)

gapminder |>
    filter(country != "Kuwait") |>  # Don't filter to 1952 any more
    ggplot(aes(y=gdpPercap, x=lifeExp, size=pop, color=continent)) + 
    geom_point() + 
    theme_bw() + 
    ylab("GDP Per Capita (Inflation Adjusted US-Dollars") + 
    xlab("Life Expectancy at Birth (Years)") + 
    theme_bw() + 
    scale_y_log10(labels = scales::dollar) + 
    scale_color_brewer(name="Continent", type="qual", palette=2) + 
    scale_size_continuous(name="Population") + 
    theme(legend.position="bottom") + 
    ggtitle("Global GDP Per Capital and Life Expectancy") + 
    transition_time(year)
```

Specifically, we have to load the `gganimate` package with a `library()` call
and then we can use `gganimate::transition_time()` to indicate that we want
"time animation" according to the `year` variable. (The other `transition_*`
functions provide other sorts of animation, *e.g.* cycling among subgroups, but 
time is the natural place to start for many animated plots.)

:::
   
5. Using the theme machinery, labels, *etc.* make this a "publication ready" 
   plot. In particular, note that when using `transition_time()` you can use
   `{frame_time}` in the title to get a dynamically changing year. 

::: {.callout-tip collapse="true" title="Solution"}

Because we already made our plot look nice earlier, the only thing left to do
here is to tweak the label

```{r}
#| label: gapminder2
#| cache: true
library(tidyverse)
library(gapminder)
library(gganimate)

gapminder |>
    filter(country != "Kuwait") |>  # Don't filter to 1952 any more
    ggplot(aes(y=gdpPercap, x=lifeExp, size=pop, color=continent)) + 
    geom_point() + 
    theme_bw() + 
    ylab("GDP Per Capita (Inflation Adjusted US-Dollars") + 
    xlab("Life Expectancy at Birth (Years)") + 
    theme_bw() + 
    scale_y_log10(labels = scales::dollar) + 
    scale_color_brewer(name="Continent", type="qual", palette=2) + 
    scale_size_continuous(name="Population") + 
    theme(legend.position="bottom") + 
    ggtitle("Global GDP Per Capital and Life Expectancy in {frame_time}") + 
    transition_time(year)
```

:::   
   
6. To perfectly reproduce the visualization from the GapMinder video, let's use 
   Dr. Rosling's preferred color scheme. The `gapminder` package provides this
   data in a vector called `country_colors`. Here we are providing our own color
   scale, so we will need to replace the default `scale_color_discrete` with
   a custom scale. 
   
   *Hint: There is an on-point example in the help page for ?country_colors.*

::: {.callout-tip collapse="true" title="Solution"}

Following `?country_colors`:

```{r}
#| label: gapminder3
#| cache: true
library(gganimate)
library(gapminder)
library(ggplot2)
gapminder |>
    ggplot(
       aes(x=lifeExp, y=gdpPercap, 
           color=country, size=pop)) + 
    geom_point() +
    scale_y_log10(labels=scales::dollar) + 
    scale_size_continuous(transform=scales::transform_log10()) + 
    guides(color="none", 
           size="none") + 
    xlab("Life Expectancy at Birth (Years)") + 
    ylab("GDP Per Capita (US$, Inflation-Adjusted)") + 
    theme_bw() + 
    transition_time(year) + 
    ggtitle("GDP vs Life Expectancy in {frame_time}") +
    labs(caption="Data from the GapMinder Project") +
    scale_color_manual(values=country_colors) + 
    facet_wrap(~continent)
```

When you have a complex data set like this, I sometimes find it valuable to
highlight a few key points to see how they evolve over time. The `ggrepel` 
package provides some nice functionality to add labels in a position that avoids
covering over the underlying data: 

```{r}
#| label: gapminder4
#| warning: false
#| cache: true
library(gganimate)
library(gapminder)
library(ggrepel)
library(ggplot2)
gapminder |>
    mutate(label = case_when(
        country %in% c("United States", "China", "United Kingdom", "Kenya") ~ country, 
        TRUE ~ NA
    )) |>
    ggplot(
       aes(x=lifeExp, y=gdpPercap, 
           color=country, size=pop, 
           label=label)) + 
    geom_point() +
    geom_label_repel() + 
    scale_y_log10(labels=scales::dollar) + 
    scale_size_continuous(transform=scales::transform_log10()) + 
    guides(color="none", 
           size="none") + 
    xlab("Life Expectancy at Birth (Years)") + 
    ylab("GDP Per Capita (US$, Inflation-Adjusted)") + 
    theme_bw() + 
    transition_time(year) + 
    ggtitle("GDP vs Life Expectancy in {frame_time}") +
    labs(caption="Data from the GapMinder Project") +
    scale_color_manual(values=country_colors) + 
    facet_wrap(~continent)
```

A few more tweaks could be used here since the labels are a bit jumpy,
but it's pretty good.

:::

```{r}
#| echo: false
## There's a weird interaction with how quarto and gganimate pick up directories
## when echoing `qmd` files in a sub-folder. I'm trying to get to the bottom
## of it, but for now, I'll just move the rendered `gif`s into place manually. 
move_to_labs_folder <- function(fname){
    file.rename(fname, file.path("labs", fname))
}

walk(Sys.glob("lab08_files/**/gapminder*"), move_to_labs_folder)
```

## Interactive Graphics {#interactive}

The `ggiraph` package can be used to add simple interactivity to any 
`ggplot2`-based visualization.[^plotly] Using `ggiraph` requires 
two basic steps: 

1) Add some _interactive_ `geom_()`s to your plot. `ggiraph` provides 
   `_interactive` variants of almost all of the basic `ggplot2` `geom` types. 
   For example, if I want to make the points in my scatter plot interactive, 
   I will switch from a simple `geom_point()` to a `geom_point_interactive()`. 
2) Passing the assembled `ggplot2` object to `girafe` (*NB*, not spelled like
   the long-necked African animal). This is a bit clunky, since we need to save
   the `ggplot` object to a variable and then pass it as a named argument to
   `girafe`. 
   
For example, we may start with a basic plot of our penguins: 

```{r}
library(ggplot2)

penguins |> 
  drop_na() |> 
  ggplot(aes(x=flipper_len, 
             y=body_mass, 
             color=species)) + 
  geom_point() + 
  theme_bw() + 
  xlab("Flipper Length (mm)") + 
  ylab("Body Mass (g)") + 
  scale_color_brewer(name="Species", 
                     palette=2, 
                     type="qual") + 
  theme(legend.position="bottom")
```

To add interactivity, I will swap to a `geom_point_interactive` plot, which 
allows a `tooltip` aesthetic that specifies the message to be shown when the 
mouse passes over a point. 

```{r}
library(ggplot2)
library(ggiraph)
library(glue)

penguins |> 
  drop_na() |> 
  mutate(sex1 = if_else(sex == "male", "M", "F"), 
         tooltip = glue("{species} Penguin ({sex1}) Observed on {island} Island in {year}\nBill Length (mm): {bill_len}\nBill Depth (mm): {bill_dep}")
  ) |>
  ggplot(aes(x=flipper_len, 
             y=body_mass, 
             color=species, 
             tooltip=tooltip)) + 
  geom_point_interactive() + 
  theme_bw() + 
  xlab("Flipper Length (mm)") + 
  ylab("Body Mass (g)") + 
  scale_color_brewer(name="Species", 
                     palette=2, 
                     type="qual") + 
  theme(legend.position="bottom")
```

On its own, this does nothing (as seen above). To make this interactive, we
work as follows: 

```{r}
library(ggplot2)
library(ggiraph)
library(glue)

penguins |> 
  drop_na() |> 
  mutate(sex1 = if_else(sex == "male", "M", "F"), 
         tooltip = glue("{species} Penguin ({sex1}) Observed on {island} Island in {year}\nBill Length (mm): {bill_len}\nBill Depth (mm): {bill_dep}")
  ) |>
  ggplot(aes(x=flipper_len, 
             y=body_mass, 
             color=species, 
             tooltip=tooltip)) + 
  geom_point_interactive() + 
  theme_bw() + 
  xlab("Flipper Length (mm)") + 
  ylab("Body Mass (g)") + 
  scale_color_brewer(name="Species", 
                     palette=2, 
                     type="qual") + 
  theme(legend.position="bottom") -> G_penguins

girafe(ggobj=G_penguins)
```

We see that, first we had to save our results to a variable `G_penguins`. I 
find this a bit easier to do with the "right-assign" idiom which puts the 
variable name after the object is built, but most folks prefer putting 
`G_penguins <- ` at the start of the pipeline. 

Then, once our plot object is created, we make it interactive by passing it as
the `ggobj` of `girafe`. As with all `ggplot2` package, we finally see the object
by printing it, which is done here implicitly. 

### Linking Across Multiple Panels

The `patchwork` package can be used to combine multiple `ggplot` objects
as separate panels in a single figure (see my `nychex` example above). It also 
plays particularly nicely with `ggiraph` and allows "connected" interactivity
across panels. Specifically, if we use the `data_id` aesthetic across multiple
panels, the corresponding element will be highlighted across each panel: 

```{r}
library(tidyverse)
library(ggiraph)
library(patchwork)

penguins_id <- penguins |> 
    drop_na() |> 
    mutate(id = row_number(), 
           tooltip=glue("Penguin #{id}\n Observed on {island} Island in {year}"))

G_BASIC <- ggplot(penguins_id, 
                  aes(data_id=id, 
                      color=species, 
                      tooltip=tooltip)) + 
    theme_bw()  + 
    scale_color_brewer(type="qual", palette=2, name="Species") + 
    theme(legend.position="bottom")

G1 <- G_BASIC + 
    geom_point_interactive(
          aes(x=body_mass, 
              y=flipper_len), 
          hover_nearest=TRUE) + 
    xlab("Body Mass (g)") + 
    ylab("Flipper Length (mm)")

G2 <- G_BASIC + 
    geom_point_interactive(
        aes(x=bill_dep, 
            y=bill_len), 
        hover_nearest=TRUE) + 
    xlab("Bill Depth (mm)") + 
    ylab("Bill Length (mm)")

# Align our two panels, combine the guides, and move the legend to the bottom
G <- G1 + G2 + plot_layout(guides="collect") & theme(legend.position="bottom")

girafe(ggobj=G,
       # Extra interactive formatting to make our highlighting even more clear
       options=list(
           opts_hover(css = "fill:cyan;stroke:black;cursor:pointer;"),
           opts_hover_inv(css="opacity:0.3;"),
           opts_selection(type = "single", css = "fill:red;stroke:black;")
       ))

```

The `ggiraph` package is quite powerful and has many additional features beyond
those demonstrated here.


[^plotly]: The `plotly` `R` package can do something similar with its 
`plotly::ggplotly()` function, but we're going to focus on `ggiraph` here 
since it has functionality that allows integration with `shiny`. This is beyond
our course, but you can see a nice example by running 
`ggiraph::run_girafe_example()` on your local machine.

### Exercise 4 - Adding Simple Interactivity with `ggiraph` {#ex04}

1. Use `ggiraph::geom_sf_interactive()` to add a tooltip to your MTA governance 
   map from above. 

::: {.callout-tip collapse="true" title="Solution"}

We can use something like the following: 

```{r}
library(ggiraph)
library(glue)

G_MTA <- ACS_DATA |>
    mutate(county_voter_power = effective_mta_vote / total_population, 
           baseline_voter_power = sum(effective_mta_vote) / sum(total_population), 
           relative_voter_power = county_voter_power / baseline_voter_power - 1, 
           pop_fmt = scales::comma(total_population),
           cvp_e6 = scales::number(county_voter_power * 1e6),
           vp_rank = rank(desc(county_voter_power)),
           tooltip = glue("{county}\n {effective_mta_vote} MTA Votes for {pop_fmt} Residents\n {cvp_e6} MTA Board Seats per Million Residents (Power Ranking: {vp_rank})")) |>
    ggplot(aes(geometry=geometry, 
               fill=relative_voter_power, 
               tooltip=tooltip)) + 
    geom_sf_interactive() + 
    theme_bw() + 
    scale_fill_fermenter(name="Relative Representation on the MTA Board [Red Regions are Over-Powered]", 
                         type="div", 
                         palette="RdBu", 
                         breaks=c(-3, -1.5, 0, 1.5, 3)) +
    theme(legend.position="bottom", 
          legend.text = element_blank()) # Disable legend ticks

girafe(ggobj=G_MTA)
```

:::

2. Use `patchwork` and the `data_id` element to combine several barplots of the
   ACS data, allowing us to compare different subpopulations (male, female, white
   non-white, *etc*) across the MTA commuter counties. 
   
::: {.callout-tip collapse="true" title="Solution"}
   
```{r}
G1 <- ACS_DATA |>
    ggplot(aes(geometry=geometry, 
               fill=total_population, 
               tooltip=county, 
               data_id=county)) + 
    geom_sf_interactive() + 
    theme_bw() + 
    scale_fill_fermenter(type="seq", 
                         palette="Reds", 
                         name="Total Population", 
                         direction=1) + 
    theme(legend.position="bottom") + 
    guides(fill="none")

G2 <- ACS_DATA |> 
    arrange(desc(-total_population)) |>
    mutate(county=factor(county, levels=county)) |> 
    ggplot(aes(x=county, 
               y=total_population, 
               fill=total_population,
               tooltip=county, 
               data_id=county)) + 
    geom_col_interactive() + 
    theme_bw() + 
    coord_flip() + 
    scale_fill_fermenter(type="seq", 
                     palette="Reds", 
                     name="Total Population", 
                     direction=1) + 
    xlab(NULL) + 
    ylab(NULL) + 
    ggtitle("Total Population") + 
    guides(fill="none")

G3 <- ACS_DATA |> 
    arrange(desc(-total_male_population)) |>
    mutate(county=factor(county, levels=county)) |> 
    ggplot(aes(x=county, 
               y=total_male_population, 
               fill=total_population,
               tooltip=county, 
               data_id=county)) + 
    geom_col_interactive() + 
    theme_bw() + 
    coord_flip() + 
    scale_fill_fermenter(type="seq", 
                         palette="Reds", 
                         direction=1) + 
    xlab(NULL) + 
    ylab(NULL) + 
    ggtitle("Total Male Population") + 
    guides(fill="none") + 
    theme(axis.text.y = element_blank(), 
          axis.ticks.y = element_blank())


G4 <- ACS_DATA |> 
    arrange(desc(-total_female_population)) |>
    mutate(county=factor(county, levels=county)) |> 
    ggplot(aes(x=county, 
               y=total_female_population, 
               fill=total_population,
               tooltip=county, 
               data_id=county)) + 
    geom_col_interactive() + 
    theme_bw() + 
    coord_flip() + 
    scale_fill_fermenter(type="seq", 
                         palette="Reds", 
                         direction=1) + 
    xlab(NULL) + 
    ylab(NULL) + 
    ggtitle("Total Female Population") + 
    guides(fill="none") + 
    theme(axis.text.y = element_blank(), 
          axis.ticks.y = element_blank())


G5 <- ACS_DATA |> 
    arrange(desc(-total_white_population)) |>
    mutate(county=factor(county, levels=county)) |> 
    ggplot(aes(x=county, 
               y=total_white_population, 
               fill=total_population,
               tooltip=county, 
               data_id=county)) + 
    geom_col_interactive() + 
    theme_bw() + 
    coord_flip() + 
    scale_fill_fermenter(type="seq", 
                         palette="Reds", 
                         direction=1) + 
    xlab(NULL) + 
    ylab(NULL) + 
    ggtitle("Total White Population") + 
    guides(fill="none") + 
    theme(axis.text.y = element_blank(), 
          axis.ticks.y = element_blank())


G6 <- ACS_DATA |> 
    arrange(desc(-(total_population - total_white_population))) |>
    mutate(county=factor(county, levels=county)) |> 
    ggplot(aes(x=county, 
               y=total_population - total_white_population, 
               fill=total_population,
               tooltip=county, 
               data_id=county)) + 
    geom_col_interactive() + 
    theme_bw() + 
    coord_flip() + 
    scale_fill_fermenter(type="seq", 
                         palette="Reds", 
                         direction=1) + 
    xlab(NULL) + 
    ylab(NULL) + 
    ggtitle("Total Non-White Population") + 
    guides(fill="none") + 
    theme(axis.text.y = element_blank(), 
          axis.ticks.y = element_blank())

G <- (G1 + G2 + plot_layout(widths=c(2, 1))) / 
     (G3 + G4) / 
     (G5 + G6) + plot_layout(guides="collect", heights = c(3, 2, 2))

girafe(ggobj=G) |> 
    girafe_options(
        opts_hover(css = "fill:cyan;stroke:black;cursor:pointer;"),
        opts_selection(type = "single", css = "fill:red;stroke:black;")
    )
```

Note that this example is _highly_ redundant and the code could be made shorter
since many graph elements are shared.

:::

## `shiny`: Creating Interactive Tools for Data Analysis

We finally turn our attention to creating interactive tools (web applications and
dashboards) for data analysis. By learning these tools, you won't necessarily
enable yourself to do analysis, but you will enable colleagues and clients to
more fully explore the results of your analyses. The main tool we will use to
do is `shiny`, a **reactive** framework for `R`. 

"Reactive" programming refers to an architecture that changes only in response
to user input. You have almost certainly worked with a reactive piece of
software in the form of Microsoft Excel (or Google Sheets or similar). The
values in the spreadsheet do not change *sua sponte*, but when you make an input
all related cells are updated. (Conceptually, all cells are updated, but
Excel is smart and tracks only those cells that actually need to be updated.)

In this week's pre-assignment, you have already read a bit about `shiny`, but
now we will put it into practice. In this exercise, we will not focus on making
things particularly attractive. If you want to improve upon the rather Spartan
appearance of standard `shiny`, tools like 
[the `flexdashboard` package](https://pkgs.rstudio.com/flexdashboard/) may help.


### Exercise 5 - Building a `shiny` tool for {{< var course.short >}} {#ex05}

Use the [course helper functions](../tips.html) to build a `shiny` tool that
can be used to verify that your mini-projects and peer feedback have been 
appropriately formatted and submitted. 

#### Step 0: Run `shiny`

Take the following prototype and confirm that it runs on your computer. Note
that this is the "single-file" format for `shiny` apps. It is somewhat more
common to split `shiny` apps across two files - backend and frontend - but
we are using a single file structure here for convenience. 

```{r}
#| eval: false
if(!require("shiny")) install.packages("shiny")
library(shiny)

if(!require("glue")) install.packages("glue")
library(glue)

if(!require("yaml")) install.packages("yaml")
library(yaml)

# Create global variables here
variables <- read_yaml("https://raw.githubusercontent.com/michaelweylandt/STA9750/refs/heads/main/_variables.yml")
course_repo  <- variables$course$repo
course_short <- variables$course$short

# Define the UI
ui <- bootstrapPage(
    textInput('github_id', 
              'What is your GitHub ID?', 
              placeholder="Type your GitHub ID here"),
    htmlOutput('gh_link')
)

# Define the server code
server <- function(input, output) {
    output$gh_link <- renderText({
        markdown(glue("My code for this course found on [GitHub](https://github.com/{input$github_id}/{course_repo})"))
    })
}

# Return a Shiny app object
shinyApp(ui = ui, server = server)
```

#### Step 1: Add a Link to your profile

Add a second output to link to the homepage you created in
[MP#00](../mini/mini00.html). Copying the treatment of `gh_link` in 
the example, you will need to 

1) Add a second output
   [`renderText`](https://shiny.posit.co/r/reference/shiny/0.11/rendertext.html)
   component in the server
2) Add a second output [`htmlOutput`](https://shiny.posit.co/r/reference/shiny/0.11/htmloutput.html) 
  component in the UI

#### Step 2: Parameterize the Mini-Project Number

Next, extend your app to handle the mini-projects. 
Modify the above example to do three things: 

- Add a [`numericInput`](https://rstudio.github.io/shiny/reference/numericInput.html) 
  that takes a number between `1` and `4` inclusive to select a mini-project.
- Add an output [`htmlOutput`](https://shiny.posit.co/r/reference/shiny/0.11/htmloutput.html) 
  element that adds a link to the relevant 
  [mini-project instructions](https://michael-weylandt.com/STA9750/miniprojects.html). 
 
  The `glue` function can be used to easily substitute values in URLs (as in
  the example). 
- Add an output [`htmlOutput`](https://shiny.posit.co/r/reference/shiny/0.11/htmloutput.html) 
  element that adds a link to *your* mini-project submission. This URL should
  take both the `numericInput` and the `textInput` user inputs to construct
  the URL. (That is, if you change the `GitHub` name used the URL should point
  to a different repo)
  
#### Step 3: Confirm Mini-Project Submission

Take one of the course helper scripts and use it to determine whether a given 
MP submission is properly formatted. To do so: 

1) Load the helpers by adding `source("https://michael-weylandt.com/STA9750/load_helpers.R")`
   to the top of your file (outside both the UI and the server). 
2) Add a reactive input that runs `mp_submission_verify` and captures the output. 
   The following code snippet will be useful:
   
```{r}
#| eval: false
O <- capture.output({
    E <- tryCatch(mp_submission_verify(N, github_id), 
                  error=function(e) e)
})

if(isTRUE(E)){
    MESSAGE <- "Mini-Project properly submitted."
} else {
    MESSAGE <- "Mini-Project not properly submitted. Issue appears to be:\n"
    MESSAGE <- paste(c(MESSAGE, O[1:2]), collapse="\n")
}
```

   where you use `MESSAGE` in creating your output. 

#### Step 4: Extensions

Time allowing, do one or more of the following: 

- Adapt `mp_feedback_locate` to provide links to all peer feedback issues. 
  Note that you will need to handle the case of later MPs, who haven't been
  assigned yet (*i.e.*, MPs #03 and #04). 
- Adapt `mp_feedback_verify` to check whether your peer feedback has been
  properly formatted. This will require a new input for the peer's GitHub ID. 
