---
lab_num: "`r lab <- 6; lab`"
lab_title: "Thoughts on Plots"
engine: knitr
format: live-html
---

{{< include _gradethis_init.qmd >}}

```{webr}
#| echo: false
webr::install(c("tidyverse", "mgcv"), quiet=TRUE)
suppressPackageStartupMessages(library(tidyverse))
```

{{< include _setup.qmd >}}

`r make_slide_link(lab)`

# Review Practice {#review}


The [Carbon Dioxide Information and Analysis Center](https://www.esd.ornl.gov/iab/iab2-15.htm) studies the effect of carbon
dioxide on global and local temperature trends. A key tool in their analysis is
the temperature "anomaly". An anomaly is the difference between observed
temperature (in a world with anthropogenic atmospheric CO2) and 'natural'
temperature (from a world without anthropogenic gases). Note that these anomalies
require significant analysis to compute and are not "simple observational" data.

Politicians have adopted the tools of temperature anomaly to set national and international emissions targets, *e.g.*, the [2 Degree
Target](https://en.wikipedia.org/wiki/2_degree_climate_target). 
Note that 2 degrees is calculated as a _global_ average: in practice, some regions
will experience a much larger change in temperature, while others may experience
a smaller change or even a negative change. 

The [`CVXR`](https://www.cvxgrp.org/CVXR/) package includes the `cdiac` dataset, 
capturing CDIAC's estimated global temperature anomalies from 1850 to 2015. In 
this question, you will explore these estimated anomalies. Note that you may
want to install the `CVXR` package before attempting this question on your
personal computer.[^1] Because CVXR does not work reliably through this online
interface, I have also posted a copy of this data [on the course website](../cdiac.csv).

```{r}
#| eval: false
#| echo: true
install.packages("CVXR")
```

```{webr}
#| warning: false
#| message: false
library(tidyverse)
if(!file.exists("cdiac.csv")){
    download.file("https://raw.githubusercontent.com/michaelweylandt/STA9750/main/cdiac.csv", 
                  destfile="cdiac.csv")
}
cdiac <- read_csv("cdiac.csv")
glimpse(cdiac)
```

As you can see, this data has `r NCOL(CVXR::cdiac)` columns, with 12 monthly
anomalies and 1 annual average anomaly, for a period covering 
`r NROW(CVXR::cdiac)` years from `r min(CVXR::cdiac$year)` to 
`r max(CVXR::cdiac$year)`. Use this data to answer the following questions: 

1. In what year was the highest annual anomaly observed? The lowest?

::: {.callout-tip collapse="true" title="Solution"}
We can select the row with the highest and lowest values using the `slice_max`
and `slice_min` functions from the `dplyr` package. If we want to specifically
get the `year` column after selecting a single row, we can use the `pull` 
function to _pull out_ that one column into a vector: 

```{webr}
cdiac |> slice_max(annual) |> pull(year)

cdiac |> slice_min(annual) |> pull(year)
```

:::

2. For how many months was 2015 the highest anomaly recorded?

   *Hint: It will be helpful to `pivot_longer` this data into a longer format.*
   *Note that you can use the `:` operator when specifying a set of columns to pivot*
   
   ```{r}
   #| eval: false
   library(tidyverse)
   library(CVXR)
   data(cdiac)
   cdiac |> 
     select(-annual) |> 
     pivot_longer(cols=jan:dec, names_to="month", values_to="anomaly")
   ```

::: {.callout-tip collapse="true" title="Solution"}

```{webr}
cdiac |> 
    select(-annual) |> 
    pivot_longer(cols=jan:dec, names_to="month", values_to="anomaly") |> 
    group_by(month) |> 
    slice_max(anomaly) |> 
    ungroup() |> 
    summarize(n_2015s = sum(year == 2015))
```
   
:::

3. For how many years did July have the largest anomaly of that year?

   *Hint: You should first pivot the data as in the previous question.*
   
::: {.callout-tip collapse="true" title="Solution"}

```{webr}
cdiac |> 
    select(-annual) |> 
    pivot_longer(cols=jan:dec, names_to="month", values_to="anomaly") |> 
    group_by(year) |> 
    slice_max(anomaly) |> 
    ungroup() |> 
    summarize(n_july = sum(month == "jul"))
```
   
:::

4. What is the probability that a given July exhibits a greater (more positive)
   anomaly than the annual average anomaly for that year?
   
   *Hint: Remember that probabilities can be estimated by constructing a Boolean
   (`TRUE`/`FALSE`) vector and taking its mean.*

::: {.callout-tip collapse="true" title="Solution"}

```{webr}
cdiac |> 
    pivot_longer(cols=jan:dec, names_to="month", values_to="anomaly") |> 
    filter(month == "jul") |> 
    mutate(july_greater = anomaly > annual) |>
    summarize(p_july_greater = mean(july_greater))
```
   
:::

5. Which month has the highest probability of exceeding the annual anomaly in
   a given year?
   
::: {.callout-tip collapse="true" title="Solution"}

```{webr}
cdiac |> 
    pivot_longer(cols=jan:dec, names_to="month", values_to="anomaly") |> 
    group_by(month) |>
    mutate(monthly_greater = anomaly > annual) |>
    summarize(p_greater = mean(monthly_greater)) |>
    slice_max(p_greater) |>
    pull(month)
```

:::

# Plotting with `ggplot2`

This week, we're going to break into project groups and do three `ggplot2` 
exercises of increasing difficulty. As you work through these with your 
teammates, be sure to reflect on what plots and what tools you will need to best
present your mini-project and course project findings. 

## Exercise 1: Basic `ggplot2` {#practice01}

In this exercise, you will create `ggplot2` graphics to analyze the `diamonds`
data from the `ggplot2` package. This data contains pricing and measurements
for over 50,000 diamonds sold in the US. (Note that these prices are rather 
out of date and pre-date the popularity of lab-grown diamonds.) Before beginning
this exercise, you might want to read about the ["4 C's of
Diamonds"](https://www.tiffany.com/engagement/the-tiffany-guide-to-diamonds/) 
commonly used to measure quality.

1. Make a scatter plot of price vs carat. 

::: {.callout-tip collapse="true" title="Solution"}

```{webr}
library(ggplot2)
ggplot(diamonds, aes(x=carat, y=price)) + geom_point()
```

:::

2. Use `stat_smooth` to see how the price-carat relationship changes by color. 

::: {.callout-tip collapse="true" title="Solution"}

```{webr}
library(ggplot2)
ggplot(diamonds, aes(x=carat, y=price)) + 
    geom_point() + 
    stat_smooth()
```

:::

3. Take your scatter plot from the previous question and iterate it to 
   "publication quality". 
   
::: {.callout-tip collapse="true" title="Solution"}

```{webr}
library(ggplot2)
ggplot(diamonds, aes(x=carat, y=price)) + 
    # Make points slightly transparent so plot is less busy
    geom_point(alpha=0.3) + 
    xlab("Weight of Round Cut Diamond (Carat)") + 
    ylab("Price (US Dollars)") + 
    labs(title="Price as a function of Carat Weight", 
         caption="Data from the ggplot2 R package.") + 
    stat_smooth(se=FALSE, 
                color="red4") +
    # Format y axis as dollars
    scale_y_continuous(labels=scales::dollar) + 
    theme_bw()
```

If you want to require your trend-line to be monotone increasing, see [this StackOverflow answer](https://stackoverflow.com/a/56559086).

:::

4. Further refine your plot by adding a two-way small-multiples design for
   color and cut. 
   
::: {.callout-tip collapse="true" title="Solution"}

```{webr}
library(ggplot2)
ggplot(diamonds, aes(x=carat, y=price)) + 
    geom_point(alpha=0.3) + 
    xlab("Weight of Round Cut Diamond (Carat)") + 
    ylab("Price (US Dollars)") + 
    labs(title="Price as a function of Carat Weight", 
         caption="Data from the ggplot2 R package.") + 
    stat_smooth(se=FALSE, 
                color="red4") +
    scale_y_continuous(labels=scales::dollar) + 
    theme_bw() + 
    facet_grid(cut ~ color)
```

:::

5. Create a [frequency polygon](https://ggplot2.tidyverse.org/reference/geom_histogram.html) 
plot of price, broken out by different diamond cuts.

   *Hint: Before trying this new `geom_`, make sure to read the documentation 
   and identify required aesthetics.*
   
::: {.callout-tip collapse="true" title="Solution"}

```{webr}
ggplot(diamonds, 
       aes(x=price, color=cut)) + 
    # Increase bin count from default 30 to show more detail
    geom_freqpoly(bins=100) + 
    # Log scale to spread-out x-axis
    scale_x_log10(labels=scales::dollar) + 
    xlab("Price (USD)") + 
    theme_bw() + 
    theme(legend.position="bottom") + 
    # Sequential palette since colors are ordered from worst to best
    scale_color_brewer(name="Color", type="seq", palette=2) + 
    ylab(NULL) # Remove y label since it's not interpretable for histograms
```

:::

6. Create a scatter plot of color by clarity. Why is this plot not useful? 

::: {.callout-tip collapse="true" title="Solution"}

```{webr}
ggplot(diamonds, aes(x=color, y=clarity)) + geom_point()
```

This plot is not useful because both the $x$- and $y$-axis are _discrete_
(categorical) values, resulting in significant overplotting. Alternative
visualizations should be used when considering data of this type.

:::
   
## Exercise 2: Trend Analysis with `ggplot2` {#practice02}

Let's return to the `cdiac` data discussed above. In this set of exercises,
you will use `geom_line` to create _line plots_ visualizing trends in this 
data. 

1. Use `geom_line` to plot the estimated annual global mean temperature (GMT)
   anomaly from 1850 to 2015. 
   
::: {.callout-tip collapse="true" title="Solution"}

```{webr}
ggplot(cdiac, aes(x=year, y=annual)) + 
    geom_point()
```

:::
   
2. Plot the GMT anomaly for each month on the same plot as different lines. 

   *Hint: You will need to use a `pivot_longer` to get the data into the
   appropriate shape. `ggplot2` expects one "data point" per row.*

::: {.callout-tip collapse="true" title="Solution"}

```{webr}
cdiac |> 
    select(-annual) |> 
    pivot_longer(cols=jan:dec, 
                 names_to="month", 
                 values_to="anomaly") |> 
    ggplot(aes(x=year, 
               y=anomaly, 
               color=month)) + 
    geom_line() + 
    theme_bw() + 
    xlab("Year") + 
    ylab("GMT Anomaly [K]") + 
    labs(caption="Data from CDIAC via CVXR") + 
    theme(legend.position="bottom") + 
    # Specify palette from name (compare to ColorBrewer site)
    scale_color_brewer(name="Month", type="qual", palette="Paired")
```

This solution has a problem, however: the months appear in _alphabetical_ order
rather than their actual order. To address this, we need to transform month
into an **factor** type with the levels in a specific order. This is a hair 
beyond the scope of this course, but I'll show you how to do so using the
`forcats` package (`tidyverse` tools for `factors`) here: 

```{webr}
library(forcats)
cdiac |> 
    select(-annual) |> 
    pivot_longer(cols=jan:dec, 
                 names_to="month", 
                 values_to="anomaly") |> 
    mutate(month = fct(month)) |> # Easy here since already in correct order
    # Optional to convert jan => January, etc.
    mutate(month = fct_relabel(month, 
                               \(x) month.name[which(tolower(month.abb) == x)])) |>
    ggplot(aes(x=year, 
               y=anomaly, 
               color=month)) + 
    geom_line() + 
    theme_bw() + 
    xlab("Year") + 
    ylab("GMT Anomaly [K]") + 
    labs(caption="Data from CDIAC via CVXR") + 
    theme(legend.position="bottom") + 
    # Specify palette from name (compare to ColorBrewer site)
    scale_color_brewer(name="Month", type="qual", palette="Paired")
```

When dealing with this sort of 'spaghetti' (overlapping lines) plot,
one of my favorite strategies is to plot some sort of _average_ in bolder line
and to plot the individual entries that go into that average semi-transparently.
For instance: 

```{webr}
cdiac |> 
    pivot_longer(jan:dec, 
                 names_to="month", 
                 values_to="anomaly") |>
    mutate(month=factor(month, 
                        ordered=TRUE, 
                        levels=tolower(month.abb), 
                        labels=month.name)) |>
    ggplot(aes(x=year)) + 
        geom_line(aes(y=anomaly, color=month), alpha=0.2) + 
        geom_line(aes(y=annual), linewidth=1.5, color="red4") + 
    theme(legend.position="bottom") + 
    scale_color_discrete(name="Month") + 
    xlab("Year") + 
    ylab("GMT Anomaly (Pre-Industrial Baseline) [K]") + 
    theme_bw() + 
    theme(legend.position="bottom") + 
    guides(color=guide_legend(override.aes=list(alpha=1)))
```

This lets us see both the overall trend and use the month-to-month variability
as a proxy for sampling variability. 

:::

3. Plot the monthly GMT anomaly series as one long line (with a point for 
   each month).
   
   *Hint: Use the `ym` function from the `lubridate` package to combine the
   year and month into a single Date object. Something like
   `time=lubridate::ym(paste(year, month))` should suffice.*
   
::: {.callout-tip collapse="true" title="Solution"}

```{webr}
cdiac |> 
    select(-annual) |> 
    pivot_longer(cols=jan:dec, 
                 names_to="month", 
                 values_to="anomaly") |> 
    mutate(time = ym(paste(year, month))) |> 
    ggplot(aes(x=time, 
               y=anomaly)) + 
    geom_line() + 
    theme_bw() + 
    xlab("Year") + 
    ylab("GMT Anomaly [K]") + 
    labs(caption="Data from CDIAC via CVXR")
```

:::

4. Now let's focus on July: use the `runmed()` function to add a second series
   to the data frame containing the running median July GMT anomaly over the 
   previous 5 years. Plot this and the raw July data on the same plot. Is 
   there evidence of an increasing warming trend? 
   
::: {.callout-tip collapse="true" title="Solution"}

```{webr}
cdiac |> 
    select(year, jul) |> 
    mutate(trend = runmed(jul, 5)) |> 
    pivot_longer(cols=c(jul, trend), 
                 names_to="series", 
                 values_to="anomaly") |> 
    mutate(series = case_when(
        series == "jul" ~ "Annual July Anomaly", 
        series == "trend" ~ "Running 5-year Median of July Anomaly"
    )) |>
    ggplot(aes(x=year, 
               y=anomaly, 
               color=series)) + 
    geom_point() + 
    geom_line() + 
    theme_bw() + 
    xlab("Year") + 
    ylab("GMT Anomaly [K]") + 
    labs(caption="Data from CDIAC via CVXR") + 
    scale_color_brewer(name=NULL, type="qual", palette=2) + 
    theme(legend.position="bottom")
```

Since the overlay of the individual years over the trend is a bit noisy, 
we can adjust the `alpha` aesthetic (transparency) to make the individual
years a bit less visible. 

```{webr}
cdiac |> 
    select(year, jul) |> 
    mutate(trend = runmed(jul, 5)) |> 
    pivot_longer(cols=c(jul, trend), 
                 names_to="series", 
                 values_to="anomaly") |> 
    mutate(
        alpha = if_else(series == "jul", 0.3, 1), 
        series = case_when(
            series == "jul" ~ "Annual July Anomaly", 
            series == "trend" ~ "Running 5-year Median of July Anomaly"
    )) |>
    ggplot(aes(x=year, 
               y=anomaly, 
               color=series, 
               alpha = alpha)) + 
    geom_point() + 
    geom_line() + 
    theme_bw() + 
    xlab("Year") + 
    ylab("GMT Anomaly [K]") + 
    labs(caption="Data from CDIAC via CVXR") + 
    scale_color_brewer(name=NULL, type="qual", palette=2) + 
    scale_alpha_identity() + 
    theme(legend.position="bottom")
```

:::

5. For each year, identify the warmest month (as measured by GMT anomaly).
   Create a bar char showing the probability a given month was the 
   hottest (greatest anomaly) in its year. 
   
   *Hint: Make sure your $x$-axis is in reasonable (chronological) order, 
   not alphabetical. The following code snippet may help:*
   
   ```{r}
   #| eval: false
   mutate(month = factor(month, 
                          levels=str_to_lower(month.abb), 
                          labels=month.name,
                          ordered=TRUE))
   ```
   This takes a column called `month` and marks it as an 
   `ordered` categorical variable (`factor`), with order given
   by `str_to_lower(month.abb)` (`month.abb` is a built-in variable
   in `R`, but it is capitalized, unlike the `cdiac` variables) and
   names to be printed from `month.name`.

::: {.callout-tip collapse="true" title="Solution"}

```{webr}
cdiac |> 
    select(-annual) |> 
    pivot_longer(jan:dec, 
                 names_to="month", 
                 values_to="anomaly") |> 
    group_by(year) |> 
    slice_max(anomaly) |> 
    ungroup() |> 
    count(month) |> 
    mutate(prob_month = n / sum(n), 
           month = factor(month, 
                          levels=str_to_lower(month.abb), 
                          labels=month.name,
                          ordered=TRUE)) |>
    ggplot(aes(x=month, y=prob_month)) + 
    geom_col() + 
    theme_bw() + 
    xlab(NULL) + # Don't need to label - obvious
    ylab("Probability of Hottest Month in Year") + 
    scale_y_continuous(labels=scales::percent) + 
    # Tweak formatting for legibility
    theme(axis.text.x = element_text(angle=90, 
                                     vjust=0.5, 
                                     hjust=1,
                                     size=rel(1.2))) 
```

:::
