---
title: "STA 9750 Week 5 In-Class Activity: Let us `JOIN` Our Tables Together"
engine: knitr
---

```{r include=FALSE}
if(!require("tidyverse")) install.packages("tidyverse")
if(!require("nycflights13")) install.packages("nycflights13")
library(tidyverse)
library(nycflights13)
```

# Welcome!

[Slides](../slides/slides05.html)

# Review Practice

The file ["births.csv"](https://github.com/michaelweylandt/STA9750/blob/main/births.csv) 
in the course repository contains daily US birth counts for 20 years from 1969
to 1988, obtained from the US Social Security Administration. Download this file
and read it into `R` and answer the following review questions with your group. 

The following code may be useful: 

```{r, eval=FALSE, echo=TRUE}
if(!file.exists("births.csv")){
    download.file("https://raw.githubusercontent.com/michaelweylandt/STA9750/main/births.csv", 
                  destfile="births.csv")
}
library(readr)
library(dplyr)

births <- read_csv("births.csv")
glimpse(births)
```

The columns here are:

- `id` : The day in the entire time series (going up to $\approx 365 * 20$ plus a few for leap day)
- `day_of_year`: the day in the year (1 to 365/366)
- `day_of_week`: the day of the week, coded as an integer
- `day`, `month`, `year`: the parts of the date as we normally think of them
- `births`: the number of births that day. 


1. How many children were born on January 1st, 1984? 

::: {.callout-tip collapse="true" title="Solution"}

```{r, eval=FALSE}
births |> filter(day==1, month==1, year==1984)
```

:::

2. How many total children were born in 1984? 

::: {.callout-tip collapse="true" title="Solution"}

```{r, eval=FALSE}
births |> filter(year==1984) |> summarize(sum(births))
```

:::

3. How many children were born each year? (Print a 20 row table)

::: {.callout-tip collapse="true" title="Solution"}

```{r, eval=FALSE}
births |> group_by(year) |> summarize(n_births = sum(births))
```

:::

4. How many more children were born each year than the preceeding? 
(The `lag` function will be useful here!)

::: {.callout-tip collapse="true" title="Solution"}

```{r, eval=FALSE}
births |> 
    group_by(year) |>
    summarize(n_births = sum(births)) |>
    mutate(increase_births = n_births - lag(n_births))
```

:::

5. On average, in what month are the most children born? 

::: {.callout-tip collapse="true" title="Solution"}

```{r, eval=FALSE}
births |> 
    group_by(month) |>
    summarize(avg_births = mean(births)) |>
    slice_max(avg_births)
```

:::

After completing these, work with your group to formulate and answer three more
advanced questions with your group.

# Multi Table Operations

This week, we are going to dive into the most useful "multi-table" `dplyr`
operations, the `join` family. We will focus on the "big three" joins: 

- `inner_join`
- `full_join`
- `left_join`

These inspired by the `SQL` joins, `INNER JOIN`, `FULL [OUTER] JOIN`, and
`LEFT JOIN`.[^1] We will apply them to the various tables in the
`nycflights13` data set. Recall the structure of each of these tables: 

```{r}
library(nycflights13)
glimpse(flights)
```

```{r}
glimpse(airlines)
```

```{r}
glimpse(airports)
```

```{r}
glimpse(planes)
```

```{r}
glimpse(weather)
```

From here, we can see that there are many relationships between these tables. For
example, the `origin` and `dest` columns of the `flights` table, representing
the origin and destination airport respectively, both correspond to the FAA
identifiers used by the `faa` column of the `airports` table. These "commonalities"
form the basis of join specifications. 

## Join Specifications

`dplyr` specifies joins using the `join_by` function. The output of the `join_by`
function, also known as a "join specification" is a series of logical tests
applied to pairs of rows. The results of these logical tests are used to identify
"matches" between rows. Joins differ primarily on how they use the outputs of
these logical tests to construct their output. 

The simplest and most useful logical test to use in a join is an equality test. 
In `dplyr`, these are simply written as

```{r, eval=FALSE}
join_by(left_name == right_name)
```

This type of test checks whether the value in the `left_name` column of the 
first (left) argument matches the value in the `right_name` column of the second
(right) argument. 

For example, if I wanted to join the `origin` column of `flights` table to the
`faa` column of the `airports` table, I might use something like the following: 

```{r eval=TRUE}
inner_join(flights, airports, join_by(origin == faa))
```

Here `origin` is taken to be a column from the first (left) table and `faa`
is taken to be a column from the second (right) table. As with other `dplyr`
functions, there is a bit of programming magic used to allow column names
to be used as variables and interpreted correctly. 

For the airport identifiers, we only need to match on the single unique ID. (We
can assume the FAA assigns unique IDs to each airport.) In other circumstances,
we need to combine several logical tests to get a true match. 

For example, suppose we want to align our flights with the weather at their 
origin airport at scheduled take off time. Here, we'd need to combine the `flights`
and `weather` table on _many_ columns: 

- `origin` to `origin`
- `year` to `year`
- `month` to `month`
- `day` to `day`
- `hour` to `hour`

In this case, we'd pass 5 equality conditions to `join_by`: 

```{r eval=TRUE}
inner_join(flights, 
           weather, 
           join_by(origin == origin,
                   year == year,
                   month == month,
                   day == day,
                   hour == hour))
```

Here we look only at those rows which match on _all 5_ tests. In this way,
`join_by` behaves like `filter`: it "passes" the _intersection_ of positive
results. 

Note that it is relatively common for matched columns to have the same name in
both tables: to support this case, `dplyr` reads a single column name as "self-equality". So the above code can be more concisely written as: 

```{r eval=TRUE}
inner_join(flights, 
           weather, 
           join_by(origin, 
                   year, 
                   month, 
                   day, 
                   hour))
```

I recommend _against_ using this short-cut. It takes hardly more time to write
your intent explicitly and it's far more robust. _Measure twice, cut once_. 

Unfortunately, it is not easy to perform an "OR" in `join_by`. We may cover this
below, time allowing. 

We now turn to specific joins. All of these joins use the `join_by` operator
but they construct results differently based on its output.

## Inner Joins

The most common and most important join in data analysis is the `inner_join`.
The inner join returns _matches_ between two tables. Conceptually, the inner join
constructs all possible pairs of rows between the two tables (so 
`{r eval=FALSE} NROW(x) * NROW(y)` total rows) and then filters down to 
those which pass the `join_by` test. In practice, more efficient algorithms 
are used to prevent wasteful computation. 

Inner joins are used when seeking matches between two tables. They are particularly
useful when both tables are "comprehensive" and we are sure that there are matches. 
For instance, we can use an `inner_join` to combine most of the tables in
`nycflights13` because they come from a comprehensive government data source.
(*E.g.*, No flights going to secret "unauthorized" airports.)

Let's start by asking what the _average_ arrival delay of flights going to
west coast airports is. We do not have enough information to answer this using
the `flights` table alone. To identify west coast airports, let's filter
`airports` on `tzone`: 

```{r}
west_coast_airports <- airports |> filter(tzone == "America/Los_Angeles")
```

We can now join this to the original flights table to find only those flights
with destination matches in `west_coast_airports`:

```{r}
inner_join(flights, west_coast_airports, join_by(dest == faa))
```

Here, we have only a subset of our original `flights` table. From this, we
can compute our relevant summary statistic: 

```{r}
inner_join(flights, west_coast_airports, join_by(dest == faa)) |>
    summarize(mean(arr_delay, na.rm=TRUE))
```

Is this any better than the following alternative approach: 

```{r}
inner_join(flights, airports, join_by(dest == faa)) |>
    filter(tzone == "America/Los_Angeles") |>
    drop_na() |>
    summarize(mean(arr_delay, na.rm=TRUE))
```

Formally, these are basically equivalent. (`filter` and `inner_join` _commute_). 
As usual, it's a matter of _communicating intent_. Here the single line 
`filter(tzone == "America/Los_Angeles")` is simple enough it probably doesn't
need a separate variable. But if, instead of a one line operation, we performed
a very complex set of filtering options, we may benefit from giving it a separate
name as opposed to trying to shoe-horn the complex filtering into a pipeline. 

Performance-wise, it is a bit better to perform `filter` before `inner_join` 
(Why? Think about the size of the result of each step.) but the difference is
rarely material. _Clarity of intent_, not _optimizing performance_, should dictate
the order in which you perform steps. 

Both approaches are also equivalent to: 

```{r}
inner_join(flights, 
           airports |> filter(tzone == "America/Los_Angeles"), 
           join_by(dest == faa)) |>
    drop_na() |>
    summarize(mean(arr_delay, na.rm=TRUE))
```

But I find this sort of "`filter` inside `join` argument" to be terribly
difficult to read: it mixes standard (inside-out) and piped (left to right)
evaluation orders in a confusing manner. Avoid this!

Work with your group to answer the following questions using `inner_join`. 

1. What is the name of the airline with the longest average departure delay?
2. What is the name of the origin airport with the longest average departure delay?
3. What is the name of the destination airport with the longest average departure delay?
4. Are average delays longer for East-coast destinations or West-coast destinations?
5. Which plane (`tailnum`) flew the most times leaving NYC? Who manufactured it?
6. Which manufacturer has the most planes flying out of NYC airports? 
7. Which manufacturer has the longest average flight? 
8. What model of plane has the smallest average delay leaving NYC?

## Left Join

Left joins are useful when you don't want to dropped unmatched columns
in one table. For instance, suppose we misplace some rows from our
`airlines` table: 

```{r}
airlines_major <- airlines |>
    filter(carrier %in% c("AA", "DL", "UA", "WN", "B6", "AS"))
```

If we inner join on `airlines_major`, we _loose_ many of the rows
in `flights`. 

```{r}
NROW(flights)
inner_join(flights, 
           airlines_major, 
           join_by(carrier == carrier)) |>
    NROW()
```

Sometimes this is what we want, but not always. If we instead use
a left join, we keep all of the rows in `flights`: 

```{r}
NROW(flights)
left_join(flights, 
          airlines_major, 
          join_by(carrier == carrier)) |>
    NROW()
```

Rows lacking a pair in `airlines_major` fill the missing columns with
`NA`. This fits our mental model of missing values in `R`: in theory,
these flights should have some carrier name, but given the data at hand, 
we don't know what it is. 

```{r}
NROW(flights)
left_join(flights, 
          airlines_major, 
          join_by(carrier == carrier)) |>
    filter(carrier %in% c("MQ", "OO", "VX")) |>
    glimpse() # Look at 'name' column
```

`left_join`s are useful if we want to join two tables, but want to avoid
dropping any rows from a 'gold standard' table.

## Outer Join

## Advanced Join Specifications

# Multi-Column Operations

# Analytical Functions

## Cumulative Operators

## `*_rank`

# Other `dplyr` functionality

## Advanced Joins

### `cross_join`

### `semi_join`

### `anti_join`

### `nest_join`

## `bind_rows` and `bind_columns`


[^1]: Note that some `SQL` engines use `LEFT OUTER JOIN` than `LEFT JOIN`. 
Because `OUTER` is a bit ambiguous, `dplyr` emphasizes `full_` vs `left_` in
its function naming. Also note the convention of `dplyr` names - lower case, 
underscore separated - and that it differs from `SQL` syntax. 

