---
title: "{{< var course.short >}} Week 2 In-Class Activity: Getting Down with Markdown"
format: 
  html:
    code-link: true
---

# [Slides](../slides/slides02.qmd)

# Welcome!

In this week's in-class activities, we are going to practice
the use of the [`quarto`](http://quarto.org) rendering
engine. 

## Quarto

### Lab Activity: Part 0

If you haven't already, [install Quarto](https://quarto.org/docs/get-started/).

### Lab Activity: Part 1

Create a 5 slide presentation showing the Houston housing market. This should include: 

- A title slide
- Three body slides, each containing a figure and some text
- A conclusion slide

Before attempting to make your slides, run the following two commands at the
`R` console: 

```{r}
#| eval: false
if(!require("tidyverse")) install.packages("tidyverse")
if(!require("quarto")) install.packages("quarto")
```

These will make sure all necessary packages are installed.

::: {.callout-warning title="Make Sure you are in Source Mode"}

When you generate your `qmd` document, make sure that you are using the `Source`
editor mode. Near the top right corner of your file, there will be two buttons
that say `Source` and `Visual`. You want to make sure `Source` is selected; 
`Visual` mode is dangerous and a bit counterproductive.

:::

Recall that in `Quarto`, each new slide is separated by a `##` (level two) header.

You may use the following code snippets to generate relevant figures.
Recall that this code needs to be between three
[back-ticks](https://en.wikipedia.org/wiki/Backtick) 
(the symbol  typically found above the `tab` key) on each end 
and start with `{r}` curly braces as well. 

You should read about 
[Quarto output options](https://quarto.org/docs/computations/execution-options.html#output-options)
to make sure the figure is displayed while hiding the code in your slides.
 
Snippet 1: 

```{r}
#| eval: false
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
txhousing |> 
  filter(city=="Houston") |> 
  group_by(year) |> 
  summarize(sales=sum(sales)) |> 
  ggplot(aes(x=year, y=sales)) + 
    geom_line() + 
    ggtitle("Annual Houses Sold in Houston, TX")
```

Snippet 2: 

```{r}
#| eval: false
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
txhousing |> 
   filter(city=="Houston") |> 
    group_by(month) |> 
    summarize(avg_price=sum(volume) / sum(sales)) |> 
    mutate(month=factor(month.abb[month], 
                 levels=month.abb, ordered=TRUE)) |>
    ggplot(aes(x=month, y=avg_price)) + 
    geom_bar(stat="identity") + 
    ggtitle("Average Price of Houses Sold in Texas by Month") + 
    xlab("Month") + 
    ylab("Average Sale Price") + 
    scale_y_continuous(labels = scales::dollar)
```

Snippet 3: 

```{r}
#| eval: false
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
txhousing |> filter(year==2015) |> 
    group_by(city) |> 
    summarize(avg_price=sum(volume) / sum(sales),
              num_sales=sum(sales)) |> 
    slice_max(num_sales, n=10) |>
    ggplot(aes(x=city, y=avg_price)) + 
    geom_bar(stat="identity") + 
    ggtitle("Average Price of Houses Sold in 2015 by City in Texas") + 
    xlab("City") + 
    ylab("Average Sale Price") + 
    scale_y_continuous(labels = scales::dollar)
```

### Lab Activity: Part 2

Review the 
[Quarto Demo Slides](https://quarto.org/docs/presentations/revealjs/demo/#/title-slide)
and add one new element to your slides from the previous section. 


### Lab Activity: Part 3 (Optional)

Create a simple PDF quarto document using the RStudio wizard. 

(Note that you may need to install `tinytex` for this to work
properly, but Quarto should install it for you automatically.)

We will not generate PDF documents in this course, so don't
worry if you can't get it working in class. This is just
intended to demonstrate the ability of Quarto to render
to non-HTML outputs. 

### Learning More

The official Quarto documentation provides three
helpful tutorials: 

- ["Hello, Quarto!"](https://quarto.org/docs/get-started/hello/rstudio.html)
- [Computations in Quarto](https://quarto.org/docs/get-started/computations/rstudio.html)
- [Authoring in Quarto](https://quarto.org/docs/get-started/authoring/rstudio.html)

These provide additional detail on using Quarto.
