---
title: "{{< var course.short >}} Week 8 In-Class Activity: More `ggplot2` + Tools for Interactive Data Analysis"
engine: knitr
---

[Week 8 Slides](../slides/slides08.html)

# Spatial Visualization with `ggplot2`

Today, we're going to dive deeper into `ggplot2`, with a focus on
visualizing spatial data. Spatial data can be [quite
complex](https://postgis.net/), but we can get pretty far with the 
**simple features** (`sf`) paradigm. Also, if we focus on relatively small
regions, we can avoid the complexities that come from Earth not being
planar. 

{{< video https://www.youtube.com/watch?v=vVX-PrBRtTY >}}

Key to this approach is a _shapefile_. A _shapefile_, conventionally stored
with a `shp` extension, gives precise coordinates outlining spatial regions. 
By plotting a polygon with those points at its boundaries, we can visualize
spatial regions. 

The `sf` package includes a `shp` file for the counties of North Carolina: 

```{r}
#| warning: false
library(sf)
nc <- read_sf(system.file("shape/nc.shp", package="sf"))
nc
```

We represent this data in "tidy" format, with each row being a county. The
"magic" is in the `geometry` column: 

```{r}
library(dplyr)
nc |> select(NAME, AREA, PERIMETER, geometry)
```

The `geometry` column is of type `MULTIPOLYGON`, essentially a list of GPS
coordinates. 

In this course, we'll mainly just use these for plotting, though it is
possible to do quite sophisticated analyses here. 

`ggplot2`'s `geom_sf` makes it easy to visualize spatial data. It has 
one required aesthetic, fittingly named `geometry`: 

```{r}
library(ggplot2)
ggplot(nc, aes(geometry = geometry)) + geom_sf()
```

Note that `ggplot2` is doing some work behind the scenes to keep the
aspect ratio reasonable and not "stretching" North Carolina north-south. 

This is nice enough, if we like maps for maps' sake, but we can do more. The
`BIR74` is the number of children born in that county between July 1, 1974
and June 30, 1978. We can use that to set the _fills_ for each county. 

```{r}
library(ggplot2)
ggplot(nc, aes(geometry = geometry, 
               fill = BIR74)) + geom_sf()
```

This type of plot is called a **chloropleth** plot, coming from the Greek
for "area-multitude". It is commonly used to visualize _areal_ geospatial
data - that is quantities associated with spatial regions - as opposed to
point data. 

Building a chloropleth plot is not too hard, if you have an appropriate
shape file. Thankfully, governments are in the business of knowing _exactly_
what the boundaries of their territories are and providing shape files.
Typically, these shape files are distributed in a "zip" format, with other
geodata files that we won't use in this course.[^3]

::: {.callout-note title="Adding Points on Chloropleths" collapse=true}

We can add the [City of
Charlotte](https://en.wikipedia.org/wiki/Charlotte,_North_Carolina),
the largest city of North Carolina on this map: 

```{r}
charlotte <- data.frame(
    x = -80 - 50/60 - 35/60^2, # West = negative
    y = +35 + 13/60 + 38/60^2  # North = positive
)

ggplot() + 
    geom_sf(data=nc, 
            aes(geometry = geometry, 
                fill = BIR74)) + 
    geom_point(data=charlotte, 
               aes(x=x, y=y),
               color="red4")
```

Note that I pass _different_ data sets and `aes` mappings to each layer
here, so I do them inside the individual `geom_`s instead of in the global
`ggplot()` call. 

:::

## Exercises 1 - Geospatial Visualizations

1. Download the [*New York City Council
   Districts (Clipped to Shoreline)*](https://www.nyc.gov/site/planning/data-maps/open-data/districts-download-metadata.page)
   shapefiles. Unzip the download and read the shape file, `nycc.shp`, using the `sf::read_sf` function. 
   
   Use `ggplot2::geom_sf` to create a basic map of NYC's city council
   districts.  This should just be a plot of the outlines of each
   district (recognizable as NYC, but otherwise showing no data).
2. Using the [`nyc_demos.csv`](./nyc_demos.csv) data file, 
   create a *chloropleth* map of NYC, where the color variable
   represents the 2010 under 5-year-old population of each district. 
   
   Note that you will need to use some `dplyr` `*_join` functions
   to combine the shape data with the demographic data before passing
   it to `ggplot2`.[^1] Which parts of NYC have the fewest young children?
   Does this seem right?
3. Under the "one person one vote" principle, the number of adult
   residents in each council district should be roughly equal. Create a
   chloropleth plot showing how many more/fewer adult residents each
   district has than the average district. Are any districts
   (significantly) over/under-represented on the NY city council?
   
   To do this, you will need to first 1) determine the number of voting age
   residents in each district; 2) compute the average number of voting age
   residents per district; 3) see how each district compares to the average
   (above/below and by how much).
4. Create a *facet plot* where each facet is a chloropleth 
   indicating the *percentage* of residents in each district identifying 
   as a member of each census-designated racial categories.[^2] Note that
   the `nyc_demos.csv` file contains *total counts* by race, so you will
   need to normalize by overall population to get percentages. 

   To get the data into suitable form for plotting, you will need to
   *pivot* the data into a longer format using the `pivot_longer`
   function from the `tidyr` package.
5. Create a visualization to compare the ratio of rental vs owner-owned
   housing per district with the age demographics of this district. What
   do you find? Is this what you would have expected?
   
## Limitations and Extensions of Chloropleths

Chloropleth plots are rightly popular but they have a major limitation: 
humans are not uniformly distributed across space! Smaller regions may
have higher populations (*cf* Connecticut vs Montana) and administrative
regions are not the same size. Because human perception is naturally drawn
to larger regions, we often need to adjust the color schemes used to 
compensate for area effects. This process can be powerful, but it is a bit
difficult to get right. 

For instance, let's see what happens if we create a chloropleth of the
`SID74` data, the number of children who die of SIDS during our sample
period in North Carolina: 

```{r}
ggplot(nc, aes(geometry = geometry, fill = SID74)) + 
    geom_sf()
```

Here, Mecklenberg County (middle, bottom) sticks out because it has a large
population, not because SIDS was particularly more common in that county. 
We can modify our plot to show the SIDS rate, rather than raw counts: 

```{r}
nc |>
    mutate(sids_rate_74 = SID74 / BIR74) |>
    ggplot(aes(geometry = geometry, 
               fill = sids_rate_74)) + 
    geom_sf()
```

We see less variation here, but there is still some. Note that this type
of data -- rate estimation from rare counts -- can be somewhat tricky to analyze, but that's not the primary focus of this class so we'll leave it
here. 

This plot still isn't perfect however: [Sampson county](https://en.wikipedia.org/wiki/Sampson_County%2C_North_Carolina)
is the largest county in NC, but it has a relatively small population. That
means that the largest area - and hence the place our eyes will most
immediately look - is assigned to a relatively unimportant county.
Relatedly, [Anson County](https://en.wikipedia.org/wiki/Anson_County%2C_North_Carolina) is a small population county, but it appears to be a "hot-spot".
This could be true, but it is more likely a noise-effect resulting from a
small population.

To address these issues, we can use a _cartogram_ which will "adjust" 
the map so that area maps to a relevant quantity. This is quite complex
mathematically, but the `cartogram` package handles it reasonably transparently for us: 

```{r}
library(cartogram)
nc |> 
    mutate(sids_rate_74 = SID74 / BIR74) |>
    st_transform(26916) |> 
    cartogram_cont(weight="BIR74") |> 
    ggplot(aes(geometry=geometry, 
               fill = sids_rate_74)) + 
    geom_sf()
```

Here the `st_transform` function species _what projection_ to equilibrate
area in. I'm not an expert on this, but the internet seems to suggest 26916
is not a bad default.

Looking at this plot, we see that counties are now adjusted to population, 
or really birth count and Anson county - while still high - is clearly
down-weighted in accordance with its population. 

## Exercises 2 - Cartograms of NYC

1. Modify your plot of NYC to adjust city council districts by area. 

2. Next, modify your analysis to use the Dorling cartogram
   (`cartogram_dorling`) which gives a more regular cartogram representation.
   
[^1]: If you get an error reading "`stat_sf()` requires the following
missing aesthetics: geometry", make sure you include `geometry=geometry`
in your `aes`   (5 points)

[^2]: These are `c("White Hispanic", "Black Hispanic", "Asian and Pacific Islander Nonhispanic", "Other Nonhispanic", "Two or More Races Nonhispanic", "Hispanic Origin")`. Note that the census has changed these over time and the categories for the 2020 and 2030 censuses will be different.

[^3]: In [MP#03](../miniprojects/mini03.html), you will learn 
how to extract the `shp` file from a `zip` archive automatically, 
but for now you can do so "by hand" by opening the zip file as
if it were a regular file. 


# `shiny`: Creating Interactive Tools for Data Analysis

We now turn our attention to creating interactive tools (web applications and
dashboards) for data analysis. By learning these tools, you won't necessarily
enable yourself to do analysis, but you will enable colleagues and clients to
more fully explore the results of your analyses. The main tool we will use to
do is `shiny`, a **reactive** framework for `R`. 

"Reactive" programmingrefers to an architecture that changes only in response
to user input. You have almost certainly worked with a reactive piece of
software in the form of Microsoft Excel (or Google Sheets or similar). The
values in the spreadsheet do not change *sua sponte*, but when you make an input
all related cells are updated. (Conceptually, all cells are updated, but
Excel is smart and tracks only those cells that actually need to be updated.)

## Exercise 3 - Building a `shiny` tool for {{< var course.short >}}

Use the [course helper functions](../tips.html) to build a `shiny` tool that
can be used to verify that your mini-projects and peer feedback have been 
appropriately formatted and submitted. 

### Step 0: Run `shiny`

Take the following prototype and confirm that it runs on your computer. Note
that this is the "single-file" format for `shiny` apps. It is somewhat more
common to split `shiny` apps across two files - backend and frontend - but
we are using a single file structure here for convenience. 

```{r}
#| eval: false
if(!require("shiny")) install.packages("shiny")
library(shiny)

if(!require("glue")) install.packages("glue")
library(glue)

if(!require("yaml")) install.packages("yaml")
library(yaml)


# Create global variables here
variables <- read_yaml("https://raw.githubusercontent.com/michaelweylandt/STA9750/refs/heads/main/_variables.yml")
course_repo  <- variables$course$repo
course_short <- variables$course$short

# Define the UI
ui <- bootstrapPage(
    textInput('github_id', 
              'What is your GitHub ID?', 
              placeholder="Type your GitHub ID here"),
    htmlOutput('gh_link')
)

# Define the server code
server <- function(input, output) {
    output$gh_link <- renderText({
        markdown(glue("My code for this course found on [GitHub](https://github.com/{input$github_id}/{course_repo})"))
    })
}

# Return a Shiny app object
shinyApp(ui = ui, server = server)
```

### Step 1: Add a Link to your profile

Add a second output to link to the homepage you created in
[MP#00](../miniprojects/mini00.html). Copying the treatment of `gh_link` in 
the example, you will need to 

1) Add a second output
   [`renderText`](https://shiny.posit.co/r/reference/shiny/0.11/rendertext.html)
   component in the server
2) Add a second output [`htmlOutput`](https://shiny.posit.co/r/reference/shiny/0.11/htmloutput.html) 
  component in the UI

### Step 2: Parameterize the Mini-Project Number

Next, extend your app to handle the mini-projects. 
Modify the above example to do three things: 

- Add a [`numericInput`](https://rstudio.github.io/shiny/reference/numericInput.html) 
  that takes a number between `1` and `4` inclusive to select a mini-project.
- Add an output [`htmlOutput`](https://shiny.posit.co/r/reference/shiny/0.11/htmloutput.html) 
  element that adds a link to the relevant 
  [mini-project instructions](https://michael-weylandt.com/STA9750/miniprojects.html). 
 
  The `glue` function can be used to easily substitute values in URLs (as in
  the example). 
- Add an output [`htmlOutput`](https://shiny.posit.co/r/reference/shiny/0.11/htmloutput.html) 
  element that adds a link to *your* mini-project submission. This URL should
  take both the `numericInput` and the `textInput` user inputs to construct
  the URL. (That is, if you change the `GitHub` name used the URL should point
  to a different repo)
  
### Step 3: Confirm Mini-Project Submission

Take one of the course helper scripts and use it to determine whether a given 
MP submission is properly formatted. To do so: 

1) Load the helpers by adding `source("https://michael-weylandt.com/STA9750/load_helpers.R")`
   to the top of your file (outside both the UI and the server). 
2) Add a reactive input that runs `mp_submission_verify` and captures the output. 
   The following code snippet will be useful:
   
```{r}
#| eval: false
O <- capture.output({
    E <- tryCatch(mp_submission_verify(N, github_id), 
                  error=function(e) e)
})

if(isTRUE(E)){
    MESSAGE <- "Mini-Project properly submitted."
} else {
    MESSAGE <- "Mini-Project not properly submitted. Issue appears to be:\n"
    MESSAGE <- paste(c(MESSAGE, O[1:2]), collapse="\n")
}
```

   where you use `MESSAGE` in creating your output. 

### Step 4: Extensions

Time allowing, do one or more of the following: 

- Adapt `mp_feedback_locate` to provide links to all peer feedback issues. 
  Note that you will need to handle the case of later MPs, who haven't been
  assigned yet (*i.e.*, MPs #03 and #04). 
- Adapt `mp_feedback_verify` to check whether your peer feedback has been
  properly formatted. This will require a new input for the peer's GitHub ID. 
