---
lab_num: "`r lab <- 7; lab`"
lab_title: "Data Import in `R`"
engine: knitr
format: live-html
course_repo: "`r yaml::read_yaml('_variables.yml')$course$repo`"
---

{{< include _setup.qmd >}}

{{< include _gradethis_init.qmd >}}

```{webr}
#| echo: false
webr::install(c("tidyverse", "mgcv", "ggpmisc"), quiet=TRUE)
suppressPackageStartupMessages(library(tidyverse))
```

# Review Practice {#review}

```{r}
#| echo: false
#| message: false
#| warning: false
library(tidyverse)

try_read_csv <- possibly(read_csv)

SSA <- try_read_csv("https://michael-weylandt.com/STA9750/labs/ssa_babynames.csv.gz")

if(is.null(SSA)){
   SSA <- try_read_csv("docs/labs/ssa_babynames.csv.gz")
}
```

Every year, the US Social Security Administration reports the number of
children born each year in each state with each name: *e.g*., I was one of
`r SSA |> filter(year==1989, state=="TX", name == "Robert", sex=="M") |> pull(number) |> scales::comma()` 
male babies named Robert born in Texas in 1989.[^ssa_babies] I have placed a 
cleaned-up CSV containing this data (through `r SSA |> pull(year) |> max()`) 
online at 
<https://michael-weylandt.com/STA9750/labs/ssa_babynames.csv.gz>. (The extra 
`gz` extension indicates that this is a [`gzip` compressed
file](https://en.wikipedia.org/wiki/Gzip) You don't need to know the details
of `gzip`, only that it is a compression scheme used to make big files smaller;
`bz` and `xz` will be used for some data sources. These are mainly *single-file* 
compression methods, as opposed to `zip` which compresses a whole directory. ) 

::: {.callout-caution title="Optional: Replication Code" collapse="true"}

The following code can be used to create your own copy of this data set.

```{r}
#| eval: false
SSA_URL <- "https://www.ssa.gov/oact/babynames/state/namesbystate.zip"
TEMPDIR <- tempdir()
TEMPZIP <- fs::path(TEMPDIR, "babynamesbystate.zip")
download.file(SSA_URL, TEMPZIP)

unzip(TEMPZIP, exdir=TEMPDIR)

fs::dir_info(TEMPDIR) |> 
    dplyr::filter(str_ends(path, ".TXT")) |> 
    dplyr::pull(path) |> 
    purrr::map(\(f) readr::read_csv(f, 
                                    show_col_types=FALSE, 
                                    col_names=c("state", "sex", "year", "name", "number"))) |> 
    dplyr::bind_rows() |> 
    readr::write_csv("labs/ssa_babynames.csv.gz") |>
    readr::write_csv("docs/labs/ssa_babynames.csv.gz")
```

By the end of today's class, you should be able to understand everything that's
happening in this little snippet.

:::

[^ssa_babies]: You might remember the aggregated version of this data from 
the [review exercise in Lab #05](./lab05.qmd#review).

The data looks something like this:

```{r}
#| echo: false
SSA |> glimpse()
```

Read this data into `R` using the `read_csv()` function from the
`readr` package; note that the `read_csv()` function can take a URL 
as an input so you don't need to directly download this data set.

Once you have downloaded the baby names data, create tables or visualizations 
that address the following questions: 

1. In 2024, what were the five most common girls' names *nationwide*?

::: {.callout-tip collapse="true" title="Solution"}

```{r}
library(tidyverse)
library(gt)
read_csv("https://michael-weylandt.com/STA9750/labs/ssa_babynames.csv.gz") |>
   filter(year == 2024, 
          sex == "F") |> 
   group_by(name) |> 
   summarize(number = sum(number)) |> 
   slice_max(number, n=5) |> 
   gt() |> 
   cols_label(name = "Name", 
              number = "Number of Girls Born Nationwide") |> 
   tab_header(title="5 Most Popular Girls Names in 2024") |> 
   tab_source_note("Data from the US SSA")
```

Since we're going to use this data set several times, it's a bit
more efficient to save it to a variable once instead of re-downloading it over
and over: 

```{r}
#| eval: false
SSA <- read_csv("https://michael-weylandt.com/STA9750/labs/ssa_babynames.csv.gz") 
```

and now we have: 

```{r}
#| eval: false
library(tidyverse)
library(gt)
SSA |>
   filter(year == 2024, 
          sex == "F") |> 
   group_by(name) |> 
   summarize(number = sum(number)) |> 
   slice_max(number, n=5) |> 
   gt() |> 
   cols_label(name = "Name", 
              number = "Number of Girls Born Nationwide") |> 
   tab_header(title="5 Most Popular Girls Names in 2024") |> 
   tab_source_note("Data from the US SSA")
```

giving the same results.

:::

2. In terms of popularity _ranking_ (1st, 2nd, 3rd, *etc*), how has the
   popularity of the name "Michael" (for boys) changed in New York?
   

::: {.callout-tip collapse="true" title="Solution"}

```{r}
SSA |> 
    filter(state == "NY", 
           sex == "M") |> 
    # Flip ranking so high number => low (1st, 2nd) rank
    # Compute rankings _for that year_
    mutate(pop_rank = dense_rank(-number), 
           .by = "year") |>
    filter(name == "Michael") |> 
    mutate(rank_smooth = runmed(pop_rank, 11)) |>
    ggplot(aes(x=year)) + 
    geom_point(aes(y=pop_rank), alpha=0.2) + 
    geom_line(aes(y=rank_smooth)) + 
    scale_y_reverse(breaks=c(1, 10, 20, 30, 40, 50)) + 
    theme_bw() + 
    xlab("Year") + 
    ylab("Popularity Ranking")
```

We see that, starting around 1925, the name "Michael" became very popular in
NY and it was the most popular male name from 1956 to 2013 in all but two
years.

:::


3. How has the *fraction* of *Juan*s born(as a fraction of all names) increased
   over time? (*Hint:* Think OLS and use the `ggpmisc::stat_poly_eq` function
   to annotate your solution.)
   

::: {.callout-tip collapse="true" title="Solution"}

```{r}
#| message: false
#| warning: false
library(ggpmisc)
SSA |> 
   summarize(n_total=sum(number), 
             n_juan = sum(number * (name == "Juan")), 
             .by = c("year")) |> 
   mutate(p_juan = n_juan / n_total) |> 
   ggplot(aes(x=year, y=1e5 * p_juan)) + 
   geom_point() + 
   xlab("Year") + 
   ylab("Juans per 100,000 Babies") + 
   theme_bw() + 
   stat_poly_eq(use_label(c("R2", "eq"))) + 
   geom_smooth(method="lm", col="red4")
```

Here, we see that, while the popularity of Juan has overall been increasing 
(with an $R^2$ of 60%, which is quite good for the social sciences), there has
been a drop-off post-2000 that a simple linear model wouldn't capture. 

Note also the use of "Juans per 100,000" on the $y$-axis here. This is a general
technique to put small probabilities into a more natural-human scale. You may
be familiar with the use of [micromorts](https://en.wikipedia.org/wiki/Micromort)
or deaths per million people as a measure of the risk associated with various 
lifestyle factors.

:::

4. Have the numbers of *Juan*s and *Jose*s born each year increased in rough
   proportion to each other?  (*Hint:* Think OLS)
   

::: {.callout-tip collapse="true" title="Solution"}
   
```{r}
library(ggpmisc)
SSA |> 
   filter(name %in% c("Juan", "Jose")) |>
   summarize(n=sum(number), 
             .by = c("year", "name")) |>
   pivot_wider(names_from=name,
               values_from=n) |>
   ggplot(aes(x=Juan, y=Jose)) +
   # Put this here b/c color doesn't impact the stat_()s
   # Only the point
   geom_point(aes(color=year)) + 
   scale_color_fermenter(palette=3,
                         direction=1,
                         breaks=c(1910, 1930, 1950,1970, 1980, 1990, 2000, 2010, 2020)) + 
   theme_bw() + 
   theme(legend.position="bottom") +
   guides(color=guide_colorbar(barwidth=20)) +
   stat_poly_line() + 
   stat_poly_eq() + 
   xlab("Number of Juan's Born Per Year") + 
   ylab("Number of Jose's Born Per Year")
```

As before, by seeing the darkest purple dots near the
center and not the top right of the graph, we see that 
there are fewer *Juan*s and *Jose*s being born than 30 years
ago.

:::

5. Were there more *Mary*s born per year in the 2000s or the 2010s? (*Hint:* Think ANOVA)


::: {.callout-tip collapse="true" title="Solution"}
   
```{r}
SSA |> 
    filter(year >= 2000, 
           year <= 2019, 
           name == "Mary") |> 
    group_by(year) |> 
    summarize(n_mary = sum(number)) |> 
    mutate(decade=if_else(year <= 2009, "2000s", "2010s")) |> 
    group_by(decade) |> 
    mutate(mean_n_mary = mean(n_mary), 
           sd_n_mary = sd(n_mary), 
           # Note n() to count number of obs per group
           
           se_n_mary = sd_n_mary / sqrt(n())) |>
    ggplot(aes(x=decade, y=n_mary))  + 
    stat_summary(fun="mean", 
                 geom="col", 
                 fill="lightblue", 
                 alpha=0.8) + 
    # After so the points appear _above_ the summary bar
    geom_jitter(height = 0) + 
    theme_bw() + 
    geom_errorbar(aes(ymin = mean_n_mary - se_n_mary, 
                      ymax = mean_n_mary + se_n_mary), 
                  color="red4", 
                  width=0.4) + 
    xlab(NULL) + 
    ylab("Number of Marys Born Per Year")
```

Because we're looking at the year as a categorical (decade)
variable instead of a numeric quantity here, we use an
ANOVA-inspired barplot instead of linear regression.[^anova] 

[^anova]: Note, however, that standard ANOVA is basically just a 
very restricted linear model; *cf.* <https://lindeloev.github.io/tests-as-linear/>. 

:::

# Using the File System {#fs}

## Exercises #01 {#exercise01}

0. Use the `getwd()` to find your current working directory. For the following
   exercises to be most useful, you will want to ensure that this is your 
   `{{< meta course_repo >}}` course directory. If you have somehow changed it
   please use the `setwd()` function to return to the proper directory.

1. In your `{{< meta course_repo >}}` directory, you have a 
   file called `index.qmd`. Use the `path` function from the 
   `fs` package to create a `path` object pointing to that
   file and then use the `path_abs` function to convert that
   relative path to an absolute path. 
   
::: {.callout-tip collapse="true" title="Solution"}
   
Note that these solutions will give different answers on your
personal computer than on this web interface. That is to be expected 
since we are going outside of `R` and interacting with the 
whole system, so differences should not surprise you. 
   
```{webr}
library(fs)
path("index.qmd") |> path_abs()
```

:::

2. Using the `fs::file_size()` function, how big is the file you found in the 
   previous section? 

::: {.callout-tip collapse="true" title="Solution"}
   
Note that these solutions will give different answers on your
personal computer as opposed to this web interface. That is to be expected 
since we are going outside of `R` and interacting with the 
whole system, so differences should not surprise you. 
   
```{webr}
library(fs)
path("index.qmd") |> path_abs() |> file_size()
```

:::

3. Use the `file_exists` function to determine whether you have started 
   [Mini-Project #03](../mini/mini03.html).

::: {.callout-tip collapse="true" title="Solution"}
   
```{webr}
library(fs)
file_exists("mp03.qmd")
```   

:::


4. A common pattern in working with files is wanting to find
   all files whose names match a certain pattern. This is known
   colloquially as _globbing_. The `dir_ls` function can be used
   to list all files in a directory. First use it to list all
   files in your `{{< meta course_repo >}}` course directory; then
   use the `glob` argument to list only the `qmd` files in that
   directory. 
   
   *Hint: The `glob` argument should be a string with asterisks in it where
   you will allow any match. For example, `*.csv` will match all CSV files, 
   `abc*` will match all files whose names start with `abc`, and 
   `abc*.csv` will match files whose names start with `abc` and end with
   `.csv` with arbitrary text in between.*
   
::: {.callout-tip collapse="true" title="Solution"}
   
```{webr}
library(fs)
dir_ls() # Default is dir_ls("."): i.e., current working directory
dir_ls(glob="*qmd")
```   

:::

5. The `dir_info` function can be used to create a data frame with information
   about all files in a directory. Use it and `dplyr` functionality to find the
   largest file in your project folder. Then return its _absolute_ path. 
   
::: {.callout-tip collapse="true" title="Solution"}
   
```{webr}
library(fs)
library(dplyr)
dir_info() |> slice_max(size) |> pull(path) |> path_abs()
```   

:::   

6. Often, we want to perform a search over all files and folders in a directory
   *recursively* - that is, looking inside subfolders, not just at files in
   the top level. The `recurse=TRUE` argument to `dir_info` will return
   information about all files contained in any level of the directory. Use
   `dir_info` to search the `data` directory you have created to store
   Mini-Project data and find the _smallest_ data file. 
   
   *Hint: You will need to pass a `path` argument to `dir_info` so that you
   search only your `data` folder and not your entire course directory.*
   
::: {.callout-tip collapse="true" title="Solution"}
   
```{webr}
library(fs)
library(dplyr)
dir_info("data", recurse=TRUE) |> slice_max(size) |> pull(path) |> path_abs()
```   

:::

7. Determine how much storage is used by the 5 largest files in any part of 
   your project directory. 

::: {.callout-tip collapse="true" title="Solution"}
   
```{webr}
library(fs)
library(dplyr)
dir_info(recurse=TRUE) |> slice_max(size, n=5) |> summarize(total_size = sum(size))
```   

:::

# HTTP and Web Access {#http}

## Exercises #02 {#exercise02}

Earlier, we used the fact that `read_csv` could download a file automatically
for us to read in the SSA Data. We are now going to perform essentially 
the same process using `httr2`, but instead we're going to look for the 
smaller `nyc_demos.csv` file in the same directory.[^gz]

[^gz]: The `ssa_babynames.csv.gz` file is compressed because it's big, but
that makes downloading *slightly* trickier than I want to demonstrate here. 
Essentially, to get `readr::read_csv` to read it correctly, we have to save
it to a file and read it there, rather than simply reading it in from memory: 

    ```{r}
#| eval: false
library(httr2)
library(readr)
request("https://michael-weylandt.com") |>
    req_url_path("STA9750", 
                 "labs",
                 "ssa_babynames.csv.gz") |>
    req_perform("ssa_babynames.csv.gz") |>
    resp_check_status()  

read_csv("ssa_babynames.csv.gz")
    ```
    
    This isn't hard *per se*, but it's just one extra thing to think about.

1. Use the `httr2` package to build a suitable `request` object. You should
   build your request in two steps: 
   
   i) Specify the domain using `request`
   
   ii) Add the path using `req_url_path`
   
::: {.callout-tip collapse="true" title="Solution"}
   
```{r}
library(httr2)
request("https://michael-weylandt.com") |>
    req_url_path("STA9750", 
                 "labs",
                 "nyc_demos.csv") 
```   

:::
   
2. Now that you have built a request, _perform_ it and check to make sure
   your request was successful. 
   
::: {.callout-tip collapse="true" title="Solution"}
   
```{r}
library(httr2)
request("https://michael-weylandt.com") |>
    req_url_path("STA9750", 
                 "labs",
                 "nyc_demos.csv") |>
    req_perform() |>
    resp_check_status()
```   

Remember - no noise is good noise! The fact `resp_check_status` didn't throw
an error means we're good to continue.

Note that we're switching from `req_` functions to `resp_` functions
after we perform the request since we're interested in what we got back
from the server, *i.e.*, the response.

:::
   
3. Because your request was successful, it will have a body. Extract this body
   as a string and pass it to `read_csv` to read the data into `R`. 

::: {.callout-tip collapse="true" title="Solution"}
   
```{r}
library(httr2)
library(readr)
request("https://michael-weylandt.com") |>
    req_url_path("STA9750", 
                 "labs",
                 "nyc_demos.csv") |>
    req_perform() |> 
    resp_check_status() |>
    resp_body_string() |> 
    read_csv()
```

Note that you can also pass a path to `req_perform` to save the result to a file
instead of holding it in memory. This is useful if you are downloading many 
files and saving them for later use.

:::

4. Modify your analysis to download **your** `mp01.qmd` file. To find the 
   appropriate URL, you will need to first find the file in the web interface 
   for your GitHub repository and then click the `Raw` button to a direct link 
   to the file. Make a request to get this data and check whether the 
   request was successful. 
   
::: {.callout-tip collapse="true" title="Solution"}
   
```{r}
#| eval: false
library(httr2)
request("https://raw.githubusercontent.com") |>
    req_url_path(YOUR GITHUB ID
                 COURSE REPO, 
                 "refs",
                 "heads",
                 "main",
                 "mp01.qmd") |>
    req_perform() |>
    resp_check_status() 
```

:::
   
5. Next, modify your analysis to check whether `mp04.qmd` has already
   been uploaded. Note that this will throw an error because a `404` is returned,
   indicating that you have not yet uploaded `mp04.qmd`. We will address this
   in the next step. 

::: {.callout-tip collapse="true" title="Solution"}
   
```{r}
#| eval: false
library(httr2)
request("https://raw.githubusercontent.com") |>
    req_url_path(YOUR GITHUB ID
                 COURSE REPO, 
                 "refs",
                 "heads",
                 "main",
                 "mp04.qmd") |>
    req_perform()
```

:::

6. Obviously, that error is a bit of a problem. Let's change how our request
   handles errors. The `req_error` function can be used to modify a request to
   change whether an error is thrown. As its argument, it takes a _function_
   that checks for an error. Since we want to never throw an error, define a
   function with one argument `x` that always returns false and pass this to
   `req_error`. (Note that `req_error` must come before `req_perform` since it
   changes how a request is performed.) Use this in conjunction with the
   `resp_is_error` function to check that your request fails. 

::: {.callout-tip collapse="true" title="Solution"}
   
```{r}
#| eval: false
library(httr2)

always_false <- function(x) FALSE

request("https://raw.githubusercontent.com") |>
    req_url_path(YOUR GITHUB ID
                 COURSE REPO, 
                 "refs",
                 "heads",
                 "main",
                 "mp04.qmd") |>
    req_error(is_error = always_false) |>
    req_perform() |>
    resp_is_error()
```

*Note*: Using an _anonymous_ function, we can write this a bit more compactly as 

```{r}
#| eval: false
library(httr2)
request("https://raw.githubusercontent.com") |>
    req_url_path(YOUR GITHUB ID
                 COURSE REPO, 
                 "refs",
                 "heads",
                 "main",
                 "mp04.qmd") |>
    req_error(is_error = \(x) FALSE) |>
    req_perform() |>
    resp_is_error()
```
:::

7. Package your code into a function that takes an integer argument and tests
   whether that mini-project is missing from your GitHub.
   
   *Hint: The `glue` package may be useful to make strings here: 
   `glue("mp0{N}.qmd")` will automatically substitute the value of the variable
   `N` for you.*

::: {.callout-tip collapse="true" title="Solution"}
   
```{r}
#| eval: false
library(httr2)
library(glue)
mp_is_missing <- function(N) {
    request("https://raw.githubusercontent.com") |>
        req_url_path(YOUR GITHUB ID
                     COURSE REPO, 
                     "refs",
                     "heads",
                     "main",
                     glue("mp0{N}.qmd")) |>
        req_error(is_error = \(x) FALSE) |>
        req_perform() |>
        resp_is_error()
}
```
:::

8. Finally, check which of your MPs have not yet been submitted. 

   Unfortunately, your function is not _vectorized_ so this is not as 
   simple as `func(1:4)`. Instead, we want to apply the same function separately
   to each element of the vector `c(1,2,3,4)`. This is a use case for the `map`
   family of functions, which we will discuss in more detail later during our
   final optional enrichment session after your final presentations. Since your 
   function returns a logical value, we'll use
   `map_lgl` here. 
   

::: {.callout-tip collapse="true" title="Solution"}

```{r}
#| eval: false
library(httr2)
library(glue)
mp_is_missing <- function(N) {
    request("https://raw.githubusercontent.com") |>
        req_url_path(YOUR GITHUB ID
                     COURSE REPO, 
                     "refs",
                     "heads",
                     "main",
                     glue("mp0{N}.qmd")) |>
        req_error(is_error = \(x) FALSE) |>
        req_perform() |>
        resp_is_error()
}

map_lgl(1:4, mp_is_missing)
```
:::

This example is maybe a bit silly, but it is essentially how I check whether
your mini-projects have been submitted on time. (I map over students rather than
the project numbers and perform more checks, but the same ideas apply.) 

# API Usage {#api}

Next, we are going to practice accessing data from a nice 
API using `httr2`. Specifically, we are going to interact with the
`cranlogs` server, which keeps records of the most popular `R` packages
based on download frequency. 

Documentation for `cranlogs` can be found [in its GitHub
`README`](https://github.com/r-hub/cranlogs.app) with a very small
example at [here](https://cranlogs.r-pkg.org/#jsonapi).[^rpkg]

[^rpkg]: In this case, there is a `cranlogs` package for interacting with
this API. This type of package is commonly called a "wrapper" because it
shields the user from the details of the API and exposes a more idiomatic
(and more useful) interface. In general, when you can find an `R` package
that wraps an API, it is a good idea to use it. For certain very complex
APIs, *e.g.*, the [API that powers Bloomberg](https://www.bloomberg.com/professional/support/api-library/) 
Financial Information Services, use of the associated 
[`R` package](https://cran.r-project.org/web/packages/Rblpapi/vignettes/rblpapi-intro.html) 
is almost mandatory because the underlying API is so complicated. For this
in-class exercise, we will use the "raw" API as practice since you can't
always assume a nice `R` package will exist. 

The `cranlogs` documentation give the following example of how the `curl` program
can call the API from the command line: 

```{bash}
#| cache: true
#| message: false
curl https://cranlogs.r-pkg.org/downloads/total/last-week/ggplot2
```

To see what type of data this URL provides us, you can simply
[open it in your browser](https://cranlogs.r-pkg.org/downloads/total/last-week/ggplot2)
and observe that it returns `JSON` data. We can emulate this action in `R` and
parse the data into an `R` object:

```{r}
library(jsonlite)
fromJSON("https://cranlogs.r-pkg.org/downloads/total/last-week/ggplot2")
```

And if we want to get download information for other packages, we can simply
modify the URL: 

```{r}
library(jsonlite)
fromJSON("https://cranlogs.r-pkg.org/downloads/total/last-week/dplyr")
```

```{r}
library(jsonlite)
fromJSON("https://cranlogs.r-pkg.org/downloads/total/last-week/readr")
```

and so on. But this quickly becomes repetitive and we would prefer a 
*programmatic* interface. This is where advanced functionality from the `httr2` 
package comes in. 

## `httr2`

Recall that `httr2` takes a "three-stage" approach to handling HTTP requests: 

1. First, we _build_ a _request_, specifying the URL to be queried, the mode of 
   that query, and any relevant information (data, passwords, *etc.*)
2. Then, we _execute_ that request to get a _response_
3. Finally, we _handle_ that response, transforming its contents into `R` as
   appropriate
   
Let's look at these one at a time. 

Note that the first few steps should feel quite similar to what we did above, 
since we're still communicating via HTTP. The *content* of that communication
will be different (JSON instead of CSV) and the server's work will be different
(adding up download statistics instead of serving a static file), but the 
*protocol*--and hence most of our `httr2` flow--remains unchanged.

### Build a Request

We can build a request using the `request` function: 

```{r}
library(httr2)
request
```

As seen here, we start a request by putting in a "base URL" - this is 
the unchanging part of the URL that won't really depend on what query we are making. 

For the `cranlogs` API, we can take the base URL to be 

> `https://cranlogs.r-pkg.org`

so our base request is: 

```{r}
my_req <- request("https://cranlogs.r-pkg.org")
print(my_req)
```

We see here that this is a `GET` request by default, indicating we would like
a response from the server, but we are not sending data to the server for 
processing. 

We then modify the _path_ of the request to point to the specific _resource_ or
_endpoint_ we want to query. For example, if we want to get the "top" `R` 
packages for the last day, we can run the following: 

```{r}
my_req <- my_req |> 
    req_url_path_append("top") |>
    req_url_path_append("last-day")
print(my_req)
```

### Execute a Request to get a Response

Now that we have built our request, we pass it to `req_perform`[^adv1] to execute 
(perform) the request: 

[^adv1]: We will only use the basic `req_perform` for now, but `httr2` 
provides options for parallel execution, delayed execution, *etc.*.

```{r}
my_resp <- req_perform(my_req)
print(my_resp)
```

The result of performing this request is a _response_ object. We see several
things in this response: 

1. We received back a "200 OK" response, indicating that our query worked perfectly
2. We received back data in a `json` format
3. Our results are currently in memory (as opposed to being saved into a file)

### Process the Response for Use in `R`

Since we know our response is in `JSON` format, we can use the `resp_body_json`
to get the "body" (content) of the response and parse it as `json`: 

```{r}
downloads_raw <- resp_body_json(my_resp)
print(downloads_raw)
```

This gives us the type of data we were looking for!

Note that `httr2` is designed for "piped" work, so we can write the entire
process as

```{r}
request("https://cranlogs.r-pkg.org") |>
    req_url_path_append("top") |>
    req_url_path_append("last-day") |>
    req_perform() |>
    resp_body_json()
```

This data is not super helpful for us, since it's in a "list of lists" format. 
This is not uncommon with `json` responses and it is usually at this point that 
we have a bit of work to do in order to make the data useable. Thankfully, API
data is typically well-structured, so this doesn't wind up being too hard. 
I personally find this type of complex `R` output a bit hard to parse, so I
instead print it as a "string" (the 'raw text' of the unparsed JSON) and use
the `prettify()` function from the `jsonlite` package to make it extra readable: 

```{r}
library(jsonlite)
my_resp |>
    resp_body_string() |>
    prettify()
```

This is the same data as before, but much easier to read. At this point, we
should pause and make an 'attack plan' for our analysis. I see several things here:

1) I really only want the `"downloads"` part of the response. 
2) Each element inside `downloads` has the same flat structure, so they can
   be easily built into a one-row data frame. 
3) The column names are the same for each `downloads` element, so we will be
   able to put them into one big easy-to-use data frame. 
   
To do these steps, we will need to use functionality from the `purrr` package, 
which we will discuss in more detail later in the course. For now, it suffices 
to run: 

```{r}
library(purrr)
library(tibble)
downloads_df <- downloads_raw |>
    pluck("downloads") |>
    map(as_tibble) |>
    list_rbind()
```

Here, we see we

1) Pulled out the `"downloads"` portion of the JSON (`pluck`)
2) Converted each row to a data frame (`map(as_tibble)`. The `as_tibble` does
   the data frame conversion and the `map` means "do this to each element 
   separately".)
3) Combined the results rowwise (`list_rbind`)

The result is a very nice little data frame: 

```{r}
downloads_df
```

## Exercises #03 {#exercise03}

Now it's your turn! In your breakout rooms, try the following: 

1. Make sure you can run all of the code above. 
2. Modify the above code to get the top 100 `R` packages. 

   This is a minor change to the request only, but you will need to read the 
   [documentation](https://github.com/r-hub/cranlogs.app) to see where and how
   the request needs to be changed. 
   
::: {.callout-note collapse="true" title="Solution"}
   
```{r}
library(httr2)
library(tidyverse)
request("https://cranlogs.r-pkg.org") |>
    req_url_path_append("top") |>
    req_url_path_append("last-day") |>
    req_url_path_append(100) |>
    req_perform() |>
    resp_body_json() |>
    pluck("downloads") |>
    map(as_tibble) |>
    list_rbind()
```

:::

3. Modify your query to get the daily downloads for the `ggplot2` package over
   the last month. This will require changes to how you process the response, so
   be sure to look at the raw JSON first. 
   
   *Hint: The `pluck` function can also take a number as input. This will say
   which list item (by position) to return.*
   
::: {.callout-note collapse="true" title="Solution"}
   
```{r}
library(httr2)
library(tidyverse)
request("https://cranlogs.r-pkg.org") |>
    req_url_path_append("downloads") |>
    req_url_path_append("daily") |>
    req_url_path_append("last-month") |>
    req_url_path_append("ggplot2") |>
    req_perform() |>
    resp_body_json() |>
    pluck(1) |>
    pluck("downloads") |>
    map(as_tibble) |>
    list_rbind()
```

:::

4. 'Functionize' your daily downloads query as a function which takes an arbitrary
    package name and gets its daily downloads.
    
    *Hint: Use a `mutate` command to add the package name
    as a new column to the resulting data frame and to
    convert the `day` column to a `Date` object (`day=as.Date(day)`)*. 
    
::: {.callout-note collapse="true" title="Solution"}
   
```{r}
library(httr2)
library(tidyverse)
get_downloads <- function(pkg){
    request("https://cranlogs.r-pkg.org") |>
        req_url_path_append("downloads") |>
        req_url_path_append("daily") |>
        req_url_path_append("last-month") |>
        req_url_path_append(pkg) |>
        req_perform() |>
        resp_body_json() |>
        pluck(1) |> 
        pluck("downloads") |>
        map(as_tibble) |>
        list_rbind() |>
        mutate(package=pkg, 
               day=as.Date(day))
}

gg_downloads <- get_downloads("ggplot2")
print(gg_downloads)
```

:::

5. Use your function to get daily downloads for the following `R` packages: 
   - `ggplot2`
   - `dplyr`
   - `httr2`
   - `tibble`
   - `purrr`
   - `tidyr`
   for the year and combine your results into a single data frame. 
   Then plot the download trends for each package using `ggplot2`.
   
   You can do this by calling your function in a loop, but you should
   also try to apply the `map() |> list_rbind()` idiom here as well. 
   
   To get values for a specific window, replace `"last-month"` with a date range
   formatted as `"YYYY-MM-DD::YYYY-MM-DD"`. For this problem, you can use
   `strftime(Sys.Date(), "%Y-01-01:%Y-%m-%d")` which evaluates to the string
   "`r strftime(Sys.Date(), "%Y-01-01:%Y-%m-%d")`", *i.e.*, from the start of
   the year to today or `paste(Sys.Date() - 365, Sys.Date(), sep=":")` to get
   "`r paste(Sys.Date() - 365, Sys.Date(), sep=":")`" for the last 365 days.
   
::: {.callout-note collapse="true" title="Solution"}
   

```{r}
library(tidyverse)
library(httr2)

PACKAGES <- c("ggplot2", "dplyr", "httr2", "tibble", "purrr", "tidyr")

date_range <- paste(Sys.Date() - 365, 
                    Sys.Date(), sep=":")

get_downloads_year <- function(pkg){
    request("https://cranlogs.r-pkg.org") |>
        req_url_path_append("downloads") |>
        req_url_path_append("daily") |>
        req_url_path_append(date_range) |>
        req_url_path_append(pkg) |>
        req_perform() |>
        resp_body_json() |>
        pluck(1) |> 
        pluck("downloads") |>
        map(as_tibble) |>
        list_rbind() |>
        mutate(package=pkg, 
               day=as.Date(day))
}

PACKAGES |> 
  map(get_downloads_year) |>
  list_rbind() |>
  ggplot(aes(x=day, 
             y=downloads, 
             color=package, 
             group=package)) +
    geom_point() +
    geom_line() +
    scale_x_date() +
    xlab("Date of Download") +
    ylab("Number of Daily Package Downloads") +
    theme_bw() +
    labs(caption="Data from cranlogs.r-pkg.org") +
    theme(legend.position="bottom") +
    scale_color_brewer(type="qual", 
                         palette=2, 
                         name="Package Name") + 
    scale_y_continuous(labels=scales::label_number())
```

Can you see the start of the semester(s) in this data? What other patterns can you spot?

:::
