---
title: "{{< var course.short >}} Week 12 In-Class Activity: Strings"
engine: knitr
format: live-html
---

{{< include ./_extensions/r-wasm/live/_knitr.qmd >}}
{{< include ./_extensions/r-wasm/live/_gradethis.qmd >}}

[Week 12 Slides](../slides/slides12.html)

## Regular Expression Practice

Complete the following exercises using functionality from the
`stringr` package. 

1. In the following sentence, extract all *plural nouns*[^1]:

```{r}
#| eval: false
todo <- "Yesterday, I needed to buy four cups of flour, a piece of Parmesan cheese, two gallons of ice cream, and a six-pack of bottled (non-alcoholic) beers."
```

```{webr}
#| exercise: plural_nouns
library(stringr)
todo <- "Yesterday, I needed to buy four cups of flour, a piece of Parmesan cheese, two gallons of ice cream, and a six-pack of bottled (non-alcoholic) beers."
str_extract_all(todo, ______, simplify=TRUE)
```


::: {.solution exercise="plural_nouns"}

```{webr}
#| exercise: plural_nouns
#| solution: true
library(stringr)
todo <- "Yesterday, I needed to buy four cups of flour, a piece of Parmesan cheese, two gallons of ice cream, and a six-pack of bottled (non-alcoholic) beers."
str_extract_all(todo, " [A-Za-z]+s[. ]", simplify=TRUE)
```
:::

```{webr}
#| exercise: plural_nouns
#| check: true
library(stringr)
gradethis::grade_this_code()
```


[^1]: While English pluralization rules are tricky, you can just find the words ending with an `s`.

2. In the following sentence, compute the total number of fruits on my shopping
list: 

```{r}
#| eval: false
shopping <- "Today, I need to purchase 3 apples, 5 limes, and 2 lemons."
```

```{webr}
#| exercise: numbers
library(stringr)
shopping <- "Today, I need to purchase 3 apples, 5 limes, and 2 lemons."
sum(as.numeric(str_extract_all(shopping, ______, simplify=TRUE)))
```

::: {.solution exercise="numbers"}

```{webr}
#| exercise: numbers
#| solution: true
library(stringr)
shopping <- "Today, I need to purchase 3 apples, 5 limes, and 2 lemons."
sum(as.numeric(str_extract_all(shopping, "\\d+", simplify=TRUE)))
```
:::

```{webr}
#| exercise: numbers
#| check: true
library(stringr)
gradethis::grade_this_code()
```

3. The following text is adapted from the [Taylor Swift wikipedia page](https://en.wikipedia.org/wiki/Taylor_Swift), with some changes made to the
punctuation to make things easier. 

> Taylor Alison Swift (born December 13, 1989) is an American singer-songwriter. A subject of widespread public interest, she has influenced the music industry and popular culture through her artistry, especially in songwriting, and entrepreneurship. She is an advocate of artists rights and womens empowerment. Swift began professional songwriting at age 14. She signed with Big Machine Records in 2005 and achieved prominence as a country pop singer with the albums Taylor Swift (2006) and Fearless (2008). Their singles 'Teardrops on My Guitar', 'Love Story', and 'You Belong with Me' were crossover successes on country and pop radio formats and brought Swift mainstream fame. She experimented with rock and electronic styles on her next albums, Speak Now (2010) and Red (2012), respectively, with the latter featuring her first Billboard Hot 100 number-one single, 'We Are Never Ever Getting Back Together'. Swift recalibrated her image from country to pop with 1989 (2014), a synth-pop album containing the chart-topping songs 'Shake It Off', 'Blank Space', and 'Bad Blood'. Media scrutiny inspired the hip-hop-influenced Reputation (2017) and its number-one single 'Look What You Made Me Do'. After signing with Republic Records in 2018, Swift released the eclectic pop album Lover (2019) and the autobiographical documentary Miss Americana (2020). She explored indie folk styles on the 2020 albums Folklore and Evermore, subdued electropop on Midnights (2022), and re-recorded four albums subtitled Taylors Version after a dispute with Big Machine. These albums spawned the number-one songs 'Cruel Summer', 'Cardigan', 'Willow', 'Anti-Hero', 'All Too Well', and 'Is It Over Now?'. Her Eras Tour (2023-2024) and its accompanying concert film became the highest-grossing tour and concert film of all time, respectively. Swift has directed videos and films such as Folklore: The Long Pond Studio Sessions (2020) and All Too Well: The Short Film (2021). Swift is one of the worlds best-selling artists, with 200 million records sold worldwide as of 2019. She is the most-streamed artist on Spotify, the highest-grossing female touring act, and the first billionaire with music as the main source of income. Six of her albums have opened with over one million sales in a week. The 2023 Time Person of the Year, Swift has appeared on lists such as Rolling Stones 100 Greatest Songwriters of All Time, Billboards Greatest of All Time Artists, and Forbes Worlds 100 Most Powerful Women. Her accolades include 14 Grammy Awards, a Primetime Emmy Award,  40 American Music Awards, 39 Billboard Music Awards, and 23 MTV Video Music Awards; she has won the Grammy Award for Album of the Year, the MTV Video Music Award for Video of the Year, and the IFPI Global Recording Artist of the Year a record four times each.

How many times does Taylor Swift's last name appear? 

```{webr}
#| exercise: swift
library(stringr)
swift <- "Taylor Alison Swift (born December 13, 1989) is an American singer-songwriter. A subject of widespread public interest, she has influenced the music industry and popular culture through her artistry, especially in songwriting, and entrepreneurship. She is an advocate of artists rights and womens empowerment. Swift began professional songwriting at age 14. She signed with Big Machine Records in 2005 and achieved prominence as a country pop singer with the albums Taylor Swift (2006) and Fearless (2008). Their singles 'Teardrops on My Guitar', 'Love Story', and 'You Belong with Me' were crossover successes on country and pop radio formats and brought Swift mainstream fame. She experimented with rock and electronic styles on her next albums, Speak Now (2010) and Red (2012), respectively, with the latter featuring her first Billboard Hot 100 number-one single, 'We Are Never Ever Getting Back Together'. Swift recalibrated her image from country to pop with 1989 (2014), a synth-pop album containing the chart-topping songs 'Shake It Off', 'Blank Space', and 'Bad Blood'. Media scrutiny inspired the hip-hop-influenced Reputation (2017) and its number-one single 'Look What You Made Me Do'. After signing with Republic Records in 2018, Swift released the eclectic pop album Lover (2019) and the autobiographical documentary Miss Americana (2020). She explored indie folk styles on the 2020 albums Folklore and Evermore, subdued electropop on Midnights (2022), and re-recorded four albums subtitled Taylors Version after a dispute with Big Machine. These albums spawned the number-one songs 'Cruel Summer', 'Cardigan', 'Willow', 'Anti-Hero', 'All Too Well', and 'Is It Over Now?'. Her Eras Tour (2023-2024) and its accompanying concert film became the highest-grossing tour and concert film of all time, respectively. Swift has directed videos and films such as Folklore: The Long Pond Studio Sessions (2020) and All Too Well: The Short Film (2021). Swift is one of the worlds best-selling artists, with 200 million records sold worldwide as of 2019. She is the most-streamed artist on Spotify, the highest-grossing female touring act, and the first billionaire with music as the main source of income. Six of her albums have opened with over one million sales in a week. The 2023 Time Person of the Year, Swift has appeared on lists such as Rolling Stones 100 Greatest Songwriters of All Time, Billboards Greatest of All Time Artists, and Forbes Worlds 100 Most Powerful Women. Her accolades include 14 Grammy Awards, a Primetime Emmy Award,  40 American Music Awards, 39 Billboard Music Awards, and 23 MTV Video Music Awards; she has won the Grammy Award for Album of the Year, the MTV Video Music Award for Video of the Year, and the IFPI Global Recording Artist of the Year a record four times each."
str_count(swift, ______)
```

::: {.solution exercise="swift"}

```{webr}
#| exercise: swift
#| solution: true
library(stringr)
swift <- "Taylor Alison Swift (born December 13, 1989) is an American singer-songwriter. A subject of widespread public interest, she has influenced the music industry and popular culture through her artistry, especially in songwriting, and entrepreneurship. She is an advocate of artists rights and womens empowerment. Swift began professional songwriting at age 14. She signed with Big Machine Records in 2005 and achieved prominence as a country pop singer with the albums Taylor Swift (2006) and Fearless (2008). Their singles 'Teardrops on My Guitar', 'Love Story', and 'You Belong with Me' were crossover successes on country and pop radio formats and brought Swift mainstream fame. She experimented with rock and electronic styles on her next albums, Speak Now (2010) and Red (2012), respectively, with the latter featuring her first Billboard Hot 100 number-one single, 'We Are Never Ever Getting Back Together'. Swift recalibrated her image from country to pop with 1989 (2014), a synth-pop album containing the chart-topping songs 'Shake It Off', 'Blank Space', and 'Bad Blood'. Media scrutiny inspired the hip-hop-influenced Reputation (2017) and its number-one single 'Look What You Made Me Do'. After signing with Republic Records in 2018, Swift released the eclectic pop album Lover (2019) and the autobiographical documentary Miss Americana (2020). She explored indie folk styles on the 2020 albums Folklore and Evermore, subdued electropop on Midnights (2022), and re-recorded four albums subtitled Taylors Version after a dispute with Big Machine. These albums spawned the number-one songs 'Cruel Summer', 'Cardigan', 'Willow', 'Anti-Hero', 'All Too Well', and 'Is It Over Now?'. Her Eras Tour (2023-2024) and its accompanying concert film became the highest-grossing tour and concert film of all time, respectively. Swift has directed videos and films such as Folklore: The Long Pond Studio Sessions (2020) and All Too Well: The Short Film (2021). Swift is one of the worlds best-selling artists, with 200 million records sold worldwide as of 2019. She is the most-streamed artist on Spotify, the highest-grossing female touring act, and the first billionaire with music as the main source of income. Six of her albums have opened with over one million sales in a week. The 2023 Time Person of the Year, Swift has appeared on lists such as Rolling Stones 100 Greatest Songwriters of All Time, Billboards Greatest of All Time Artists, and Forbes Worlds 100 Most Powerful Women. Her accolades include 14 Grammy Awards, a Primetime Emmy Award,  40 American Music Awards, 39 Billboard Music Awards, and 23 MTV Video Music Awards; she has won the Grammy Award for Album of the Year, the MTV Video Music Award for Video of the Year, and the IFPI Global Recording Artist of the Year a record four times each."
str_count(swift, "Swift")
```
:::

```{webr}
#| exercise: swift
#| check: true
library(stringr)
gradethis::grade_this_code()
```


4. In the above quote, how many *different years* (strings of exactly 4 digits)
   appear?
   
```{webr}
#| exercise: years
library(stringr)
swift <- "Taylor Alison Swift (born December 13, 1989) is an American singer-songwriter. A subject of widespread public interest, she has influenced the music industry and popular culture through her artistry, especially in songwriting, and entrepreneurship. She is an advocate of artists rights and womens empowerment. Swift began professional songwriting at age 14. She signed with Big Machine Records in 2005 and achieved prominence as a country pop singer with the albums Taylor Swift (2006) and Fearless (2008). Their singles 'Teardrops on My Guitar', 'Love Story', and 'You Belong with Me' were crossover successes on country and pop radio formats and brought Swift mainstream fame. She experimented with rock and electronic styles on her next albums, Speak Now (2010) and Red (2012), respectively, with the latter featuring her first Billboard Hot 100 number-one single, 'We Are Never Ever Getting Back Together'. Swift recalibrated her image from country to pop with 1989 (2014), a synth-pop album containing the chart-topping songs 'Shake It Off', 'Blank Space', and 'Bad Blood'. Media scrutiny inspired the hip-hop-influenced Reputation (2017) and its number-one single 'Look What You Made Me Do'. After signing with Republic Records in 2018, Swift released the eclectic pop album Lover (2019) and the autobiographical documentary Miss Americana (2020). She explored indie folk styles on the 2020 albums Folklore and Evermore, subdued electropop on Midnights (2022), and re-recorded four albums subtitled Taylors Version after a dispute with Big Machine. These albums spawned the number-one songs 'Cruel Summer', 'Cardigan', 'Willow', 'Anti-Hero', 'All Too Well', and 'Is It Over Now?'. Her Eras Tour (2023-2024) and its accompanying concert film became the highest-grossing tour and concert film of all time, respectively. Swift has directed videos and films such as Folklore: The Long Pond Studio Sessions (2020) and All Too Well: The Short Film (2021). Swift is one of the worlds best-selling artists, with 200 million records sold worldwide as of 2019. She is the most-streamed artist on Spotify, the highest-grossing female touring act, and the first billionaire with music as the main source of income. Six of her albums have opened with over one million sales in a week. The 2023 Time Person of the Year, Swift has appeared on lists such as Rolling Stones 100 Greatest Songwriters of All Time, Billboards Greatest of All Time Artists, and Forbes Worlds 100 Most Powerful Women. Her accolades include 14 Grammy Awards, a Primetime Emmy Award,  40 American Music Awards, 39 Billboard Music Awards, and 23 MTV Video Music Awards; she has won the Grammy Award for Album of the Year, the MTV Video Music Award for Video of the Year, and the IFPI Global Recording Artist of the Year a record four times each."
str_count(swift, ______)
```

::: {.solution exercise="years"}

```{webr}
#| exercise: years
#| solution: true
library(stringr)
swift <- "Taylor Alison Swift (born December 13, 1989) is an American singer-songwriter. A subject of widespread public interest, she has influenced the music industry and popular culture through her artistry, especially in songwriting, and entrepreneurship. She is an advocate of artists rights and womens empowerment. Swift began professional songwriting at age 14. She signed with Big Machine Records in 2005 and achieved prominence as a country pop singer with the albums Taylor Swift (2006) and Fearless (2008). Their singles 'Teardrops on My Guitar', 'Love Story', and 'You Belong with Me' were crossover successes on country and pop radio formats and brought Swift mainstream fame. She experimented with rock and electronic styles on her next albums, Speak Now (2010) and Red (2012), respectively, with the latter featuring her first Billboard Hot 100 number-one single, 'We Are Never Ever Getting Back Together'. Swift recalibrated her image from country to pop with 1989 (2014), a synth-pop album containing the chart-topping songs 'Shake It Off', 'Blank Space', and 'Bad Blood'. Media scrutiny inspired the hip-hop-influenced Reputation (2017) and its number-one single 'Look What You Made Me Do'. After signing with Republic Records in 2018, Swift released the eclectic pop album Lover (2019) and the autobiographical documentary Miss Americana (2020). She explored indie folk styles on the 2020 albums Folklore and Evermore, subdued electropop on Midnights (2022), and re-recorded four albums subtitled Taylors Version after a dispute with Big Machine. These albums spawned the number-one songs 'Cruel Summer', 'Cardigan', 'Willow', 'Anti-Hero', 'All Too Well', and 'Is It Over Now?'. Her Eras Tour (2023-2024) and its accompanying concert film became the highest-grossing tour and concert film of all time, respectively. Swift has directed videos and films such as Folklore: The Long Pond Studio Sessions (2020) and All Too Well: The Short Film (2021). Swift is one of the worlds best-selling artists, with 200 million records sold worldwide as of 2019. She is the most-streamed artist on Spotify, the highest-grossing female touring act, and the first billionaire with music as the main source of income. Six of her albums have opened with over one million sales in a week. The 2023 Time Person of the Year, Swift has appeared on lists such as Rolling Stones 100 Greatest Songwriters of All Time, Billboards Greatest of All Time Artists, and Forbes Worlds 100 Most Powerful Women. Her accolades include 14 Grammy Awards, a Primetime Emmy Award,  40 American Music Awards, 39 Billboard Music Awards, and 23 MTV Video Music Awards; she has won the Grammy Award for Album of the Year, the MTV Video Music Award for Video of the Year, and the IFPI Global Recording Artist of the Year a record four times each."
str_count(swift, "\\d{4}")
```
:::

```{webr}
#| exercise: years
#| check: true
library(stringr)
gradethis::grade_this_code()
```

5. Extract the names of all songs mentioned in the biography above. (Note 
   that song names are surrounded by single quotes.)

   You will need to use a [_lazy_ regular expression](https://stringr.tidyverse.org/articles/regular-expressions.html#repetition). 
   
   
```{webr}
#| exercise: albums
library(stringr)
swift <- "Taylor Alison Swift (born December 13, 1989) is an American singer-songwriter. A subject of widespread public interest, she has influenced the music industry and popular culture through her artistry, especially in songwriting, and entrepreneurship. She is an advocate of artists rights and womens empowerment. Swift began professional songwriting at age 14. She signed with Big Machine Records in 2005 and achieved prominence as a country pop singer with the albums Taylor Swift (2006) and Fearless (2008). Their singles 'Teardrops on My Guitar', 'Love Story', and 'You Belong with Me' were crossover successes on country and pop radio formats and brought Swift mainstream fame. She experimented with rock and electronic styles on her next albums, Speak Now (2010) and Red (2012), respectively, with the latter featuring her first Billboard Hot 100 number-one single, 'We Are Never Ever Getting Back Together'. Swift recalibrated her image from country to pop with 1989 (2014), a synth-pop album containing the chart-topping songs 'Shake It Off', 'Blank Space', and 'Bad Blood'. Media scrutiny inspired the hip-hop-influenced Reputation (2017) and its number-one single 'Look What You Made Me Do'. After signing with Republic Records in 2018, Swift released the eclectic pop album Lover (2019) and the autobiographical documentary Miss Americana (2020). She explored indie folk styles on the 2020 albums Folklore and Evermore, subdued electropop on Midnights (2022), and re-recorded four albums subtitled Taylors Version after a dispute with Big Machine. These albums spawned the number-one songs 'Cruel Summer', 'Cardigan', 'Willow', 'Anti-Hero', 'All Too Well', and 'Is It Over Now?'. Her Eras Tour (2023-2024) and its accompanying concert film became the highest-grossing tour and concert film of all time, respectively. Swift has directed videos and films such as Folklore: The Long Pond Studio Sessions (2020) and All Too Well: The Short Film (2021). Swift is one of the worlds best-selling artists, with 200 million records sold worldwide as of 2019. She is the most-streamed artist on Spotify, the highest-grossing female touring act, and the first billionaire with music as the main source of income. Six of her albums have opened with over one million sales in a week. The 2023 Time Person of the Year, Swift has appeared on lists such as Rolling Stones 100 Greatest Songwriters of All Time, Billboards Greatest of All Time Artists, and Forbes Worlds 100 Most Powerful Women. Her accolades include 14 Grammy Awards, a Primetime Emmy Award,  40 American Music Awards, 39 Billboard Music Awards, and 23 MTV Video Music Awards; she has won the Grammy Award for Album of the Year, the MTV Video Music Award for Video of the Year, and the IFPI Global Recording Artist of the Year a record four times each."
str_extract_all(swift, ______, simplify=TRUE)
```

::: {.solution exercise="albums"}

```{webr}
#| exercise: albums
#| solution: true
library(stringr)
swift <- "Taylor Alison Swift (born December 13, 1989) is an American singer-songwriter. A subject of widespread public interest, she has influenced the music industry and popular culture through her artistry, especially in songwriting, and entrepreneurship. She is an advocate of artists rights and womens empowerment. Swift began professional songwriting at age 14. She signed with Big Machine Records in 2005 and achieved prominence as a country pop singer with the albums Taylor Swift (2006) and Fearless (2008). Their singles 'Teardrops on My Guitar', 'Love Story', and 'You Belong with Me' were crossover successes on country and pop radio formats and brought Swift mainstream fame. She experimented with rock and electronic styles on her next albums, Speak Now (2010) and Red (2012), respectively, with the latter featuring her first Billboard Hot 100 number-one single, 'We Are Never Ever Getting Back Together'. Swift recalibrated her image from country to pop with 1989 (2014), a synth-pop album containing the chart-topping songs 'Shake It Off', 'Blank Space', and 'Bad Blood'. Media scrutiny inspired the hip-hop-influenced Reputation (2017) and its number-one single 'Look What You Made Me Do'. After signing with Republic Records in 2018, Swift released the eclectic pop album Lover (2019) and the autobiographical documentary Miss Americana (2020). She explored indie folk styles on the 2020 albums Folklore and Evermore, subdued electropop on Midnights (2022), and re-recorded four albums subtitled Taylors Version after a dispute with Big Machine. These albums spawned the number-one songs 'Cruel Summer', 'Cardigan', 'Willow', 'Anti-Hero', 'All Too Well', and 'Is It Over Now?'. Her Eras Tour (2023-2024) and its accompanying concert film became the highest-grossing tour and concert film of all time, respectively. Swift has directed videos and films such as Folklore: The Long Pond Studio Sessions (2020) and All Too Well: The Short Film (2021). Swift is one of the worlds best-selling artists, with 200 million records sold worldwide as of 2019. She is the most-streamed artist on Spotify, the highest-grossing female touring act, and the first billionaire with music as the main source of income. Six of her albums have opened with over one million sales in a week. The 2023 Time Person of the Year, Swift has appeared on lists such as Rolling Stones 100 Greatest Songwriters of All Time, Billboards Greatest of All Time Artists, and Forbes Worlds 100 Most Powerful Women. Her accolades include 14 Grammy Awards, a Primetime Emmy Award,  40 American Music Awards, 39 Billboard Music Awards, and 23 MTV Video Music Awards; she has won the Grammy Award for Album of the Year, the MTV Video Music Award for Video of the Year, and the IFPI Global Recording Artist of the Year a record four times each."
str_extract_all(swift, "'.*?'", simplify=TRUE)
```
:::

```{webr}
#| exercise: albums
#| check: true
library(stringr)
gradethis::grade_this_code()
```

## Scraping Practice I: Cocktail Recipies (Part 2)

Last week, we began to scrape [Hadley's Cocktails](https://cocktails.hadley.nz/)
with an (eventual) goal of creating a "spreadsheet" of recipes by ingredients.

We found the following: 

```{r}
#| eval: false
library(rvest)
BASE_URL <- "https://cocktails.hadley.nz/"

PAGES <- read_html(BASE_URL) |> 
    html_elements("nav a") |> 
    html_attr("href")

read_article <- function(article){
    title <- article |> html_element("h2") |> html_text()
    ingredients <- article |> html_elements("li") |> html_text()
    
    data.frame(title=title, ingredient=ingredients)
}

read_page <- function(stub){
    URL <- paste0(BASE_URL, stub)
    COCKTAILS <- read_html(URL) |> html_elements("article")
    
    map_df(COCKTAILS, read_article)
}

RECIPES_LONG <- map(PAGES, read_page) |> list_rbind()
```

Take this output and use `stringr` and `tidyr` to complete the transition to
a well-formated "wide" set of recipies. 

1. Clean up the `title` column. 
2. Split the `ingredient` column into three new columns: 
   1. Amount 
   2. Unit
   3. Ingredient Name
   
   The following functions may be useful to you: 
```{r}
#| eval: false
get_number_part <- function(x){
    library(dplyr)
    library(stringr)
    x |>
        str_replace_all("[A-Za-z,']", "") |>
        str_trim() |>
        case_match(
            "1" ~ 1, 
            "2" ~ 2,
            "2½" ~ 2.5,
            "1½" ~ 1.5,
            "½" ~ 0.5, 
            "¾" ~ 0.75,
            "1¾" ~ 1.75,
            "¼" ~ 0.25,
            "1¼" ~ 1.25,
            "3" ~ 3,
            "4" ~ 4, 
            "5" ~ 5, 
            "6" ~ 6, 
            "7" ~ 7, 
            "8" ~ 8
        )
}

get_cocktail_unit <- function(x){
    library(dplyr)
    library(stringr)
    case_when(
        str_detect(x, "oz") ~ "oz",
        str_detect(x, "dash") ~ "dash",
        str_detect(x, "drop") ~ "drop",
        str_detect(x, " t ") ~ "t", 
        str_detect(x, "chunk") ~ "chunk",
        str_detect(x, "leaves") ~ "leaves",
        str_detect(x, " cm ") ~ "cm"
    )
}
```

   You will need to create a third helper function to pull out
   the actual ingredient name. You can implement this with a (somewhat complex)
   `str_replace_all`. 
   
   You should wind up with a table that looks something like
   
| Cocktail | Ingredient | Unit | Amount |
|----------|------------|------|--------|
| Bachelor | rum, dark  | oz   | 1      |
| Bachelor | Meletti    | oz   | 1      |

3. Combine the ingredient and unit columns so that `Meletti` and ` oz ` in 
   two separate columns becomes `Meletti (oz)` in a single column. 
4. Use a `pivot_*` function to create a new _wide_ table with each ingredient
   as a column. 
   
   Which `pivot` operation do you want to use here? How should the empty cells
   be treated?
   
::: {.callout-note title="Solution" collapse="true"}

```{r}
#| eval: false

get_ingredient_name <- function(x){
    str_remove_all(x, "[0-9¼½¾]|( oz )|( dashes )|( dash )|( drops )|( drop )|( t )|( chunks )|( chunk )|( leaves )|( leaf )|( cm )")
}

RECIPES_LONG |>
    # Remove duplicates from import process
    # When we import each ingredient page, we get duplicates
    # as drinks are listed on multiple ingredient pages.
    distinct() |>
    mutate(title  = str_trim(title), 
           amount = get_number_part(ingredient), 
           unit  = get_cocktail_unit(ingredient), 
           ingredient = get_ingredient_name(ingredient)) |>
    # For some ingredients, e.g. a lemon twist, the implied
    # but unstated quantity is 1
    mutate(amount = case_when(
        is.na(amount) ~ 1, 
        TRUE ~ amount), 
        ingredient = str_to_title(ingredient)) |>
    rename(Cocktail = title, 
           Amount = amount, 
           Unit = unit, 
           Ingredient = ingredient) |>
    mutate(Ingredient = case_when(
        is.na(Unit) ~ Ingredient, # Handle unit-less ingredients
        TRUE ~ paste0(Ingredient, " (", Unit, ")"))) |>
    # Spanish Coffee lists orange liqueur twice, so let's 
    # add up repeated ingredients before pivoting.
    # (I think this is the only one)
    group_by(Cocktail, Ingredient) |>
    summarize(Amount = sum(Amount)) |>
    ungroup() |>
    pivot_wider(id_cols = Cocktail, 
                names_from = Ingredient, 
                values_from = Amount, 
                values_fill = 0) |>
    select("Cocktail", sort(tidyselect::peek_vars()))
```

:::
## Scraping Practice II: Quotes

Next, let's analyze the website [https://quotes.toscrape.com](https://quotes.toscrape.com),
a website designed to practice web-scraping. 

Scrape the contents of that website--note that quotes continue for multiple 
pages--and answer the following questions. You can check most of these "by hand"
but you need to compute your answers in code!

1. How many quotes are on this website (all pages)? 
2. How many quotes are tagged "Death"?
3. What is the longest quote (by number of characters)? The `nchar` function
   will be helpful. 
4. How many quotes are by (or at least are attributed to) Albert Einstein?
5. Of all authors quoted, who has the earliest (estimated) birthday? 

## Basic Statistical Modeling

`R` has a number of powerful built-in statistical methods, many of which you will
explore in other courses, *e.g.*, `lm` and `glm` in your regression course. 
Time allowing, we're going to explore an alternative approach to statistics using
_computational inference_. While these techniques can be implemented by hand, 
we're going to use the [`infer` package](https://infer.netlify.app/).

**Time Allowing**:

- Discussion of computational inference
- Demonstration with `infer` package. 
