---
lab_num: "`r lab <- 11; lab`"
lab_title: "Predictive Modeling - Inference and Prediction in `R`"
engine: knitr 
---

{{< include _setup.qmd >}}

# Review Exercise {#review}

For this final warm-up exercise, you are going to use a combination of the
skills learned over the semester to create a map of the countries of the world,
colored by population density.

1) Using the data at <https://www.scrapethissite.com/pages/simple/>, determine 
   the _capital_ of the _most densely populated_ country in the world. 

   To answer this: 

   - Scrape the webpage into a four-column data frame with columns:
     `name`, `capital`, `population`, `area`
   - Ensure that the population and area columns are properly
     converted to numeric values
   - Compute population density as `population/area`
   - Select the country with the highest density and determine its capital
   

::: {.callout-tip title="Solution" collapse="true"}

```{r}
#| message: false
library(tidyverse)
library(httr2)
library(rvest)

parse_country <- function(country_element){
    name <- country_element |> 
        html_element("h3") |> 
        html_text2()
    
    capital <- country_element |> 
        html_element(".country-capital") |> 
        html_text2()
    
    population <- country_element |> 
        html_element(".country-population") |> 
        html_text2() |> 
        as.numeric()
    
    area <- country_element |> 
        html_element(".country-area") |> 
        html_text2() |> 
        as.numeric()
    
    data.frame(name, capital, population, area)
}

COUNTRIES <- read_html("https://www.scrapethissite.com/pages/simple/") |>
    html_elements(".country") |>
    map(parse_country) |>
    list_rbind() |>
    mutate(density = population / area)
    

COUNTRIES |> slice_max(density, n=1) |> pull(capital)
```

:::

2) When dealing with countries, it is dangerous to use raw names as these are not
   entirely standardized. *E.g.*, do we live in the "United States" or the "United
   States of America"? To address this, we want to use a standard format like
   the ISO 3-digit country codes. 
   
   The website <https://peric.github.io/GetCountries/> provides this information
   interactively, but is a bit difficult to access programmatically. This page
   actually loads a `JSON` object that will provide all of the information we
   need. Use the Network tools in your web browser's developer tools to identify
   the request that loads this data. (You may need to hit the `Generate` button
   again after opening the Network tools to see the request being made.)
   
   Read this data into `R` and create a data frame with (at least) columns for
   the 3 digit ISO country code, the continent, the country name, 
   the population, and the area in square kilometers. Make sure the population 
   and area have been properly converted into numeric values. 

::: {.callout-tip title="Solution" collapse="true"}

```{r}
#| message: false
#| cache: true
library(tidyverse)
library(jsonlite)

COUNTRIES <- read_json("https://secure.geonames.org/countryInfoJSON?username=dperic", 
          simplify=TRUE) |>
    pluck("geonames") |>
    select(iso_code = isoAlpha3, 
           name = countryName, 
           continent = continentName,
           area = areaInSqKm, 
           population = population) |>
    mutate(area = as.numeric(area), 
           population = as.numeric(population)) |> 
   filter(area > 0) # Ignore countries / political entities without land area
```

:::
   
3) Next, we are going to use an API to access data on country boundaries provided
   by the [GeoBoundaries](https://www.geoboundaries.org) project. Specifically,
   we want to download shape files for all countries of the world. We do not want
   to use the _high-resolution_ shapefiles as they are highly detailed and will
   slow down processing. Thankfully, the GeoBoundaries project already provides
   simplified representations of each country more suitable for plotting. 
   
   Review the API documentation at <https://www.geoboundaries.org/api.html#api>
   and make a call for `gbOpen` data for `ALL` countries at `ADM0` level (country
   boundaries). The result will be a JSON object with elements for each country.
   Parse this JSON object in `R` and pull out (at least) the
   `boundaryName`, `boundaryISO`, and `simplifiedGeometryGeoJSON` columns 
   into a data frame. 

::: {.callout-tip title="Solution" collapse="true"}

```{r}
#| cache: true
library(tidyverse)
library(jsonlite)

API_REFERENCE <- read_json("https://www.geoboundaries.org/api/current/gbOpen/ALL/ADM0/", 
                           simplify = TRUE) |>
    select(country = boundaryName, 
           iso_code = boundaryISO,
           continent = Continent, 
           simple_geojson = simplifiedGeometryGeoJSON)
```



:::

4) Join the result of the previous two sections together using an `inner_join`. 
   How many countries were in each data set? Which countries are lost in the
   inner join? What is their total population? 
   
   *Hint*: An `anti_join` may be useful here. 
   
   In real work, you would want to do some additional manual cleaning to better
   align these two data sets, but we are not going to do so today to save time.
   There is at least one _major_ missing country in this merge. 

::: {.callout-tip title="Solution" collapse="true"}

```{r}
library(tidyverse)

JOINED <- inner_join(COUNTRIES, API_REFERENCE, join_by(iso_code))

UNJOINED <- anti_join(COUNTRIES, API_REFERENCE, join_by(iso_code))

UNJOINED |> summarize(n_missing = n(), 
                      pop_missing = sum(population))
```

Wow! That's over a billion people. 

```{r}
UNJOINED |> slice_max(population, n=1)
```

So India is missing - that's a problem!

:::

5)  **Optional** In the previous section, we found that one major country was
    missing. After some poking around, it turns out that the missing country
    is linked in the `ADM1` listing and not the `ADM0` listing. This appears
    to be just a manual error, but the [related GitHub
    issue](https://github.com/wmgeolab/geoBoundaries/issues/4265) has not yet
    been addressed, so we need to handle it ourselves. 
   
    Do this by making a second request to request the `ADM1` listings and
    then filtering out any duplicates. Next join this to the response you
    had from before, now using a `left_join()` to avoid dropping results
    from intermediate tables. Finally, use the `coalesce()` function to handle
    redundant columns. Note that the `coalesce()` column takes a set of
    column names and gets the first non-`NA` value so it is great for things
    like this: 
   
    ```{r}
    coalesce_df <- tribble(
    ~a, ~b, ~c, 
    1, 2, 3, 
    4, NA, 6, 
    NA, NA, 9)
    coalesce_df |> mutate(fixed_val = coalesce(a, b, c))
    ```


::: {.callout-tip title="Solution" collapse="true"}

```{r}
#| cache: true

library(tidyverse)

API_REFERENCE_ADM0 <- "https://www.geoboundaries.org/api/current/gbOpen/ALL/ADM0/" |>
    read_json(simplify = TRUE) |>
    select(country = boundaryName, 
           iso_code = boundaryISO,
           simple_geojson = simplifiedGeometryGeoJSON) |>
    distinct()


API_REFERENCE_ADM1 <- "https://www.geoboundaries.org/api/current/gbOpen/ALL/ADM1/" |>
    read_json(simplify = TRUE) |>
    select(country = boundaryName, 
           iso_code = boundaryISO,
           simple_geojson = simplifiedGeometryGeoJSON) |>
    distinct()

JOINED <- COUNTRIES |> 
   left_join(API_REFERENCE_ADM0, 
             join_by(iso_code)) |> 
   left_join(API_REFERENCE_ADM1, 
             join_by(iso_code)) |>
   mutate(country = coalesce(country.x, country.y), 
          simple_geojson = coalesce(simple_geojson.x, simple_geojson.y)) |>
   select(-country.x, -country.y, -simple_geojson.x, -simple_geojson.y) |>
   filter_out(is.na(simple_geojson))
```

:::


6)  Make a series of web requests to read the `geojson` files for each of 
    the countries using the URLs obtained from the API. Since you will want to
    make this request sequentially, use the `map()` function to call   
    `sf::st_read` repeatedly. You will also want to use `unnest()` to remove 
    the resulting list-column. 
   
::: {.callout-tip title="Solution" collapse="true"}

```{r}
#| cache: true
#| message: false
library(tidyverse)
library(sf)

JOINED_GEOMETRY <- JOINED |>
    mutate(geometry = map(simple_geojson, st_read, quiet=TRUE)) |>
    unnest(geometry) |>
    ungroup()
```

:::

7) Create a chloropleth visualization with each country colored according to
   its population density. 
   
   Note that this will take a little while due to the high-resolution
   shapefiles used above. If you want a faster plot, you can use 
   `sf::st_simplify()` to reduce the resolution of the boundaries.
   
::: {.callout-tip title="Solution" collapse="true"}

```{r draw_map}
#| cache: true
#| warning: false
JOINED_GEOMETRY |>
    mutate(density = population / area) |>
    ggplot(aes(fill=density, geometry=geometry)) + 
    geom_sf() + 
    theme_bw() + 
    theme(legend.position="bottom") + 
    scale_fill_distiller(name="Population Density", 
                         palette=4, 
                         direction=1,
                         trans=scales::pseudo_log_trans(), 
                         breaks=c(0, 3, 10, 30, 100, 300, 1000, 3000, 10000, 30000)) + 
    guides(fill = guide_colorbar(barwidth=unit(10, "cm")))
```

:::

# Classical Statistical Models in `R` {#lm}

Using the countries data downloaded above, use statistical models like `lm()`
and `mgcv::gam()` to answer the following: 

1. Recall the `flights` data from the `nycflights13` package. Load this package
   and fit a linear model to predict air time as a function of distance. 
   
   The intercept term of this model can be interpreted as the "fixed cost" of
   a flight - the unavoidable time taken to travel between the gate and the
   runway. What is the estimate of this cost based on the model you fit?
   
::: {.callout-tip title="Solution" collapse="true"}

```{r}
#| message: false
library(tidyverse)
library(nycflights13)
library(broom)
air_time_model <- lm(air_time ~ distance, data=flights) 

air_time_model |> tidy()
```

From this, we see that the 'fixed cost' of air travel is approximately
`r round(coef(air_time_model)["(Intercept)"], 1)` minutes. 

:::

2. The same linear model can be used to estimate flight speed. In particular, 
   the distance coefficient (which has units of minutes / mile) can be 
   extracted and transformed into miles per hour. What is the average speed of
   a plane leaving NYC?

::: {.callout-tip title="Solution" collapse="true"}

```{r}
#| message: false
library(tidyverse)
library(nycflights13)
library(broom)
air_time_model <- lm(air_time ~ distance, data=flights) 

air_time_model |> 
    tidy() |> 
    filter(term == "distance") |> 
    mutate(hours_per_mile = estimate / 60, 
           miles_per_hour = 1/hours_per_mile) |> 
    pull(miles_per_hour) |>
    round(2) |>
    paste("miles per hour")
```

:::

3. Modify your analysis above to estimate a different fixed cost for each of the
   three NYC-area airports. (Maintain a single estimate of speed for all 
   airports.) Which seems to be the most efficient? 
   
   Use `broom::tidy()` and some regular expressions to create a table that looks
   like this: 
   
   | Airport | Estimated Fixed Cost |
   |---------|----------------------|
   | EWR     | NNN.NN               |
   | LGA     | NNN.NN               |
   | JFK     | NNN.NN               |
   
   *Hint*: Your estimated coefficients will be easier to interpret if you remove
   the intercept term from your model by including `-1` at the end of the
   formula. 

::: {.callout-tip title="Solution" collapse="true"}

```{r}
#| message: false
library(tidyverse)
library(nycflights13)
library(broom)
library(gt)
lm(air_time ~ distance + origin - 1, data=flights) |> 
    tidy() |> 
    filter(str_detect(term, "origin")) |> 
    mutate(term=str_remove(term, "origin")) |> 
    gt(id="tbl_airport1") |>
    cols_hide(c(statistic, p.value)) |>
    cols_label(term="Airport", 
               estimate="Estimated Fixed Transit Cost (Minutes)") |>
    cols_merge(c(estimate, std.error), 
               pattern="{1}±{2}") |>
    fmt_number(c(estimate, std.error))
```

:::

4. Now modify your model to allow different estimated flight speeds for each
   airport. Which airport seems to have the highest average flight speed?
   
   *Hint*: Use the `*` operator to allow for _interactions_ between airport
   and distance. This will estimate different slopes for each group. 
   
   *Hint*: If you find your coefficients hard to interpret, consider adding
   `-distance` to the end of your formula. What does this do? Why does it help?
   
::: {.callout-tip title="Solution" collapse="true"}

```{r}
library(tidyverse)
library(nycflights13)
library(broom)
library(gt)
lm(air_time ~ distance * origin - 1 - distance,
    data=flights) |>
    tidy() |>
    mutate(airport = str_sub(term, start=-3), 
           effect = if_else(str_detect(term, "distance"), "speed", "fixed")) |>
    select(estimate, airport, effect) |>
    pivot_wider(names_from=effect, values_from=estimate) |>
    mutate(speed = 60 / speed) |>
    gt(id="tbl_airport2") |>
    cols_label(airport=md("**Origin Airport**"),
               fixed=md("Gate Taxi Time (Minutes)"),
               speed=md("Estimated Air Speed (mph)")) |>
   fmt_number(c(fixed, speed), decimals = 1)
```

:::

5. Use the `augment` function to compute both the RMSE (Root Mean Squared Error)
   and the MedAE (Median Absolute Error) of the model you fit in the
   previous section. 
   
::: {.callout-tip title="Solution" collapse="true"}
   
```{r}
library(tidyverse)
library(nycflights13)
library(broom)
lm(air_time ~ distance * origin - 1 - distance, data=flights) |>
    augment() |>
    summarize(rmse = sqrt(mean(.resid^2)), 
              medae = median(abs(.resid)))
```

:::

6. In which month did the model have the lowest RMSE? 

   *Hint*: Pass the `newdata` argument to `augment` to get back 
   additional columns you need. 

::: {.callout-tip title="Solution" collapse="true"}

```{r}
lm(air_time ~ distance * origin - 1 - distance, data=flights) |>
    augment(newdata=flights) |>
    group_by(month) |>
    summarize(rmse = sqrt(mean(.resid^2, na.rm=TRUE))) |>
    slice_min(rmse)
```

So our model seems to be most predictive in October.

:::

7. The `yardstick` package provides tools for evaluating model performance. 
   The `rsq` function computes $R^2$ statistics. It takes two column names as
   arguments (in addition to the data frame being analyzed): `truth`,
   the observed values ($y$), and `estimate`, the model predictions ($\hat{y}$).
   It can be used like this: 
   
```{r}
#| echo: false
#| message: false
conflicted::conflicts_prefer(datasets::penguins)
library(yardstick)
```

```{r}
library(broom)
library(yardstick)
lm(body_mass ~ flipper_len * species, data=penguins) |> 
    augment() |> 
    rsq(truth=body_mass, estimate=.fitted)
```
   
   Use it to compute the $R^2$ of your model on a month-by-month and 
   per-airport basis. At what airport and in what month does this model exhibit
   the lowest $R^2$? What does that tell you? How does this compare to the 
   overall (full-sample) $R^2$ of the model that `glance` provides? 
   
::: {.callout-tip title="Solution" collapse="true"}

```{r}
lm(air_time ~ distance * origin - 1 - distance, data=flights) |>
    augment(newdata=flights) |>
    group_by(month, origin) |>
    rsq(truth=air_time,
        estimate=.fitted) |>
    slice_min(.estimate)
```

Our model performs (relatively) worst at LGA in January, suggesting that other
effects (*e.g.* weather) play the largest role here, but the effect is quite
minimal. 

```{r}
lm(air_time ~ distance * origin - 1 - distance, data=flights) |>
    glance()
```

So the drop in $R^2$ is 99.5% to 95.6%: this isn't nothing...

:::

# Computational Inference {#inference}

**Using the countries data downloaded above, test the null hypothesis that the 
_standard deviation of population density is equal in African and non-African
countries_.** 

Before doing so, let's review the concept of _bootstrapping_, a variant of the
computationally-powered inference you reviewed in this week's pre-assignment 
that attempts to relax the requirement that we specify a null distribution 
precisely.

### Refresher

As discussed [in this week's pre-assignment](../pre/pa11.qmd), the core of
computationally-intensive approach to statistical inference is to get _many_ 
'pseudo-samples' from a null distribution, to compute the test statistic on 
those 'pseudo-samples', and to compare the distribution of the test statistic 
against the observed value. Doing so requires us to have some sense of a null 
distribution: it can be a parametric posited *a priori* or, as we do here, it 
can be based on re-sampling the data at hand. 

In this case, we don't want to posit a full model for country density: we only 
want to see if our data supports the claim that the standard deviation of 
population density is the same within Africa and outside of Africa. Because we
don't require a full model to define our null hypothesis, you might be tempted
to use a 'working model' based on a normal distribution. A quick plot of the
data suggests that this data is _quite_ heavy tailed (there are some very small,
very densely populated areas like Monaco, Hong Kong, and Singapore) so this
strategy of 'get a good enough' model isn't likely to be productive. 

But we don't _exactly_ need a model for country density - we really just need the
ability to _sample_ from such a null model that doesn't distinguish Africa and 
non-Africa. Is there a way we can do that instead without the full overhead of
a 'proper' model?

We turn to the technique of _bootstrapping_. In essence, we use the data
we have _as a rough model_ of the (unknown) distribution of interest. How does 
this work? 

First, note that IID sampling + something like density estimation is a universal
estimator. No matter the distribution (within reason), if we have enough data
the sample density estimate (or histogram or empirical CDF) will converge to
the corresponding true distribution. While this seems intuitive, not everything 
in statistics is like this: *e.g.*, no matter how much data you have, linear
regression will never get the 'true' relationship if it is actually non-linear.
Here, we're replying the fact that things like a density estimate or the ECDF
are _non-parametric_ and can fit essentially any shape. 

Let $\hat{\mathbb{P}}_n$ be the 'estimated CDF' based on the observed data: 
we can evaluate $\hat{\mathbb{P}}_n$ by: 

$$\hat{\mathbb{P}}_n(x) = \frac{1}{n} \sum_{i=1}^n 1_{x \leq x_i} $$

That is, to estimate the CDF at any point $x$, we simply look at the fraction
of points that fall below $x$. The law of large numbers[^gc] tells us that

$$\hat{\mathbb{P}}_n(x) \to \mathbb{P}(x) \text{ for all } x$$ or just

$$\hat{\mathbb{P}}_n \to \mathbb{P}$$


where $\mathbb{P}$ is the true distribution used to generate the data. This
suggests to us that, while we might be interested in the variance of our
test statistic $T$ under distribution $\mathbb{P}$, we can approximate it 
using $\hat{\mathbb{P}}_n$. That is, we hope that

$$\text{Variance}_{\hat{\mathbb{P}}_n}(T) \to \text{Variance}_{\mathbb{P}}(T)$$

for any test statistic $T$ as the sample size $n \to \infty$. Notably, this 
should work for almost _any_ distribution $\mathbb{P}$ and any test statistic
$T$. Similar results hold for quantiles of $T$ as well:

$$q\text{-Quantile}_{\hat{\mathbb{P}}_n}(T) \to q\text{-Quantile}_{\mathbb{P}}(T) \text{ for all } q \in (0, 1)$$

So this means that we can approximate $\text{Quantile}_{\mathbb{P}}(T)$ if we
can estimate $\text{Quantile}_{\hat{\mathbb{P}}_n}(T)$. This leaves us with 
just one question: how do we sample from $\hat{\mathbb{P}}_n$? 

Note that $\hat{\mathbb{P}}_n$ is the CDF of a discrete distribution with equal 
jumps at each observation: that is, it's uniform over the observed samples. So 
how can we sample from a uniform distribution? We simply pick one of its values
at random. And how do we get a second sample, we just get another value at
random once again. 

::: {.callout-info title="Bootstrap Resampling"}

Bootstrap sampling is often described as _sampling with replacement_. This is
true, but I don't like it as an explanation because it seems a bit magical. 

The true idea of bootstrapping is that we are just getting $n$ samples from a 
(uniform) distribution over the original data. When we take any number of
samples from a discrete distribution, there is always a possibility of ties, 
regardless of how we came up with that distribution. "Bootstrap sampling with
replacement" is then just weirdly redundant: regular sampling of any 
distribution always allows with replacement."[^cont] (You would never specify 
'flipping a coin three times _with replacement_'!)

To recap: in bootstrapping

- We build a discrete uniform distribution from the original data
- We sample from that distribution to estimate quantities of interest

The fact we get samples _with replacement_ is not because we are bootstrapping
*per se*, it's just because we are IID sampling from a _distribution_ and the 
values that distribution don't change based on what we saw previously.

:::

[^cont]: Even sampling from a continuous distribution _theoretically_ allows 
us to get repeated values; in practice, these are 'probability zero' types of
events and never happen. But it's not _actually_ impossible - just _never going
to happen_!

Before we apply this to statistical hypothesis testing, let's just use this
idea to estimate the _quantiles_ of the standard deviation of the country 
density. The point estimate is straightforward: 

```{r}
COUNTRIES |> 
   mutate(density = population / area) |> 
   summarize(sd(density))
```

But how can we get a sense of the variance of this quantity? Ideally, we would
have _many_ sets of countries and could compute standard deviation _within_ each
set and then compute a variance _across_ sets. Resampling will get us there: 

```{r}
COUNTRIES |> 
   mutate(density = population / area) |> 
   mutate(density_resampled = sample(density, replace=TRUE)) |> 
   summarize(sd(density_resampled))
```

A bit different - but not too much. But one 'alternate world value' isn't enough
to get quantiles. We need many! There are several ways to code this in `dplyr`, 
but one of the easiest is to use the `expand_grid()` function to induce 
replicates: 

```{r}
#| echo: false
#| warning: false
#| message: false
conflicted::conflict_prefer("filter", "dplyr")
```

```{r}
COUNTRIES |> 
    filter(area > 0) |> 
    mutate(density = population / area) |> 
    expand_grid(replicate=1:100) |> 
    mutate(density_resampled = sample(density, replace=TRUE)) |> 
    group_by(replicate) |> 
    summarize(density_sd_resampled = sd(density_resampled)) |> 
    summarize(density_sd_q95 = quantile(density_sd_resampled, 0.95), 
              density_sd_q05 = quantile(density_sd_resampled, 0.05))
```

To understand this better, let's apply it to a smaller example with 9 penguins:

```{r}
small_penguins <- penguins |> 
   drop_na() |> 
   group_by(species) |> 
   slice_sample(n=3)

small_penguins
```

If we use `expand_grid()`, we induce _copies_ of each data set: 

```{r}
small_penguins_eg <- small_penguins |> expand_grid(replicate = 1:3)

small_penguins_eg
```

We then use `sample(..., replace=TRUE)` to perform bootstrap sampling on the 
`body_mass` column: 

```{r}
small_penguins_eg_bs <- small_penguins_eg |> 
   mutate(body_mass_bs = sample(body_mass, replace=TRUE))
```

And finally, we use the replicate column to get multiple samples we can use 
to get a sense of variance: 

```{r}
small_penguins_eg_bs |> 
   group_by(replicate) |> 
   summarize(avg_body_mass = mean(body_mass_bs)) |> 
   summarize(sd_avg_body_mass = sd(avg_body_mass))
```

With only three samples, this doesn't give us much strength, so let's just let
the number of bootstrap samples go to 5000. 

```{r}
small_penguins |> 
   expand_grid(replicate = 1:5000) |>
   mutate(body_mass_bs = sample(body_mass, replace=TRUE)) |>
   group_by(replicate) |> 
   summarize(avg_body_mass = mean(body_mass_bs)) |> 
   summarize(sd_avg_body_mass = sd(avg_body_mass))
```

Usually, around 1000 bootstrap samples is enough to account for the randomness 
from the `sample()` in the third line.

Returning to our population density SD example from above: 

```{r}
COUNTRIES |> 
    filter(area > 0) |> 
    mutate(density = population / area) |> 
    expand_grid(replicate=1:2000) |> 
    mutate(density_resampled = sample(density, replace=TRUE)) |> 
    group_by(replicate) |> 
    summarize(density_sd_resampled = sd(density_resampled)) |> 
    summarize(density_sd_q95 = quantile(density_sd_resampled, 0.95), 
              density_sd_q05 = quantile(density_sd_resampled, 0.05))
```

This gives us a sense of the _range_ of plausible standard deviations - or, more
precisely, the range of standard deviations that are consistent with the 
observations we have seen. 

We can plot these as well: 

```{r}
COUNTRIES |> 
    filter(area > 0) |> 
    mutate(density = population / area, 
           true_density_sd = sd(density)) |> 
    expand_grid(replicate=1:200) |> 
    mutate(density_resampled = sample(density, replace=TRUE)) |> 
    group_by(replicate) |> 
    summarize(density_sd_resampled = sd(density_resampled), 
              true_density_sd = first(true_density_sd)) |> 
    mutate(density_sd_q95 = quantile(density_sd_resampled, 0.95), 
           density_sd_q05 = quantile(density_sd_resampled, 0.05)) |> 
    ggplot(aes(x=density_sd_resampled)) + 
    geom_density() + 
    theme_bw() + 
    xlab("Standard Deviation of Density") + 
    ylab("Bootstrap Sampling Density") + 
    geom_area(stat="density", 
              aes(fill = after_stat(x) > quantile(x, 0.05) & 
                         after_stat(x) < quantile(x, 0.95))) + 
    scale_fill_manual(name="Central 90%", 
                      values=c("TRUE"="#4682B477",
                               "FALSE"=NA)) + 
    guides(fill="none") + 
    geom_vline(aes(xintercept=true_density_sd), 
               color="red4", 
               linewidth=2, 
               linetype=2) + 
    ggtitle("Sampling Distribution of Standard Deviation of Country Densities", 
            "True value (Red) lies in the middle of the sampling distribution")
```

We see here that, unsurpsingly, the bootstrap distribution is located around
the 'true' value, with a second mode that likely corresponds to the times that
the highest- and lowest-density countries are randomly excluded. This second
mode provides us strong evidence that country populations and areas are not 
normally distributed (if that weren't already obvious).

In this case, if we, for whatever reason, had a null belief that the 'true'
standard deviation was near 3000, it would be clear that this is quite 
incompatible with the observed value (red line) but also with the variance 
around that observed value (blue area). 

[^gc]: More precisely, a *uniform* law of large numbers called the 
Glivenko-Cantelli theorem. 


### Boostrap Testing in Action

Now it's your turn! Using the countries data downloaded above, test the null 
hypothesis that the **standard deviation of population density is equal in 
African and non-African countries**.

1. Create a new (Boolean) variable to determine whether a country is in 
   Africa or not. 
   
::: {.callout-tip title="Solution" collapse="true"}

```{r}
COUNTRIES <- COUNTRIES |>
   mutate(in_africa = (continent == "Africa"), 
          density = population / area)
```

:::
   
2. Compute the standard deviation of density for both African and non-African
   countries, then compute the _difference_ in standard deviations. This will be
   the 'observed' value we want to compare against the bootstrap samples. 
   

::: {.callout-tip title="Solution" collapse="true"}

```{r}
delta_sd_observed <- COUNTRIES |> 
    group_by(in_africa) |> 
    summarize(sd_density = sd(density)) |> 
    mutate(in_africa = if_else(in_africa, "africa", "nonaf")) |>
    pivot_wider(names_from=in_africa,
                values_from=sd_density) |> 
    mutate(delta_sd = nonaf - africa) |> 
    pull(delta_sd)
```

:::
   
3. Modify the strategy above to compute standard deviations for African and
   non-African countries in each replicate. 
   
   *Hint*: When forming replicates, we are assuming that the distribution of
   African and non-African countries is the same, so we do not need to modify
   our bootstrap sampling in any significant way. Here, the two-group structure
   is only used to define our test statistic. 
   
::: {.callout-tip title="Solution" collapse="true"}

```{r}
COUNTRIES |> 
    expand_grid(replicate=1:1000) |> 
    mutate(density_resampled = sample(density, replace=TRUE)) |> 
    group_by(replicate, in_africa) |> 
    summarize(density_sd_resampled = sd(density_resampled))
```

:::

3. Now compute the _difference_ in SDs for each replicate and construct 
   a density plot for the middle 95% of the distribution of differences.  

::: {.callout-tip title="Solution" collapse="true"}

```{r}
COUNTRIES |> 
    expand_grid(replicate=1:1000) |> 
    mutate(density_resampled = sample(density, replace=TRUE)) |> 
    group_by(replicate, in_africa) |> 
    summarize(sd_density_resampled = sd(density_resampled)) |>
    mutate(in_africa = if_else(in_africa, "africa", "nonaf")) |>
    pivot_wider(names_from=in_africa,
                values_from=sd_density_resampled) |> 
    mutate(delta_sd_resampled = nonaf - africa) |>
    ggplot(aes(x=delta_sd_resampled)) + 
    geom_density() + 
    theme_bw() + 
    xlab("Standard Deviation of Difference in Population Densities") + 
    ylab("Bootstrap Sampling Density") + 
    geom_area(stat="density", 
              aes(fill = after_stat(x) > quantile(x, 0.05) & 
                         after_stat(x) < quantile(x, 0.95))) + 
    scale_fill_manual(name="Central 90%", 
                      values=c("TRUE"="#4682B477",
                               "FALSE"=NA)) + 
    guides(fill="none")
```

:::
   
4. Add a vertical line for your observed value (computed above). If this
   falls _outside_ the shaded area, we can _reject_ the null of no difference
   between African and non-African countries. 
   
::: {.callout-tip title="Solution" collapse="true"}

```{r}
COUNTRIES |> 
    expand_grid(replicate=1:1000) |> 
    mutate(density_resampled = sample(density, replace=TRUE)) |> 
    group_by(replicate, in_africa) |> 
    summarize(sd_density_resampled = sd(density_resampled)) |>
    mutate(in_africa = if_else(in_africa, "africa", "nonaf")) |>
    pivot_wider(names_from=in_africa,
                values_from=sd_density_resampled) |> 
    mutate(delta_sd_resampled = nonaf - africa) |>
    ggplot(aes(x=delta_sd_resampled)) + 
    geom_density() + 
    theme_bw() + 
    xlab("Standard Deviation of Difference in Population Densities") + 
    ylab("Bootstrap Sampling Density") + 
    geom_area(stat="density", 
              aes(fill = after_stat(x) > quantile(x, 0.05) & 
                         after_stat(x) < quantile(x, 0.95))) + 
    scale_fill_manual(name="Central 90%", 
                      values=c("TRUE"="#4682B477",
                               "FALSE"=NA)) + 
    guides(fill="none") + 
    geom_vline(xintercept=delta_sd_observed, 
               color="red4", 
               linewidth=2)
```

From here, we see that while there is a wider spread among non-African countries
than there is among African countries, it is not inconsistent with 
'pure randomness' that we could get if we randomly swapped the labels on this
data. In particular, we see that we can get differences in standard deviations 
of $\pm 2000$, so the observed difference of `r delta_sd_observed |> round()` 
is not that big in context.[^perm]

:::

[^perm]: We also could have _permuted_ the labels of Africa / non-Africa here
to get a similar type of test. In finite samples, this test is likely to be just 
a bit more reliable (since we don't rely on the 'enough data to estimate a 
true distribution' argument that powers bootstrapping), but the differences 
aren't huge. In practice, this comes down to _shuffling_ rather than 
_resampling_, which can also be done with the `sample` function, but now with
`sample(..., replace=FALSE)`. Permutation testing is a bit simpler and better
when possible, but not everything is easily done with a permutation test.

::: {.callout-tip title="All In One Go" collapse="true"}

If you want to avoid having `delta_sd_observed` as its own variable, this
example shows how it can be handled as its own column to get the same results:

```{r}
COUNTRIES |> 
    expand_grid(replicate=1:1000) |> 
    mutate(density_resampled = sample(density, replace=TRUE)) |> 
    group_by(replicate, in_africa) |> 
    summarize(sd_density_resampled = sd(density_resampled), 
              sd_density = sd(density), 
              .groups="drop") |>
    mutate(in_africa = if_else(in_africa, "africa", "nonafr")) |>
    pivot_wider(names_from=in_africa, 
                values_from=c(sd_density, sd_density_resampled)) |>
    mutate(delta_sd_resampled = sd_density_resampled_nonafr - sd_density_resampled_africa, 
           delta_sd = sd_density_nonafr - sd_density_africa ) |>
    ggplot(aes(x=delta_sd_resampled)) + 
    geom_density() + 
    theme_bw() + 
    xlab("Standard Deviation of Difference in Population Densities") + 
    ylab("Bootstrap Sampling Density") + 
    geom_area(stat="density", 
              aes(fill = after_stat(x) > quantile(x, 0.05) & 
                         after_stat(x) < quantile(x, 0.95))) + 
    scale_fill_manual(name="Central 90%", 
                      values=c("TRUE"="#4682B477",
                               "FALSE"=NA)) + 
    guides(fill="none") + 
    geom_vline(aes(xintercept=delta_sd), 
               color="red4", 
               linewidth=2) + 
   ggtitle("Comparison of Population Density Variance Among African and Non-African Countries", 
           "Insignificant Evidence that Non-African Countries Exhibit Higher Density Variance")
```

:::

I like to code these types of analyses 'by hand' and think that it is a useful
exercise to clarify the logic of statistical testing; sometimes, particularly
if you want to stay closer to a 'standard' test with standard test statitics, 
it may be easier to rely on a pre-packaged implementation. The `infer` package
makes this type of computational inference quite easy and supports both 
parametric, bootstrap, and permutation based approaches to inference. See
[this vignette](https://infer.tidymodels.org/articles/observed_stat_examples.html)
for more.

# Tidymodels {#tidymodels}

The `tidymodels` package provides an elegant interface for applying many 
"black box" machine learning models in a standardized manner. This is especially
useful when you want to compare many different models to see which performs the
best on a prediction task.[^sta9890]

[^sta9890]: As someone who has dedicated large part of my career to the 
subtle _differences_ between these models, this sort of 'easy swappability'
rubs me the wrong way, but it is _immensely_ useful in applied work. If you do 
want to learn what these models actually do - and when one should be preferred
over the other - I'd encourage you to take STA 9890 and STA 9891, which cover
these types of "ML" models in more detail.

In these notes, we're going to use `tidymodels` for two tasks that may be
useful as you finish your final projects: 

i)  Building a model that takes in several features and uses them to predict
    an output of interest
ii) Using both feature importance scores and a _leave-one-covariate-out_ (LOCO)
    strategy to determine which features are most important. 
    
## Tidymodels Basics

`tidymodels` provides a unified framework for building _predictive_ models. 
In this context, we are _solely_ focused on out-of-sample prediction and aren't
paying attention to things like coefficients or statistical significance. 
This single-minded focus makes it possible for `tidymodels` to "abstract" over
different types of machine-learning models, using them only as prediction 
machines.[^bb]

[^bb]: This type of framework uses models as 'black-boxes' because the
user is not required to, nor able to, interrogate the parameters of the model.
Indeed, for some of the most powerful and useful models we will consider, the
model does not have parameters in the classical sense. 

We will demonstrate the use of the `tidymodels` ecosystem[^tm] to build a predictive 
model on a data set of NYS public school test results. I have prepared two files
containing relevant data: 

[^tm]: Like the `tidyverse` packages, we will typically load the `tidymodels`
packages through a helper package `tidymodels` that loads several packages for
us. Running `library(tidymodels)` is functionally equivalent to something like: 

    ```{r}
#| message: false
#| warning: false
library(dials)
library(broom)
library(hardhat)
library(parsnip)
library(recipes)
library(rsample)
library(tune)
library(workflows)
library(yardstick)
    ```

- [School-Level Covariates (`schools_covariates.csv`)](../schools_covariates.csv)
- [School-Level Test Results (`schools_test_results.csv`)](../schools_test_results.csv)

```{r}
#| echo: false
#| message: false
library(tidyverse)
COVARIATES <- read_csv("schools_covariates.csv")
RESULTS <- read_csv("schools_test_results.csv")
```

The first file, `schools_covariates.csv`, has data on `r NROW(COVARIATES)` 
schools, covering information about class size, the number of teachers and 
support staff, the percentage of students who quality for discounted or free
lunches, and funding levels.[^cov] The latter file, `school_test_results.csv`,
has `r NROW(RESULTS)` test results, indexed by the school, the student subgroup
(*e.g.*., all male students or all non-white students), and the test subject.
The variable we seek to predict is `PERCENT_PROFICIENT`, the fraction of 
students who achieved a 'proficient' score on that test.[^scores]

[^cov]: There are `r NCOL(COVARIATES)` school covariates in total, but we won't 
use them all in this exercise. 

[^scores]: While these are percentages, they are scaled up by 100, so a value
of "50" indicates that half (50%) of students were proficient. 

::: {.callout-note title="Additional Background" collapse="true"}

Persuant to the [Every Student Succeeds Act (ESSA) of
2015](https://en.wikipedia.org/wiki/Every_Student_Succeeds_Act) and its 
predecessor, the [No Child Left Behind Act (NCLB) of
2001](https://en.wikipedia.org/wiki/No_Child_Left_Behind_Act), all public 
schools in New York State are required to administer standardized examinations
annually to their students. Students' results on these tests are used for a
variety of purposes, including assessing teacher and school performance, 
providing additional recognition to high-performing schools, and identifying
struggling students or subgroups. The 
[New York State Education Department (NYSED)](https://www.nysed.gov/)
makes aggregate test results available through its annual 
["School Report Card"](https://data.nysed.gov/) data
releases. For this exercise, we will predict the percentage of students at each 
school who score in the "Proficient" range (*i.e.*, at or above grade level) on 
a wide battery of standardized tests during the 2024 testing period. 

New York administers two primary sets of standardized exams:[^alts]

- Grade 3-8 Exams: 
  - English Language Arts ("ELA") given every year 
  - Mathematics given every year
  - Science, typically given only in grade 5 and grade 8
- [Regents Exams](https://en.wikipedia.org/wiki/New_York_Regents_Examinations) 
  for high school students

The Regents exams cover a broader variety of topics and, historically, have been
a graduation requirement in NYS. For each exam, a "proficiency" level is 
determined: *e.g.*, if a exam gives scores from "1" to "4", levels "3" and "4"
may be considered "proficient". Your task is to predict the fraction of students
who achieve proficiency. 

The organization of public schools in New York State is, like everything in New
York State, rather confusing. As of the time of writing, NYSED reports: 

- 730 Districts
- 4,394 Public Schools
- 373 Charter Schools
- 215,701 Public School Teachers
- 2,421,491 Public School Students

in the state. These districts range from very small, *e.g.*, [Newcomb
CSD](https://data.nysed.gov/profile.php?instid=800000051865) with 47 students to
very large, *e.g.*, [NYC Public 
Schools](https://data.nysed.gov/profile.php?instid=7889678368) with just under 
a million students. 

Outside of NYC, public schools are organized into local school districts, which
typically have (at least) one elementary school, one middle school, and one
high school.[^small] These districts are further organized into [Boards of
Cooperative Educational Serices "BOCES"](https://www.boces.org/) which provide
shared resources to groups of smaller districts.  Local school districts are 
principally funded by property taxes, with more affluent districts having a 
larger tax base and providing higher per-student funding levels. Local school
districts are run by an elected school board, following statewide standards. To
address differences in funding, NYSED provides supplemental funding to districts
based on their assessed ["Needs/Resource
Capacity"](https://p12.nysed.gov/irs/accountability/2011-12/NeedResourceCapacityIndex.pdf),
though [work is ongoing](https://rockinst.org/wp-content/uploads/2024/12/2024-12-Foundation-Aid-Report.pdf)
to update and refine the current funding formula. 

Within NYC, public schools are run by the [NYC Department of Education (NYC
DOE)](https://en.wikipedia.org/wiki/New_York_City_Department_of_Education), 
under a commissioner and [school
board](https://en.wikipedia.org/wiki/New_York_City_Panel_for_Educational_Policy)
appointed by the mayor and borough presidents. NYCDOE is subdivided into 
[32 geographic school districts](https://www.schools.nyc.gov/docs/default-source/default-document-library/nycdoe-district-map.pdf)
and two citywide districts providing specialized services for students with 
disabilities or students in alternative education programs.

[^alts]: We are ignoring alternative assessments like the
[NYSAA](https://www.nysed.gov/state-assessment/new-york-state-alternate-assessment)
for this exercise. 

[^small]: Very small districts may consolidate into two (or even one) school
serving a wider variety of grades or arrange for their students to be educated
within a neighboring district. 

:::

Let's read these files into `R`, join them together, and remove some identifiers
that aren't relevant to us: 

```{r}
#| message: false
library(tidyverse)
COVARIATES <- read_csv("schools_covariates.csv")
RESULTS <- read_csv("schools_test_results.csv")

DATA <- inner_join(RESULTS, 
                   COVARIATES, 
                   join_by(SCHOOL == SCHOOL)) |> 
    select(-ASSESSMENT_ID)
```

In the interest of speed, we're only going to use a few features and a subset
of the entire data set, though all of the code shown here can be scaled up to 
use all of the rows and columns in this data set: 

```{r feature_engineering_subsample_data}
#| cache: true
DATA <- DATA |> 
    mutate(
        # Lunch discount as a proxy for family income
        LUNCH_DISCOUNT = PERCENT_FREE_LUNCH + PERCENT_REDUCED_LUNCH, 
        # Student-teacher-ratio
        STUDENT_TEACHER_RATIO = N_PUPILS / NUMBER_OF_TEACHERS, 
        # Does more funding actually help?
        FUNDING_PER_STUDENT = LOCAL_FUNDING_PER_PUPIL + 
            FEDERAL_FUNDING_PER_PUPIL) |>
    select(
        # Outcome of interest
        PERCENT_PROFICIENT, 
        # Population of interest: which test and which group of students?
        SUBGROUP_NAME, 
        ASSESSMENT_NAME, 
        # Three constructed features
        LUNCH_DISCOUNT, 
        STUDENT_TEACHER_RATIO, 
        FUNDING_PER_STUDENT) |> 
    slice_sample(n=10000)
```


::: {.callout-caution title="Simplified Modeling"}

The model demonstrated here ignores much of the interesting structure (schools
within districts within regions, *etc*) of this data that a more effective model
would take advantage of. The simple model we show here is intended to 
demonstrate the use of `tidymodels`, not to be the last word in modeling this
type of data.

:::

## Prediction Models

For this demonstration, we are going to build a _random forest_ (RF) predictor. 
Random forests are a powerful and flexible machine learning tool that can be
fit efficiently and don't require too much tuning.

`tidymodels` requires us to specify a few separate components of our model 
fitting pipeline: 

- A _recipe_ describing how the prediction task is specified and any data
  pre-processing that is performed. This is the stage in which we specify which
  variables are to be used as predictors and which is our outcome. We also 
  perform steps like data standardization, imputing missing values, handling
  categorical features, removing 'low information features', *etc.* here. 
- A _model_ specification indicating what type of model we want to use. At this
  stage, in addition to specifying the _type_ of model, we also have the option
  to specify any 'knobs' (hyperparameters) that control model behavior and, for
  models with multiple implementations, specify which package we want to use.
- The range of tuning parameters (if any) to consider
- Data used to tune and evaluate the model

Let us build these each in turn. We first build a `recipe()` for our task, a 
set of steps that define the roles of each variable (predictor or response) and
any pre-processing, if necessary. In this case, we don't need much 
pre-processing as we're using a model family that's quite robust. 

```{r}
#| message: false
library(tidymodels)

schools_proficiency_task <- recipe(PERCENT_PROFICIENT ~ ., data=DATA)
```

Here, we simply specify that we want to predict one column (`PERCENT_PROFICIENT`)
using all other features (`.`). At this point, we have specified our goal, but
we still need to specify the manner (model) by which we hope to achieve that goal.

For this task, we're going to specify a _random forest_ model: 

```{r}
rf_model <- rand_forest() |> 
    set_mode("regression") |> 
    set_engine("ranger")
```

Here, we have specified a basic random forest and set two properties: 

- the `mode` specifies what type of random forest we are fitting. Here we are
  predicting a continuous response rather than a categorical or discrete one,
  so our task is 'regression' in the jargon of machine learning. 
- the `engine` specifies which `R` package we want to use. Here, we're using the
  [`ranger` package](https://imbs-hl.github.io/ranger/) which provides a 
  high-performance and high-quality implementation of the random forest framework.
  
At this point, we can define a `workflow()` object that will put these two together:

```{r}
schools_rf <- workflow() |> 
    add_recipe(schools_proficiency_task) |> 
    add_model(rf_model)
```

Finally, we can use this `workflow`  to fit our model to data: 

```{r fit_full_rf}
#| cache: true
rf_fitted <- schools_rf |> fit(data=DATA)
```

At this point, we have a model that could be used to make predictions: 

```{r}
augment(rf_fitted, new_data=DATA) |> head(n=5)
```

Indeed, we could even compute a set of accuracy measures for this problem:[^err]

[^err]: It is not clear to me why we have to specify the `estimate` and `truth`
columns automatically here. It seems like the kind of thing that should be 
inferred from the `recipe`, but it's the way it is (for now). 

```{r}
metrics <- metric_set(
    # Root Mean Squared Error
    rmse, 
    # R-Squared
    rsq, 
    # Mean Absolute Error
    mae, 
    # Mean Absolute Percentage Error, 
    mape)

augment(rf_fitted, new_data=DATA) |> 
    metrics(truth=PERCENT_PROFICIENT, 
            estimate=.pred)
```

So that last row -- "percent error" -- goes a bit janky just because we have
`r DATA |> filter(PERCENT_PROFICIENT == 0) |> NROW()` tests on which 0% of
students were at grade level.  We can exclude these from our analysis and use
other `tidyverse` techniques to see how our model performs on different tests: 

```{r}
library(gt)
rf_fitted |> 
    augment(new_data=DATA) |> 
    filter(PERCENT_PROFICIENT > 0) |> 
    group_by(ASSESSMENT_NAME) |> 
    metrics(truth=PERCENT_PROFICIENT, 
            estimate=.pred) |> 
    select(-.estimator) |> 
    pivot_wider(names_from=.metric, values_from=.estimate) |>
    mutate(mape = mape/100) |> 
    gt(id="tbl_exams_1", rowname_col = "ASSESSMENT_NAME") |>
    cols_label(rmse="Root Mean Squared Error", 
               rsq="R^2", 
               mae="Mean Absolute Error", 
               mape="Mean Absolute Percent Error") |>
    fmt_percent(c(rsq, mape)) |> 
    fmt_number(c(rmse, mae))
```

This is good, as far as it goes, but we have violated *several* ML 
best practices. In fact, the `new_data` argument to `augment()` should have
been a bit of a warning sign. Here, we are evaluating the model on the data
used to train it. 

A better practice is to hold some of our data aside and use it for model
evaluation. This can be done in several ways, but here we're going to use a
technique known as 
[cross-validation](https://en.wikipedia.org/wiki/Cross-validation_(statistics))
wherein we split the data into several pieces and fit the model several times,
holding out one piece for evaluation each time. 

This is conceptually straightforward, but the book-keeping can be a bit 
of a pain. Thankfully, `tidymodels` is available to help us: 

```{r fit_rf_cv}
#| cache: true
schools_cv <- vfold_cv(DATA, v=5)

schools_rf_cv <- schools_rf |> fit_resamples(schools_cv)
```

Note that here we use `fit_resamples()`  instead of a plain `fit()` to
account for this repeated fitting and evaluation structure. In this case, 
we also see that several metrics were automatically computed for us

```{r}
schools_rf_cv |> select(id, .metrics) |> unnest(.metrics) |> head(n=5)
```

If we want to use all four metrics from before, we can pass the
`metrics` argument to `fit_resamples()`. This gives us a sense of 
error bars on our metrics, which we can compute by hand or with the
help of the `collect_metrics()`: 

```{r fit_rf_cv2}
#| cache: true
schools_rf |> fit_resamples(schools_cv, metrics = metrics) |> collect_metrics()
```

If we want to get group-wise statistics as before, we still need to do
a bit of work by hand. 

```{r fit_rf_cv3}
#| cache: true
#| message: false
#| warning: false
schools_rf |> 
    fit_resamples(schools_cv, 
                  metrics = metrics, 
                  control = control_resamples(save_pred = TRUE)) |>
    collect_predictions() |> 
    arrange(.row) |> 
    inner_join(DATA |> mutate(.row=row_number()), 
               join_by(.row, 
                       PERCENT_PROFICIENT)) |> 
    filter(PERCENT_PROFICIENT > 0) |> 
    group_by(ASSESSMENT_NAME, id) |>
    metrics(truth=PERCENT_PROFICIENT, 
            estimate=.pred) |> 
    group_by(ASSESSMENT_NAME, 
             .metric) |> 
    summarize(mean_metric = mean(.estimate), 
              sd_metric = sd(.estimate), 
              se_metric = sd_metric / sqrt(n())) |>
    ungroup() |>
    select(-sd_metric) |> 
    pivot_wider(values_from=c(mean_metric, se_metric), 
                names_from=.metric) |> 
    mutate(mean_metric_mape = mean_metric_mape / 100, 
           se_metric_mape = se_metric_mape / 100) |>
    gt(id="tbl_exams_2", 
       rowname_col = "ASSESSMENT_NAME") |> 
    cols_label(mean_metric_rmse="Root Mean Squared Error", 
               mean_metric_rsq="R^2",
               mean_metric_mae="Mean Absolute Error", 
               mean_metric_mape="Mean Absolute Percent Error") |>
    fmt_percent(c(mean_metric_rsq, se_metric_rsq, 
                  mean_metric_mape, se_metric_mape)) |>
    fmt_number(c(mean_metric_rmse, se_metric_rmse, 
                  mean_metric_mae, se_metric_mae)) |>
    cols_merge(c(mean_metric_rmse, se_metric_rmse), pattern="{1}±{2}") |>
    cols_merge(c(mean_metric_rsq, se_metric_rsq), pattern="{1}±{2}") |>
    cols_merge(c(mean_metric_mae, se_metric_mae), pattern="{1}±{2}") |>
    cols_merge(c(mean_metric_mape, se_metric_mape), pattern="{1}±{2}") |> 
    sub_missing() |>
    tab_header("Prediction Error Metrics by Exam", 
               "CV Prediction Error of a Random Forest Model Predicting Student Exam Proficiency Using Funding, SES, and Student-Teacher Ratios")
```

Not surprisingly, our performance here is worse than we saw above, because we 
are now getting 'out-of-sample' results.

This is rather messy, but it gives us a good sense of how well our model can
do on different exams. In general, we see that we are 10-15% off (RMSE) using
these three features, so our model isn't great yet. 

## LOCO Inference

With our model in hand, we now want to ask which of our three big features 
is most important. In general, when assessing feature importance we have a
choice between methods _intrinsic_ to the model (*e.g.*, regression
coefficients for OLS) or methods _extrinsic_ to the model. Extrinsic methods
are necessarily less specific, but useful because they can be compared across
different models. 

Here, we're going to demonstrate the use of the _extrinsic_ model known as 
*Leave-One-Covariate-Out* (LOCO) feature importance. In brief, LOCO compares
the model with and without some feature and compares the predictic performance.
If the performance changes significantly, the feature was important; if the
performance doesn't change much, _either_ the model didn't really use the
feature or the value of that feature could be replaced by other correlated
features; either way, a small change (low LOCO score) indicates that the
feature is not that important. 

The following code can be used to compute the LOCO scores here: 

```{r}
calc_loco <- function(workflow_obj, data, metrics = NULL, 
                      loco_groups = NULL, v = 5) {
    # Step 1: Extract LOCO Features
    # 
    # We want loco-groups to be a list of features (or groups of features) we can
    # iterate over. This does a bit of pre-processing to automatically produce
    # the list if not provided or to convert a vector to a list of singletons
    # if a character vector is provided
    if (is.null(loco_groups)) {
        # If the loco_groups aren't provided, let's just pull out each predictor
        # from the recipe component of the workflow object and use those as our
        # features to LOCO
        loco_groups <- workflow_obj |> 
            extract_recipe(estimated=FALSE) |> 
            summary() |> 
            filter(role == "predictor") |> 
            pull(variable) |> 
            set_names() |> 
            as.list()
    } else if (is.character(loco_groups)) {
        loco_groups <- loco_groups |> 
            set_names() |> 
            as.list()
    }
    
    # Step 2 - Set up resampling. If given an rset (resampling set) object, just
    # use that. Otherwise, automatically perform v-fold CV (v from above) internally
    folds <- if (inherits(data, "data.frame") && !inherits(data, "rset")) {
        vfold_cv(data, v = v)
    } else {
        data
    }
    
    # Step 3. Implementing LOCO. 
    #
    # This function will take a list of variables (if provided) and temporarily
    # remove them from the predictors (marking them as 'LOCO' role instead) 
    # and re-fit the model. tidymodels will automaticaly ignore LOCO-marked 
    # variables, giving us the predictions we need to get our LOCO scores. 
    # 
    # This function only takes variables to drop - everything else is captured
    # dynamically from the enclosing environment
    run_loco <- function(drop_vars) {
        if(is.null(drop_vars)) {
            message("Running Full Model (LOCO Baseline)")
            loco_wflow <- workflow_obj
        } else {
            message(paste("Running LOCO Model - Omitting", drop_vars))
            wflow_recipe <- workflow_obj |> 
                extract_recipe(estimated = FALSE) |> 
                update_role(all_of(drop_vars),
                            new_role="LOCO")
            loco_wflow <- workflow_obj |> update_recipe(wflow_recipe)
        }
        
        # Fit model to resampled data and get results
        loco_wflow |>
            fit_resamples(resamples = folds, 
                          metrics = metrics) |> 
            collect_metrics() |> 
            select(-.estimator, -.config) |> 
            mutate(direction = map_chr(.metric, \(m) get(m) |> attr("direction")))
    }
    
    # Step 4. Showtime. First fit the full model, then iterate through loco_groups
    # performing each LOCO
    full_results <- run_loco(NULL) |> 
        mutate(LOCO_feature="FULL_MODEL") |>
        select(-n, -std_err, -LOCO_feature)
    loco_results <- imap(loco_groups, \(x, y) run_loco(x) |> mutate(LOCO_feature=y)) |>
        list_rbind() |> 
        select(-n, -std_err)
    
    inner_join(loco_results, 
               full_results, 
               join_by(.metric, direction), 
               suffix=c(".LOCO", ".FULL")) |> 
        mutate(loco_value = mean.FULL - mean.LOCO, 
               loco_value = if_else(direction == "minimize", -1, 1) * loco_value) |> 
    rename(loco_score = loco_value, 
           metric=.metric, 
           full_model=mean.FULL, 
           loco_model=mean.LOCO)
}

```

We can apply this to get LOCO scores for this model: 

```{r fit_loco}
#| cache: true
schools_loco <- calc_loco(schools_rf, schools_cv)
```

which gives us: 

```{r}
#| message: false
#| warning: false
#| echo: false
schools_loco |> 
    mutate(across(where(is.numeric), \(x) round(x, 2))) |> 
    mutate(LOCO = glue("{loco_score} ({full_model} -> {loco_model})")) |>
    select(metric, LOCO, LOCO_feature) |>
    pivot_wider(names_from=metric, 
                values_from=c(LOCO)) |> 
    gt(id="tbl_loco") |> 
    cols_label(LOCO_feature="Dropped Feature", 
               rmse="Root Mean Squared Error", 
               rsq="R^2")
```

From this, we see that knowing the subject and grade (via `ASSESSMENT_NAME`) 
is the most important information in predicting performance, followed very 
closely by family socioeconomic status (`LUNCH_DISCOUNT`). By contrast, 
student demographics (via `SUBGROUP_NAME`) or school properties 
(`FUNDING_PER_STUDENT` and `STUDENT_TEACHER_RATIO`) seem much less important. 

While powerful - this LOCO framework has a few limitations to be aware of: 

1) It can only measure importance *vis-a-vis* the type of model being fit, here
   a random forest and is not an abstract 'model-free' measure of importance.[^mf] 
   Theoretically, some other model could take better advantage of a feature that
   random forests can't really make much use of in this data. 
2) LOCO scores are only meaningful insofar as the model itself is predictive. 
   Knowing how useful a feature is to a useless model is not actually valuable.
3) Fitting LOCO can be quite computationally intensive: if we have $p$ features
   and use 5-fold CV, we have to fit $5(p+1)$ models. If these are slow, this
   can start to add up. To improve performance, we can either parallelize at
   the model fitting level (`set_engine("ranger", num.threads=4)`) or at the 
   cross-validation level by registering a parallelization system and letting
   `tidymodels` pick it up automatically. 
   
[^mf]: Scores like LOCO are sometimes called 'model-free', but I prefer to call
them 'model-agnostic' or 'model-flexible'. They can be applied to any single
model at a time, but their results still depend on the model choice. 

## LOCO in Action

Apply this LOCO technique to the Hotel Prediction case study from the 
[`tidymodels` tutorial](https://www.tidymodels.org/start/case-study/).

For your convenience, I have copied and simplified the relevant parts of the 
case study below: 

```{r}
#| cache: true
#| message: false
library(tidyverse)
library(tidymodels)

hotels <- read_csv("https://tidymodels.org/start/case-study/hotels.csv") |>
  mutate(across(where(is.character), as.factor))

hotel_splits <- initial_validation_split(hotels, 
                                         strata = children,
                                         prop = c(0.60, 0.15))

# TODO: Clarify how tidymodels is using these various splits.
hotel_train <- training(hotel_splits)
hotel_val   <- validation(hotel_splits)
hotel_test  <- testing(hotel_splits)
hotel_other <- bind_rows(hotel_train, hotel_val)

val_set <- validation_set(hotel_splits)

holidays <- c("AllSouls", "AshWednesday", "ChristmasEve", "Easter", 
              "ChristmasDay", "GoodFriday", "NewYearsDay", "PalmSunday")

hotel_recipe <- recipe(children ~ ., data = hotel_other) |> 
  step_date(arrival_date) |> 
  step_holiday(arrival_date, holidays = holidays) |> 
  step_rm(arrival_date) 

hotel_recipe_one_hot <- hotel_recipe |> 
    step_dummy(all_nominal_predictors())|> 
    step_zv(all_predictors()) |> 
    step_normalize(all_predictors())

lr_mod <- logistic_reg(penalty = tune(), 
                       mixture = 1) |> 
  set_engine("glmnet")

lr_workflow <- workflow() |> 
  add_model(lr_mod) |> 
  add_recipe(hotel_recipe_one_hot)

lr_reg_grid <- tibble(penalty = 10^seq(-4, -1, length.out = 30))

lr_res <- lr_workflow |> 
  tune_grid(val_set,
            grid = lr_reg_grid,
            control = control_grid(save_pred = TRUE),
            metrics = metric_set(roc_auc))

rf_mod <- rand_forest(mtry = tune(), 
              min_n = tune(), 
              trees = 200) |> 
  set_engine("ranger", 
             num.threads = parallel::detectCores()) |> 
  set_mode("classification")

rf_workflow <- workflow() |> 
  add_model(rf_mod) |> 
  add_recipe(hotel_recipe)

rf_res <- rf_workflow |> 
  tune_grid(val_set,
            grid = 25,
            control = control_grid(save_pred = TRUE),
            metrics = metric_set(roc_auc))

lr_best_params <- lr_res |> select_best(metric="roc_auc")
lr_best_workflow <- lr_workflow |> finalize_workflow(lr_best_params)

rf_best_params <- rf_res |> select_best(metric="roc_auc")
rf_best_workflow <- rf_workflow |> finalize_workflow(rf_best_params)
```

1. Apply LOCO inference to the `lr_best_workflow` to determine the most 
   important features to logistic regression. For data, use `hotel_train`. 
   
::: {.callout-tip title="Solution" collapse="true"}

```{r lr_loco}
#| message: false
#| cache: true
lr_loco_results <- calc_loco(lr_best_workflow, hotel_train, v=3)
```

In a table, 

```{r}
lr_loco_results |> 
    select(-full_model, -loco_model, -direction) |> 
    pivot_wider(names_from=metric, values_from=loco_score) |> 
    arrange(desc(roc_auc)) |>
    gt(id="tbl_loco_lr") |> 
    cols_label(accuracy="Classification Accuracy", 
               brier_class="Brier Score", 
               roc_auc="Area Under the ROC Curve", 
               LOCO_feature="Feature") |> 
    tab_header("LOCO Feature Importance", 
               "Larger values signify larger (adverse) shifts in model performance") |>
    tab_spanner("Performance Measures", 
                -LOCO_feature) |>
    fmt_percent(accuracy) |>
    fmt_number(c(brier_class, roc_auc), 
               decimals=4)
```

:::

Note here that these values are a bit random due to the internal
cross-validation used for LOCO scores, but most of the scores are 0
across several features, consistent with the fact we are performing sparse
logistic regression, *i.e.*, logistic regression with several coefficients 
set to zero.

2. Compare the LOCO feature scores computed above with the regression 
   coefficients of the fitted model. Do they align? If no, why not? 
   
   *Hint*: Use `tidy()` on the result of fitting the workflow to get the 
   coefficients of the selected linear model. 

::: {.callout-tip title="Solution" collapse="true"}

This function is a bit complex, will help coloring columns in a predictable 
and standardized way:

```{r}
color_cols_centered <- function(gt_tbl, ..., palette="RdBu", .multiplier=1.01){
    tbl_body <- gt_tbl |> extract_body()
    ## tidyselect magic to handle the dots and allow helpers and `-` selectors
    column_quosure <- rlang::enquos(...)
    col_names <- names(tidyselect::eval_select(expr(c(!!!column_quosure)), tbl_body))
    
    center_column <- function(gt_tbl, col_name){
        x <- tbl_body |> pull(!!col_name)
        
        suppressWarnings(x_numeric <- as.numeric(x))
        
        if(all(is.na(x_numeric))){
            warning(glue::glue("Could not treat column {col_name} as numeric. 
                               Consider using color_col_centered before fmt_*() functions"))
            return(gt_tbl)
        }
        
        limit <- max(abs(x_numeric), na.rm=TRUE) * .multiplier
        
        gt_tbl |> 
            data_color(!!col_name, 
                       palette=palette, 
                       domain=c(-limit, limit))
    }
    
    reduce(col_names, center_column, .init=gt_tbl)
}
```

With that in hand:

```{r}
#| message: false
lr_coefs <- lr_best_workflow |> fit(hotel_train) |> tidy()

left_join(lr_loco_results, 
           lr_coefs, 
           join_by(LOCO_feature == term)) |>
    select(LOCO_feature, loco_score, metric, coefficient=estimate) |>
    pivot_wider(names_from=metric, 
                values_from=loco_score) |> 
    mutate(abs_coefficient=abs(coefficient)) |>
    arrange(desc(abs_coefficient)) |> 
    gt(id="tbl_loco_lr_coefs") |>
    cols_label(accuracy="Classification Accuracy", 
               brier_class="Brier Score", 
               roc_auc="Area Under the ROC Curve", 
               LOCO_feature="Feature", 
               coefficient="Logistic Regression Coefficient") |>
    cols_hide(abs_coefficient) |>
    tab_header("LOCO Feature Importance", 
               "Larger values signify larger (adverse) shifts in model performance") |>
    tab_spanner("LOCO Metric Measures", 
                -c(LOCO_feature, coefficient)) |>
    data_color(abs_coefficient,
               target_columns=coefficient, 
               method="numeric", 
               palette="Greens", 
               na_color="white") |>
    color_cols_centered(-c(coefficient, LOCO_feature), palette="RdBu") |>
    fmt_percent(accuracy) |>
    fmt_number(c(brier_class, roc_auc, coefficient), 
               decimals=4)  |>
    sub_missing()
```

Note that we have `NA` values for the regression coefficients for categorical
features. Under the hood, `tidymodels` has encoded these as numerical features
for us to use in modeling, but it is not evident how to construct a single
'coefficient' for display in this table.

A visualization is perhaps a bit more helpful here: 

```{r}
#| warning: false
#| message: false
library(ggrepel)
library(broom)

lr_coefs <- lr_best_workflow |> fit(hotel_train) |> tidy()

inner_join(lr_loco_results, 
           lr_coefs, 
           join_by(LOCO_feature == term)) |>
    select(LOCO_feature, loco_score, metric, coefficient=estimate) |>
    filter(metric == "roc_auc") |>
    ggplot(aes(x=abs(coefficient), 
                y=loco_score, 
                label=LOCO_feature)) + 
    geom_point() + 
    theme_bw() + 
    xlab("Absolute Value of Regression Coefficient") + 
    ylab("LOCO Score [ROC AUC]") + 
    geom_smooth(method="lm", se=FALSE) + 
    geom_text_repel()
```


The question of consistency among importance scores is subtle. One important
factor to be aware of is that predictive importance is a combination of 'how big
is the effect' and 'how much does that covariate change'. For example, in 
comparing the importance of obesity on health outcomes vs that of a rare but 
highly fatal infection, *e.g.*, Ebola. The direct impact of the infection is
larger on an individual, but for America writ large, obesity is a larger 
problem. Any comparison of 'importance' needs to distinguish these two factors
and this is well beyond the scope of this course.[^importance]

[^importance]: For this an _much_ more on the difficulty of having a single
universal measure of 'importance', see 
[this paper](https://arxiv.org/pdf/2505.15728), which demonstrates challenges
of consistency empirically, or [this
paper](https://www.annualreviews.org/content/journals/10.1146/annurev-statistics-040120-030919)
which wrestles with the problem from a more theoretical point-of-view.


:::

3. Apply LOCO inference to the `rf_best_workflow` to determine the most 
   important features to a random forest model. For data, use `hotel_train`. 
   
::: {.callout-tip title="Solution" collapse="true"}

```{r rf_loco}
#| message: false
#| cache: true
rf_loco_results <- calc_loco(rf_best_workflow, hotel_train, v=3)
```

As before, it is interesting to compare these results with the regression 
coefficients and LOCO scores arising from our logistic regression model:

```{r}
lr_loco2 <- lr_loco_results |> select(metric, lr_loco_score=loco_score, LOCO_feature)
rf_loco2 <- rf_loco_results |> select(metric, rf_loco_score=loco_score, LOCO_feature)

inner_join(lr_loco2, rf_loco2, join_by(LOCO_feature, metric)) |>
    left_join(lr_coefs, join_by(LOCO_feature == term)) |>
    select(-penalty) |> 
    rename(coefficient=estimate) |>
    pivot_wider(names_from=metric, 
                values_from=c(lr_loco_score, rf_loco_score)) |>
    mutate(abs_coefficient=abs(coefficient)) |>
    arrange(desc(abs_coefficient)) |>
    gt(id="tbl_lr_rf_loco") |> 
    cols_label_with(contains("accuracy"), \(x) "Classification Accuracy") |>
    cols_label_with(contains("roc_auc"), \(x) "AUC ROC") |>
    cols_label_with(contains("brier_class"), \(x) "Brier Score") |>
    cols_label(coefficient="Logistic Regression Coefficient", 
               LOCO_feature="Feature") |> 
    tab_spanner("Logistic Regression LOCO", 
                contains("lr")) |>
    tab_spanner("Random Forest LOCO", 
                contains("rf")) |>
    cols_hide(abs_coefficient) |>
    data_color(abs_coefficient,
               target_columns=coefficient, 
               method="numeric", 
               palette="Greens", 
               na_color="white") |>
    color_cols_centered(-c(coefficient, LOCO_feature), 
                        palette="RdBu") |>
    fmt_percent(contains("accuracy")) |>
    fmt_number(matches("brier_class|roc_auc|coefficient"), 
               decimals=4) |>
    sub_missing() |>
    tab_header("Comparison of Regression Coefficients and LOCO Feature Importance Scores")
```

:::

## Going Deeper with `tidymodels`

To learn more about `tidymodels`, review the introductory case study at
<https://www.tidymodels.org/start/>. 
